### 2019-11-20 paper.rmd
add paper

---
title             : "Getting started with Latent Profile Analysis (LPA) in the psychological sciences: A tutorial using the tidyLPA R package that provides an interface to open-source and commercial software"
shorttitle        : "Title"

author: 
  - name          : "Joshua Rosenberg"
    affiliation   : "1"
    corresponding : yes
    address       : "1122 Volunteer Blvd., Knoxville, TN, 37996"
    email         : "jmrosenberg@utk.edu"
  - name          : "Caspar van Lissa"
    affiliation   : "2"
  - name          : "Jennifer Schmidt"
    affiliation   : "3"
  - name          : "Patrick Beymer"
    affiliation   : "3"
  - name          : "Daniel Anderson"
    affiliation   : "4"
  - name          : "Matthew Schell"
    affiliation   : "3"
  - name          : "Rebecca Steingut"
    affiliation   : "5"
  - name          : "Stephanie Wormington"
    affiliation   : "6"

affiliation:
  - id            : "1"
    institution   : "University of Tennessee, Knoxville"
  - id            : "2"
    institution   : "Utrecht University"
  - id            : "3"
    institution   : "Michigan State University"
  - id            : "4"
    institution   : "University of Oregon"
  - id            : "5"
    institution   : "Columbia University"
  - id            : "6"
    institution   : "University of Virginia"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |

keywords          : "Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust"
wordcount         : "4,882 (need to shorten to ~ 3,000)"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

The questions on the Scimo per and post survey are from Ryan and Deci's "Learning Climate Questionnaire".  You can read a short description here:

https://selfdeterminationtheory.org/pas-learning-climate/

It is 15 questions, but they identify 6 which can be used as  short-form.  Questions 1, 2, 4, 7, 10, 14.  I have attached the PDF of the survey since you have to create an account on their website do download it.  Below are the question text, along with the variable name in the Scimo data:

I feel that my instructor provides me choices and options.  t_choices1
I feel understood by my instructor.  t_understand1
My instructor conveyed confidence in my ability to do well in the course.  t_confidence1
My instructor encouraged me to ask questions.  t_encourage1
My instructor listens to how I would like to do things.  t_listen1
My instructor tries to understand how I see things before suggesting a new way to do things.  t_triesunder1

Also, is this tutorial going to be a journal paper, or something on github, etc. as a kind of manual for using LPA?  If so, and you think I would be a good fit, I would love to help out on it if you have author space!

Let me know if you need anything else.

<!-- tidyLPA Tutorial for AMPPS -->
<!-- https://www.psychologicalscience.org/publications/ampps/ampps-submission-guidelines#NEMP -->
<!-- Tutorials -->
<!-- Tutorials provide hands-on, practical guidance for researchers. Any topic that could enhance research practices or methods might be suitable for a Tutorial, provided that the material covered in the Tutorial would be relevant to and useful for the journal’s broad readership; Tutorials of more narrow interest or of relevance to only one subfield or literature typically are not appropriate for AMPPS. AMPPS welcomes Tutorials that focus on helping researchers learn to use statistical tools, improve their statistical practices and intuitions, better their data-management and lab practices, enhance the reliability and reproducibility of their research, engage in transparent and open practices, and so on. Tutorials often include dynamic, interactive content and should provide concrete guidance rather than solely abstract principles. Some Tutorials will be solicited by the editorial team, and the team welcomes suggestions and proposals for Tutorials. -->
<!-- Most Tutorials should be brief (< 3,000 words), but they may be longer if necessary to explain the content fully and make it accessible to and usable by readers. -->
<!-- The introduction to a Tutorial typically should be no more than one to two paragraphs long (< 500 words) and should not include an extensive literature review. The introduction should explain the motivation for the Tutorial and highlight how learning the contents will benefit readers. -->
<!-- Tutorials should have a brief summary of their contents, rather than a General Discussion section. -->
<!-- Tutorials should be accompanied by publicly available code and all resources necessary for researchers (and reviewers) to follow the text. -->
<!-- Tutorials can include a list of additional resources (e.g., citations and links) for readers who would like to learn more. -->

# Background

Person-oriented analysis involves a statistical approach to exploring psychological constructs in a way that captures and simplifies complexity. This approach can be used to consider the way in which psychological constructs, for example, are experienced together and at once. Thus, conceptually, this approach emphasizes how individuals go about interacting, socializing, and behaving in this way, experiencing these activities as a whole person. Person-oriented analysis, developed within developmental science, focuses on how individuals have experiences in a holistic way and also how these ways can be identified through common groups, or profiles.(Bergman & El-Khouri, 1997; Magnusson & Cairns, 1996). One way to distinguish the approach is to consider it in contrast to variable-centered approaches, such as those for which the covariability between variables, rather than the relations between profiles of individuals, is the focus (Bergman & Trost, 2006).

Though studies examining learning from a person-oriented perspective are not very common, some examples include studies of intrinsic and extrinsic motivation (Corpus & Wormington, 2014; Hayenga & Corpus, 2010), profiles of achievement goals (see Wormington & Linnenbrink-Garcia, advance online publication, for a review), and epistemic cognition (Trevors, Kendeou, Braten, & Braasch, 2017). Recently, too, there have been excellent reviews of person-oriented analyses in motivation and educational psychology (Linnenbrink-Garcia & Wormington, 2017).

In this article, our goal is to review person-oriented analysis focused on getting started (on both knowing about and doing person-oriented analyses) in an informed way. More specifically, we start with an overview of the methodology, describe specific software tools, explain analytic choices made in the course of their use, and conclude with general recommendations One challenge facing those trying to learn about and apply person-oriented analysis in their research is the multitude of terms used for similar analytic approaches or ideas. Throughout this manuscript, we aim to point these out and explain when and why a variety of terms have been used. Thus, this manuscript is written for an informed and curious reader at any level interested in understanding this particular approach.

## Overview of Methodologies

At present, there are a number of approaches to carrying out person-oriented analysis when the indicators, of variables, used to find the groups are continuous. This section is focused upon two approaches: cluster analysis and Latent Profile Analysis (LPA). We focus on these because they are relatively simple and accessible forms of person-oriented analysis that can also accommodate cross sectional and longitudinal data alike. Just as a brief point of contrast, when the indicators for the groups are dichotomous (or categorical and can therefore be “dummy-coded” to be dichotomous), an approach closely related to LPA, Latent Class Analysis, can be used (cite PSU folks).

## Cluster Analysis

Cluster analysis is an approach that has been frequently applied to analysis of topics in educational psychology (such as motivation and engagement) from a more holistic, whole-person approach. One particular way in which cluster analysis has been used has been as part of an approach termed I-States-as-Objects Analysis (ISOA). The purpose of ISOA is to identify profiles at specific time points, and then to explore how individuals do or do not shift between profiles across time points (needs cite). In their early work, Bergman and Magnusson (1999) and colleagues (1999) described a two-step approach to cluster analysis as part of their I-States-as-Objects analysis. In this approach, all of the observations are clustered. The key idea behind the approach is that each observation represents not a permanent group, but rather a status, or an i-state. These i-states, are considered independently for the purposes of constructing profiles. Then, shifts between these i-states across time (say, two or three time points) are evaluated.

Commonly, cluster analysis carried out as part of ISOA in fact involves two distinct clustering steps. This is because different clustering algorithms have different strengths--as well as weaknesses. In ISOA, the first step is Ward’s hierarchical cluster analysis (Bergman & El-Khouri, 1999). In this step, observations are split, step-by-step, based upon their multivariate similarity. The analyst provides the number of groups which will be used and interpreted, though technically most algorithms will continue splitting the groups until each observation makes up its own. This type of clustering is deterministic: The same results will be obtained each time it is run. In the second step of the cluster analysis, an iterative, k-means cluster analysis is carried out (Bergman and El-Khouri, 1999). For this step, the algorithm partitions the observations into groups, the success of the clustering based on a criterion such as the variability in the data explained by the groups, and then “moves,” when possible, the observations that are not in the group they are most similar to to the one that they are most similar to. When no more such moves are possible, the algorithm stops. On its own, k-means is not deterministic: Depending on the starting values, it can obtain different results the different times that it is run. So, the results from the hierarchical cluster analysis, the first step, are used as start values to initialize the k-means clustering.

The rationale behind using these two approaches is that the hierarchical clustering provides reasonable guesses for the profile membership, and then the iterative clustering optimizes the clustering solution. To sum up, then, cluster analysis as part of ISOA in fact involves two separate cluster analyses. The merits of this approach are to leverage the benefits of each: the iterative clustering starts with reasonable cluster solutions generated from the hierarchical clustering and optimizes them.

## Latent Profile Analysis

Another approach what is becoming increasingly common for carrying out person-oriented analysis is Latent Profile Analysis (LPA). LPA is a type of finite mixture model (Harring and Hodis, 2017). The goal of this approach is estimate the parameters for a number of distributions (often multivariate) from a single dataset. The approach can be thought of as seeking an answer to the question, Do the data in a sample come from more than one population? Thus, such an approach is model-based, and some descriptions in the literature refer to it as model-based clustering (Hennig, Meila, Murtagh, & Rocci, 2015; Scrucca, Fop, Murphy, & Raftery, 2017). Thus, one distinction between LPA and cluster analytic approaches is that LPA is model-based; instead of using algorithms to group together cases, LPA seeks to estimate parameters - in terms of variances and covariances and how they are the same or different across profiles - that best characterize the different distributions. Then, this approach seeks to assign to each observation a probability that the observation is a sample from the population associated with each profile (or mixture component).

Because LPA is model-based, a number of different model parameterizations can be estimated. These models differ in terms of whether--and how--parameters are estimated across the profiles. These parameters are the means for the different profiles, which, in this approach, always are estimated freely across the profiles; the variances for the variables used to create the profiles, which can be estimated freely or can be estimated to be the same, or equal, across profiles; and the covariances of the variables used to create the profiles, which can be freely-estimated, estimated to be equal, or fixed to be zero. One challenge facing the analyst using LPA is that these parameters and the distinct model parameterizations that can be estimated is the different terminology used. As one example, Scrucca et al. (2017) refer to these parameterizations not in terms of whether and how parameters are estimated, but rather in terms of the geometric properties of the distributions that result from particular parameterizations. Muthen and Muthen (1997-2017) and others (Pastor et al., 2007) commonly refer to local independence to mean that the covariances are fixed to zero (also described as the specification of the covariance matrix as “diagonal,” because only the diagonal components, or the variances, are estimated). More information on the model parameterizations are discussed in the context of the software tool tidyLPA that we have developed.

 This section outlines several LPA models and guidelines/examples for when they might be most appropriate in person-centered analysis. In general, as more parameters are estimated (i.e., those that are fixed to zero are estimated as being equal across profiles; or those estimated as being equal across profiles are freely-estimated across them), the model becomes more complex; the model may fit better, but also be overfit, meaning that the profiles identified may be challenging to replicate with another, separate data set. Even still, flexibility in terms of which models can be estimated also has affordances. For example, the varying means, equal variances, and covariances fixed to 0. A researcher might choose this model specification if she wants to model the variables to be used to create profiles that are independent. This model is very simple, as no covariances are estimated and the variances are estimated to be the same across profiles.  As we estimate more parameters (and decrease the degrees of freedom), we are more likely to fit the data, but less likely to be able to replicate the model with a second set of data.  In other words, more parameters may mean a loss of external validity. As we progress toward more complex models (with increasingly complex parameterization, i.e. models 3 and 4 above), then we are more likely to fit the data better.

    A number of notes regarding person-oriented analyses are important to raise at the same time as their promise. Bauer (2007) notes that many samples of data can be usefully broken down into profiles, and that the addition of profiles will likely be suggested for reasons other than the samples coming from more than one distribution (i.e., due to non-normality in the variables measured). Bauer also cautions that profiles should not be reified; that profiles do not necessarily exist outside of the analysis that they should be interpreted more as useful interpretative devices. These cautions suggest that, in general, parsimony, interpretability, and a general sense that the profiles are not necessarily real, but are rather helpful analytic tools, should be both priorities for the analyst and the reader of studies using this approach.

# Specific Software Tools

SPSS is a common tool to carry out cluster analyses. SPSS provides functionality to carry out both hierarchical and k-means cluster analysis. While somewhat straightforward to carry out, particularly in SPSS’s graphical user interface (GUI), there are some challenges to use of this approach. The GUI in SPSS can be challenging, even for the most able analyst, to be able to document every step with syntax, and so reproducing the entire analysis efficiently can be a challenge, both for the analyst exploring various solutions and for the reviewer looking to replicate this work. Additionally, SPSS is commercial software (and is expensive), and so analysts without the software cannot carry out this analysis.

 Accordingly, we worked to develop tools, initially used by the research team and now shared more widely, to provide basic functionality for each of the two primary approaches described in this section, cluster analysis and Latent Profile Analysis (LPA). In addition to being freely-available and open-source (i.e., anyone can inspect the source code), the tools described lend themselves to an open science. It is especially important to have an open transparent methodology when using person-centered approaches because of the high dependence on researcher judgment at multiple points in the process.

Accordingly, we set out to develop packages that a) provided sensible defaults and were easy to use, but provided the option to access and modify all of the inputs to the model (i.e., low barrier, high ceiling), b) interfaced to existing tools,, and are able to translate between what existing tools are capable of and what researchers and analysts carrying-out person-oriented analyses would like to specify, c) made it easy to carry-out fully-reproducible analyses and d) were well-documented. Both of these packages--one for two-step cluster analysis and one for LPA--are designed to do one thing well, namely, quickly and easily estimate profiles using cluster analysis or a model-based approach (Latent Profile Analysis).

One way that tidyLPA is designed to be easy to use is that it assumes a “tidy” data structure (Wickham, 2014). This means that it emphasizes the use of a data frame as both the primary input and output of functions for the package. Because data is passed to and returned (in amended form, i.e., with the latent profile probabilities and classes appended to the data) from the function, it makes it easy to create plots or use results in subsequent analyses. Another noteworthy feature of tidyLPA is that it provides the same functionality through two different tools, one that is open-source and available through R, the mclust package (Scrucca et al., 2017) and one that is available through the commercial software MPlus (Muthen & Muthen, 1997-2017). Moreover, as both tools use the same maximum likelihood estimation procedure, they are benchmarked to produce the same output (see here). Also, note that we have described the model specifications with descriptions of what is estimated in terms of the variances and covariances, the common names for the models (i.e., class-varying unrestricted), and the covariance matrix associated with the parameterizationfor the six models that are possible to be estimated on the website for tidyLPA (see here).

# Analytic Choices

There are a number of analytic choices that need to be made when carrying out person-oriented analyses. Because such person-oriented approaches are often more subjective (in practice) than other approaches (Linnenbrink-Garcia and Wormington, 2017), there is no one rule for determining the solution obtained. This solution is obtained on the basis of multiple decisions, such as the number of profiles selected or the modeling decisions such as what specific options are used for the cluster analysis (i.e., the distance metric used to calculate the similarity of the observations as part of the Ward’s hierarchical clustering) or what parameters are estimated and how as part of LPA.

Given the subjectivity involved, it is important that researchers be transparent and work as part of a team to obtain clustering solutions. Transparency about the design and analytic choices is important so that readers can appropriately interpret the report. Researchers can enhance transparency and reproducibility by sharing detailed descriptions of methodology and document it through the use of syntax (and, if possible, data) that we share with others. Working as part of a team can help to serve as a check on several of the choices researchers make, such as over-fitting or under-fitting the model to the data. Each decision depends on multiple factors and balancing tensions. We discuss each of the key decisions listed in an analysis.

## How the Data are Centering or Scaled

The data can be transformed by centering or scaling. Typically, these are done after the profiles are created, so that differences between profiles can be more easily explored. They can also be done prior to the analysis, which can be helpful for obtaining solutions when the variables are on very different scales..

## Whether to Use Cluster Analysis or LPA

Whether to use cluster analysis or LPA depends in part on the nature of the data being investigated (cite the 2010 debate on this topic between Vermunt and others; MacLachlan, 2011; Steinley & Brusco, 2011a; Steinley & Brusco, 2011b; Vermunt, 2011). In some cases, cluster analysis and LPA solutions are very similar (when a constrained LPA model is fit). Broadly, if the variables used can be considered independent of one another, then cluster analysis may be justified in terms of reaching an optimal solution. However, LPA provides other benefits, such as the availability of information criteria (i.e., the Aikake Information Criteria [AIC] and the Bayesian Information Criteria [BIC] and statistical tests for the number of profiles (when a model is chosen, with a bootstrapped likelihood-ratio test). LPA also results in probabilistic assignments; this information can be used in subsequent analyses rather than the simple assignment (Collins & Lanza, 2010).

## How to Choose the Number of Profiles

In the case of choosing the number of profiles (and the specification of the model / profile solution), multiple criteria, such as the BIC or the proportion of variance explained are recommended for decision-making, but also interpretability in light of theory, parsimony, and evidence from cross-validation should be considered. For cluster analysis, the proportion of variance explained can be helpful for selecting a number of candidate profiles. Then, the analyst should focus on the interpretability of the solution as well as concerns of parsimony; if two profile solutions are similar, then the more simple solution may be preferred.

## How to Choose Options for the Clustering Procedure or to Select the Model Parameterization

This applies most to LPA, but in cluster analysis, there are also options about the distance metric used. In LPA, we can determine which parameters - variance and covariance - are estimated to be the same or different across all profiles, and, in the case of covariance, whether it is fixed to zero. Of course, for the profiles we are most interested in, the mean is allowed to vary across profiles. The models with more parameters freely-estimated use more degrees of freedom; thus, similar to the number of profiles selected, the balance between adding more parameters (and having a more complex model) and interpretability and parsimony must be balanced. If a more complex model fits better and is interpretable and is justifiable in terms of the data collected, then it may be preferred to simpler models. If the two models are similar in terms of their fit, then the more simple, parsimonious model should be selected.

# More General Recommendations

This section highlights three general recommendations. First, scholars taking a person-oriented approach should emphasize reproducibility in carrying out analyses. This is in part due to the exploratory and interpretative nature of person-oriented approaches. To this end, presenting multiple models, as in Pastor et al. (2007), should be encouraged, rather than presenting one solution. In addition, having multiple analysts review the solutions found is encouraged. As part of this recommendation, we suggest that researchers consider how their analyses can be reproduced by other analysts outside of the research team: Sharing code and data is an important part of this work. Also related, researchers should consider how profiles are replicated across samples.

A second general recommendation considers more flexibly incorporating time, when data are collected across multiple time points, into analyses. ISOA groups all time points and doesn’t make distinctions. Other approaches perform analysis separately at different timepoints (Corpus & Wormington, 2014). Some integrate time as a part of the profiles, i.e. growth mixture modeling (groups of patterns), i.e. within-person growth modeling, where there are individual growth patterns. Research to date has yet to consider additional challenges in applying person centered approaches to longitudinal data. For instance, Schmidt et al (2018) use an Experience Sampling Method (ESM) approach to collecting data and used a person-centered approach to generate profiles of students in science class. This work does not account for student-level effects. In other words, it did not model the shared variance of multiple observations of the same student.

Third, best practices in within-person or longitudinal research call for modeling the nesting structure. However, researchers have yet to successfully incorporate this practice into the person centered approach. One way to approach this is to use cross-classified mixed effects models, as in Strati, Schmidt, and Maier (2017) and in Rosenberg (2018). In such approaches, dependencies in terms of, for example, individuals responding to (ESM) surveys at the same time and repeated responses being associated with the same individuals can both be modeled, although the effects of these two sources of dependencies are not nested as in very common uses of multi-level models, but rather are cross-classified. West, Welch, & Galecki (2014) have a description of the use of multi-level models with cross-classified data and tools (including those freely available through R) that can be used to estimate them.

# tidyLPA update

## Rationale for the update

**tidyLPA** has received a major update. Your old function calls might not work as expected anymore. The new version of tidyLPA aims to make Latent Profile Analysis (LPA) even more accessible to a broad audience, by means of a user-friendly interface, and to fit more seemlessly into a 'tidyverse' workflow.

As before, **tidyLPA** offers a parallel workflow for the open-source package Mclust, and the commercial package *Mplus*.

These changes necessitated some simplification of the existing functions. Advanced options are still available, but require slightly more work on the user's part. The upside of this is that, for most users, the tidyLPA workflow and documentation are substantially simplified.

## Major changes

There have been major changes to the two primary functions, `estimate_profiles()` and `compare_solutions()`

* `estimate_profiles()` now estimates any number of profiles, for any number of models, using any algorithm. It has a simplified user interface. This makes it a more general function, of which the former version (which could only estimate one number of profiles for one model) is a special, restricted case.

* `compare_solutions()` now compares solutions estimated by estimate_profiles(), suggesting the optimal model and number of classes. This replaces its dual role of estimating AND comparing solutions.
    `* `compare_solutions()` gives you the same set of fit indices for Mclust and Mplus. It provides a broader range of fit indices than before
    * `compare_solutions()` calls `AHP()``, a function which uses the Automatic Hierarchical Process to determine the optimal number of classes, based on several reputable fit indices

* `plot_profiles()` now incorporates best practices for plotting latent profile analyses: Visualizing cluster centroids, standard errors, cluster variances, and classification uncertainty by means of weighted raw data

* `impute_missing()` offers two easy-to-use single imputation options, for use with the Mclust package (which does not natively handle missing data). Mplus handles partially missing data by default.

The best way to understand these changes to **tidyLPA**, though, is by explaining the new workflow.

## Example analysis using estimate_profiles()

A full analysis might look like the following:

```{r}
library(tidyverse)
library(tidyLPA)

pisaUSA15[1:100, ] %>%
    select(broad_interest, enjoyment, self_efficacy) %>%
    single_imputation() %>%
    estimate_profiles(3)
```

You can use Mplus simply by changing the package argument for `estimate_profiles()`:

```{r, eval = FALSE}
pisaUSA15[1:100, ] %>%
    select(broad_interest, enjoyment, self_efficacy) %>%
    single_imputation() %>%
    estimate_profiles(3, package = "MplusAutomation")
```

A simple summary of the analysis is printed to the console (and its posterior probability). The resulting object can be further passed down a pipeline to other functions, such as `plot`, `compare_solutions`, `get_data`, `get_fit`, etc. This is the "tidy" part, in that the function can be embedded in a tidy analysis pipeline.

If you have Mplus installed, you can call the version of this function that uses MPlus in the same way, by adding the argument `package = "MplusAutomation`.

We can plot the profiles by piping the output to `plot_profiles()`.

```{r, eval = FALSE}
pisaUSA15[1:100, ] %>%
    select(broad_interest, enjoyment, self_efficacy) %>%
    single_imputation() %>%
    scale() %>%
    estimate_profiles(3) %>% 
    plot_profiles()
```

## Example analysis using compare_solutions()

The function `compare_solutions()` compares the fit of several estimated models,
with varying numbers of profiles and model specifications:

```{r, eval = TRUE}
pisaUSA15[1:100, ] %>%
    select(broad_interest, enjoyment, self_efficacy) %>%
    single_imputation() %>%
    estimate_profiles(1:3, 
                      variances = c("equal", "varying"),
                      covariances = c("zero", "varying")) %>%
    compare_solutions(statistics = c("AIC", "BIC"))
```

## Deprecated functionality:

* `compare_solutions()` no longer estimates profiles. It merely compares solutions estimated by `estimate_profiles()`

* all Mplus-specific functions are deprecated. `estimate_profiles()` offers a unified interface to Mclust and Mplus. These Mplus-specific functions are deprecated:

    * `estimate_profiles_mplus()`
        * `compare_profiles_mplus()`
        * `plot_profiles_mplus()`

* estimate_profiles and compare_solutions no longer select variables, passed as unquoted variable names, from 'df'. Instead, they use all variables in df. To select variables prior to analysis, use:

    * The dplyr function select(df, Your, Selected, Variable, Names)
        * The base R function df[, c("Your", "Selected", "Variable", "Names")]

* estimate_profiles and compare_solutions no longer center and scale the raw data. To center and scale variables prior to analysis, use scale(df, center = TRUE, scale = TRUE)

* estimate_profiles no longer accepts arguments specific to Mplus or Mclust. Instead, you can specify additional arguments directly to the core functions `mplusModeler()` and `Mclust()`. See the help file for these functions for more details. Deprecated arguments, whose functionality is supplanted by passing arguments directly to these core functions, are:

For Mclust:

    * prior_control

For Mplus:

    * starts
    * m_iterations
    * st_iterations
    * convergence_criterion
    * optseed
    * cluster_ID
    * include_VLMR
    * include_BLRT (but BLRT is returned by default now)

# tidyLPA function with _mplus() appended require Mplus to use

```{r, eval = FALSE}
estimate_profiles_mplus(data, var1, var2, var3, ariances = “free”, covariances = “equal”) %>% 
    plot_profiles_mplus()
```

# Conclusion

Person-oriented analysis ia way to consider how psychological constructs are experienced (and can be analyzed) together and at once. Though described in contrast to a variable-centered approach, scholars have pointed out how person-oriented approaches are complementary to variable-centered analyses (Marsh, Ludtke, Trautwein, & Morin, 2009). A person-oriented approach can help us to consider multiple variables together and at once and in a more dynamic way, reflected in the methodological approaches for cluster analysis and LPA that identify profiles of individuals responses.

This manuscript provided an outline of how to get started with person-oriented analyses in an informed way. We provided a general overview of the methodology and described tools to carry out such an analysis. We also described specific tools, emphasizing freely-available open-source options that we have developed. Because of the inherently exploratory nature of person-oriented analysis, carrying out the analysis in a trustworthy and open way is particularly important. In this way, the interpretative aspect of settling on a solution shares some features of quantitative and qualitative research: The systematic nature of quantitative research methods (focused upon agreed-upon criteria such as likelihood-ratio tests) and qualitative research methods (focused upon the trustworthiness of both the analysis and the analyst) are important to consider when carrying out person-oriented analysis. Lastly, we made some general recommendations for future directions--and also highlighted some situations for which person-oriented approaches may not be the best and some cautions raised in past research regarding how such approaches are used.

In conclusion, as use of person-oriented approaches expand, new questions and opportunities for carrying out research in a more holistic, dynamic way will be presented. Analyzing constructs together and at once is appealing to researchers, particularly those carrying out research in fields such as education for which communicating findings to stakeholders in a way that has the chance to impact practice is important. Our aim was not to suggest that such an approach is always the goal or should always be carried out, but rather to describe how researchers may get started in an informed way as researchers seek to understand how individuals interact, behave, and learn in ways that embraces the complexity of these experiences.

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```
Bergman, L. R., & Magnusson, D. (1997). A person-oriented approach in research on developmental psychopathology. Development and psychopathology, 9(2), 291-319.  
Bergman, L. R., & Trost, K. (2006). The person-oriented versus the variable-oriented approach: Are they complementary, opposites, or exploring different worlds?. Merrill-Palmer Quarterly, 52(3), 601-632.
Collins, L. M., & Lanza, S. T. (2010). Latent class and latent transition analysis: With applications in the social, behavioral, and health sciences (Vol. 718). John Wiley & Sons.
Corpus, J. H., & Wormington, S. V. (2014). Profiles of intrinsic and extrinsic motivations in elementary school: A longitudinal analysis. The Journal of Experimental Education, 82(4), 480-501.
Corpus, J. H., & Wormington, S. V. (2014). Profiles of intrinsic and extrinsic motivations in elementary school: A longitudinal analysis. The Journal of Experimental Education, 82(4), 480-501.
Harring, J. R., & Hodis, F. A. (2016). Mixture modeling: Applications in educational psychology. Educational Psychologist, 51(3-4), 354-367.
Hayenga, A. O., & Corpus, J. H. (2010). Profiles of intrinsic and extrinsic motivations: A person-centered approach to motivation and achievement in middle school. Motivation and Emotion, 34(4), 371-383.
Linnenbrink-Garcia, L., & Wormington, S. V. (2017). Key challenges and potential solutions for studying the complexity of motivation in schooling: An integrative, dynamic person-oriented perspective. British Journal of Educational Psychology monograph series II. Psychological aspects of education: Current trends—the role of competence beliefs in teaching and learning. Chichester, UK: Wiley.
Magnusson, D., & Cairns, R. B. (1996). Developmental science: Toward a unified framework. Cambridge, England: Cambridge University Press.  
Marsh, H. W., Lüdtke, O., Trautwein, U., & Morin, A. J. (2009). Classical latent profile analysis of academic self-concept dimensions: Synergy of person-and variable-centered approaches to theoretical models of self-concept. Structural Equation Modeling, 16(2), 191-225.
McLachlan, G. J. (2011). Commentary on Steinley and Brusco (2011): Recommendations and cautions.
Muthen, L. K., & Muthen, B. O. (1997-2017). Mplus User’s Guide. Los Angeles, CA: Muthén & Muthén.
Rosenberg, J. M., Schmidt, J. A., Beymer, P. N., & Steingut, R. R. (2018). Interface to mclust to easily carry out Latent Profile Analysis [Statistical software for R]. https://github.com/jrosen48/tidyLPA
Rosenberg, J. M., Schmidt, J. A., Beymer, P. N., & Steingut, R. R. (2017). prcr: Person-Centered Analysis. R package version 0.1.5. https://CRAN.R-project.org/package=prcr
Scrucca L., Fop M., Murphy T. B. and Raftery A. E. (2017) mclust 5: clustering, classification and density estimation using Gaussian finite mixture models The R Journal 8/1, pp. 205-233
Hennig, C., Meila, M., Murtagh, F., & Rocci, R. (Eds.). (2015). Handbook of cluster analysis. CRC Press.
Steinley, D., & Brusco, M. J. (2011). Evaluating mixture modeling for clustering: Recommendations and cautions. Psychological Methods, 16(1), 63.
Steinley, D., & Brusco, M. J. (2011). K-means clustering and mixture model clustering: Reply to McLachlan (2011) and Vermunt (2011).
Trevors, G. J., Kendeou, P., Bråten, I., & Braasch, J. L. (2017). Adolescents’ epistemic profiles in the service of knowledge revision. Contemporary Educational Psychology, 49, 107-120.
Vermunt, J. K. (2011). K-means may perform as well as mixture model clustering but may also be much worse: Comment on Steinley and Brusco (2011).
West, B. T., Welch, K. B., & Galecki, A. T. (2014). Linear mixed models: a practical guide using statistical software. Chapman and Hall/CRC.
Wickham, H. (2014). Tidy data. Journal of Statistical Software, 59(10), 1-23.

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup

### 2019-11-26 paper.rmd
update

@@
 ---
-title             : "Getting started with Latent Profile Analysis (LPA) in the psychological sciences: A tutorial using the tidyLPA R package that provides an interface to open-source and commercial software"
+title             : "Doing Latent Profile Analysis in a more rigorous and reproducible way: A tutorial using the tidyLPA R package"
 shorttitle        : "Title"
@@
     affiliation   : "3"
-  - name          : "Rebecca Steingut"
-    affiliation   : "5"
-  - name          : "Stephanie Wormington"
-    affiliation   : "6"
 
@@
   - id            : "5"
-    institution   : "Columbia University"
-  - id            : "6"
-    institution   : "University of Virginia"
 
@@
 
-The questions on the Scimo per and post survey are from Ryan and Deci's "Learning Climate Questionnaire".  You can read a short description here:
+<!-- The questions on the Scimo per and post survey are from Ryan and Deci's "Learning Climate Questionnaire".  You can read a short description here: -->
 
-https://selfdeterminationtheory.org/pas-learning-climate/
+<!-- https://selfdeterminationtheory.org/pas-learning-climate/ -->
 
-It is 15 questions, but they identify 6 which can be used as  short-form.
+<!-- It is 15 questions, but they identify 6 which can be used as  short-form.
 Questions 1, 2, 4, 7, 10, 14.
 I have attached the PDF of the survey since you have to create an account on their website do download it.
-Below are the question text, along with the variable name in the Scimo data:
+Below are the question text, along with the variable name in the Scimo data: -->
 
-I feel that my instructor provides me choices and options.  t_choices1
-I feel understood by my instructor.  t_understand1
-My instructor conveyed confidence in my ability to do well in the course.  t_confidence1
-My instructor encouraged me to ask questions.  t_encourage1
-My instructor listens to how I would like to do things.  t_listen1
-My instructor tries to understand how I see things before suggesting a new way to do things.  t_triesunder1
-
-Also, is this tutorial going to be a journal paper, or something on github, etc.
-as a kind of manual for using LPA?
-If so, and you think I would be a good fit, I would love to help out on it if you have author space!
-
-Let me know if you need anything else.
+<!-- I feel that my instructor provides me choices and options.  t_choices1 -->
+<!-- I feel understood by my instructor.  t_understand1 -->
+<!-- My instructor conveyed confidence in my ability to do well in the course.  t_confidence1 -->
+<!-- My instructor encouraged me to ask questions.  t_encourage1 -->
+<!-- My instructor listens to how I would like to do things.  t_listen1 -->
+<!-- My instructor tries to understand how I see things before suggesting a new way to do things.  t_triesunder1 -->
 
@@
 
-Person-oriented analysis involves a statistical approach to exploring psychological constructs in a way that captures and simplifies complexity.
-This approach can be used to consider the way in which psychological constructs, for example, are experienced together and at once.
-Thus, conceptually, this approach emphasizes how individuals go about interacting, socializing, and behaving in this way, experiencing these activities as a whole person.
-Person-oriented analysis, developed within developmental science, focuses on how individuals have experiences in a holistic way and also how these ways can be identified through common groups, or profiles.(Bergman & El-Khouri, 1997; Magnusson & Cairns, 1996).
-One way to distinguish the approach is to consider it in contrast to variable-centered approaches, such as those for which the covariability between variables, rather than the relations between profiles of individuals, is the focus (Bergman & Trost, 2006).
+Person-oriented analysis involves a statistical approach to exploring
+psychological constructs in a way that captures and simplifies complexity. This
+approach can be used to consider the way in which psychological constructs, for
+example, are experienced together and at once. Thus, conceptually, this approach
+emphasizes how individuals go about interacting, socializing, and behaving in
+this way, experiencing these activities as a whole person. Person-oriented
+analysis, developed within developmental science, focuses on how individuals
+have experiences in a holistic way and also how these ways can be identified
+through common groups, or profiles.(Bergman & El-Khouri, 1997; Magnusson &
+Cairns, 1996). One way to distinguish the approach is to consider it in contrast
+to variable-centered approaches, such as those for which the covariability
+between variables, rather than the relations between profiles of individuals, is
+the focus (Bergman & Trost, 2006).
 
-Though studies examining learning from a person-oriented perspective are not very common, some examples include studies of intrinsic and extrinsic motivation (Corpus & Wormington, 2014; Hayenga & Corpus, 2010), profiles of achievement goals (see Wormington & Linnenbrink-Garcia, advance online publication, for a review), and epistemic cognition (Trevors, Kendeou, Braten, & Braasch, 2017).
-Recently, too, there have been excellent reviews of person-oriented analyses in motivation and educational psychology (Linnenbrink-Garcia & Wormington, 2017).
+Though studies examining learning from a person-oriented perspective are not
+very common, some examples include studies of intrinsic and extrinsic motivation
+(Corpus & Wormington, 2014; Hayenga & Corpus, 2010), profiles of achievement
+goals (see Wormington & Linnenbrink-Garcia, advance online publication, for a
+review), and epistemic cognition (Trevors, Kendeou, Braten, & Braasch, 2017).
+Recently, too, there have been excellent reviews of person-oriented analyses in
+motivation and educational psychology (Linnenbrink-Garcia & Wormington, 2017).
 
-In this article, our goal is to review person-oriented analysis focused on getting started (on both knowing about and doing person-oriented analyses) in an informed way.
-More specifically, we start with an overview of the methodology, describe specific software tools, explain analytic choices made in the course of their use, and conclude with general recommendations One challenge facing those trying to learn about and apply person-oriented analysis in their research is the multitude of terms used for similar analytic approaches or ideas.
-Throughout this manuscript, we aim to point these out and explain when and why a variety of terms have been used.
-Thus, this manuscript is written for an informed and curious reader at any level interested in understanding this particular approach.
+In this article, our goal is to review person-oriented analysis focused on
+getting started (on both knowing about and doing person-oriented analyses) in an
+informed way. More specifically, we start with an overview of the methodology,
+describe specific software tools, explain analytic choices made in the course of
+their use, and conclude with general recommendations One challenge facing those
+trying to learn about and apply person-oriented analysis in their research is
+the multitude of terms used for similar analytic approaches or ideas. Throughout
+this manuscript, we aim to point these out and explain when and why a variety of
+terms have been used. Thus, this manuscript is written for an informed and
+curious reader at any level interested in understanding this particular
+approach.
 
@@
 
-At present, there are a number of approaches to carrying out person-oriented analysis when the indicators, of variables, used to find the groups are continuous.
-This section is focused upon two approaches: cluster analysis and Latent Profile Analysis (LPA).
-We focus on these because they are relatively simple and accessible forms of person-oriented analysis that can also accommodate cross sectional and longitudinal data alike.
-Just as a brief point of contrast, when the indicators for the groups are dichotomous (or categorical and can therefore be “dummy-coded” to be dichotomous), an approach closely related to LPA, Latent Class Analysis, can be used (cite PSU folks).
+At present, there are a number of approaches to carrying out person-oriented
+analysis when the indicators, of variables, used to find the groups are
+continuous. This section is focused upon two approaches: cluster analysis and
+Latent Profile Analysis (LPA). We focus on these because they are relatively
+simple and accessible forms of person-oriented analysis that can also
+accommodate cross sectional and longitudinal data alike. Just as a brief point
+of contrast, when the indicators for the groups are dichotomous (or categorical
+and can therefore be “dummy-coded” to be dichotomous), an approach closely
+related to LPA, Latent Class Analysis, can be used (cite PSU folks).
 
@@
 
-Cluster analysis is an approach that has been frequently applied to analysis of topics in educational psychology (such as motivation and engagement) from a more holistic, whole-person approach.
-One particular way in which cluster analysis has been used has been as part of an approach termed I-States-as-Objects Analysis (ISOA).
-The purpose of ISOA is to identify profiles at specific time points, and then to explore how individuals do or do not shift between profiles across time points (needs cite).
-In their early work, Bergman and Magnusson (1999) and colleagues (1999) described a two-step approach to cluster analysis as part of their I-States-as-Objects analysis.
-In this approach, all of the observations are clustered.
-The key idea behind the approach is that each observation represents not a permanent group, but rather a status, or an i-state.
-These i-states, are considered independently for the purposes of constructing profiles.
-Then, shifts between these i-states across time (say, two or three time points) are evaluated.
+Cluster analysis is an approach that has been frequently applied to analysis of
+topics in educational psychology (such as motivation and engagement) from a more
+holistic, whole-person approach. One particular way in which cluster analysis
+has been used has been as part of an approach termed I-States-as-Objects
+Analysis (ISOA). The purpose of ISOA is to identify profiles at specific time
+points, and then to explore how individuals do or do not shift between profiles
+across time points (needs cite). In their early work, Bergman and Magnusson
+(1999) and colleagues (1999) described a two-step approach to cluster analysis
+as part of their I-States-as-Objects analysis. In this approach, all of the
+observations are clustered. The key idea behind the approach is that each
+observation represents not a permanent group, but rather a status, or an
+i-state. These i-states, are considered independently for the purposes of
+constructing profiles. Then, shifts between these i-states across time (say, two
+or three time points) are evaluated.
 
-Commonly, cluster analysis carried out as part of ISOA in fact involves two distinct clustering steps.
-This is because different clustering algorithms have different strengths--as well as weaknesses.
-In ISOA, the first step is Ward’s hierarchical cluster analysis (Bergman & El-Khouri, 1999).
-In this step, observations are split, step-by-step, based upon their multivariate similarity.
-The analyst provides the number of groups which will be used and interpreted, though technically most algorithms will continue splitting the groups until each observation makes up its own.
-This type of clustering is deterministic: The same results will be obtained each time it is run.
-In the second step of the cluster analysis, an iterative, k-means cluster analysis is carried out (Bergman and El-Khouri, 1999).
-For this step, the algorithm partitions the observations into groups, the success of the clustering based on a criterion such as the variability in the data explained by the groups, and then “moves,” when possible, the observations that are not in the group they are most similar to to the one that they are most similar to.
-When no more such moves are possible, the algorithm stops.
-On its own, k-means is not deterministic: Depending on the starting values, it can obtain different results the different times that it is run.
-So, the results from the hierarchical cluster analysis, the first step, are used as start values to initialize the k-means clustering.
+Commonly, cluster analysis carried out as part of ISOA in fact involves two
+distinct clustering steps. This is because different clustering algorithms have
+different strengths--as well as weaknesses. In ISOA, the first step is Ward’s
+hierarchical cluster analysis (Bergman & El-Khouri, 1999). In this step,
+observations are split, step-by-step, based upon their multivariate similarity.
+The analyst provides the number of groups which will be used and interpreted,
+though technically most algorithms will continue splitting the groups until each
+observation makes up its own. This type of clustering is deterministic: The same
+results will be obtained each time it is run. In the second step of the cluster
+analysis, an iterative, k-means cluster analysis is carried out (Bergman and
+El-Khouri, 1999). For this step, the algorithm partitions the observations into
+groups, the success of the clustering based on a criterion such as the
+variability in the data explained by the groups, and then “moves,” when
+possible, the observations that are not in the group they are most similar to to
+the one that they are most similar to. When no more such moves are possible, the
+algorithm stops. On its own, k-means is not deterministic: Depending on the
+starting values, it can obtain different results the different times that it is
+run. So, the results from the hierarchical cluster analysis, the first step, are
+used as start values to initialize the k-means clustering.
 
-The rationale behind using these two approaches is that the hierarchical clustering provides reasonable guesses for the profile membership, and then the iterative clustering optimizes the clustering solution.
-To sum up, then, cluster analysis as part of ISOA in fact involves two separate cluster analyses.
-The merits of this approach are to leverage the benefits of each: the iterative clustering starts with reasonable cluster solutions generated from the hierarchical clustering and optimizes them.
+The rationale behind using these two approaches is that the hierarchical
+clustering provides reasonable guesses for the profile membership, and then the
+iterative clustering optimizes the clustering solution. To sum up, then, cluster
+analysis as part of ISOA in fact involves two separate cluster analyses. The
+merits of this approach are to leverage the benefits of each: the iterative
+clustering starts with reasonable cluster solutions generated from the
+hierarchical clustering and optimizes them.
 
@@
 
-Another approach what is becoming increasingly common for carrying out person-oriented analysis is Latent Profile Analysis (LPA).
-LPA is a type of finite mixture model (Harring and Hodis, 2017).
-The goal of this approach is estimate the parameters for a number of distributions (often multivariate) from a single dataset.
-The approach can be thought of as seeking an answer to the question, Do the data in a sample come from more than one population?
-Thus, such an approach is model-based, and some descriptions in the literature refer to it as model-based clustering (Hennig, Meila, Murtagh, & Rocci, 2015; Scrucca, Fop, Murphy, & Raftery, 2017).
-Thus, one distinction between LPA and cluster analytic approaches is that LPA is model-based; instead of using algorithms to group together cases, LPA seeks to estimate parameters - in terms of variances and covariances and how they are the same or different across profiles - that best characterize the different distributions.
-Then, this approach seeks to assign to each observation a probability that the observation is a sample from the population associated with each profile (or mixture component).
+Another approach what is becoming increasingly common for carrying out
+person-oriented analysis is Latent Profile Analysis (LPA). LPA is a type of
+finite mixture model (Harring and Hodis, 2017). The goal of this approach is
+estimate the parameters for a number of distributions (often multivariate) from
+a single dataset. The approach can be thought of as seeking an answer to the
+question, Do the data in a sample come from more than one population? Thus, such
+an approach is model-based, and some descriptions in the literature refer to it
+as model-based clustering (Hennig, Meila, Murtagh, & Rocci, 2015; Scrucca, Fop,
+Murphy, & Raftery, 2017). Thus, one distinction between LPA and cluster analytic
+approaches is that LPA is model-based; instead of using algorithms to group
+together cases, LPA seeks to estimate parameters - in terms of variances and
+covariances and how they are the same or different across profiles - that best
+characterize the different distributions. Then, this approach seeks to assign to
+each observation a probability that the observation is a sample from the
+population associated with each profile (or mixture component).
 
-Because LPA is model-based, a number of different model parameterizations can be estimated.
-These models differ in terms of whether--and how--parameters are estimated across the profiles.
-These parameters are the means for the different profiles, which, in this approach, always are estimated freely across the profiles; the variances for the variables used to create the profiles, which can be estimated freely or can be estimated to be the same, or equal, across profiles; and the covariances of the variables used to create the profiles, which can be freely-estimated, estimated to be equal, or fixed to be zero.
-One challenge facing the analyst using LPA is that these parameters and the distinct model parameterizations that can be estimated is the different terminology used.
-As one example, Scrucca et al.
-(2017) refer to these parameterizations not in terms of whether and how parameters are estimated, but rather in terms of the geometric properties of the distributions that result from particular parameterizations.
-Muthen and Muthen (1997-2017) and others (Pastor et al., 2007) commonly refer to local independence to mean that the covariances are fixed to zero (also described as the specification of the covariance matrix as “diagonal,” because only the diagonal components, or the variances, are estimated).
-More information on the model parameterizations are discussed in the context of the software tool tidyLPA that we have developed.
+Because LPA is model-based, a number of different model parameterizations can be
+estimated. These models differ in terms of whether--and how--parameters are
+estimated across the profiles. These parameters are the means for the different
+profiles, which, in this approach, always are estimated freely across the
+profiles; the variances for the variables used to create the profiles, which can
+be estimated freely or can be estimated to be the same, or equal, across
+profiles; and the covariances of the variables used to create the profiles,
+which can be freely-estimated, estimated to be equal, or fixed to be zero. One
+challenge facing the analyst using LPA is that these parameters and the distinct
+model parameterizations that can be estimated is the different terminology used.
+As one example, Scrucca et al. (2017) refer to these parameterizations not in
+terms of whether and how parameters are estimated, but rather in terms of the
+geometric properties of the distributions that result from particular
+parameterizations. Muthen and Muthen (1997-2017) and others (Pastor et
+al., 2007) commonly refer to local independence to mean that the covariances are
+fixed to zero (also described as the specification of the covariance matrix as
+“diagonal,” because only the diagonal components, or the variances, are
+estimated). More information on the model parameterizations are discussed in the
+context of the software tool tidyLPA that we have developed.
 
- This section outlines several LPA models and guidelines/examples for when they might be most appropriate in person-centered analysis.
-In general, as more parameters are estimated (i.e., those that are fixed to zero are estimated as being equal across profiles; or those estimated as being equal across profiles are freely-estimated across them), the model becomes more complex; the model may fit better, but also be overfit, meaning that the profiles identified may be challenging to replicate with another, separate data set.
-Even still, flexibility in terms of which models can be estimated also has affordances.
-For example, the varying means, equal variances, and covariances fixed to 0.
-A researcher might choose this model specification if she wants to model the variables to be used to create profiles that are independent.
-This model is very simple, as no covariances are estimated and the variances are estimated to be the same across profiles.
-As we estimate more parameters (and decrease the degrees of freedom), we are more likely to fit the data, but less likely to be able to replicate the model with a second set of data.
-In other words, more parameters may mean a loss of external validity.
-As we progress toward more complex models (with increasingly complex parameterization, i.e.
-models 3 and 4 above), then we are more likely to fit the data better.
+This section outlines several LPA models and guidelines/examples for when they
+might be most appropriate in person-centered analysis. In general, as more
+parameters are estimated (i.e., those that are fixed to zero are estimated as
+being equal across profiles; or those estimated as being equal across profiles
+are freely-estimated across them), the model becomes more complex; the model may
+fit better, but also be overfit, meaning that the profiles identified may be
+challenging to replicate with another, separate data set. Even still,
+flexibility in terms of which models can be estimated also has affordances. For
+example, the varying means, equal variances, and covariances fixed to 0. A
+researcher might choose this model specification if she wants to model the
+variables to be used to create profiles that are independent. This model is very
+simple, as no covariances are estimated and the variances are estimated to be
+the same across profiles. As we estimate more parameters (and decrease the
+degrees of freedom), we are more likely to fit the data, but less likely to be
+able to replicate the model with a second set of data. In other words, more
+parameters may mean a loss of external validity. As we progress toward more
+complex models (with increasingly complex parameterization, i.e. models 3 and 4
+above), then we are more likely to fit the data better.
 
@@
 
-SPSS is a common tool to carry out cluster analyses.
-SPSS provides functionality to carry out both hierarchical and k-means cluster analysis.
-While somewhat straightforward to carry out, particularly in SPSS’s graphical user interface (GUI), there are some challenges to use of this approach.
-The GUI in SPSS can be challenging, even for the most able analyst, to be able to document every step with syntax, and so reproducing the entire analysis efficiently can be a challenge, both for the analyst exploring various solutions and for the reviewer looking to replicate this work.
-Additionally, SPSS is commercial software (and is expensive), and so analysts without the software cannot carry out this analysis.
+SPSS is a common tool to carry out cluster analyses. SPSS provides functionality
+to carry out both hierarchical and k-means cluster analysis. While somewhat
+straightforward to carry out, particularly in SPSS’s graphical user interface
+(GUI), there are some challenges to use of this approach. The GUI in SPSS can be
+challenging, even for the most able analyst, to be able to document every step
+with syntax, and so reproducing the entire analysis efficiently can be a
+challenge, both for the analyst exploring various solutions and for the reviewer
+looking to replicate this work. Additionally, SPSS is commercial software (and
+is expensive), and so analysts without the software cannot carry out this
+analysis.
 
- Accordingly, we worked to develop tools, initially used by the research team and now shared more widely, to provide basic functionality for each of the two primary approaches described in this section, cluster analysis and Latent Profile Analysis (LPA).
-In addition to being freely-available and open-source (i.e., anyone can inspect the source code), the tools described lend themselves to an open science.
-It is especially important to have an open transparent methodology when using person-centered approaches because of the high dependence on researcher judgment at multiple points in the process.
+Accordingly, we worked to develop tools, initially used by the research team and
+now shared more widely, to provide basic functionality for each of the two
+primary approaches described in this section, cluster analysis and Latent
+Profile Analysis (LPA). In addition to being freely-available and open-source
+(i.e., anyone can inspect the source code), the tools described lend themselves
+to an open science. It is especially important to have an open transparent
+methodology when using person-centered approaches because of the high dependence
+on researcher judgment at multiple points in the process.
 
-Accordingly, we set out to develop packages that a) provided sensible defaults and were easy to use, but provided the option to access and modify all of the inputs to the model (i.e., low barrier, high ceiling), b) interfaced to existing tools,, and are able to translate between what existing tools are capable of and what researchers and analysts carrying-out person-oriented analyses would like to specify, c) made it easy to carry-out fully-reproducible analyses and d) were well-documented.
-Both of these packages--one for two-step cluster analysis and one for LPA--are designed to do one thing well, namely, quickly and easily estimate profiles using cluster analysis or a model-based approach (Latent Profile Analysis).
+Accordingly, we set out to develop packages that a) provided sensible defaults
+and were easy to use, but provided the option to access and modify all of the
+inputs to the model (i.e., low barrier, high ceiling), b) interfaced to existing
+tools,, and are able to translate between what existing tools are capable of and
+what researchers and analysts carrying-out person-oriented analyses would like
+to specify, c) made it easy to carry-out fully-reproducible analyses and d) were
+well-documented. Both of these packages--one for two-step cluster analysis and
+one for LPA--are designed to do one thing well, namely, quickly and easily
+estimate profiles using cluster analysis or a model-based approach (Latent
+Profile Analysis).
 
-One way that tidyLPA is designed to be easy to use is that it assumes a “tidy” data structure (Wickham, 2014).
-This means that it emphasizes the use of a data frame as both the primary input and output of functions for the package.
-Because data is passed to and returned (in amended form, i.e., with the latent profile probabilities and classes appended to the data) from the function, it makes it easy to create plots or use results in subsequent analyses.
-Another noteworthy feature of tidyLPA is that it provides the same functionality through two different tools, one that is open-source and available through R, the mclust package (Scrucca et al., 2017) and one that is available through the commercial software MPlus (Muthen & Muthen, 1997-2017).
-Moreover, as both tools use the same maximum likelihood estimation procedure, they are benchmarked to produce the same output (see here).
-Also, note that we have described the model specifications with descriptions of what is estimated in terms of the variances and covariances, the common names for the models (i.e., class-varying unrestricted), and the covariance matrix associated with the parameterizationfor the six models that are possible to be estimated on the website for tidyLPA (see here).
+One way that tidyLPA is designed to be easy to use is that it assumes a “tidy”
+data structure (Wickham, 2014). This means that it emphasizes the use of a data
+frame as both the primary input and output of functions for the package. Because
+data is passed to and returned (in amended form, i.e., with the latent profile
+probabilities and classes appended to the data) from the function, it makes it
+easy to create plots or use results in subsequent analyses. Another noteworthy
+feature of tidyLPA is that it provides the same functionality through two
+different tools, one that is open-source and available through R, the mclust
+package (Scrucca et al., 2017) and one that is available through the commercial
+software MPlus (Muthen & Muthen, 1997-2017). Moreover, as both tools use the
+same maximum likelihood estimation procedure, they are benchmarked to produce
+the same output (see here). Also, note that we have described the model
+specifications with descriptions of what is estimated in terms of the variances
+and covariances, the common names for the models (i.e., class-varying
+unrestricted), and the covariance matrix associated with the parameterizationfor
+the six models that are possible to be estimated on the website for tidyLPA (see
+here).
 
@@
 
-There are a number of analytic choices that need to be made when carrying out person-oriented analyses.
-Because such person-oriented approaches are often more subjective (in practice) than other approaches (Linnenbrink-Garcia and Wormington, 2017), there is no one rule for determining the solution obtained.
-This solution is obtained on the basis of multiple decisions, such as the number of profiles selected or the modeling decisions such as what specific options are used for the cluster analysis (i.e., the distance metric used to calculate the similarity of the observations as part of the Ward’s hierarchical clustering) or what parameters are estimated and how as part of LPA.
+There are a number of analytic choices that need to be made when carrying out
+person-oriented analyses. Because such person-oriented approaches are often more
+subjective (in practice) than other approaches (Linnenbrink-Garcia and
+Wormington, 2017), there is no one rule for determining the solution obtained.
+This solution is obtained on the basis of multiple decisions, such as the number
+of profiles selected or the modeling decisions such as what specific options are
+used for the cluster analysis (i.e., the distance metric used to calculate the
+similarity of the observations as part of the Ward’s hierarchical clustering) or
+what parameters are estimated and how as part of LPA.
 
-Given the subjectivity involved, it is important that researchers be transparent and work as part of a team to obtain clustering solutions.
-Transparency about the design and analytic choices is important so that readers can appropriately interpret the report.
-Researchers can enhance transparency and reproducibility by sharing detailed descriptions of methodology and document it through the use of syntax (and, if possible, data) that we share with others.
-Working as part of a team can help to serve as a check on several of the choices researchers make, such as over-fitting or under-fitting the model to the data.
-Each decision depends on multiple factors and balancing tensions.
-We discuss each of the key decisions listed in an analysis.
+Given the subjectivity involved, it is important that researchers be transparent
+and work as part of a team to obtain clustering solutions. Transparency about
+the design and analytic choices is important so that readers can appropriately
+interpret the report. Researchers can enhance transparency and reproducibility
+by sharing detailed descriptions of methodology and document it through the use
+of syntax (and, if possible, data) that we share with others. Working as part of
+a team can help to serve as a check on several of the choices researchers make,
+such as over-fitting or under-fitting the model to the data. Each decision
+depends on multiple factors and balancing tensions. We discuss each of the key
+decisions listed in an analysis.
 
@@
 
-The data can be transformed by centering or scaling.
-Typically, these are done after the profiles are created, so that differences between profiles can be more easily explored.
-They can also be done prior to the analysis, which can be helpful for obtaining solutions when the variables are on very different scales..
+The data can be transformed by centering or scaling. Typically, these are done
+after the profiles are created, so that differences between profiles can be more
+easily explored. They can also be done prior to the analysis, which can be
+helpful for obtaining solutions when the variables are on very different
+scales..
 
@@
 
-Whether to use cluster analysis or LPA depends in part on the nature of the data being investigated (cite the 2010 debate on this topic between Vermunt and others; MacLachlan, 2011; Steinley & Brusco, 2011a; Steinley & Brusco, 2011b; Vermunt, 2011).
-In some cases, cluster analysis and LPA solutions are very similar (when a constrained LPA model is fit).
-Broadly, if the variables used can be considered independent of one another, then cluster analysis may be justified in terms of reaching an optimal solution.
-However, LPA provides other benefits, such as the availability of information criteria (i.e., the Aikake Information Criteria [AIC] and the Bayesian Information Criteria [BIC] and statistical tests for the number of profiles (when a model is chosen, with a bootstrapped likelihood-ratio test).
-LPA also results in probabilistic assignments; this information can be used in subsequent analyses rather than the simple assignment (Collins & Lanza, 2010).
+Whether to use cluster analysis or LPA depends in part on the nature of the data
+being investigated (cite the 2010 debate on this topic between Vermunt and
+others; MacLachlan, 2011; Steinley & Brusco, 2011a; Steinley & Brusco, 2011b;
+Vermunt, 2011). In some cases, cluster analysis and LPA solutions are very
+similar (when a constrained LPA model is fit). Broadly, if the variables used
+can be considered independent of one another, then cluster analysis may be
+justified in terms of reaching an optimal solution. However, LPA provides other
+benefits, such as the availability of information criteria (i.e., the Aikake
+Information Criteria [AIC] and the Bayesian Information Criteria [BIC] and
+statistical tests for the number of profiles (when a model is chosen, with a
+bootstrapped likelihood-ratio test). LPA also results in probabilistic
+assignments; this information can be used in subsequent analyses rather than the
+simple assignment (Collins & Lanza, 2010).
 
@@
 
-In the case of choosing the number of profiles (and the specification of the model / profile solution), multiple criteria, such as the BIC or the proportion of variance explained are recommended for decision-making, but also interpretability in light of theory, parsimony, and evidence from cross-validation should be considered.
-For cluster analysis, the proportion of variance explained can be helpful for selecting a number of candidate profiles.
-Then, the analyst should focus on the interpretability of the solution as well as concerns of parsimony; if two profile solutions are similar, then the more simple solution may be preferred.
+In the case of choosing the number of profiles (and the specification of the
+model / profile solution), multiple criteria, such as the BIC or the proportion
+of variance explained are recommended for decision-making, but also
+interpretability in light of theory, parsimony, and evidence from
+cross-validation should be considered. For cluster analysis, the proportion of
+variance explained can be helpful for selecting a number of candidate profiles.
+Then, the analyst should focus on the interpretability of the solution as well
+as concerns of parsimony; if two profile solutions are similar, then the more
+simple solution may be preferred.
 
@@
 
-This applies most to LPA, but in cluster analysis, there are also options about the distance metric used.
-In LPA, we can determine which parameters - variance and covariance - are estimated to be the same or different across all profiles, and, in the case of covariance, whether it is fixed to zero.
-Of course, for the profiles we are most interested in, the mean is allowed to vary across profiles.
-The models with more parameters freely-estimated use more degrees of freedom; thus, similar to the number of profiles selected, the balance between adding more parameters (and having a more complex model) and interpretability and parsimony must be balanced.
-If a more complex model fits better and is interpretable and is justifiable in terms of the data collected, then it may be preferred to simpler models.
-If the two models are similar in terms of their fit, then the more simple, parsimonious model should be selected.
+This applies most to LPA, but in cluster analysis, there are also options about
+the distance metric used. In LPA, we can determine which parameters - variance
+and covariance - are estimated to be the same or different across all profiles,
+and, in the case of covariance, whether it is fixed to zero. Of course, for the
+profiles we are most interested in, the mean is allowed to vary across profiles.
+The models with more parameters freely-estimated use more degrees of freedom;
+thus, similar to the number of profiles selected, the balance between adding
+more parameters (and having a more complex model) and interpretability and
+parsimony must be balanced. If a more complex model fits better and is
+interpretable and is justifiable in terms of the data collected, then it may be
+preferred to simpler models. If the two models are similar in terms of their
+fit, then the more simple, parsimonious model should be selected.
 
@@
 
-This section highlights three general recommendations.
-First, scholars taking a person-oriented approach should emphasize reproducibility in carrying out analyses.
-This is in part due to the exploratory and interpretative nature of person-oriented approaches.
-To this end, presenting multiple models, as in Pastor et al.
-(2007), should be encouraged, rather than presenting one solution.
+This section highlights three general recommendations. First, scholars taking a
+person-oriented approach should emphasize reproducibility in carrying out
+analyses. This is in part due to the exploratory and interpretative nature of
+person-oriented approaches. To this end, presenting multiple models, as in
+Pastor et al. (2007), should be encouraged, rather than presenting one solution.
 In addition, having multiple analysts review the solutions found is encouraged.
-As part of this recommendation, we suggest that researchers consider how their analyses can be reproduced by other analysts outside of the research team: Sharing code and data is an important part of this work.
-Also related, researchers should consider how profiles are replicated across samples.
+As part of this recommendation, we suggest that researchers consider how their
+analyses can be reproduced by other analysts outside of the research team:
+Sharing code and data is an important part of this work. Also related,
+researchers should consider how profiles are replicated across samples.
 
-A second general recommendation considers more flexibly incorporating time, when data are collected across multiple time points, into analyses.
-ISOA groups all time points and doesn’t make distinctions.
-Other approaches perform analysis separately at different timepoints (Corpus & Wormington, 2014).
-Some integrate time as a part of the profiles, i.e.
-growth mixture modeling (groups of patterns), i.e.
-within-person growth modeling, where there are individual growth patterns.
-Research to date has yet to consider additional challenges in applying person centered approaches to longitudinal data.
-For instance, Schmidt et al (2018) use an Experience Sampling Method (ESM) approach to collecting data and used a person-centered approach to generate profiles of students in science class.
-This work does not account for student-level effects.
-In other words, it did not model the shared variance of multiple observations of the same student.
+A second general recommendation considers more flexibly incorporating time, when
+data are collected across multiple time points, into analyses. ISOA groups all
+time points and doesn’t make distinctions. Other approaches perform analysis
+separately at different timepoints (Corpus & Wormington, 2014). Some integrate
+time as a part of the profiles, i.e. growth mixture modeling (groups of
+patterns), i.e. within-person growth modeling, where there are individual growth
+patterns. Research to date has yet to consider additional challenges in applying
+person centered approaches to longitudinal data. For instance, Schmidt et al
+(2018) use an Experience Sampling Method (ESM) approach to collecting data and
+used a person-centered approach to generate profiles of students in science
+class. This work does not account for student-level effects. In other words, it
+did not model the shared variance of multiple observations of the same student.
 
-Third, best practices in within-person or longitudinal research call for modeling the nesting structure.
-However, researchers have yet to successfully incorporate this practice into the person centered approach.
-One way to approach this is to use cross-classified mixed effects models, as in Strati, Schmidt, and Maier (2017) and in Rosenberg (2018).
-In such approaches, dependencies in terms of, for example, individuals responding to (ESM) surveys at the same time and repeated responses being associated with the same individuals can both be modeled, although the effects of these two sources of dependencies are not nested as in very common uses of multi-level models, but rather are cross-classified.
-West, Welch, & Galecki (2014) have a description of the use of multi-level models with cross-classified data and tools (including those freely available through R) that can be used to estimate them.
+Third, best practices in within-person or longitudinal research call for
+modeling the nesting structure. However, researchers have yet to successfully
+incorporate this practice into the person centered approach. One way to approach
+this is to use cross-classified mixed effects models, as in Strati, Schmidt, and
+Maier (2017) and in Rosenberg (2018). In such approaches, dependencies in terms
+of, for example, individuals responding to (ESM) surveys at the same time and
+repeated responses being associated with the same individuals can both be
+modeled, although the effects of these two sources of dependencies are not
+nested as in very common uses of multi-level models, but rather are
+cross-classified. West, Welch, & Galecki (2014) have a description of the use of
+multi-level models with cross-classified data and tools (including those freely
+available through R) that can be used to estimate them.
 
@@
 
-**tidyLPA** has received a major update.
-Your old function calls might not work as expected anymore.
-The new version of tidyLPA aims to make Latent Profile Analysis (LPA) even more accessible to a broad audience, by means of a user-friendly interface, and to fit more seemlessly into a 'tidyverse' workflow.
+**tidyLPA** has received a major update. Your old function calls might not work
+as expected anymore. The new version of tidyLPA aims to make Latent Profile
+Analysis (LPA) even more accessible to a broad audience, by means of a
+user-friendly interface, and to fit more seemlessly into a 'tidyverse' workflow.
 
-As before, **tidyLPA** offers a parallel workflow for the open-source package Mclust, and the commercial package *Mplus*.
+As before, **tidyLPA** offers a parallel workflow for the open-source package
+Mclust, and the commercial package *Mplus*.
 
 These changes necessitated some simplification of the existing functions.
-Advanced options are still available, but require slightly more work on the user's part.
-The upside of this is that, for most users, the tidyLPA workflow and documentation are substantially simplified.
+Advanced options are still available, but require slightly more work on the
+user's part. The upside of this is that, for most users, the tidyLPA workflow
+and documentation are substantially simplified.
 
@@
 
-There have been major changes to the two primary functions, `estimate_profiles()` and `compare_solutions()`
+There have been major changes to the two primary functions,
+`estimate_profiles()` and `compare_solutions()`
 
-* `estimate_profiles()` now estimates any number of profiles, for any number of models, using any algorithm.
-It has a simplified user interface.
-This makes it a more general function, of which the former version (which could only estimate one number of profiles for one model) is a special, restricted case.
+  - `estimate_profiles()` now estimates any number of profiles, for any number
+    of models, using any algorithm. It has a simplified user interface. This
+    makes it a more general function, of which the former version (which could
+    only estimate one number of profiles for one model) is a special, restricted
+    case.
 
-* `compare_solutions()` now compares solutions estimated by estimate_profiles(), suggesting the optimal model and number of classes. This replaces its dual role of estimating AND comparing solutions.
-    `* `compare_solutions()` gives you the same set of fit indices for Mclust and Mplus. It provides a broader range of fit indices than before
-    * `compare_solutions()` calls `AHP()``, a function which uses the Automatic Hierarchical Process to determine the optimal number of classes, based on several reputable fit indices
+  - `compare_solutions()` now compares solutions estimated by
+    estimate\_profiles(), suggesting the optimal model and number of classes.
+    This replaces its dual role of estimating AND comparing solutions. ` *
+     `compare\_solutions()\` gives you the same set of fit indices for Mclust
+    and Mplus. It provides a broader range of fit indices than before
+      - `compare_solutions()` calls \`AHP()\`\`, a function which uses the
+        Automatic Hierarchical Process to determine the optimal number of
+        classes, based on several reputable fit indices
 
-* `plot_profiles()` now incorporates best practices for plotting latent profile analyses: Visualizing cluster centroids, standard errors, cluster variances, and classification uncertainty by means of weighted raw data
+  - `plot_profiles()` now incorporates best practices for plotting latent
+    profile analyses: Visualizing cluster centroids, standard errors, cluster
+    variances, and classification uncertainty by means of weighted raw data
 
-* `impute_missing()` offers two easy-to-use single imputation options, for use with the Mclust package (which does not natively handle missing data). Mplus handles partially missing data by default.
+  - `impute_missing()` offers two easy-to-use single imputation options, for use
+    with the Mclust package (which does not natively handle missing data). Mplus
+    handles partially missing data by default.
 
-The best way to understand these changes to **tidyLPA**, though, is by explaining the new workflow.
+The best way to understand these changes to **tidyLPA**, though, is by
+explaining the new workflow.
 
-## Example analysis using estimate_profiles()
+## Example analysis using estimate\_profiles()
 
@@
 
-```{r}
+``` {r}
 library(tidyverse)
 library(tidyLPA)
-
-pisaUSA15[1:100, ] %>%
-    select(broad_interest, enjoyment, self_efficacy) %>%
-    single_imputation() %>%
-    estimate_profiles(3)
 ```
 
-You can use Mplus simply by changing the package argument for `estimate_profiles()`:
+pisaUSA15[1:100, ] %\>% select(broad\_interest, enjoyment, self\_efficacy)
+%\>% single\_imputation() %\>% estimate\_profiles(3)
 
-```{r, eval = FALSE}
+``` 
+```
+
+You can use Mplus simply by changing the package argument for
+`estimate_profiles()`:
+
+``` {r, eval = FALSE}
 pisaUSA15[1:100, ] %>%
@@
 
-A simple summary of the analysis is printed to the console (and its posterior probability).
-The resulting object can be further passed down a pipeline to other functions, such as `plot`, `compare_solutions`, `get_data`, `get_fit`, etc.
-This is the "tidy" part, in that the function can be embedded in a tidy analysis pipeline.
+A simple summary of the analysis is printed to the console (and its posterior
+probability). The resulting object can be further passed down a pipeline to
+other functions, such as `plot`, `compare_solutions`, `get_data`, `get_fit`,
+etc. This is the "tidy" part, in that the function can be embedded in a tidy
+analysis pipeline.
 
-If you have Mplus installed, you can call the version of this function that uses MPlus in the same way, by adding the argument `package = "MplusAutomation`.
+If you have Mplus installed, you can call the version of this function that uses
+MPlus in the same way, by adding the argument `package = "MplusAutomation`.
 
@@
 
-```{r, eval = FALSE}
+``` {r, eval = FALSE}
 pisaUSA15[1:100, ] %>%
@@
 
-## Example analysis using compare_solutions()
+## Example analysis using compare\_solutions()
 
@@
 
-```{r, eval = TRUE}
+``` {r, eval = TRUE}
 pisaUSA15[1:100, ] %>%
@@
 
-* `compare_solutions()` no longer estimates profiles. It merely compares solutions estimated by `estimate_profiles()`
+  - `compare_solutions()` no longer estimates profiles. It merely compares
+    solutions estimated by `estimate_profiles()`
 
-* all Mplus-specific functions are deprecated. `estimate_profiles()` offers a unified interface to Mclust and Mplus. These Mplus-specific functions are deprecated:
+  - all Mplus-specific functions are deprecated. `estimate_profiles()` offers a
+    unified interface to Mclust and Mplus. These Mplus-specific functions are
+    deprecated:
 
@@
 
-* estimate_profiles and compare_solutions no longer select variables, passed as unquoted variable names, from 'df'. Instead, they use all variables in df. To select variables prior to analysis, use:
+  - estimate\_profiles and compare\_solutions no longer select variables, passed
+    as unquoted variable names, from 'df'. Instead, they use all variables in
+    df. To select variables prior to analysis, use:
 
@@
 
-* estimate_profiles and compare_solutions no longer center and scale the raw data. To center and scale variables prior to analysis, use scale(df, center = TRUE, scale = TRUE)
+  - estimate\_profiles and compare\_solutions no longer center and scale the raw
+    data. To center and scale variables prior to analysis, use scale(df, center
+    = TRUE, scale = TRUE)
 
-* estimate_profiles no longer accepts arguments specific to Mplus or Mclust.
-Instead, you can specify additional arguments directly to the core functions `mplusModeler()` and `Mclust()`.
-See the help file for these functions for more details.
-Deprecated arguments, whose functionality is supplanted by passing arguments directly to these core functions, are:
+  - estimate\_profiles no longer accepts arguments specific to Mplus or Mclust.
+    Instead, you can specify additional arguments directly to the core functions
+    `mplusModeler()` and `Mclust()`. See the help file for these functions for
+    more details. Deprecated arguments, whose functionality is supplanted by
+    passing arguments directly to these core functions, are:
 
@@
 
-# tidyLPA function with _mplus() appended require Mplus to use
+# tidyLPA function with \_mplus() appended require Mplus to use
 
-```{r, eval = FALSE}
+``` {r, eval = FALSE}
 estimate_profiles_mplus(data, var1, var2, var3, ariances = “free”, covariances = “equal”) %>% 
@@
 
-Person-oriented analysis ia way to consider how psychological constructs are experienced (and can be analyzed) together and at once.
-Though described in contrast to a variable-centered approach, scholars have pointed out how person-oriented approaches are complementary to variable-centered analyses (Marsh, Ludtke, Trautwein, & Morin, 2009).
-A person-oriented approach can help us to consider multiple variables together and at once and in a more dynamic way, reflected in the methodological approaches for cluster analysis and LPA that identify profiles of individuals responses.
+Person-oriented analysis ia way to consider how psychological constructs are
+experienced (and can be analyzed) together and at once. Though described in
+contrast to a variable-centered approach, scholars have pointed out how
+person-oriented approaches are complementary to variable-centered analyses
+(Marsh, Ludtke, Trautwein, & Morin, 2009). A person-oriented approach can help
+us to consider multiple variables together and at once and in a more dynamic
+way, reflected in the methodological approaches for cluster analysis and LPA
+that identify profiles of individuals responses.
 
-This manuscript provided an outline of how to get started with person-oriented analyses in an informed way.
-We provided a general overview of the methodology and described tools to carry out such an analysis.
-We also described specific tools, emphasizing freely-available open-source options that we have developed.
-Because of the inherently exploratory nature of person-oriented analysis, carrying out the analysis in a trustworthy and open way is particularly important.
-In this way, the interpretative aspect of settling on a solution shares some features of quantitative and qualitative research: The systematic nature of quantitative research methods (focused upon agreed-upon criteria such as likelihood-ratio tests) and qualitative research methods (focused upon the trustworthiness of both the analysis and the analyst) are important to consider when carrying out person-oriented analysis.
-Lastly, we made some general recommendations for future directions--and also highlighted some situations for which person-oriented approaches may not be the best and some cautions raised in past research regarding how such approaches are used.
+This manuscript provided an outline of how to get started with person-oriented
+analyses in an informed way. We provided a general overview of the methodology
+and described tools to carry out such an analysis. We also described specific
+tools, emphasizing freely-available open-source options that we have developed.
+Because of the inherently exploratory nature of person-oriented analysis,
+carrying out the analysis in a trustworthy and open way is particularly
+important. In this way, the interpretative aspect of settling on a solution
+shares some features of quantitative and qualitative research: The systematic
+nature of quantitative research methods (focused upon agreed-upon criteria such
+as likelihood-ratio tests) and qualitative research methods (focused upon the
+trustworthiness of both the analysis and the analyst) are important to consider
+when carrying out person-oriented analysis. Lastly, we made some general
+recommendations for future directions--and also highlighted some situations for
+which person-oriented approaches may not be the best and some cautions raised in
+past research regarding how such approaches are used.
 
-In conclusion, as use of person-oriented approaches expand, new questions and opportunities for carrying out research in a more holistic, dynamic way will be presented.
-Analyzing constructs together and at once is appealing to researchers, particularly those carrying out research in fields such as education for which communicating findings to stakeholders in a way that has the chance to impact practice is important.
-Our aim was not to suggest that such an approach is always the goal or should always be carried out, but rather to describe how researchers may get started in an informed way as researchers seek to understand how individuals interact, behave, and learn in ways that embraces the complexity of these experiences.
+In conclusion, as use of person-oriented approaches expand, new questions and
+opportunities for carrying out research in a more holistic, dynamic way will be
+presented. Analyzing constructs together and at once is appealing to
+researchers, particularly those carrying out research in fields such as
+education for which communicating findings to stakeholders in a way that has the
+chance to impact practice is important. Our aim was not to suggest that such an
+approach is always the goal or should always be carried out, but rather to
+describe how researchers may get started in an informed way as researchers seek
+to understand how individuals interact, behave, and learn in ways that embraces
+the complexity of these experiences.
 

### 2019-11-26 paper.rmd
minor edits

@@
 ---
-title             : "Doing Latent Profile Analysis in a more rigorous and reproducible way: A tutorial using the tidyLPA R package"
+title             : "Getting started with Latent Profile Analysis (LPA) in the psychological sciences: A tutorial using the tidyLPA R package that provides an interface to open-source and commercial software"
 shorttitle        : "Title"
@@
     affiliation   : "3"
+  - name          : "Rebecca Steingut"
+    affiliation   : "5"
+  - name          : "Stephanie Wormington"
+    affiliation   : "6"
 
@@
   - id            : "5"
+    institution   : "Columbia University"
+  - id            : "6"
+    institution   : "University of Virginia"
 
@@
 
-<!-- The questions on the Scimo per and post survey are from Ryan and Deci's "Learning Climate Questionnaire".  You can read a short description here: -->
+The questions on the Scimo per and post survey are from Ryan and Deci's "Learning Climate Questionnaire".  You can read a short description here:
 
-<!-- https://selfdeterminationtheory.org/pas-learning-climate/ -->
+https://selfdeterminationtheory.org/pas-learning-climate/
 
-<!-- It is 15 questions, but they identify 6 which can be used as  short-form.
+It is 15 questions, but they identify 6 which can be used as  short-form.
 Questions 1, 2, 4, 7, 10, 14.
 I have attached the PDF of the survey since you have to create an account on their website do download it.
-Below are the question text, along with the variable name in the Scimo data: -->
+Below are the question text, along with the variable name in the Scimo data:
 
-<!-- I feel that my instructor provides me choices and options.  t_choices1 -->
-<!-- I feel understood by my instructor.  t_understand1 -->
-<!-- My instructor conveyed confidence in my ability to do well in the course.  t_confidence1 -->
-<!-- My instructor encouraged me to ask questions.  t_encourage1 -->
-<!-- My instructor listens to how I would like to do things.  t_listen1 -->
-<!-- My instructor tries to understand how I see things before suggesting a new way to do things.  t_triesunder1 -->
+I feel that my instructor provides me choices and options.  t_choices1
+I feel understood by my instructor.  t_understand1
+My instructor conveyed confidence in my ability to do well in the course.  t_confidence1
+My instructor encouraged me to ask questions.  t_encourage1
+My instructor listens to how I would like to do things.  t_listen1
+My instructor tries to understand how I see things before suggesting a new way to do things.  t_triesunder1
+
+Also, is this tutorial going to be a journal paper, or something on github, etc.
+as a kind of manual for using LPA?
+If so, and you think I would be a good fit, I would love to help out on it if you have author space!
+
+Let me know if you need anything else.
 

### 2019-11-26 paper.rmd
re-write intro

--- 2019-11-26 paper.rmd
@@
 
-Person-oriented analysis involves a statistical approach to exploring
-psychological constructs in a way that captures and simplifies complexity. This
-approach can be used to consider the way in which psychological constructs, for
-example, are experienced together and at once. Thus, conceptually, this approach
-emphasizes how individuals go about interacting, socializing, and behaving in
-this way, experiencing these activities as a whole person. Person-oriented
-analysis, developed within developmental science, focuses on how individuals
-have experiences in a holistic way and also how these ways can be identified
-through common groups, or profiles.(Bergman & El-Khouri, 1997; Magnusson &
-Cairns, 1996). One way to distinguish the approach is to consider it in contrast
-to variable-centered approaches, such as those for which the covariability
-between variables, rather than the relations between profiles of individuals, is
-the focus (Bergman & Trost, 2006).
+For teachers, psychologists, researchers, and friends and family member, people 
+are highly-complex and not easily able to be characterized by a single 
+characteristic or personality trait. In the social sciences, broadly, and in 
+psychology, particularly, a statistical method that can be used to describe how 
+people, in their individual particularities, may have similarities on the basis 
+of some set of measures through which they can be grouped in meaningful, 
+distinctive ways. This approach, which has a provenance in developmental 
+approaches (Bergman & El-Khouri, 1997; Magnusson & Cairns, 1996), though is now 
+widely used in educational, social, personality, and other domains of psychology, 
+is an example of a general mixture model (Harring & Hodis, 2016; Pastor, Barron, 
+Miller, & Davis, 2007).
 
-Though studies examining learning from a person-oriented perspective are not
-very common, some examples include studies of intrinsic and extrinsic motivation
-(Corpus & Wormington, 2014; Hayenga & Corpus, 2010), profiles of achievement
-goals (see Wormington & Linnenbrink-Garcia, advance online publication, for a
-review), and epistemic cognition (Trevors, Kendeou, Braten, & Braasch, 2017).
-Recently, too, there have been excellent reviews of person-oriented analyses in
-motivation and educational psychology (Linnenbrink-Garcia & Wormington, 2017).
+In this tutorial, we aim to describe one of the most commonly-used--and relevant 
+to psychologists--application of the general mixture model, to cases for which 
+all of the variables for which (relatively) homogeneous groups are identified 
+from among a (relatively) heterogeneous sample are continuos, *latent profile 
+analysis* (LPA). After describing the method and some examples of its use, we 
+provide a tutorial for carrying out LPA in the context of a freely-available, 
+open-source statistical software package we created for R (R Core Team, 2019), 
+tidyLPA. Finally, we offer some informed recommendations for researchers aiming
+to use LPA in their applied work, and conclude with reflections on the role of 
+statistical software--especially software that is freely-available, open-source, 
+and highly-performant--in the psychological sciences.
 
-In this article, our goal is to review person-oriented analysis focused on
-getting started (on both knowing about and doing person-oriented analyses) in an
-informed way. More specifically, we start with an overview of the methodology,
-describe specific software tools, explain analytic choices made in the course of
-their use, and conclude with general recommendations One challenge facing those
-trying to learn about and apply person-oriented analysis in their research is
-the multitude of terms used for similar analytic approaches or ideas. Throughout
-this manuscript, we aim to point these out and explain when and why a variety of
-terms have been used. Thus, this manuscript is written for an informed and
-curious reader at any level interested in understanding this particular
-approach.
+<!-- Though studies examining learning from a person-oriented perspective are not -->
+<!-- very common, some examples include studies of intrinsic and extrinsic motivation -->
+<!-- (Corpus & Wormington, 2014; Hayenga & Corpus, 2010), profiles of achievement -->
+<!-- goals (see Wormington & Linnenbrink-Garcia, advance online publication, for a -->
+<!-- review), and epistemic cognition (Trevors, Kendeou, Braten, & Braasch, 2017). -->
+<!-- Recently, too, there have been excellent reviews of person-oriented analyses in -->
+<!-- motivation and educational psychology (Linnenbrink-Garcia & Wormington, 2017). -->
 
@@
 
-    A number of notes regarding person-oriented analyses are important to raise at the same time as their promise.
-Bauer (2007) notes that many samples of data can be usefully broken down into profiles, and that the addition of profiles will likely be suggested for reasons other than the samples coming from more than one distribution (i.e., due to non-normality in the variables measured).
-Bauer also cautions that profiles should not be reified; that profiles do not necessarily exist outside of the analysis that they should be interpreted more as useful interpretative devices.
-These cautions suggest that, in general, parsimony, interpretability, and a general sense that the profiles are not necessarily real, but are rather helpful analytic tools, should be both priorities for the analyst and the reader of studies using this approach.
+A number of notes regarding person-oriented analyses are important to raise at
+the same time as their promise. Bauer (2007) notes that many samples of data can
+be usefully broken down into profiles, and that the addition of profiles will
+likely be suggested for reasons other than the samples coming from more than one
+distribution (i.e., due to non-normality in the variables measured). Bauer also
+cautions that profiles should not be reified; that profiles do not necessarily
+exist outside of the analysis that they should be interpreted more as useful
+interpretative devices. These cautions suggest that, in general, parsimony,
+interpretability, and a general sense that the profiles are not necessarily
+real, but are rather helpful analytic tools, should be both priorities for the
+analyst and the reader of studies using this approach.
 

### 2019-11-26 paper.rmd
add rough draft

@@
 
-# Background
+# Introduction
 
@@
 
-## Overview of Methodologies
+<!-- ## Cluster Analysis -->
 
-At present, there are a number of approaches to carrying out person-oriented
-analysis when the indicators, of variables, used to find the groups are
-continuous. This section is focused upon two approaches: cluster analysis and
-Latent Profile Analysis (LPA). We focus on these because they are relatively
-simple and accessible forms of person-oriented analysis that can also
-accommodate cross sectional and longitudinal data alike. Just as a brief point
-of contrast, when the indicators for the groups are dichotomous (or categorical
-and can therefore be “dummy-coded” to be dichotomous), an approach closely
-related to LPA, Latent Class Analysis, can be used (cite PSU folks).
+<!-- Cluster analysis is an approach that has been frequently applied to analysis of -->
+<!-- topics in educational psychology (such as motivation and engagement) from a more -->
+<!-- holistic, whole-person approach. One particular way in which cluster analysis -->
+<!-- has been used has been as part of an approach termed I-States-as-Objects -->
+<!-- Analysis (ISOA). The purpose of ISOA is to identify profiles at specific time -->
+<!-- points, and then to explore how individuals do or do not shift between profiles -->
+<!-- across time points (needs cite). In their early work, Bergman and Magnusson -->
+<!-- (1999) and colleagues (1999) described a two-step approach to cluster analysis -->
+<!-- as part of their I-States-as-Objects analysis. In this approach, all of the -->
+<!-- observations are clustered. The key idea behind the approach is that each -->
+<!-- observation represents not a permanent group, but rather a status, or an -->
+<!-- i-state. These i-states, are considered independently for the purposes of -->
+<!-- constructing profiles. Then, shifts between these i-states across time (say, two -->
+<!-- or three time points) are evaluated. -->
 
-## Cluster Analysis
+<!-- Commonly, cluster analysis carried out as part of ISOA in fact involves two -->
+<!-- distinct clustering steps. This is because different clustering algorithms have -->
+<!-- different strengths--as well as weaknesses. In ISOA, the first step is Ward’s -->
+<!-- hierarchical cluster analysis (Bergman & El-Khouri, 1999). In this step, -->
+<!-- observations are split, step-by-step, based upon their multivariate similarity. -->
+<!-- The analyst provides the number of groups which will be used and interpreted, -->
+<!-- though technically most algorithms will continue splitting the groups until each -->
+<!-- observation makes up its own. This type of clustering is deterministic: The same -->
+<!-- results will be obtained each time it is run. In the second step of the cluster -->
+<!-- analysis, an iterative, k-means cluster analysis is carried out (Bergman and -->
+<!-- El-Khouri, 1999). For this step, the algorithm partitions the observations into -->
+<!-- groups, the success of the clustering based on a criterion such as the -->
+<!-- variability in the data explained by the groups, and then “moves,” when -->
+<!-- possible, the observations that are not in the group they are most similar to to -->
+<!-- the one that they are most similar to. When no more such moves are possible, the -->
+<!-- algorithm stops. On its own, k-means is not deterministic: Depending on the -->
+<!-- starting values, it can obtain different results the different times that it is -->
+<!-- run. So, the results from the hierarchical cluster analysis, the first step, are -->
+<!-- used as start values to initialize the k-means clustering. -->
 
-Cluster analysis is an approach that has been frequently applied to analysis of
-topics in educational psychology (such as motivation and engagement) from a more
-holistic, whole-person approach. One particular way in which cluster analysis
-has been used has been as part of an approach termed I-States-as-Objects
-Analysis (ISOA). The purpose of ISOA is to identify profiles at specific time
-points, and then to explore how individuals do or do not shift between profiles
-across time points (needs cite). In their early work, Bergman and Magnusson
-(1999) and colleagues (1999) described a two-step approach to cluster analysis
-as part of their I-States-as-Objects analysis. In this approach, all of the
-observations are clustered. The key idea behind the approach is that each
-observation represents not a permanent group, but rather a status, or an
-i-state. These i-states, are considered independently for the purposes of
-constructing profiles. Then, shifts between these i-states across time (say, two
-or three time points) are evaluated.
+<!-- The rationale behind using these two approaches is that the hierarchical -->
+<!-- clustering provides reasonable guesses for the profile membership, and then the -->
+<!-- iterative clustering optimizes the clustering solution. To sum up, then, cluster -->
+<!-- analysis as part of ISOA in fact involves two separate cluster analyses. The -->
+<!-- merits of this approach are to leverage the benefits of each: the iterative -->
+<!-- clustering starts with reasonable cluster solutions generated from the -->
+<!-- hierarchical clustering and optimizes them. -->
 
-Commonly, cluster analysis carried out as part of ISOA in fact involves two
-distinct clustering steps. This is because different clustering algorithms have
-different strengths--as well as weaknesses. In ISOA, the first step is Ward’s
-hierarchical cluster analysis (Bergman & El-Khouri, 1999). In this step,
-observations are split, step-by-step, based upon their multivariate similarity.
-The analyst provides the number of groups which will be used and interpreted,
-though technically most algorithms will continue splitting the groups until each
-observation makes up its own. This type of clustering is deterministic: The same
-results will be obtained each time it is run. In the second step of the cluster
-analysis, an iterative, k-means cluster analysis is carried out (Bergman and
-El-Khouri, 1999). For this step, the algorithm partitions the observations into
-groups, the success of the clustering based on a criterion such as the
-variability in the data explained by the groups, and then “moves,” when
-possible, the observations that are not in the group they are most similar to to
-the one that they are most similar to. When no more such moves are possible, the
-algorithm stops. On its own, k-means is not deterministic: Depending on the
-starting values, it can obtain different results the different times that it is
-run. So, the results from the hierarchical cluster analysis, the first step, are
-used as start values to initialize the k-means clustering.
+# Latent Profile Analysis
 
-The rationale behind using these two approaches is that the hierarchical
-clustering provides reasonable guesses for the profile membership, and then the
-iterative clustering optimizes the clustering solution. To sum up, then, cluster
-analysis as part of ISOA in fact involves two separate cluster analyses. The
-merits of this approach are to leverage the benefits of each: the iterative
-clustering starts with reasonable cluster solutions generated from the
-hierarchical clustering and optimizes them.
-
-## Latent Profile Analysis
-
-Another approach what is becoming increasingly common for carrying out
-person-oriented analysis is Latent Profile Analysis (LPA). LPA is a type of
-finite mixture model (Harring and Hodis, 2017). The goal of this approach is
-estimate the parameters for a number of distributions (often multivariate) from
-a single dataset. The approach can be thought of as seeking an answer to the
-question, Do the data in a sample come from more than one population? Thus, such
-an approach is model-based, and some descriptions in the literature refer to it
-as model-based clustering (Hennig, Meila, Murtagh, & Rocci, 2015; Scrucca, Fop,
-Murphy, & Raftery, 2017). Thus, one distinction between LPA and cluster analytic
-approaches is that LPA is model-based; instead of using algorithms to group
-together cases, LPA seeks to estimate parameters - in terms of variances and
-covariances and how they are the same or different across profiles - that best
-characterize the different distributions. Then, this approach seeks to assign to
-each observation a probability that the observation is a sample from the
-population associated with each profile (or mixture component).
+The goal of LPA is estimate the parameters for a number of distributions
+(typically multivariate) from a single dataset. Thus, such an approach is
+model-based, and some descriptions in the literature refer to it as model-based
+clustering (Hennig, Meila, Murtagh, & Rocci, 2015; Scrucca, Fop, Murphy, &
+Raftery, 2017). Thus, one distinction between LPA and other, similar cluster
+analytic approaches is that LPA is model-based; instead of using algorithms to
+group together cases, LPA seeks to estimate parameters - in terms of variances
+and covariances and how they are the same or different across profiles - that
+best characterize the different distributions. Then, this approach seeks to
+assign to each observation a probability that the observation is a sample from
+the population associated with each profile (or mixture component).
 
@@
 profiles; and the covariances of the variables used to create the profiles,
-which can be freely-estimated, estimated to be equal, or fixed to be zero. One
-challenge facing the analyst using LPA is that these parameters and the distinct
-model parameterizations that can be estimated is the different terminology used.
-As one example, Scrucca et al. (2017) refer to these parameterizations not in
-terms of whether and how parameters are estimated, but rather in terms of the
-geometric properties of the distributions that result from particular
-parameterizations. Muthen and Muthen (1997-2017) and others (Pastor et
-al., 2007) commonly refer to local independence to mean that the covariances are
-fixed to zero (also described as the specification of the covariance matrix as
-“diagonal,” because only the diagonal components, or the variances, are
-estimated). More information on the model parameterizations are discussed in the
-context of the software tool tidyLPA that we have developed.
+which can be freely-estimated, estimated to be equal, or fixed to be zero.
 
-This section outlines several LPA models and guidelines/examples for when they
-might be most appropriate in person-centered analysis. In general, as more
-parameters are estimated (i.e., those that are fixed to zero are estimated as
-being equal across profiles; or those estimated as being equal across profiles
-are freely-estimated across them), the model becomes more complex; the model may
-fit better, but also be overfit, meaning that the profiles identified may be
-challenging to replicate with another, separate data set. Even still,
-flexibility in terms of which models can be estimated also has affordances. For
-example, the varying means, equal variances, and covariances fixed to 0. A
-researcher might choose this model specification if she wants to model the
-variables to be used to create profiles that are independent. This model is very
-simple, as no covariances are estimated and the variances are estimated to be
-the same across profiles. As we estimate more parameters (and decrease the
-degrees of freedom), we are more likely to fit the data, but less likely to be
-able to replicate the model with a second set of data. In other words, more
-parameters may mean a loss of external validity. As we progress toward more
-complex models (with increasingly complex parameterization, i.e. models 3 and 4
-above), then we are more likely to fit the data better.
+We wish to note that challenge facing the analyst using LPA is that these
+parameters and the distinct model parameterizations that can be estimated is the
+different terminology used. As one example, Scrucca et al. (2017) refer to these
+parameterizations not in terms of whether and how parameters are estimated, but
+rather in terms of the geometric properties of the distributions that result
+from particular parameterizations. Muthen and Muthen (1997-2017) and others
+(Pastor et al., 2007) commonly refer to local independence to mean that the
+covariances are fixed to zero (also described as the specification of the
+covariance matrix as “diagonal,” because only the diagonal components, or the
+variances, are estimated).
 
-A number of notes regarding person-oriented analyses are important to raise at
-the same time as their promise. Bauer (2007) notes that many samples of data can
-be usefully broken down into profiles, and that the addition of profiles will
-likely be suggested for reasons other than the samples coming from more than one
-distribution (i.e., due to non-normality in the variables measured). Bauer also
-cautions that profiles should not be reified; that profiles do not necessarily
-exist outside of the analysis that they should be interpreted more as useful
-interpretative devices. These cautions suggest that, in general, parsimony,
-interpretability, and a general sense that the profiles are not necessarily
-real, but are rather helpful analytic tools, should be both priorities for the
-analyst and the reader of studies using this approach.
+In general, as more parameters are estimated (i.e., those that are fixed to zero
+are estimated as being equal across profiles; or those estimated as being equal
+across profiles are freely-estimated across them), the model becomes more
+complex; the model may fit better, but also be overfit, meaning that the
+profiles identified may be challenging to replicate with another, separate data
+set. Even still, flexibility in terms of which models can be estimated also has
+affordances. For example, the varying means, equal variances, and covariances
+fixed to 0. A researcher might choose this model specification if she wants to
+model the variables to be used to create profiles that are independent. This
+model is very simple, as no covariances are estimated and the variances are
+estimated to be the same across profiles. As we estimate more parameters (and
+decrease the degrees of freedom), we are more likely to fit the data, but less
+likely to be able to replicate the model with a second set of data. In other
+words, more parameters may mean a loss of external validity. As we progress
+toward more complex models (with increasingly complex parameterization), then we
+are more likely to fit the data better. More information on the model
+parameterizations are discussed in the context of the software tool tidyLPA that
+we have developed.
+
+A note of caution is warranted about LPA in the context of their potential.
+Bauer (2007) notes that many samples of data can be usefully broken down into
+profiles, and that the addition of profiles will likely be suggested for reasons
+other than the samples coming from more than one distribution (i.e., due to
+non-normality in the variables measured). Bauer also cautions that profiles
+should not be reified; that profiles do not necessarily exist outside of the
+analysis that they should be interpreted more as useful interpretative devices.
+These cautions suggest that, in general, parsimony, interpretability, and a
+general sense that the profiles are not necessarily real, but are rather helpful
+analytic tools, should be both priorities for the analyst and the reader of
+studies using this approach.
 
@@
 
-SPSS is a common tool to carry out cluster analyses. SPSS provides functionality
-to carry out both hierarchical and k-means cluster analysis. While somewhat
+## Other Tools
+
+SPSS is a common tool to carry out cluster analyses (but not LPA). While somewhat
 straightforward to carry out, particularly in SPSS’s graphical user interface
@@
 
-Accordingly, we worked to develop tools, initially used by the research team and
-now shared more widely, to provide basic functionality for each of the two
-primary approaches described in this section, cluster analysis and Latent
-Profile Analysis (LPA). In addition to being freely-available and open-source
-(i.e., anyone can inspect the source code), the tools described lend themselves
-to an open science. It is especially important to have an open transparent
-methodology when using person-centered approaches because of the high dependence
-on researcher judgment at multiple points in the process.
+Another common tool is MPlus (Muthen & Muthen, 20xx). MPlus is a commercial tool
+that provides functionality for many latent variable (and multi-level) models.
+We will speak more to MPlus in the section on tidyLPA, as our software provides
+an interface to both it and an open-source tool.
 
-Accordingly, we set out to develop packages that a) provided sensible defaults
-and were easy to use, but provided the option to access and modify all of the
+In R, a number of tools *can* be used to carry out LPA. OpenMx can be used for
+this purpose (and to specify almost any model possible to specify within a
+latent variable modeling approach). However, while OpenMx is very flexible, it
+can also be challenging to use. Other tools in R allow for estimating Gaussian
+mixture models, or models of multivariate Gaussian (or normal) distributions. In
+this framework, the term “mixture component” has a similar meaning to a profile.
+While much more constraining than the latent variable modeling framework, the
+approach is often similar or the same: the EM algorithm is used to (aim to)
+obtain the maximum likelihood estimates for the parameters being estimated. Like
+in the latent variable modeling framework, different models can be specified.
+
+In addition to following the same general approach, using tools that are
+designed for Gaussian mixture modeling have other benefits, some
+efficiency-related (see RMixMod, which uses compiled C++ code) and others in
+terms of ease-of-use (i.e., the plot methods built-in to RMixMod, mclust, and
+other tools). However, they also have some drawbacks, in that it can be
+difficult to translate between the model specifications, which are often
+described in terms of the geometric properties of the multivariate distributions
+being estimated (i.e., “spherical, equal volume”), rather than in terms of
+whether and how the means, variances, and covariances are estimated. They also
+may use different default settings (than those encountered in MPlus) in terms of
+the expectation-maximization algorithm, which can make comparing results across
+tools challenging.
+
+## tidyLPA
+
+Because of the limitations in other tools, we set out to develop a tool that a) 
+provided sensible defaults and were easy to use, but provided the option to access and modify all of the
 inputs to the model (i.e., low barrier, high ceiling), b) interfaced to existing
@@
 to specify, c) made it easy to carry-out fully-reproducible analyses and d) were
-well-documented. Both of these packages--one for two-step cluster analysis and
-one for LPA--are designed to do one thing well, namely, quickly and easily
-estimate profiles using cluster analysis or a model-based approach (Latent
-Profile Analysis).
+well-documented.
+
+This package focuses on models that are commonly specified as part of LPA.
+Because MPlus is so widely-used, it can be helpful to compare output from other
+software to MPlus. The functions in tidyLPA that use mclust have been
+benchmarked to MPlus for a series of simple models (with small datasets and for
+models with small numbers of profiles. This R Markdown output contains
+information on how mclust and Mplus compare. The R Markdown to generate the
+output is also available here, and, as long as you have purchased MPlus (and
+installed MplusAutomation), can be used to replicate all of the results for the
+benchmark. Note that most of the output is identical, though there are some
+differences in the hundreths decimal places for some. Because of differences in
+settings for the EM algorithm and particularly for the start values (random
+starts for MPlus and starting values from hierarchical clustering for mclust),
+differences may be expected for more complex data and models.
 
@@
 same maximum likelihood estimation procedure, they are benchmarked to produce
-the same output (see here). Also, note that we have described the model
+the same output. Also, note that we have described the model
 specifications with descriptions of what is estimated in terms of the variances
@@
 
-# Analytic Choices
+## Installation
+
+You can install tidyLPA from CRAN with:
+
+``` {r, eval = F}
+install.packages("tidyLPA")
+```
+
+You can also install the development version of tidyLPA from GitHub with:
+
+``` {r gh-installation, eval = FALSE}
+install.packages("devtools")
+devtools::install_github("data-edu/tidyLPA")
+```
+
+## Examples
+
+<!-- This is where we need to use an example from our own data -->
+
+### Mclust via tidyLPA
+
+Here is a brief example using the built-in `pisaUSA15` data set and variables
+for broad interest, enjoyment, and self-efficacy. Note that we first type the
+name of the data frame, followed by the unquoted names of the variables used to
+create the profiles. We also specify the number of profiles and the model. See
+`?estimate_profiles` for more details.
+
+In these examples, we pass the results of one function to the next by *piping*
+(using the `%>%` operator, loaded from the `dplyr` package). We pass the data to
+a function that selects relevant variables, and then to `estimate_profiles`:
+
+``` {r, message = F}
+library(tidyLPA)
+library(dplyr)
+```
+
+``` {r}
+pisaUSA15[1:100, ] %>%
+    select(broad_interest, enjoyment, self_efficacy) %>%
+    single_imputation() %>%
+    estimate_profiles(3)
+```
+
+### Mplus via tidyLPA
+
+We can use Mplus simply by changing the package argument for
+`estimate_profiles()`:
+
+``` {r, eval = FALSE}
+pisaUSA15[1:100, ] %>%
+    select(broad_interest, enjoyment, self_efficacy) %>%
+    single_imputation() %>%
+    estimate_profiles(3, package = "MplusAutomation")
+```
+
+A simple summary of the analysis is printed to the console (and its posterior
+probability). The resulting object can be further passed down a pipeline to
+other functions, such as `plot`, `compare_solutions`, `get_data`, `get_fit`,
+etc. This is the "tidy" part, in that the function can be embedded in a tidy
+analysis pipeline.
+
+If you have Mplus installed, you can call the version of this function that uses
+MPlus in the same way, by adding the argument `package = "MplusAutomation`.
+
+### Plotting the profiles
+
+We can plot the profiles by piping the output to `plot_profiles()`.
+
+``` {r}
+pisaUSA15[1:100, ] %>%
+    select(broad_interest, enjoyment, self_efficacy) %>%
+    single_imputation() %>%
+    scale() %>%
+    estimate_profiles(3) %>% 
+    plot_profiles()
+```
+
+## Model specification
+
+In addition to the number of profiles (specified with the `n_profiles`
+argument), the model can be specified in terms of whether and how the variable
+variances and covariances are estimated.
+
+The models are specified by passing arguments to the `variance` and `covariance`
+arguments. The possible values for these arguments are:
+
+  - `variances`: "equal" and "zero"
+  - `covariances`: "varying", "equal", and "zero"
+
+If no values are specified for these, then the variances are constrained to be
+equal across classes, and covariances are fixed to 0 (conditional independence
+of the indicators).
+
+These arguments allow for four models to be specified:
+
+  - Equal variances and covariances fixed to 0 (Model 1)
+  - Varying variances and covariances fixed to 0 (Model 2)
+  - Equal variances and equal covariances (Model 3)
+  - Varying variances and varying covariances (Model 6)
+
+Two additional models (Models 4 and 5) can be fit using MPlus. More information
+on the models can be found in the
+[vignette](https://data-edu.github.io/tidyLPA/articles/Introduction_to_tidyLPA.html).
+
+Here is an example of specifying a model with varying variances and covariances
+(Model 6; not run here):
+
+``` {r, eval = F}
+pisaUSA15[1:100, ] %>%
+    select(broad_interest, enjoyment, self_efficacy) %>%
+    single_imputation() %>%
+    estimate_profiles(3, 
+                      variances = "varying",
+                      covariances = "varying")
+```
+
+## Comparing a wide range of solutions
+
+The function `compare_solutions()` compares the fit of several estimated models,
+with varying numbers of profiles and model specifications:
+
+``` {r, eval = FALSE}
+pisaUSA15[1:100, ] %>%
+    select(broad_interest, enjoyment, self_efficacy) %>%
+    single_imputation() %>%
+    estimate_profiles(1:3, 
+                      variances = c("equal", "varying"),
+                      covariances = c("zero", "varying")) %>%
+    compare_solutions(statistics = c("AIC", "BIC"))
+```
+
+## Accessing values in the output
+
+A few helper functions are available to make it easier to work with the output
+of an analysis.
+
+`get_data()` returns the data:
+
+``` {r}
+m <- pisaUSA15[1:100, ] %>%
+    select(broad_interest, enjoyment, self_efficacy) %>%
+    single_imputation() %>%
+    estimate_profiles(3:4)
+```
+
+get\_data(m)
+
+``` 
+```
+
+We note that `get_data()` returns data in wide format when applied to an object
+of class tidyProfile (one element of a tidyLPA object), or when applied to a
+tidyLPA object of length one. `get_data()` returns long format when applied to a
+tidyLPA object containing multiple tidyProfile analyses (because then the wide
+format does not make sense).
+
+To transform data in the wide format into the long format, the `gather()`
+function from the **tidyr** package can be used, e.g.:
+
+``` {r}
+get_data(m) %>% 
+    tidyr::gather(Class_prob, Probability, contains("CPROB"))
+```
+
+# Recommendations Regarding Analytic Choices
 
@@
 
-## Whether to Use Cluster Analysis or LPA
+<!-- ## Whether to Use Cluster Analysis or LPA -->
 
-Whether to use cluster analysis or LPA depends in part on the nature of the data
-being investigated (cite the 2010 debate on this topic between Vermunt and
-others; MacLachlan, 2011; Steinley & Brusco, 2011a; Steinley & Brusco, 2011b;
-Vermunt, 2011). In some cases, cluster analysis and LPA solutions are very
-similar (when a constrained LPA model is fit). Broadly, if the variables used
-can be considered independent of one another, then cluster analysis may be
-justified in terms of reaching an optimal solution. However, LPA provides other
-benefits, such as the availability of information criteria (i.e., the Aikake
-Information Criteria [AIC] and the Bayesian Information Criteria [BIC] and
-statistical tests for the number of profiles (when a model is chosen, with a
-bootstrapped likelihood-ratio test). LPA also results in probabilistic
-assignments; this information can be used in subsequent analyses rather than the
-simple assignment (Collins & Lanza, 2010).
+<!-- Whether to use cluster analysis or LPA depends in part on the nature of the data -->
+<!-- being investigated (cite the 2010 debate on this topic between Vermunt and -->
+<!-- others; MacLachlan, 2011; Steinley & Brusco, 2011a; Steinley & Brusco, 2011b; -->
+<!-- Vermunt, 2011). In some cases, cluster analysis and LPA solutions are very -->
+<!-- similar (when a constrained LPA model is fit). Broadly, if the variables used -->
+<!-- can be considered independent of one another, then cluster analysis may be -->
+<!-- justified in terms of reaching an optimal solution. However, LPA provides other -->
+<!-- benefits, such as the availability of information criteria (i.e., the Aikake -->
+<!-- Information Criteria [AIC] and the Bayesian Information Criteria [BIC] and -->
+<!-- statistical tests for the number of profiles (when a model is chosen, with a -->
+<!-- bootstrapped likelihood-ratio test). LPA also results in probabilistic -->
+<!-- assignments; this information can be used in subsequent analyses rather than the -->
+<!-- simple assignment (Collins & Lanza, 2010). -->
 
-## How to Choose the Number of Profiles
+## How to Choose the Number of Profiles/Classes
 
@@
 interpretability in light of theory, parsimony, and evidence from
-cross-validation should be considered. For cluster analysis, the proportion of
-variance explained can be helpful for selecting a number of candidate profiles.
-Then, the analyst should focus on the interpretability of the solution as well
-as concerns of parsimony; if two profile solutions are similar, then the more
-simple solution may be preferred.
+cross-validation should be considered.
 
-## How to Choose Options for the Clustering Procedure or to Select the Model Parameterization
+<!-- ## How to Choose Options for the Clustering Procedure or to Select the Model Parameterization -->
 
-This applies most to LPA, but in cluster analysis, there are also options about
-the distance metric used. In LPA, we can determine which parameters - variance
-and covariance - are estimated to be the same or different across all profiles,
-and, in the case of covariance, whether it is fixed to zero. Of course, for the
-profiles we are most interested in, the mean is allowed to vary across profiles.
-The models with more parameters freely-estimated use more degrees of freedom;
-thus, similar to the number of profiles selected, the balance between adding
-more parameters (and having a more complex model) and interpretability and
-parsimony must be balanced. If a more complex model fits better and is
-interpretable and is justifiable in terms of the data collected, then it may be
-preferred to simpler models. If the two models are similar in terms of their
-fit, then the more simple, parsimonious model should be selected.
+<!-- This applies most to LPA, but in cluster analysis, there are also options about -->
+<!-- the distance metric used. In LPA, we can determine which parameters - variance -->
+<!-- and covariance - are estimated to be the same or different across all profiles, -->
+<!-- and, in the case of covariance, whether it is fixed to zero. Of course, for the -->
+<!-- profiles we are most interested in, the mean is allowed to vary across profiles. -->
+<!-- The models with more parameters freely-estimated use more degrees of freedom; -->
+<!-- thus, similar to the number of profiles selected, the balance between adding -->
+<!-- more parameters (and having a more complex model) and interpretability and -->
+<!-- parsimony must be balanced. If a more complex model fits better and is -->
+<!-- interpretable and is justifiable in terms of the data collected, then it may be -->
+<!-- preferred to simpler models. If the two models are similar in terms of their -->
+<!-- fit, then the more simple, parsimonious model should be selected. -->
 
-# More General Recommendations
+## More General Recommendations
 
@@
 available through R) that can be used to estimate them.
-
-# tidyLPA update
-
-## Rationale for the update
-
-**tidyLPA** has received a major update. Your old function calls might not work
-as expected anymore. The new version of tidyLPA aims to make Latent Profile
-Analysis (LPA) even more accessible to a broad audience, by means of a
-user-friendly interface, and to fit more seemlessly into a 'tidyverse' workflow.
-
-As before, **tidyLPA** offers a parallel workflow for the open-source package
-Mclust, and the commercial package *Mplus*.
-
-These changes necessitated some simplification of the existing functions.
-Advanced options are still available, but require slightly more work on the
-user's part. The upside of this is that, for most users, the tidyLPA workflow
-and documentation are substantially simplified.
-
-## Major changes
-
-There have been major changes to the two primary functions,
-`estimate_profiles()` and `compare_solutions()`
-
-  - `estimate_profiles()` now estimates any number of profiles, for any number
-    of models, using any algorithm. It has a simplified user interface. This
-    makes it a more general function, of which the former version (which could
-    only estimate one number of profiles for one model) is a special, restricted
-    case.
-
-  - `compare_solutions()` now compares solutions estimated by
-    estimate\_profiles(), suggesting the optimal model and number of classes.
-    This replaces its dual role of estimating AND comparing solutions. ` *
-     `compare\_solutions()\` gives you the same set of fit indices for Mclust
-    and Mplus. It provides a broader range of fit indices than before
-      - `compare_solutions()` calls \`AHP()\`\`, a function which uses the
-        Automatic Hierarchical Process to determine the optimal number of
-        classes, based on several reputable fit indices
-
-  - `plot_profiles()` now incorporates best practices for plotting latent
-    profile analyses: Visualizing cluster centroids, standard errors, cluster
-    variances, and classification uncertainty by means of weighted raw data
-
-  - `impute_missing()` offers two easy-to-use single imputation options, for use
-    with the Mclust package (which does not natively handle missing data). Mplus
-    handles partially missing data by default.
-
-The best way to understand these changes to **tidyLPA**, though, is by
-explaining the new workflow.
-
-## Example analysis using estimate\_profiles()
-
-A full analysis might look like the following:
-
-``` {r}
-library(tidyverse)
-library(tidyLPA)
-```
-
-pisaUSA15[1:100, ] %\>% select(broad\_interest, enjoyment, self\_efficacy)
-%\>% single\_imputation() %\>% estimate\_profiles(3)
-
-``` 
-```
-
-You can use Mplus simply by changing the package argument for
-`estimate_profiles()`:
-
-``` {r, eval = FALSE}
-pisaUSA15[1:100, ] %>%
-    select(broad_interest, enjoyment, self_efficacy) %>%
-    single_imputation() %>%
-    estimate_profiles(3, package = "MplusAutomation")
-```
-
-A simple summary of the analysis is printed to the console (and its posterior
-probability). The resulting object can be further passed down a pipeline to
-other functions, such as `plot`, `compare_solutions`, `get_data`, `get_fit`,
-etc. This is the "tidy" part, in that the function can be embedded in a tidy
-analysis pipeline.
-
-If you have Mplus installed, you can call the version of this function that uses
-MPlus in the same way, by adding the argument `package = "MplusAutomation`.
-
-We can plot the profiles by piping the output to `plot_profiles()`.
-
-``` {r, eval = FALSE}
-pisaUSA15[1:100, ] %>%
-    select(broad_interest, enjoyment, self_efficacy) %>%
-    single_imputation() %>%
-    scale() %>%
-    estimate_profiles(3) %>% 
-    plot_profiles()
-```
-
-## Example analysis using compare\_solutions()
-
-The function `compare_solutions()` compares the fit of several estimated models,
-with varying numbers of profiles and model specifications:
-
-``` {r, eval = TRUE}
-pisaUSA15[1:100, ] %>%
-    select(broad_interest, enjoyment, self_efficacy) %>%
-    single_imputation() %>%
-    estimate_profiles(1:3, 
-                      variances = c("equal", "varying"),
-                      covariances = c("zero", "varying")) %>%
-    compare_solutions(statistics = c("AIC", "BIC"))
-```
-
-## Deprecated functionality:
-
-  - `compare_solutions()` no longer estimates profiles. It merely compares
-    solutions estimated by `estimate_profiles()`
-
-  - all Mplus-specific functions are deprecated. `estimate_profiles()` offers a
-    unified interface to Mclust and Mplus. These Mplus-specific functions are
-    deprecated:
-
-    * `estimate_profiles_mplus()`
-        * `compare_profiles_mplus()`
-        * `plot_profiles_mplus()`
-
-  - estimate\_profiles and compare\_solutions no longer select variables, passed
-    as unquoted variable names, from 'df'. Instead, they use all variables in
-    df. To select variables prior to analysis, use:
-
-    * The dplyr function select(df, Your, Selected, Variable, Names)
-        * The base R function df[, c("Your", "Selected", "Variable", "Names")]
-
-  - estimate\_profiles and compare\_solutions no longer center and scale the raw
-    data. To center and scale variables prior to analysis, use scale(df, center
-    = TRUE, scale = TRUE)
-
-  - estimate\_profiles no longer accepts arguments specific to Mplus or Mclust.
-    Instead, you can specify additional arguments directly to the core functions
-    `mplusModeler()` and `Mclust()`. See the help file for these functions for
-    more details. Deprecated arguments, whose functionality is supplanted by
-    passing arguments directly to these core functions, are:
-
-For Mclust:
-
-    * prior_control
-
-For Mplus:
-
-    * starts
-    * m_iterations
-    * st_iterations
-    * convergence_criterion
-    * optseed
-    * cluster_ID
-    * include_VLMR
-    * include_BLRT (but BLRT is returned by default now)
-
-# tidyLPA function with \_mplus() appended require Mplus to use
-
-``` {r, eval = FALSE}
-estimate_profiles_mplus(data, var1, var2, var3, ariances = “free”, covariances = “equal”) %>% 
-    plot_profiles_mplus()
-```
 

### 2019-12-01 paper.rmd
update front matter

@@
 
-<!-- The questions on the Scimo per and post survey are from Ryan and Deci's "Learning Climate Questionnaire".  You can read a short description here: -->
-
-<!-- https://selfdeterminationtheory.org/pas-learning-climate/ -->
-
-<!-- It is 15 questions, but they identify 6 which can be used as  short-form.
-Questions 1, 2, 4, 7, 10, 14.
-I have attached the PDF of the survey since you have to create an account on their website do download it.
-Below are the question text, along with the variable name in the Scimo data: -->
-
-<!-- I feel that my instructor provides me choices and options.  t_choices1 -->
-<!-- I feel understood by my instructor.  t_understand1 -->
-<!-- My instructor conveyed confidence in my ability to do well in the course.  t_confidence1 -->
-<!-- My instructor encouraged me to ask questions.  t_encourage1 -->
-<!-- My instructor listens to how I would like to do things.  t_listen1 -->
-<!-- My instructor tries to understand how I see things before suggesting a new way to do things.  t_triesunder1 -->
-
-<!-- tidyLPA Tutorial for AMPPS -->
-<!-- https://www.psychologicalscience.org/publications/ampps/ampps-submission-guidelines#NEMP -->
-<!-- Tutorials -->
-<!-- Tutorials provide hands-on, practical guidance for researchers.
-Any topic that could enhance research practices or methods might be suitable for a Tutorial, provided that the material covered in the Tutorial would be relevant to and useful for the journal’s broad readership; Tutorials of more narrow interest or of relevance to only one subfield or literature typically are not appropriate for AMPPS.
-AMPPS welcomes Tutorials that focus on helping researchers learn to use statistical tools, improve their statistical practices and intuitions, better their data-management and lab practices, enhance the reliability and reproducibility of their research, engage in transparent and open practices, and so on.
-Tutorials often include dynamic, interactive content and should provide concrete guidance rather than solely abstract principles.
-Some Tutorials will be solicited by the editorial team, and the team welcomes suggestions and proposals for Tutorials.
--->
-<!-- Most Tutorials should be brief (< 3,000 words), but they may be longer if necessary to explain the content fully and make it accessible to and usable by readers. -->
-<!-- The introduction to a Tutorial typically should be no more than one to two paragraphs long (< 500 words) and should not include an extensive literature review.
-The introduction should explain the motivation for the Tutorial and highlight how learning the contents will benefit readers.
--->
-<!-- Tutorials should have a brief summary of their contents, rather than a General Discussion section. -->
-<!-- Tutorials should be accompanied by publicly available code and all resources necessary for researchers (and reviewers) to follow the text. -->
-<!-- Tutorials can include a list of additional resources (e.g., citations and links) for readers who would like to learn more. -->
-
 # Introduction
@@
 of some set of measures through which they can be grouped in meaningful,
-distinctive ways. This approach, which has a provenance in developmental 
-approaches (Bergman & El-Khouri, 1997; Magnusson & Cairns, 1996), though is now 
-widely used in educational, social, personality, and other domains of psychology, 
-is an example of a general mixture model (Harring & Hodis, 2016; Pastor, Barron, 
-Miller, & Davis, 2007).
+distinctive ways. This approach, which has a provenance in developmental (or
+*person-oriented*) approaches (Bergman & El-Khouri, 1997; Magnusson &
+Cairns, 1996; see Linnenbrink-Garcia & Wormington, 2017, for a recent review),
+though is now widely used in educational, social, personality, and other domains
+of psychology, is an example of a general mixture model (Harring & Hodis, 2016;
+Pastor, Barron, Miller, & Davis, 2007).
 
@@
 from among a (relatively) heterogeneous sample are continuos, *latent profile
-analysis* (LPA). After describing the method and some examples of its use, we 
-provide a tutorial for carrying out LPA in the context of a freely-available, 
-open-source statistical software package we created for R (R Core Team, 2019), 
-tidyLPA. Finally, we offer some informed recommendations for researchers aiming
-to use LPA in their applied work, and conclude with reflections on the role of 
+analysis* (LPA). We note that scholars using the approach, especially in
+developmental and educational psychology, use cluster analytic methods. Such an
+approach has some similarities with LPA (CITE paper showing how k-means is
+similar to the simplest model specification of LPA?), though LPA also has a
+number of benefits (Linnenbrink-Garcia & Wormington, 2017), and we we focus on
+LPA in this paper.
+
+After describing the method and some examples of its use, we provide a tutorial
+for carrying out LPA in the context of a freely-available, open-source
+statistical software package we created for R (R Core Team, 2019), tidyLPA.
+Finally, we offer some informed recommendations for researchers aiming to use
+LPA in their applied work, and conclude with reflections on the role of
 statistical software--especially software that is freely-available, open-source,
 and highly-performant--in the psychological sciences.
-
-<!-- Though studies examining learning from a person-oriented perspective are not -->
-<!-- very common, some examples include studies of intrinsic and extrinsic motivation -->
-<!-- (Corpus & Wormington, 2014; Hayenga & Corpus, 2010), profiles of achievement -->
-<!-- goals (see Wormington & Linnenbrink-Garcia, advance online publication, for a -->
-<!-- review), and epistemic cognition (Trevors, Kendeou, Braten, & Braasch, 2017). -->
-<!-- Recently, too, there have been excellent reviews of person-oriented analyses in -->
-<!-- motivation and educational psychology (Linnenbrink-Garcia & Wormington, 2017). -->
-
-<!-- ## Cluster Analysis -->
-
-<!-- Cluster analysis is an approach that has been frequently applied to analysis of -->
-<!-- topics in educational psychology (such as motivation and engagement) from a more -->
-<!-- holistic, whole-person approach. One particular way in which cluster analysis -->
-<!-- has been used has been as part of an approach termed I-States-as-Objects -->
-<!-- Analysis (ISOA). The purpose of ISOA is to identify profiles at specific time -->
-<!-- points, and then to explore how individuals do or do not shift between profiles -->
-<!-- across time points (needs cite). In their early work, Bergman and Magnusson -->
-<!-- (1999) and colleagues (1999) described a two-step approach to cluster analysis -->
-<!-- as part of their I-States-as-Objects analysis. In this approach, all of the -->
-<!-- observations are clustered. The key idea behind the approach is that each -->
-<!-- observation represents not a permanent group, but rather a status, or an -->
-<!-- i-state. These i-states, are considered independently for the purposes of -->
-<!-- constructing profiles. Then, shifts between these i-states across time (say, two -->
-<!-- or three time points) are evaluated. -->
-
-<!-- Commonly, cluster analysis carried out as part of ISOA in fact involves two -->
-<!-- distinct clustering steps. This is because different clustering algorithms have -->
-<!-- different strengths--as well as weaknesses. In ISOA, the first step is Ward’s -->
-<!-- hierarchical cluster analysis (Bergman & El-Khouri, 1999). In this step, -->
-<!-- observations are split, step-by-step, based upon their multivariate similarity. -->
-<!-- The analyst provides the number of groups which will be used and interpreted, -->
-<!-- though technically most algorithms will continue splitting the groups until each -->
-<!-- observation makes up its own. This type of clustering is deterministic: The same -->
-<!-- results will be obtained each time it is run. In the second step of the cluster -->
-<!-- analysis, an iterative, k-means cluster analysis is carried out (Bergman and -->
-<!-- El-Khouri, 1999). For this step, the algorithm partitions the observations into -->
-<!-- groups, the success of the clustering based on a criterion such as the -->
-<!-- variability in the data explained by the groups, and then “moves,” when -->
-<!-- possible, the observations that are not in the group they are most similar to to -->
-<!-- the one that they are most similar to. When no more such moves are possible, the -->
-<!-- algorithm stops. On its own, k-means is not deterministic: Depending on the -->
-<!-- starting values, it can obtain different results the different times that it is -->
-<!-- run. So, the results from the hierarchical cluster analysis, the first step, are -->
-<!-- used as start values to initialize the k-means clustering. -->
-
-<!-- The rationale behind using these two approaches is that the hierarchical -->
-<!-- clustering provides reasonable guesses for the profile membership, and then the -->
-<!-- iterative clustering optimizes the clustering solution. To sum up, then, cluster -->
-<!-- analysis as part of ISOA in fact involves two separate cluster analyses. The -->
-<!-- merits of this approach are to leverage the benefits of each: the iterative -->
-<!-- clustering starts with reasonable cluster solutions generated from the -->
-<!-- hierarchical clustering and optimizes them. -->
 

### 2019-12-02 paper.rmd
rework front end and move content to notes

@@
 ---
-title             : "Doing Latent Profile Analysis in a more rigorous and reproducible way: A tutorial using the tidyLPA R package"
-shorttitle        : "Title"
+title             : "Carrying out Latent Profile Analysis in Accordance With Best Practices: A Tutorial Using the tidyLPA R package"
+shorttitle        : "tidyLPA"
 
@@
 statistical software package we created for R (R Core Team, 2019), tidyLPA.
-Finally, we offer some informed recommendations for researchers aiming to use
-LPA in their applied work, and conclude with reflections on the role of
-statistical software--especially software that is freely-available, open-source,
-and highly-performant--in the psychological sciences.
+Finally, we offer some ideas about best practices and informed recommendations
+for researchers aiming to use LPA in their applied work, and conclude with
+reflections on the role of statistical software--especially software that is
+freely-available, open-source, and highly-performant--in the psychological
+sciences.
 
@@
 
-A note of caution is warranted about LPA in the context of their potential.
-Bauer (2007) notes that many samples of data can be usefully broken down into
-profiles, and that the addition of profiles will likely be suggested for reasons
-other than the samples coming from more than one distribution (i.e., due to
-non-normality in the variables measured). Bauer also cautions that profiles
-should not be reified; that profiles do not necessarily exist outside of the
-analysis that they should be interpreted more as useful interpretative devices.
-These cautions suggest that, in general, parsimony, interpretability, and a
-general sense that the profiles are not necessarily real, but are rather helpful
-analytic tools, should be both priorities for the analyst and the reader of
-studies using this approach.
+<!-- I think we move this to the discussion in light of best practices -->
+
+A note of caution is warranted about LPA in the context of their characteristics
+and strengths. Bauer (2007) notes that many samples of data can be usefully
+broken down into profiles, and that the addition of profiles will likely be
+suggested for reasons other than the samples coming from more than one
+distribution (i.e., due to non-normality in the variables measured). Bauer also
+cautions that profiles should not be reified; that profiles do not necessarily
+exist outside of the analysis that they should be interpreted more as useful
+interpretative devices. These cautions suggest that, in general, parsimony,
+interpretability, and a general sense that the profiles are not necessarily
+real, but are rather helpful analytic tools, should be both priorities for the
+analyst and the reader of studies using this approach.
 
@@
 
-SPSS is a common tool to carry out cluster analyses (but not LPA). While somewhat
-straightforward to carry out, particularly in SPSS’s graphical user interface
-(GUI), there are some challenges to use of this approach. The GUI in SPSS can be
-challenging, even for the most able analyst, to be able to document every step
-with syntax, and so reproducing the entire analysis efficiently can be a
-challenge, both for the analyst exploring various solutions and for the reviewer
-looking to replicate this work. Additionally, SPSS is commercial software (and
-is expensive), and so analysts without the software cannot carry out this
+SPSS is a common tool to carry out cluster analyses (but not LPA). While
+somewhat straightforward to carry out, particularly in SPSS’s graphical user
+interface (GUI), there are some challenges to use of this approach. The GUI in
+SPSS can be challenging, even for the most able analyst, to be able to document
+every step with syntax, and so reproducing the entire analysis efficiently can
+be a challenge, both for the analyst exploring various solutions and for the
+reviewer looking to replicate this work. Additionally, SPSS is commercial
+software (and is expensive), and so analysts access to it cannot carry out this
 analysis.
 
-Another common tool is MPlus (Muthen & Muthen, 20xx). MPlus is a commercial tool
+Another common tool is MPlus (Muthen & Muthen, 2019). MPlus is a commercial tool
 that provides functionality for many latent variable (and multi-level) models.
@@
 mixture models, or models of multivariate Gaussian (or normal) distributions. In
-this framework, the term “mixture component” has a similar meaning to a profile.
-While much more constraining than the latent variable modeling framework, the
-approach is often similar or the same: the EM algorithm is used to (aim to)
-obtain the maximum likelihood estimates for the parameters being estimated. Like
-in the latent variable modeling framework, different models can be specified.
-
-In addition to following the same general approach, using tools that are
-designed for Gaussian mixture modeling have other benefits, some
-efficiency-related (see RMixMod, which uses compiled C++ code) and others in
-terms of ease-of-use (i.e., the plot methods built-in to RMixMod, mclust, and
-other tools). However, they also have some drawbacks, in that it can be
-difficult to translate between the model specifications, which are often
-described in terms of the geometric properties of the multivariate distributions
-being estimated (i.e., “spherical, equal volume”), rather than in terms of
-whether and how the means, variances, and covariances are estimated. They also
-may use different default settings (than those encountered in MPlus) in terms of
-the expectation-maximization algorithm, which can make comparing results across
-tools challenging.
+addition to following the same general approach, using tools that are designed
+for Gaussian mixture modeling have other benefits, some efficiency-related (see
+RMixMod, which uses compiled C++ code) and others in terms of ease-of-use (i.e.,
+the plot methods built-in to RMixMod, mclust, and other tools). However, they
+also have some drawbacks, in that it can be difficult to translate between the
+model specifications, which are often described in terms of the geometric
+properties of the multivariate distributions being estimated (i.e., “spherical,
+equal volume”), rather than in terms of whether and how the means, variances,
+and covariances are estimated. They also may use different default settings
+(than those encountered in MPlus) in terms of the expectation-maximization
+algorithm, which can make comparing results across tools challenging.
 
@@
 Because of the limitations in other tools, we set out to develop a tool that a)
-provided sensible defaults and were easy to use, but provided the option to access and modify all of the
-inputs to the model (i.e., low barrier, high ceiling), b) interfaced to existing
-tools,, and are able to translate between what existing tools are capable of and
-what researchers and analysts carrying-out person-oriented analyses would like
-to specify, c) made it easy to carry-out fully-reproducible analyses and d) were
-well-documented.
+provided sensible defaults and were easy to use, but provided the option to
+access and modify all of the inputs to the model (i.e., low barrier, high
+ceiling), b) interfaced to existing tools, and are able to translate between
+what existing tools are capable of and what researchers and analysts
+carrying-out person-oriented analyses would like to specify, c) made it easy to
+carry-out fully-reproducible analyses and d) were well-documented.
 
@@
 
+``` {r, message = F}
+library(tidyLPA)
+library(dplyr)
+```
+
 ## Examples
@@
 a function that selects relevant variables, and then to `estimate_profiles`:
-
-``` {r, message = F}
-library(tidyLPA)
-library(dplyr)
-```
 
@@
 
-A simple summary of the analysis is printed to the console (and its posterior
-probability). The resulting object can be further passed down a pipeline to
-other functions, such as `plot`, `compare_solutions`, `get_data`, `get_fit`,
+A simple summary of the analysis is printed to the console. 
+The resulting object can be further passed down a pipeline to
+other functions, such as `compare_solutions`, `get_data`, `get_fit`,
 etc. This is the "tidy" part, in that the function can be embedded in a tidy

### 2019-12-08 paper.rmd
revise intro

@@
 keywords          : "Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust"
-wordcount         : "4,882 (need to shorten to ~ 3,000)"
+wordcount         : ""
 
@@
 
-For teachers, psychologists, researchers, and friends and family member, people
-are highly-complex and not easily able to be characterized by a single
-characteristic or personality trait. In the social sciences, broadly, and in
-psychology, particularly, a statistical method that can be used to describe how
+In statistics classes, textbooks, and workshops, an example like the following is common:
+
+> Grades are normally distributed, with $\mu$ = 75, $\sigma$ = 5.
+
+Students, may (reasonably) ask, though, are grades *really* normally distributed?
+
+```{r, include = FALSE}
+library(simstudy)
+
+def <- defData(varname = "x1", formula = 65, variance = 30)
+def <- defData(def, varname = "x2", formula = 85, variance = 23)
+def <- defData(def, varname = "xMix", 
+               formula = "x1 | .5 + x2 | .5", 
+               dist = "mixture")
+
+set.seed(2716)
+dx <- genData(1000, def)
+```
+
+Are their one distribution?
+
+```{r, echo - FALSe}
+library(ggplot2)
+ggplot(dx, aes(x = xMix)) +
+  geom_density() +
+  papaja::theme_apa() +
+  ylim(0, .10)
+```
+
+Or two?
+
+```{r, echo - FALSe}
+library(tidyverse)
+dx %>% 
+  gather(key, val, x1:x2) %>% 
+  ggplot(aes(x = val, fill = key)) +
+  geom_density(alpha = .3) +
+  papaja::theme_apa() +
+  ylim(0, .10)
+```
+
+This kind of distribution (that is *bimodal*) is not exclusive to grades. For teachers, psychologists, researchers, and friends and family members, people
+are highly-complex and not easily able to be characterized by *one*
+characteristic or personality trait--and its distribution. In the social sciences, broadly, and in psychology, particularly, a statistical method that can be used to describe how
 people, in their individual particularities, may have similarities on the basis
@@
 *person-oriented*) approaches (Bergman & El-Khouri, 1997; Magnusson &
-Cairns, 1996; see Linnenbrink-Garcia & Wormington, 2017, for a recent review),
-though is now widely used in educational, social, personality, and other domains
-of psychology, is an example of a general mixture model (Harring & Hodis, 2016;
-Pastor, Barron, Miller, & Davis, 2007).
+Cairns, 1996; see Linnenbrink-Garcia & Wormington, 2017, for a recent review) is
+an example of a general mixture model (Harring & Hodis, 2016; Pastor, Barron, Miller, & Davis, 2007).
 

### 2019-12-08 paper.rmd
update citations

@@
 
-bibliography      : ["r-references.bib"]
+bibliography      : ["r-references.bib", "citations.bib"]
 
@@
 
-```{r, echo - FALSe}
-library(ggplot2)
+```{r, echo - FALSE}
+library(tidyverse)
 ggplot(dx, aes(x = xMix)) +
@@
 
-```{r, echo - FALSe}
-library(tidyverse)
+```{r, echo = FALSE}
 dx %>% 
@@
 carrying-out person-oriented analyses would like to specify, c) made it easy to
-carry-out fully-reproducible analyses and d) were well-documented.
+carry-out fully-reproducible analyses and d) were well-documented. To do so, we 
+created tidyLPA [@R-tidyLPA].
 
@@
 ```
-Bergman, L. R., & Magnusson, D. (1997). A person-oriented approach in research on developmental psychopathology. Development and psychopathology, 9(2), 291-319.  
-Bergman, L.
-R., & Trost, K.
-(2006).
-The person-oriented versus the variable-oriented approach: Are they complementary, opposites, or exploring different worlds?.
-Merrill-Palmer Quarterly, 52(3), 601-632.
-Collins, L. M., & Lanza, S. T. (2010). Latent class and latent transition analysis: With applications in the social, behavioral, and health sciences (Vol. 718). John Wiley & Sons.
-Corpus, J. H., & Wormington, S. V. (2014). Profiles of intrinsic and extrinsic motivations in elementary school: A longitudinal analysis. The Journal of Experimental Education, 82(4), 480-501.
-Corpus, J. H., & Wormington, S. V. (2014). Profiles of intrinsic and extrinsic motivations in elementary school: A longitudinal analysis. The Journal of Experimental Education, 82(4), 480-501.
-Harring, J. R., & Hodis, F. A. (2016). Mixture modeling: Applications in educational psychology. Educational Psychologist, 51(3-4), 354-367.
-Hayenga, A.
-O., & Corpus, J.
-H.
-(2010).
-Profiles of intrinsic and extrinsic motivations: A person-centered approach to motivation and achievement in middle school.
-Motivation and Emotion, 34(4), 371-383.
-Linnenbrink-Garcia, L., & Wormington, S.
-V.
-(2017).
-Key challenges and potential solutions for studying the complexity of motivation in schooling: An integrative, dynamic person-oriented perspective.
-British Journal of Educational Psychology monograph series II.
-Psychological aspects of education: Current trends—the role of competence beliefs in teaching and learning.
-Chichester, UK: Wiley.
-Magnusson, D., & Cairns, R. B. (1996). Developmental science: Toward a unified framework. Cambridge, England: Cambridge University Press.  
-Marsh, H.
-W., Lüdtke, O., Trautwein, U., & Morin, A.
-J.
-(2009).
-Classical latent profile analysis of academic self-concept dimensions: Synergy of person-and variable-centered approaches to theoretical models of self-concept.
-Structural Equation Modeling, 16(2), 191-225.
-McLachlan, G. J. (2011). Commentary on Steinley and Brusco (2011): Recommendations and cautions.
-Muthen, L. K., & Muthen, B. O. (1997-2017). Mplus User’s Guide. Los Angeles, CA: Muthén & Muthén.
-Rosenberg, J.
-M., Schmidt, J.
-A., Beymer, P.
-N., & Steingut, R.
-R.
-(2018).
-Interface to mclust to easily carry out Latent Profile Analysis [Statistical software for R].
-https://github.com/jrosen48/tidyLPA
-Rosenberg, J. M., Schmidt, J. A., Beymer, P. N., & Steingut, R. R. (2017). prcr: Person-Centered Analysis. R package version 0.1.5. https://CRAN.R-project.org/package=prcr
-Scrucca L., Fop M., Murphy T. B. and Raftery A. E. (2017) mclust 5: clustering, classification and density estimation using Gaussian finite mixture models The R Journal 8/1, pp. 205-233
-Hennig, C., Meila, M., Murtagh, F., & Rocci, R. (Eds.). (2015). Handbook of cluster analysis. CRC Press.
-Steinley, D., & Brusco, M. J. (2011). Evaluating mixture modeling for clustering: Recommendations and cautions. Psychological Methods, 16(1), 63.
-Steinley, D., & Brusco, M. J. (2011). K-means clustering and mixture model clustering: Reply to McLachlan (2011) and Vermunt (2011).
-Trevors, G. J., Kendeou, P., Bråten, I., & Braasch, J. L. (2017). Adolescents’ epistemic profiles in the service of knowledge revision. Contemporary Educational Psychology, 49, 107-120.
-Vermunt, J. K. (2011). K-means may perform as well as mixture model clustering but may also be much worse: Comment on Steinley and Brusco (2011).
-West, B. T., Welch, K. B., & Galecki, A. T. (2014). Linear mixed models: a practical guide using statistical software. Chapman and Hall/CRC.
-Wickham, H. (2014). Tidy data. Journal of Statistical Software, 59(10), 1-23.
 

### 2019-12-11 paper.rmd
remove commented out sections

@@
 ``` {r gh-installation, eval = FALSE}
-install.packages("devtools")
+# install.packages("devtools")
 devtools::install_github("data-edu/tidyLPA")
@@
 
-get\_data(m)
-
-``` 
+```{r}
+get_data(m)
 ```
@@
 # Recommendations Regarding Analytic Choices
+
+## GRoLTs checklist
+
+```{r, echo = FALSE}
+t <- read_csv("GRoLTS-applicability-to-LPA.csv")
+t
+```
 
@@
 helpful for obtaining solutions when the variables are on very different
-scales..
-
-<!-- ## Whether to Use Cluster Analysis or LPA -->
-
-<!-- Whether to use cluster analysis or LPA depends in part on the nature of the data -->
-<!-- being investigated (cite the 2010 debate on this topic between Vermunt and -->
-<!-- others; MacLachlan, 2011; Steinley & Brusco, 2011a; Steinley & Brusco, 2011b; -->
-<!-- Vermunt, 2011). In some cases, cluster analysis and LPA solutions are very -->
-<!-- similar (when a constrained LPA model is fit). Broadly, if the variables used -->
-<!-- can be considered independent of one another, then cluster analysis may be -->
-<!-- justified in terms of reaching an optimal solution. However, LPA provides other -->
-<!-- benefits, such as the availability of information criteria (i.e., the Aikake -->
-<!-- Information Criteria [AIC] and the Bayesian Information Criteria [BIC] and -->
-<!-- statistical tests for the number of profiles (when a model is chosen, with a -->
-<!-- bootstrapped likelihood-ratio test). LPA also results in probabilistic -->
-<!-- assignments; this information can be used in subsequent analyses rather than the -->
-<!-- simple assignment (Collins & Lanza, 2010). -->
+scales.
 
@@
 
-<!-- ## How to Choose Options for the Clustering Procedure or to Select the Model Parameterization -->
+## How to Choose Options for the Clustering Procedure or to Select the Model Parameterization
 
-<!-- This applies most to LPA, but in cluster analysis, there are also options about -->
-<!-- the distance metric used. In LPA, we can determine which parameters - variance -->
-<!-- and covariance - are estimated to be the same or different across all profiles, -->
-<!-- and, in the case of covariance, whether it is fixed to zero. Of course, for the -->
-<!-- profiles we are most interested in, the mean is allowed to vary across profiles. -->
-<!-- The models with more parameters freely-estimated use more degrees of freedom; -->
-<!-- thus, similar to the number of profiles selected, the balance between adding -->
-<!-- more parameters (and having a more complex model) and interpretability and -->
-<!-- parsimony must be balanced. If a more complex model fits better and is -->
-<!-- interpretable and is justifiable in terms of the data collected, then it may be -->
-<!-- preferred to simpler models. If the two models are similar in terms of their -->
-<!-- fit, then the more simple, parsimonious model should be selected. -->
+In LPA, we can determine which parameters - variance
+and covariance - are estimated to be the same or different across all profiles,
+and, in the case of covariance, whether it is fixed to zero. Of course, for the
+profiles we are most interested in, the mean is allowed to vary across profiles.
+The models with more parameters freely-estimated use more degrees of freedom;
+thus, similar to the number of profiles selected, the balance between adding
+more parameters (and having a more complex model) and interpretability and
+parsimony must be balanced. If a more complex model fits better and is
+interpretable and is justifiable in terms of the data collected, then it may be
+preferred to simpler models. If the two models are similar in terms of their
+fit, then the more simple, parsimonious model should be selected.
 

### 2019-12-27 paper.rmd
update GRoLTS checklist and present in a table in the paper
add GRoLTS items as headers

@@
 
-This kind of distribution (that is *bimodal*) is not exclusive to grades. For teachers, psychologists, researchers, and friends and family members, people
-are highly-complex and not easily able to be characterized by *one*
-characteristic or personality trait--and its distribution. In the social sciences, broadly, and in psychology, particularly, a statistical method that can be used to describe how
-people, in their individual particularities, may have similarities on the basis
-of some set of measures through which they can be grouped in meaningful,
+This kind of distribution (that is *bimodal*) is not exclusive to grades. For
+teachers, psychologists, researchers, and friends and family members, people are
+highly-complex and not easily able to be characterized by *one* characteristic
+or personality trait--and its distribution. In the social sciences, broadly, and
+in psychology, particularly, a statistical method that can be used to describe
+how people, in their individual particularities, may have similarities on the
+basis of some set of measures through which they can be grouped in meaningful,
 distinctive ways. This approach, which has a provenance in developmental (or
@@
 Cairns, 1996; see Linnenbrink-Garcia & Wormington, 2017, for a recent review) is
-an example of a general mixture model (Harring & Hodis, 2016; Pastor, Barron, Miller, & Davis, 2007).
+an example of a general mixture model (Harring & Hodis, 2016; Pastor, Barron,
+Miller, & Davis, 2007).
 
@@
 t <- read_csv("GRoLTS-applicability-to-LPA.csv")
-t
+tp <- t %>% 
+  filter(use == 1) %>% 
+  mutate(`LPA Number` = 1:nrow(.)) %>% 
+  select(`GRoLTS Number` = Number, `LPA Number`, `Checklist Item`)
+
+tp %>% 
+  DT::datatable()
 ```
+
+We walk through each of these in turn.
+
+### Is the missing data mechanism reported?
+### Is a description provided of what variables are related to attrition/missing data?
+
+### Is a description provided of how missing data in the analyses were dealt with?
+
+### Is information about the distribution of the observed variables included?
+
+### Is the software mentioned?
+
+### Are alternative specifications of the between-class differences in variance–covariance matrix structure considered and clearly documented?
+
+### Is information reported about the number of random start values and final iterations included?
+
+### Are the model comparison (and selection) tools described from a statistical perspective?
+
+### Are the total number of fitted models reported, including a one-class solution?
+
+### Are the number of cases per class reported for each model (absolute sample size, or proportion)?
+
+### If classification of cases in a trajectory is the goal, is entropy reported?
+
+### Are all of the estimates plotted?
+
+### Are the raw data plotted along with the estimates?
+
+### Are characteristics of the final class solution numerically described (i.e., means, SD/SE, n, CI, etc.)?
+
+### Are the syntax files available (either in the appendix, supplementary materials, or from the authors)?
 

### 2019-12-27 paper.rmd
add some implications for software developers

@@
 ### Is the missing data mechanism reported?
+
 ### Is a description provided of what variables are related to attrition/missing data?
@@
 
+# Implications
+
+A step away from the tutorial side of this work, we wish to note some implications 
+for the developers of statistical software for psychological scientists.
+
+First, we note that developing a tool that emphasized both rigor and access has
+been, so far, well-received. This suggests that there is an audience (or a
+market) for tools to carry out sophisticated tools that may be otherwise
+out-of-reach. Tools do not need to be hard to use, a point made by developers
+of, for example, tools to carry out highly-sophisticated Bayesian methods using
+the most cutting-edge tools (Makowski et al., 2019, Burkner, 2018). We think
+that tidyLPA speaks to this, too.
+
+Another implication concerns how we developed tidyLPA. We used a modern software
+development workflow, including git/GitHub, a suite of tests, and continuous
+integration tools. We also note that we emphasized communicating that we were
+developing the software; when it was updated; and how we made major updates.
+
 # Conclusion

### 2019-12-27 paper.rmd
update refs

@@
 of, for example, tools to carry out highly-sophisticated Bayesian methods using
-the most cutting-edge tools (Makowski et al., 2019, Burkner, 2018). We think
+the most cutting-edge tools [@burkner2017, @makowski2019]. We think
 that tidyLPA speaks to this, too.

### 2020-01-13 paper.tex

--- 2019-12-27 paper.rmd
@@
----
-title             : "Carrying out Latent Profile Analysis in Accordance With Best Practices: A Tutorial Using the tidyLPA R package"
-shorttitle        : "tidyLPA"
+\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[man]{apa6}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provides euro and other symbols
+\else % if luatex or xelatex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+\fi
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+% use microtype if available
+\IfFileExists{microtype.sty}{%
+\usepackage[]{microtype}
+\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\IfFileExists{parskip.sty}{%
+\usepackage{parskip}
+}{% else
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
+}
+\usepackage{hyperref}
+\hypersetup{
+            pdftitle={Carrying out Latent Profile Analysis in Accordance With Best Practices: A Tutorial Using the tidyLPA R package},
+            pdfauthor={Joshua Rosenberg, Caspar van Lissa, Jennifer Schmidt, Patrick Beymer, Daniel Anderson, \& Matthew Schell},
+            pdfkeywords={Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust},
+            pdfborder={0 0 0},
+            breaklinks=true}
+\urlstyle{same}  % don't use monospace font for urls
+\usepackage{graphicx,grffile}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\setlength{\emergencystretch}{3em}  % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{0}
+% Redefines (sub)paragraphs to behave more like sections
+\ifx\paragraph\undefined\else
+\let\oldparagraph\paragraph
+\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+\let\oldsubparagraph\subparagraph
+\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-author: 
-  - name          : "Joshua Rosenberg"
-    affiliation   : "1"
-    corresponding : yes
-    address       : "1122 Volunteer Blvd., Knoxville, TN, 37996"
-    email         : "jmrosenberg@utk.edu"
-  - name          : "Caspar van Lissa"
-    affiliation   : "2"
-  - name          : "Jennifer Schmidt"
-    affiliation   : "3"
-  - name          : "Patrick Beymer"
-    affiliation   : "3"
-  - name          : "Daniel Anderson"
-    affiliation   : "4"
-  - name          : "Matthew Schell"
-    affiliation   : "3"
+% set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 
-affiliation:
-  - id            : "1"
-    institution   : "University of Tennessee, Knoxville"
-  - id            : "2"
-    institution   : "Utrecht University"
-  - id            : "3"
-    institution   : "Michigan State University"
-  - id            : "4"
-    institution   : "University of Oregon"
-  - id            : "5"
+\usepackage{etoolbox}
+\makeatletter
+\providecommand{\subtitle}[1]{% add subtitle to \maketitle
+  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
+}
+\makeatother
 
-authornote: |
-  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
+\title{Carrying out Latent Profile Analysis in Accordance With Best Practices: A Tutorial Using the tidyLPA R package}
+\author{Joshua Rosenberg\textsuperscript{1}, Caspar van Lissa\textsuperscript{2}, Jennifer Schmidt\textsuperscript{3}, Patrick Beymer\textsuperscript{3}, Daniel Anderson\textsuperscript{4}, \& Matthew Schell\textsuperscript{3}}
+\date{}
+
+\authornote{Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
 
@@
 
-abstract: |
+Correspondence concerning this article should be addressed to Joshua Rosenberg, 1122 Volunteer Blvd., Knoxville, TN, 37996. E-mail: \href{mailto:jmrosenberg@utk.edu}{\nolinkurl{jmrosenberg@utk.edu}}}
 
-keywords          : "Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust"
-wordcount         : ""
+\abstract{
 
-bibliography      : ["r-references.bib", "citations.bib"]
+}
 
-floatsintext      : no
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : no
-mask              : no
-draft             : no
+\begin{document}
+\maketitle
 
-documentclass     : "apa6"
-classoption       : "man"
-output            : papaja::apa6_pdf
----
-
-```{r setup, include = FALSE}
-library("papaja")
-```
-
-```{r analysis-preferences}
-# Seed for random number generation
-set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
-```
-
-# Introduction
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 
-> Grades are normally distributed, with $\mu$ = 75, $\sigma$ = 5.
+\begin{quote}
+Grades are normally distributed, with \(\mu\) = 75, \(\sigma\) = 5.
+\end{quote}
 
-Students, may (reasonably) ask, though, are grades *really* normally distributed?
-
-```{r, include = FALSE}
-library(simstudy)
-
-def <- defData(varname = "x1", formula = 65, variance = 30)
-def <- defData(def, varname = "x2", formula = 85, variance = 23)
-def <- defData(def, varname = "xMix", 
-               formula = "x1 | .5 + x2 | .5", 
-               dist = "mixture")
-
-set.seed(2716)
-dx <- genData(1000, def)
-```
+Students, may (reasonably) ask, though, are grades \emph{really} normally distributed?
 
@@
 
-```{r, echo - FALSE}
-library(tidyverse)
-ggplot(dx, aes(x = xMix)) +
-  geom_density() +
-  papaja::theme_apa() +
-  ylim(0, .10)
-```
+\includegraphics{paper_files/figure-latex/echo - FALSE-1.pdf}
 
@@
 
-```{r, echo = FALSE}
-dx %>% 
-  gather(key, val, x1:x2) %>% 
-  ggplot(aes(x = val, fill = key)) +
-  geom_density(alpha = .3) +
-  papaja::theme_apa() +
-  ylim(0, .10)
-```
+\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
 
-This kind of distribution (that is *bimodal*) is not exclusive to grades. For
+This kind of distribution (that is \emph{bimodal}) is not exclusive to grades. For
 teachers, psychologists, researchers, and friends and family members, people are
-highly-complex and not easily able to be characterized by *one* characteristic
+highly-complex and not easily able to be characterized by \emph{one} characteristic
 or personality trait--and its distribution. In the social sciences, broadly, and
@@
 distinctive ways. This approach, which has a provenance in developmental (or
-*person-oriented*) approaches (Bergman & El-Khouri, 1997; Magnusson &
-Cairns, 1996; see Linnenbrink-Garcia & Wormington, 2017, for a recent review) is
-an example of a general mixture model (Harring & Hodis, 2016; Pastor, Barron,
-Miller, & Davis, 2007).
+\emph{person-oriented}) approaches (Bergman \& El-Khouri, 1997; Magnusson \&
+Cairns, 1996; see Linnenbrink-Garcia \& Wormington, 2017, for a recent review) is
+an example of a general mixture model (Harring \& Hodis, 2016; Pastor, Barron,
+Miller, \& Davis, 2007).
 
@@
 all of the variables for which (relatively) homogeneous groups are identified
-from among a (relatively) heterogeneous sample are continuos, *latent profile
-analysis* (LPA). We note that scholars using the approach, especially in
+from among a (relatively) heterogeneous sample are continuos, \emph{latent profile
+analysis} (LPA). We note that scholars using the approach, especially in
 developmental and educational psychology, use cluster analytic methods. Such an
@@
 similar to the simplest model specification of LPA?), though LPA also has a
-number of benefits (Linnenbrink-Garcia & Wormington, 2017), and we we focus on
+number of benefits (Linnenbrink-Garcia \& Wormington, 2017), and we we focus on
 LPA in this paper.
@@
 
-# Latent Profile Analysis
+\hypertarget{latent-profile-analysis}{%
+\section{Latent Profile Analysis}\label{latent-profile-analysis}}
 
@@
 model-based, and some descriptions in the literature refer to it as model-based
-clustering (Hennig, Meila, Murtagh, & Rocci, 2015; Scrucca, Fop, Murphy, &
+clustering (Hennig, Meila, Murtagh, \& Rocci, 2015; Scrucca, Fop, Murphy, \&
 Raftery, 2017). Thus, one distinction between LPA and other, similar cluster
@@
 covariances are fixed to zero (also described as the specification of the
-covariance matrix as “diagonal,” because only the diagonal components, or the
+covariance matrix as ``diagonal,'' because only the diagonal components, or the
 variances, are estimated).
@@
 
-<!-- I think we move this to the discussion in light of best practices -->
-
 A note of caution is warranted about LPA in the context of their characteristics
@@
 
-# Specific Software Tools
+\hypertarget{specific-software-tools}{%
+\section{Specific Software Tools}\label{specific-software-tools}}
 
-## Other Tools
+\hypertarget{other-tools}{%
+\subsection{Other Tools}\label{other-tools}}
 
 SPSS is a common tool to carry out cluster analyses (but not LPA). While
-somewhat straightforward to carry out, particularly in SPSS’s graphical user
+somewhat straightforward to carry out, particularly in SPSS's graphical user
 interface (GUI), there are some challenges to use of this approach. The GUI in
@@
 
-Another common tool is MPlus (Muthen & Muthen, 2019). MPlus is a commercial tool
+Another common tool is MPlus (Muthen \& Muthen, 2019). MPlus is a commercial tool
 that provides functionality for many latent variable (and multi-level) models.
@@
 
-In R, a number of tools *can* be used to carry out LPA. OpenMx can be used for
+In R, a number of tools \emph{can} be used to carry out LPA. OpenMx can be used for
 this purpose (and to specify almost any model possible to specify within a
@@
 model specifications, which are often described in terms of the geometric
-properties of the multivariate distributions being estimated (i.e., “spherical,
-equal volume”), rather than in terms of whether and how the means, variances,
+properties of the multivariate distributions being estimated (i.e., ``spherical,
+equal volume''), rather than in terms of whether and how the means, variances,
 and covariances are estimated. They also may use different default settings
@@
 
-## tidyLPA
+\hypertarget{tidylpa}{%
+\subsection{tidyLPA}\label{tidylpa}}
 
@@
 carry-out fully-reproducible analyses and d) were well-documented. To do so, we
-created tidyLPA [@R-tidyLPA].
+created tidyLPA (Rosenberg, Beymer, Anderson, Lissa, \& Schmidt, 2018).
 
@@
 
-One way that tidyLPA is designed to be easy to use is that it assumes a “tidy”
+One way that tidyLPA is designed to be easy to use is that it assumes a ``tidy''
 data structure (Wickham, 2014). This means that it emphasizes the use of a data
@@
 package (Scrucca et al., 2017) and one that is available through the commercial
-software MPlus (Muthen & Muthen, 1997-2017). Moreover, as both tools use the
+software MPlus (Muthen \& Muthen, 1997-2017). Moreover, as both tools use the
 same maximum likelihood estimation procedure, they are benchmarked to produce
@@
 
-## Installation
+\hypertarget{installation}{%
+\subsection{Installation}\label{installation}}
 
@@
 
-``` {r, eval = F}
-install.packages("tidyLPA")
-```
-
 You can also install the development version of tidyLPA from GitHub with:
 
-``` {r gh-installation, eval = FALSE}
-# install.packages("devtools")
-devtools::install_github("data-edu/tidyLPA")
-```
+\hypertarget{examples}{%
+\subsection{Examples}\label{examples}}
 
-``` {r, message = F}
-library(tidyLPA)
-library(dplyr)
-```
+\hypertarget{mclust-via-tidylpa}{%
+\subsubsection{Mclust via tidyLPA}\label{mclust-via-tidylpa}}
 
-## Examples
-
-<!-- This is where we need to use an example from our own data -->
-
-### Mclust via tidyLPA
-
-Here is a brief example using the built-in `pisaUSA15` data set and variables
+Here is a brief example using the built-in \texttt{pisaUSA15} data set and variables
 for broad interest, enjoyment, and self-efficacy. Note that we first type the
@@
 create the profiles. We also specify the number of profiles and the model. See
-`?estimate_profiles` for more details.
+\texttt{?estimate\_profiles} for more details.
 
-In these examples, we pass the results of one function to the next by *piping*
-(using the `%>%` operator, loaded from the `dplyr` package). We pass the data to
-a function that selects relevant variables, and then to `estimate_profiles`:
+In these examples, we pass the results of one function to the next by \emph{piping}
+(using the \texttt{\%\textgreater{}\%} operator, loaded from the \texttt{dplyr} package). We pass the data to
+a function that selects relevant variables, and then to \texttt{estimate\_profiles}:
 
-``` {r}
-pisaUSA15[1:100, ] %>%
-    select(broad_interest, enjoyment, self_efficacy) %>%
-    single_imputation() %>%
-    estimate_profiles(3)
-```
+\begin{verbatim}
+## tidyLPA analysis using mclust: 
+## 
+##  Model Classes AIC    BIC    Entropy prob_min prob_max n_min n_max BLRT_p
+##  1     3       629.10 665.57 0.82    0.83     0.96     0.03  0.70  0.01
+\end{verbatim}
 
-### Mplus via tidyLPA
+\hypertarget{mplus-via-tidylpa}{%
+\subsubsection{Mplus via tidyLPA}\label{mplus-via-tidylpa}}
 
 We can use Mplus simply by changing the package argument for
-`estimate_profiles()`:
-
-``` {r, eval = FALSE}
-pisaUSA15[1:100, ] %>%
-    select(broad_interest, enjoyment, self_efficacy) %>%
-    single_imputation() %>%
-    estimate_profiles(3, package = "MplusAutomation")
-```
+\texttt{estimate\_profiles()}:
 
@@
 The resulting object can be further passed down a pipeline to
-other functions, such as `compare_solutions`, `get_data`, `get_fit`,
-etc. This is the "tidy" part, in that the function can be embedded in a tidy
+other functions, such as \texttt{compare\_solutions}, \texttt{get\_data}, \texttt{get\_fit},
+etc. This is the ``tidy'' part, in that the function can be embedded in a tidy
 analysis pipeline.
@@
 If you have Mplus installed, you can call the version of this function that uses
-MPlus in the same way, by adding the argument `package = "MplusAutomation`.
+MPlus in the same way, by adding the argument \texttt{package\ =\ "MplusAutomation}.
 
-### Plotting the profiles
+\hypertarget{plotting-the-profiles}{%
+\subsubsection{Plotting the profiles}\label{plotting-the-profiles}}
 
-We can plot the profiles by piping the output to `plot_profiles()`.
+We can plot the profiles by piping the output to \texttt{plot\_profiles()}.
 
-``` {r}
-pisaUSA15[1:100, ] %>%
-    select(broad_interest, enjoyment, self_efficacy) %>%
-    single_imputation() %>%
-    scale() %>%
-    estimate_profiles(3) %>% 
-    plot_profiles()
-```
+\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1.pdf}
 
-## Model specification
+\hypertarget{model-specification}{%
+\subsection{Model specification}\label{model-specification}}
 
-In addition to the number of profiles (specified with the `n_profiles`
+In addition to the number of profiles (specified with the \texttt{n\_profiles}
 argument), the model can be specified in terms of whether and how the variable
@@
 
-The models are specified by passing arguments to the `variance` and `covariance`
+The models are specified by passing arguments to the \texttt{variance} and \texttt{covariance}
 arguments. The possible values for these arguments are:
 
-  - `variances`: "equal" and "zero"
-  - `covariances`: "varying", "equal", and "zero"
+\begin{itemize}
+\tightlist
+\item
+  \texttt{variances}: ``equal'' and ``zero''
+\item
+  \texttt{covariances}: ``varying'', ``equal'', and ``zero''
+\end{itemize}
 
@@
 
-  - Equal variances and covariances fixed to 0 (Model 1)
-  - Varying variances and covariances fixed to 0 (Model 2)
-  - Equal variances and equal covariances (Model 3)
-  - Varying variances and varying covariances (Model 6)
+\begin{itemize}
+\tightlist
+\item
+  Equal variances and covariances fixed to 0 (Model 1)
+\item
+  Varying variances and covariances fixed to 0 (Model 2)
+\item
+  Equal variances and equal covariances (Model 3)
+\item
+  Varying variances and varying covariances (Model 6)
+\end{itemize}
 
@@
 on the models can be found in the
-[vignette](https://data-edu.github.io/tidyLPA/articles/Introduction_to_tidyLPA.html).
+\href{https://data-edu.github.io/tidyLPA/articles/Introduction_to_tidyLPA.html}{vignette}.
 
@@
 
-``` {r, eval = F}
-pisaUSA15[1:100, ] %>%
-    select(broad_interest, enjoyment, self_efficacy) %>%
-    single_imputation() %>%
-    estimate_profiles(3, 
-                      variances = "varying",
-                      covariances = "varying")
-```
+\hypertarget{comparing-a-wide-range-of-solutions}{%
+\subsection{Comparing a wide range of solutions}\label{comparing-a-wide-range-of-solutions}}
 
-## Comparing a wide range of solutions
-
-The function `compare_solutions()` compares the fit of several estimated models,
+The function \texttt{compare\_solutions()} compares the fit of several estimated models,
 with varying numbers of profiles and model specifications:
 
-``` {r, eval = FALSE}
-pisaUSA15[1:100, ] %>%
-    select(broad_interest, enjoyment, self_efficacy) %>%
-    single_imputation() %>%
-    estimate_profiles(1:3, 
-                      variances = c("equal", "varying"),
-                      covariances = c("zero", "varying")) %>%
-    compare_solutions(statistics = c("AIC", "BIC"))
-```
-
-## Accessing values in the output
+\hypertarget{accessing-values-in-the-output}{%
+\subsection{Accessing values in the output}\label{accessing-values-in-the-output}}
 
@@
 
-`get_data()` returns the data:
+\texttt{get\_data()} returns the data:
 
-``` {r}
-m <- pisaUSA15[1:100, ] %>%
-    select(broad_interest, enjoyment, self_efficacy) %>%
-    single_imputation() %>%
-    estimate_profiles(3:4)
-```
+\begin{verbatim}
+## # A tibble: 700 x 9
+##    model_number classes_number broad_interest enjoyment self_efficacy Class
+##           <dbl>          <int>          <dbl>     <dbl>         <dbl> <dbl>
+##  1            1              3            3.8       4            1        1
+##  2            1              3            3         3            2.75     3
+##  3            1              3            1.8       2.8          3.38     3
+##  4            1              3            1.4       1            2.75     2
+##  5            1              3            1.8       2.2          2        3
+##  6            1              3            1.6       1.6          1.88     3
+##  7            1              3            3         3.8          2.25     1
+##  8            1              3            2.6       2.2          2        3
+##  9            1              3            1         2.8          2.62     3
+## 10            1              3            2.2       2            1.75     3
+## # … with 690 more rows, and 3 more variables: Class_prob <int>,
+## #   Probability <dbl>, id <int>
+\end{verbatim}
 
-```{r}
-get_data(m)
-```
-
-We note that `get_data()` returns data in wide format when applied to an object
+We note that \texttt{get\_data()} returns data in wide format when applied to an object
 of class tidyProfile (one element of a tidyLPA object), or when applied to a
-tidyLPA object of length one. `get_data()` returns long format when applied to a
+tidyLPA object of length one. \texttt{get\_data()} returns long format when applied to a
 tidyLPA object containing multiple tidyProfile analyses (because then the wide
@@
 
-To transform data in the wide format into the long format, the `gather()`
-function from the **tidyr** package can be used, e.g.:
+To transform data in the wide format into the long format, the \texttt{gather()}
+function from the \textbf{tidyr} package can be used, e.g.:
 
-``` {r}
-get_data(m) %>% 
-    tidyr::gather(Class_prob, Probability, contains("CPROB"))
-```
+\begin{verbatim}
+## # A tibble: 700 x 9
+##    model_number classes_number broad_interest enjoyment self_efficacy Class
+##           <dbl>          <int>          <dbl>     <dbl>         <dbl> <dbl>
+##  1            1              3            3.8       4            1        1
+##  2            1              3            3         3            2.75     3
+##  3            1              3            1.8       2.8          3.38     3
+##  4            1              3            1.4       1            2.75     2
+##  5            1              3            1.8       2.2          2        3
+##  6            1              3            1.6       1.6          1.88     3
+##  7            1              3            3         3.8          2.25     1
+##  8            1              3            2.6       2.2          2        3
+##  9            1              3            1         2.8          2.62     3
+## 10            1              3            2.2       2            1.75     3
+## # … with 690 more rows, and 3 more variables: Class_prob <int>,
+## #   Probability <dbl>, id <int>
+\end{verbatim}
 
-# Recommendations Regarding Analytic Choices
+\hypertarget{recommendations-regarding-analytic-choices}{%
+\section{Recommendations Regarding Analytic Choices}\label{recommendations-regarding-analytic-choices}}
 
-## GRoLTs checklist
+\hypertarget{grolts-checklist}{%
+\subsection{GRoLTs checklist}\label{grolts-checklist}}
 
-```{r, echo = FALSE}
-t <- read_csv("GRoLTS-applicability-to-LPA.csv")
-tp <- t %>% 
-  filter(use == 1) %>% 
-  mutate(`LPA Number` = 1:nrow(.)) %>% 
-  select(`GRoLTS Number` = Number, `LPA Number`, `Checklist Item`)
-
-tp %>% 
-  DT::datatable()
-```
+\includegraphics{paper_files/figure-latex/unnamed-chunk-13-1.pdf}
 
@@
 
-### Is the missing data mechanism reported?
-### Is a description provided of what variables are related to attrition/missing data?
+\hypertarget{is-the-missing-data-mechanism-reported}{%
+\subsubsection{Is the missing data mechanism reported?}\label{is-the-missing-data-mechanism-reported}}
 
-### Is a description provided of how missing data in the analyses were dealt with?
+\hypertarget{is-a-description-provided-of-what-variables-are-related-to-attritionmissing-data}{%
+\subsubsection{Is a description provided of what variables are related to attrition/missing data?}\label{is-a-description-provided-of-what-variables-are-related-to-attritionmissing-data}}
 
-### Is information about the distribution of the observed variables included?
+\hypertarget{is-a-description-provided-of-how-missing-data-in-the-analyses-were-dealt-with}{%
+\subsubsection{Is a description provided of how missing data in the analyses were dealt with?}\label{is-a-description-provided-of-how-missing-data-in-the-analyses-were-dealt-with}}
 
-### Is the software mentioned?
+\hypertarget{is-information-about-the-distribution-of-the-observed-variables-included}{%
+\subsubsection{Is information about the distribution of the observed variables included?}\label{is-information-about-the-distribution-of-the-observed-variables-included}}
 
-### Are alternative specifications of the between-class differences in variance–covariance matrix structure considered and clearly documented?
+\hypertarget{is-the-software-mentioned}{%
+\subsubsection{Is the software mentioned?}\label{is-the-software-mentioned}}
 
-### Is information reported about the number of random start values and final iterations included?
+\hypertarget{are-alternative-specifications-of-the-between-class-differences-in-variancecovariance-matrix-structure-considered-and-clearly-documented}{%
+\subsubsection{Are alternative specifications of the between-class differences in variance--covariance matrix structure considered and clearly documented?}\label{are-alternative-specifications-of-the-between-class-differences-in-variancecovariance-matrix-structure-considered-and-clearly-documented}}
 
-### Are the model comparison (and selection) tools described from a statistical perspective?
+\hypertarget{is-information-reported-about-the-number-of-random-start-values-and-final-iterations-included}{%
+\subsubsection{Is information reported about the number of random start values and final iterations included?}\label{is-information-reported-about-the-number-of-random-start-values-and-final-iterations-included}}
 
-### Are the total number of fitted models reported, including a one-class solution?
+\hypertarget{are-the-model-comparison-and-selection-tools-described-from-a-statistical-perspective}{%
+\subsubsection{Are the model comparison (and selection) tools described from a statistical perspective?}\label{are-the-model-comparison-and-selection-tools-described-from-a-statistical-perspective}}
 
-### Are the number of cases per class reported for each model (absolute sample size, or proportion)?
+\hypertarget{are-the-total-number-of-fitted-models-reported-including-a-one-class-solution}{%
+\subsubsection{Are the total number of fitted models reported, including a one-class solution?}\label{are-the-total-number-of-fitted-models-reported-including-a-one-class-solution}}
 
-### If classification of cases in a trajectory is the goal, is entropy reported?
+\hypertarget{are-the-number-of-cases-per-class-reported-for-each-model-absolute-sample-size-or-proportion}{%
+\subsubsection{Are the number of cases per class reported for each model (absolute sample size, or proportion)?}\label{are-the-number-of-cases-per-class-reported-for-each-model-absolute-sample-size-or-proportion}}
 
-### Are all of the estimates plotted?
+\hypertarget{if-classification-of-cases-in-a-trajectory-is-the-goal-is-entropy-reported}{%
+\subsubsection{If classification of cases in a trajectory is the goal, is entropy reported?}\label{if-classification-of-cases-in-a-trajectory-is-the-goal-is-entropy-reported}}
 
-### Are the raw data plotted along with the estimates?
+\hypertarget{are-all-of-the-estimates-plotted}{%
+\subsubsection{Are all of the estimates plotted?}\label{are-all-of-the-estimates-plotted}}
 
-### Are characteristics of the final class solution numerically described (i.e., means, SD/SE, n, CI, etc.)?
+\hypertarget{are-the-raw-data-plotted-along-with-the-estimates}{%
+\subsubsection{Are the raw data plotted along with the estimates?}\label{are-the-raw-data-plotted-along-with-the-estimates}}
 
-### Are the syntax files available (either in the appendix, supplementary materials, or from the authors)?
+\hypertarget{are-characteristics-of-the-final-class-solution-numerically-described-i.e.-means-sdse-n-ci-etc.}{%
+\subsubsection{Are characteristics of the final class solution numerically described (i.e., means, SD/SE, n, CI, etc.)?}\label{are-characteristics-of-the-final-class-solution-numerically-described-i.e.-means-sdse-n-ci-etc.}}
+
+\hypertarget{are-the-syntax-files-available-either-in-the-appendix-supplementary-materials-or-from-the-authors}{%
+\subsubsection{Are the syntax files available (either in the appendix, supplementary materials, or from the authors)?}\label{are-the-syntax-files-available-either-in-the-appendix-supplementary-materials-or-from-the-authors}}
 
@@
 used for the cluster analysis (i.e., the distance metric used to calculate the
-similarity of the observations as part of the Ward’s hierarchical clustering) or
+similarity of the observations as part of the Ward's hierarchical clustering) or
 what parameters are estimated and how as part of LPA.
@@
 
-## How the Data are Centering or Scaled
+\hypertarget{how-the-data-are-centering-or-scaled}{%
+\subsection{How the Data are Centering or Scaled}\label{how-the-data-are-centering-or-scaled}}
 
@@
 
-## How to Choose the Number of Profiles/Classes
+\hypertarget{how-to-choose-the-number-of-profilesclasses}{%
+\subsection{How to Choose the Number of Profiles/Classes}\label{how-to-choose-the-number-of-profilesclasses}}
 
@@
 
-## How to Choose Options for the Clustering Procedure or to Select the Model Parameterization
+\hypertarget{how-to-choose-options-for-the-clustering-procedure-or-to-select-the-model-parameterization}{%
+\subsection{How to Choose Options for the Clustering Procedure or to Select the Model Parameterization}\label{how-to-choose-options-for-the-clustering-procedure-or-to-select-the-model-parameterization}}
 
@@
 
-## More General Recommendations
+\hypertarget{more-general-recommendations}{%
+\subsection{More General Recommendations}\label{more-general-recommendations}}
 
@@
 data are collected across multiple time points, into analyses. ISOA groups all
-time points and doesn’t make distinctions. Other approaches perform analysis
-separately at different timepoints (Corpus & Wormington, 2014). Some integrate
-time as a part of the profiles, i.e. growth mixture modeling (groups of
-patterns), i.e. within-person growth modeling, where there are individual growth
+time points and doesn't make distinctions. Other approaches perform analysis
+separately at different timepoints (Corpus \& Wormington, 2014). Some integrate
+time as a part of the profiles, i.e.~growth mixture modeling (groups of
+patterns), i.e.~within-person growth modeling, where there are individual growth
 patterns. Research to date has yet to consider additional challenges in applying
@@
 nested as in very common uses of multi-level models, but rather are
-cross-classified. West, Welch, & Galecki (2014) have a description of the use of
+cross-classified. West, Welch, \& Galecki (2014) have a description of the use of
 multi-level models with cross-classified data and tools (including those freely
@@
 
-# Conclusion
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
 
@@
 person-oriented approaches are complementary to variable-centered analyses
-(Marsh, Ludtke, Trautwein, & Morin, 2009). A person-oriented approach can help
+(Marsh, Ludtke, Trautwein, \& Morin, 2009). A person-oriented approach can help
 us to consider multiple variables together and at once and in a more dynamic
@@
 
-# References
-```{r create_r-references}
-r_refs(file = "r-references.bib")
-```
+\hypertarget{references}{%
+\section{References}\label{references}}
 
@@
 
-<div id = "refs"></div>
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-R-tidyLPA}{}%
+Rosenberg, J.
+M., Beymer, P.
+N., Anderson, D.
+J., Lissa, C.
+van, \& Schmidt, J.
+A.
+(2018).
+TidyLPA: An r package to easily carry out latent profile analysis (lpa) using open-source or commercial software.
+\emph{Journal of Open Source Software}, \emph{3}(30), 978.
+doi:\href{https://doi.org/10.21105/joss.00978}{10.21105/joss.00978}
+
 \endgroup
+
+\end{document}

### 2020-01-16 paper.rmd
update

@@
 
-authornote: |
-  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-
-  Enter author note here.
-
-abstract: |
-
+abstract: 
 keywords          : "Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust"
@@
 output            : papaja::apa6_pdf
+always_allow_html: true
 ---
@@
 ```{r setup, include = FALSE}
-library("papaja")
+library(papaja)
 ```
@@
 set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
+library(tidyverse)
+knitr::opts_chunk$set(cache.extra = knitr::rand_seed, cache = TRUE)
 ```
@@
 
-```{r, include = FALSE}
+```{r, include = FALSE, eval = FALSE}
 library(simstudy)
@@
 
-```{r, echo - FALSE}
-library(tidyverse)
+```{r, echo = FALSE, eval = FALSE}
+
 ggplot(dx, aes(x = xMix)) +
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, eval = FALSE}
 dx %>% 
@@
 
-<!-- This is where we need to use an example from our own data -->
+Here, we use a simulated dataset, based on Curry et al., 2019, contains data on moral relevance and judgment across the seven domains of the *Morality As Cooperation* scale.
 
 ### Mclust via tidyLPA
-
-Here is a brief example using the built-in `pisaUSA15` data set and variables
-for broad interest, enjoyment, and self-efficacy. Note that we first type the
-name of the data frame, followed by the unquoted names of the variables used to
-create the profiles. We also specify the number of profiles and the model. See
-`?estimate_profiles` for more details.
 
@@
 ``` {r}
-pisaUSA15[1:100, ] %>%
-    select(broad_interest, enjoyment, self_efficacy) %>%
-    single_imputation() %>%
+curry_mac %>% 
+  select(KinshipR:PossessionR, 
+         KinshipJ:PossessionJ) %>% 
   estimate_profiles(3)
@@
 ``` {r, eval = FALSE}
-pisaUSA15[1:100, ] %>%
-    select(broad_interest, enjoyment, self_efficacy) %>%
-    single_imputation() %>%
+curry_mac %>% 
+  select(KinshipR:PossessionR, 
+         KinshipJ:PossessionJ) %>% 
   estimate_profiles(3, package = "MplusAutomation")
@@
 
-``` {r}
-pisaUSA15[1:100, ] %>%
-    select(broad_interest, enjoyment, self_efficacy) %>%
-    single_imputation() %>%
-    scale() %>%
+```{r}
+curry_mac %>% 
+  select(KinshipR:PossessionR, 
+         KinshipJ:PossessionJ) %>% 
   estimate_profiles(3) %>% 
@@
 ``` {r, eval = F}
-pisaUSA15[1:100, ] %>%
-    select(broad_interest, enjoyment, self_efficacy) %>%
-    single_imputation() %>%
+curry_mac %>% 
+  select(KinshipR:PossessionR, 
+         KinshipJ:PossessionJ) %>% 
   estimate_profiles(3, 
@@
 
-``` {r, eval = FALSE}
-pisaUSA15[1:100, ] %>%
-    select(broad_interest, enjoyment, self_efficacy) %>%
-    single_imputation() %>%
-    estimate_profiles(1:3, 
-                      variances = c("equal", "varying"),
-                      covariances = c("zero", "varying")) %>%
+``` {r, compare-mclust, eval = FALSE}
+curry_mac %>% 
+  select(KinshipR:PossessionR, 
+         KinshipJ:PossessionJ) %>% 
+  estimate_profiles(models = c(1, 2, 3, 6),
+                    n_profiles = 1:9) %>%
+  compare_solutions(statistics = c("AIC", "BIC"))
+```
+
+We can also run the same process using mplus:
+
+``` {r, compare-mplus, eval = FALSE}
+curry_mac %>% 
+  select(KinshipR:PossessionR, 
+         KinshipJ:PossessionJ) %>% 
+  estimate_profiles(models = c(1, 2, 3, 6),
+                    n_profiles = 1:9, package = "mplus") %>%
   compare_solutions(statistics = c("AIC", "BIC"))

### 2020-01-16 paper.tex
update

@@
 
-\authornote{Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-
-Enter author note here.
+\authornote{
 
@@
 
-\includegraphics{paper_files/figure-latex/echo - FALSE-1.pdf}
-
 Or two?
-
-\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
 
@@
 
+Here, we use a simulated dataset, based on Curry et al., 2019, contains data on moral relevance and judgment across the seven domains of the \emph{Morality As Cooperation} scale.
+
 \hypertarget{mclust-via-tidylpa}{%
 \subsubsection{Mclust via tidyLPA}\label{mclust-via-tidylpa}}
-
-Here is a brief example using the built-in \texttt{pisaUSA15} data set and variables
-for broad interest, enjoyment, and self-efficacy. Note that we first type the
-name of the data frame, followed by the unquoted names of the variables used to
-create the profiles. We also specify the number of profiles and the model. See
-\texttt{?estimate\_profiles} for more details.
 
@@
 ##  Model Classes AIC       BIC       Entropy prob_min prob_max n_min n_max BLRT_p
-##  1     3       629.10 665.57 0.82    0.83     0.96     0.03  0.70  0.01
+##  1     3       158594.55 158898.38 0.82    0.91     0.92     0.15  0.53  0.01
 \end{verbatim}
@@
 
-\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1.pdf}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}
 
@@
 with varying numbers of profiles and model specifications:
+
+We can also run the same process using mplus:
 
@@
 
-\includegraphics{paper_files/figure-latex/unnamed-chunk-13-1.pdf}
+\hypertarget{htmlwidget-a6cf205f9fe062a59d12}{}
 
@@
 
+\hypertarget{implications}{%
+\section{Implications}\label{implications}}
+
+A step away from the tutorial side of this work, we wish to note some implications
+for the developers of statistical software for psychological scientists.
+
+First, we note that developing a tool that emphasized both rigor and access has
+been, so far, well-received. This suggests that there is an audience (or a
+market) for tools to carry out sophisticated tools that may be otherwise
+out-of-reach. Tools do not need to be hard to use, a point made by developers
+of, for example, tools to carry out highly-sophisticated Bayesian methods using
+the most cutting-edge tools (Bürkner \& others, 2017, @makowski2019). We think
+that tidyLPA speaks to this, too.
+
+Another implication concerns how we developed tidyLPA. We used a modern software
+development workflow, including git/GitHub, a suite of tests, and continuous
+integration tools. We also note that we emphasized communicating that we were
+developing the software; when it was updated; and how we made major updates.
+
 \hypertarget{conclusion}{%
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-burkner2017}{}%
+Bürkner, P.-C., \& others. (2017). Brms: An r package for bayesian multilevel models using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.
+
+\leavevmode\hypertarget{ref-makowski2019}{}%
+Makowski, D., Ben-Shachar, M. S., Chen, S., \& Lüdecke, D. (2019). Indices of effect existence and significance in the bayesian framework. \emph{Frontiers in Psychology}, \emph{10}, 2767.
+
 \leavevmode\hypertarget{ref-R-tidyLPA}{}%

### 2020-01-28 paper.rmd
update

@@
 
-``` {r}
+``` {r, curry-mac-mclust}
 curry_mac %>% 
@@
 
-``` {r, eval = FALSE}
+``` {r, eval = FALSE, curry-mac-mplus}
 curry_mac %>% 
@@
 
-```{r}
-curry_mac %>% 
+```{r, curry_mac_p3}
+curry_mac_three <- curry_mac %>% 
   select(KinshipR:PossessionR, 
@@
   plot_profiles()
+
+curry_mac_three +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 ```
@@
 ``` {r, compare-mclust, eval = FALSE}
-curry_mac %>% 
+curry_mac_solutions <- curry_mac %>% 
   select(KinshipR:PossessionR, 
@@
 
+The results of this analysis suggests that the Model 6, two or seven class models are best.
+
+```{r, m6p7-solution}
+curry_mac_m6p7 <- curry_mac %>% 
+  select(KinshipR:PossessionR, 
+         KinshipJ:PossessionJ) %>% 
+  estimate_profiles(3, 
+                    variances = "varying",
+                    covariances = "varying")
+
+plot_profiles(curry_mac_m6p7)
+```
+
+```{r, m6p2-solution}
+curry_mac_m6p2 <- curry_mac %>% 
+  select(KinshipR:PossessionR, 
+         KinshipJ:PossessionJ) %>% 
+  estimate_profiles(3, 
+                    variances = "varying",
+                    covariances = "varying")
+
+plot_profiles(curry_mac_m6p7)
+```
+
 We can also run the same process using mplus:
 
-``` {r, compare-mplus, eval = FALSE}
+``` {r, compare-mplus, eval = TRUE}
 curry_mac %>% 
@@
 
-``` {r}
+``` {r, pisa-get-data}
 m <- pisaUSA15[1:100, ] %>%

### 2020-01-30 paper.rmd
update

@@
   compare_solutions(statistics = c("AIC", "BIC"))
+
+3Best model according to AIC is Model 6 with 7 classes.
+# Best model according to BIC is Model 6 with 2 classes.
+
+#An analytic hierarchy process, based on the fit indices AIC, AWE, BIC, CLC, and KIC (Akogul & Erisoglu, 2017), suggests the best solution is Model 6 with 2 classes.
 ```
@@
   compare_solutions(statistics = c("AIC", "BIC"))
+
+# Best model according to AIC is Model 6 with 4 classes.
+# Best model according to BIC is Model 6 with 2 classes.
+
+# An analytic hierarchy process, based on the fit indices AIC, AWE, BIC, CLC, and KIC (Akogul & Erisoglu, 2017), suggests the best solution is Model 6 with 2 classes.
+```
+
+```{r}
+
 ```

### 2020-02-04 paper.rmd
update

@@
 
-``` {r, eval = FALSE, curry-mac-mplus}
+``` {r, curry-mac-mplus, eval = FALSE}
 curry_mac %>% 
@@
 
-``` {r, eval = F}
+``` {r, curry-mac-m6, eval = F}
 curry_mac %>% 
@@
 ``` {r, compare-mclust, eval = FALSE}
-curry_mac_solutions <- curry_mac %>% 
+curry_mac_mclust_solutions <- curry_mac %>% 
   select(KinshipR:PossessionR, 
@@
 
-3Best model according to AIC is Model 6 with 7 classes.
+write_rds(curry_mac_mclust_solutions, "cm_mclust.rds")
+
+# Best model according to AIC is Model 6 with 7 classes.
 # Best model according to BIC is Model 6 with 2 classes.
@@
 #An analytic hierarchy process, based on the fit indices AIC, AWE, BIC, CLC, and KIC (Akogul & Erisoglu, 2017), suggests the best solution is Model 6 with 2 classes.
+```
+
+````{r, just-one}
+curry_mac_mclust_j <- curry_mac %>% 
+  select(KinshipR:PossessionR) %>% 
+  estimate_profiles(models = c(1, 2, 3, 6),
+                    n_profiles = 1:9) %>%
+  compare_solutions(statistics = c("AIC", "BIC"))
+
+write_rds(curry_mac_mclust_j, "cm_mclust_r.rds")
+
+curry_mac_mclust_j <- curry_mac %>% 
+  select(KinshipJ:PossessionJ) %>% 
+  estimate_profiles(models = c(1, 2, 3, 6),
+                    n_profiles = 1:9) %>%
+  compare_solutions(statistics = c("AIC", "BIC"))
+
+write_rds(curry_mac_mclust_j, "cm_mclust_j.rds")
+
+curry_mac_mclust_j <- curry_mac %>% 
+  select(KinshipR:PossessionR) %>% 
+  estimate_profiles(models = c(1, 2, 3, 6),
+                    n_profiles = 1:9,
+                    package = "mplus") %>%
+  compare_solutions(statistics = c("AIC", "BIC"))
+
+write_rds(curry_mac_mclust_j, "cm_mclust_r.rds")
+
+curry_mac_mclust_j <- curry_mac %>% 
+  select(KinshipJ:PossessionJ) %>% 
+  estimate_profiles(models = c(1, 2, 3, 6),
+                    n_profiles = 1:9,
+                    package = "mplus") %>%
+  compare_solutions(statistics = c("AIC", "BIC"))
+
+write_rds(curry_mac_mclust_j, "cm_mclust_j.rds")
 ```
@@
 ``` {r, compare-mplus, eval = TRUE}
-curry_mac %>% 
+curry_mac_mplus_solutions <- curry_mac %>% 
   select(KinshipR:PossessionR, 
@@
   compare_solutions(statistics = c("AIC", "BIC"))
+
+write_rds(curry_mac_mplus_solutions, "cm_mplus.rds")
 
@@
 
-```{r}
+```{r, m6ps-mplus, error = TRUE}
+curry_mac_m6p2_mplus <- curry_mac %>% 
+  select(KinshipR:PossessionR, 
+         KinshipJ:PossessionJ) %>% 
+  estimate_profiles(n_profiles = 2, 
+                    variances = "varying",
+                    covariances = "varying", 
+                    package = "mplus")
 
+# equivalent to model 6
+
+plot_profiles(curry_mac_m6p2_mplus)
+```
+
+```{r, m6p2-mclust}
+curry_mac_m6p2 <- curry_mac %>% 
+  select(KinshipR:PossessionR, 
+         KinshipJ:PossessionJ) %>% 
+  estimate_profiles(n_profiles = 2, 
+                    variances = "varying",
+                    covariances = "varying")
+
+plot_profiles(curry_mac_m6p2)
 ```

### 2020-02-04 paper.tex
update

@@
 
-\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}
+\includegraphics{paper_files/figure-latex/curry_mac_p3-1.pdf}
+
+\begin{verbatim}
+## Warning: Using shapes for an ordinal variable is not advised
+\end{verbatim}
+
+\includegraphics{paper_files/figure-latex/curry_mac_p3-2.pdf}
 
@@
 
+\begin{verbatim}
+## Warning: 
+## One or more analyses resulted in warnings! Examine these analyses carefully: model_3_class_3, model_3_class_6, model_3_class_8
+
+## Warning: 
+## One or more analyses resulted in warnings! Examine these analyses carefully: model_3_class_3, model_3_class_6, model_3_class_8
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: 
+## One or more analyses resulted in warnings!
+Examine these analyses carefully: model_1_class_5, model_1_class_6, model_1_class_8, model_1_class_9, model_2_class_6, model_2_class_7, model_2_class_8, model_2_class_9, model_3_class_1, model_3_class_2, model_3_class_3, model_3_class_4, model_3_class_5, model_3_class_6, model_3_class_7, model_3_class_8, model_3_class_9, model_6_class_1, model_6_class_2, model_6_class_3, model_6_class_4, model_6_class_5, model_6_class_6, model_6_class_7, model_6_class_8, model_6_class_9
+
+## Warning: 
+## One or more analyses resulted in warnings!
+Examine these analyses carefully: model_1_class_5, model_1_class_6, model_1_class_8, model_1_class_9, model_2_class_6, model_2_class_7, model_2_class_8, model_2_class_9, model_3_class_1, model_3_class_2, model_3_class_3, model_3_class_4, model_3_class_5, model_3_class_6, model_3_class_7, model_3_class_8, model_3_class_9, model_6_class_1, model_6_class_2, model_6_class_3, model_6_class_4, model_6_class_5, model_6_class_6, model_6_class_7, model_6_class_8, model_6_class_9
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: 
+## One or more analyses resulted in warnings!
+Examine these analyses carefully: model_1_class_6, model_1_class_7, model_1_class_8, model_1_class_9, model_2_class_5, model_2_class_6, model_2_class_7, model_2_class_8, model_2_class_9, model_3_class_1, model_3_class_2, model_3_class_3, model_3_class_4, model_3_class_5, model_3_class_6, model_3_class_7, model_3_class_8, model_3_class_9, model_6_class_1, model_6_class_2, model_6_class_3, model_6_class_4, model_6_class_5, model_6_class_6, model_6_class_7, model_6_class_8, model_6_class_9
+
+## Warning: 
+## One or more analyses resulted in warnings!
+Examine these analyses carefully: model_1_class_6, model_1_class_7, model_1_class_8, model_1_class_9, model_2_class_5, model_2_class_6, model_2_class_7, model_2_class_8, model_2_class_9, model_3_class_1, model_3_class_2, model_3_class_3, model_3_class_4, model_3_class_5, model_3_class_6, model_3_class_7, model_3_class_8, model_3_class_9, model_6_class_1, model_6_class_2, model_6_class_3, model_6_class_4, model_6_class_5, model_6_class_6, model_6_class_7, model_6_class_8, model_6_class_9
+\end{verbatim}
+
+The results of this analysis suggests that the Model 6, two or seven class models are best.
+
+\includegraphics{paper_files/figure-latex/m6p7-solution-1.pdf}
+
+\includegraphics{paper_files/figure-latex/m6p2-solution-1.pdf}
+
 We can also run the same process using mplus:
+
+\begin{verbatim}
+## Warning: 
+## One or more analyses resulted in warnings!
+Examine these analyses carefully: model_1_class_5, model_1_class_6, model_1_class_7, model_1_class_8, model_1_class_9, model_2_class_5, model_2_class_8, model_2_class_9, model_3_class_1, model_3_class_2, model_3_class_3, model_3_class_4, model_3_class_5, model_3_class_6, model_3_class_7, model_3_class_8, model_3_class_9, model_6_class_1, model_6_class_2, model_6_class_3, model_6_class_4, model_6_class_5, model_6_class_6, model_6_class_7, model_6_class_8, model_6_class_9
+
+## Warning: 
+## One or more analyses resulted in warnings!
+Examine these analyses carefully: model_1_class_5, model_1_class_6, model_1_class_7, model_1_class_8, model_1_class_9, model_2_class_5, model_2_class_8, model_2_class_9, model_3_class_1, model_3_class_2, model_3_class_3, model_3_class_4, model_3_class_5, model_3_class_6, model_3_class_7, model_3_class_8, model_3_class_9, model_6_class_1, model_6_class_2, model_6_class_3, model_6_class_4, model_6_class_5, model_6_class_6, model_6_class_7, model_6_class_8, model_6_class_9
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: 
+## One or more analyses resulted in warnings! Examine these analyses carefully: model_6_class_2
+\end{verbatim}
+
+\begin{verbatim}
+## Error in plot_profiles.default(x = list(df_plot = structure(list(Class = structure(c(1L, : Could not match raw data to model estimates.
+\end{verbatim}
+
+\includegraphics{paper_files/figure-latex/m6p2-mclust-1.pdf}
 
@@
 ##  2            1              3            3         3            2.75     3
-##  3            1              3            1.8       2.8          3.38     3
+##  3            1              3            1.8       2.8          3.38     2
 ##  4            1              3            1.4       1            2.75     2
-##  5            1              3            1.8       2.2          2        3
-##  6            1              3            1.6       1.6          1.88     3
-##  7            1              3            3         3.8          2.25     1
+##  5            1              3            1.8       2.2          2        2
+##  6            1              3            1.6       1.6          1.88     2
+##  7            1              3            3         3.8          2.25     3
 ##  8            1              3            2.6       2.2          2        3
-##  9            1              3            1         2.8          2.62     3
-## 10            1              3            2.2       2            1.75     3
+##  9            1              3            1         2.8          2.62     2
+## 10            1              3            2.2       2            1.75     2
 ## # … with 690 more rows, and 3 more variables: Class_prob <int>,
@@
 
-\hypertarget{htmlwidget-a6cf205f9fe062a59d12}{}
+\hypertarget{htmlwidget-5db543d8ea5b00cc330c}{}
 

### 2020-08-23 paper.rmd
work on examples

@@
 ---
-title             : "Carrying out Latent Profile Analysis in Accordance With Best Practices: A Tutorial Using the tidyLPA R package"
+title             : "Decisions Decisions Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package"
 shorttitle        : "tidyLPA"
@@
 
+We can see a number of statistics for each of the models that were estimated, including the 
+AIC (Aikake information criterion; based on -2 log-likelihood, and penalized by number of parameters) and BIC (BIC: Bayesian information criterion; based on -2 log-likelihood, and penalized by number of parameters adjusted by sample size).
+Both are _penalized_ likelihood statistics; the log-likelihood will always decrease with the addition of additional
+paremeters estimated by the model, whereas the AIC and BIC will account for the number of parameters; lower values of
+either indicate better fit.
+
+Entropy.
+`prob_min` refers to the minimum of the diagonal of the average latent class probabilities for most likely class membership, by assigned class.
+The minimum should be as high as possible, reflecting greater classification certainty (cases are assigned to classes they have a high probability of belonging to; see Jung & Wickrama, 2008); `prob_max`
+refers to the maximum.
+Finally.
+`n_min` refers to the proportion of the sample assigned to the smallest class (based on most likely class membership); `n_max` to the largest.
+Finally, `BLRT_p` refers to the p-value for the bootstrapped likelihood ratio test; a value greater than .05 indicates that the model fit better than one with one fewer class.
+
 ``` {r, curry-mac-mclust}
-curry_mac %>% 
+m1 <- curry_mac %>% 
   select(KinshipR:PossessionR, 
          KinshipJ:PossessionJ) %>% 
-  estimate_profiles(3)
+  estimate_profiles(1:10)
+```
+
+Additional fit statistics (other penalized likelihood values, such as the CAIC, SABIC, and ICL), and the bootstrapped LRT 
+test statistic, can be obtained with the `get_fit()` function:
+
+```{r}
+get_fit(m1)
 ```
@@
 We can use Mplus simply by changing the package argument for
-`estimate_profiles()`:
+`estimate_profiles()` to "mplus" or "MplusAutomation" (either will work):
 
@@
          KinshipJ:PossessionJ) %>% 
-  estimate_profiles(3, package = "MplusAutomation")
+  estimate_profiles(3, package = "mplus")
 ```
@@
 The resulting object can be further passed down a pipeline to
-other functions, such as `compare_solutions`, `get_data`, `get_fit`,
-etc. This is the "tidy" part, in that the function can be embedded in a tidy
-analysis pipeline.
-
-If you have Mplus installed, you can call the version of this function that uses
-MPlus in the same way, by adding the argument `package = "MplusAutomation`.
+other functions, such as `compare_solutions`, `get_data`, and `get_fit`.
 

### 2020-08-23 paper.rmd
add model types to paper

@@
 (using the `%>%` operator, loaded from the `dplyr` package). We pass the data to
-a function that selects relevant variables, and then to `estimate_profiles`:
+a function that selects relevant variables, and then to `estimate_profiles`.
+
+A simple summary of the analysis is printed to the console. 
+The resulting object can be further passed down a pipeline to
+other functions, such as `compare_solutions`, `get_data`, and `get_fit`.
+
+``` {r, curry-mac-mclust}
+m1 <- curry_mac %>% 
+  select(KinshipR:PossessionR, 
+         KinshipJ:PossessionJ) %>% 
+  estimate_profiles(1:10)
+```
 
@@
 Finally, `BLRT_p` refers to the p-value for the bootstrapped likelihood ratio test; a value greater than .05 indicates that the model fit better than one with one fewer class.
-
-``` {r, curry-mac-mclust}
-m1 <- curry_mac %>% 
-  select(KinshipR:PossessionR, 
-         KinshipJ:PossessionJ) %>% 
-  estimate_profiles(1:10)
-```
-
 Additional fit statistics (other penalized likelihood values, such as the CAIC, SABIC, and ICL), and the bootstrapped LRT 
@@
          KinshipJ:PossessionJ) %>% 
-  estimate_profiles(3, package = "mplus")
+  estimate_profiles(1:10, package = "mplus")
 ```
 
-A simple summary of the analysis is printed to the console. 
-The resulting object can be further passed down a pipeline to
-other functions, such as `compare_solutions`, `get_data`, and `get_fit`.
+These are the very basics; we will return to what we do with this output later in the tutorial.
 
-### Plotting the profiles
+### Specifying the _model_
 
-We can plot the profiles by piping the output to `plot_profiles()`.
+We passed the number of profiles, but no details about the model we wished to estimate.
 
-```{r, curry_mac_p3}
-curry_mac_three <- curry_mac %>% 
-  select(KinshipR:PossessionR, 
-         KinshipJ:PossessionJ) %>% 
-  estimate_profiles(3) %>% 
-  plot_profiles()
+There are a number of different types of models for LPA that estimate different parameters; choosing 
+between them is an important consideration, and one that has a bearing on how we will interpret the results we saw above.
 
-curry_mac_three +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
-
-## Model specification
-
-In addition to the number of profiles (specified with the `n_profiles`
+Thus, in addition to the number of profiles (specified with the `n_profiles`
 argument), the model can be specified in terms of whether and how the variable
-variances and covariances are estimated.
-
-The models are specified by passing arguments to the `variance` and `covariance`
+variances and covariances are estimated. The models are specified by passing arguments to the `variance` and `covariance`
 arguments. The possible values for these arguments are:
@@
 - `covariances`: "varying", "equal", and "zero"
+
+In general, the approach to choosing the model is similar to choosing the number of profiles, requiring **deciding on the basis of evidence from multiple sources**, including information criteria, statistical tests, and concerns of interpretability and parsimony.
+The article by [Pastor and colleagues (2007)](https://www.sciencedirect.com/science/article/pii/S0361476X06000543) has helpful information on the model specifications.
+Here, the six models that are possible to specify in LPA are described in terms of how the variables used to create the profiles are estimated.
+
+Note that *p* represents different profiles and each parameterization is represented by a 4 x 4 covariance matrix and therefore would represent the parameterization for a four-profile solution.
+In all of the models, the means are estimated freely in the different profiles.
+Imagine that each row and column represents a different variable, i.e., the first row (and column) represents broad interest, the second enjoyment, the third self-efficacy, and the fourth another variable, i.e., future goals and plans.
+
+### 1. Equal variances, and covariances fixed to 0 (model 1)
+
+In this model, which corresponds to the mclust model wit the name "EEI", the variances are estimated to be equal across profiles, indicated by the absence of a p subscript for any of the diagonal elements of the matrix.
+The covariances are constrained to be zero, as indicated by the 0's between every combination of the variables.
+
+It is specified with `variances = "equal"` and `covariances = "zero"`.
+
+This model is highly constrained but also parsimonious: the profiles are estimated in such a way that the variables' variances are identical for each of the profiles, and the relationships between the variables are not estimated.
+In this way, less degrees of freedom are taken used to explain the observations that make up the data.
+However, estimating more parameters--as in the other models--may better explain the data, justifying the addition in complexity that their addition involves (and their reduction in degrees of freedom).
+This model is sometimes referred to as a *class-invariant* parameterization.
+
+$$
+\left[ \begin{matrix} { \sigma  }_{ 1 }^{ 2 } & 0 & 0 & 0 \\ 0 & { \sigma  }_{ 2 }^{ 2 } & 0 & 0 \\ 0 & 0 & { \sigma  }_{ 3 }^{ 2 } & 0 \\ 0 & 0 & 0 & { \sigma  }_{ 4 }^{ 2 } \end{matrix} \right] 
+$$
+
+### 2. Varying variances and covariances fixed to 0 (model 2)
+
+This model corresponds to the mclust model "VVI" and allows for the variances to be freely estimated across profiles. The covariances are constrained to zero.
+
+It is specified with `variances = "varying"` and `covariances = "zero"`.
+
+Thus, it is more flexible (and less parsimonious) than model 1, but in terms of the covariances, is more constrained than model 2.
+This model is sometimes referred to as a *class-varying diagonal* parameterization.
+
+$$ 
+\left[ \begin{matrix} { \sigma  }_{ 1p }^{ 2 } & 0 & 0 & 0 \\ 0 & { \sigma  }_{ 2p }^{ 2 } & 0 & 0 \\ 0 & 0 & { \sigma  }_{ 3p }^{ 2 } & 0 \\ 0 & 0 & 0 & { \sigma  }_{ 4p }^{ 2 } \end{matrix} \right] 
+$$
+
+### 3. Equal variances and equal covariances (model 3)
+
+This model corresponds to the mclust model "EEE".
+In this model, the variances are still constrained to be the same across the profiles, although now the covariances are estimated (but like the variances, are constrained to be the same across profiles).
+
+It is specified with `variances = "equal"` and `covariances = "equal"`.
+
+Thus, this model is the first to estimate the covariance (or correlations) of the variables used to create the profiles, thus adding more information that can be used to better understand the characteristics of the profiles (and, potentially, better explain the data).
+This model is sometimes referred to as a *class-invariant unrestricted* parameterization.
+
+$$
+\left[ \begin{matrix} { \sigma  }_{ 1 }^{ 2 } & { \sigma  }_{ 21 } & { \sigma  }_{ 31 } & { \sigma  }_{ 41 } \\ { \sigma  }_{ 12 } & { \sigma  }_{ 2 }^{ 2 } & { \sigma  }_{ 23 } & { \sigma  }_{ 24 } \\ { \sigma  }_{ 13 } & { \sigma  }_{ 12 } & { \sigma  }_{ 3 }^{ 2 } & { \sigma  }_{ 33 } \\ { \sigma  }_{ 14 } & { \sigma  }_{ 12 } & { \sigma  }_{ 12 } & { \sigma  }_{ 4 }^{ 2 } \end{matrix} \right]
+$$
+
+### 4. Varying means, varying variances, and equal covariances (model 4)
+
+This model, which specifies for the variances to be freely estimated across the profiles and for the covariances to be estimated to be equal across profiles, extends model 3.
+
+It is specified with `variances = "varying"` and `covariances = "equal"`.
+
+Unfortunately, this model cannot be specified with mclust, though it can be with MPlus; this model *can* be used with the functions to interface to MPlus described below.
+
+$$
+\left[ \begin{matrix} { \sigma  }_{ 1p }^{ 2 } & { \sigma  }_{ 21 } & { \sigma  }_{ 31 } & { \sigma  }_{ 41 } \\ { \sigma  }_{ 12 } & { \sigma  }_{ 2p }^{ 2 } & { \sigma  }_{ 23 } & { \sigma  }_{ 24 } \\ { \sigma  }_{ 13 } & { \sigma  }_{ 12 } & { \sigma  }_{ 3p }^{ 2 } & { \sigma  }_{ 33 } \\ { \sigma  }_{ 14 } & { \sigma  }_{ 12 } & { \sigma  }_{ 12 } & { \sigma  }_{ 4p }^{ 2 } \end{matrix} \right]
+$$
+
+### 5. Varying means, equal variances, and varying covariances (model 5)
+
+This model specifies the variances to be equal across the profiles, but allows the covariances to be freely estimated across the profiles.
+
+It is specified with `variances = "equal"` and `covariances = "varying"`.
+
+Like model 4, this model cannot be specified with mclust, though it can be with MPlus. Again, this model *can* be used with the functions to interface to MPlus described below.
+
+$$
+\left[ \begin{matrix} { \sigma  }_{ 1 }^{ 2 } & { \sigma  }_{ 21p } & { \sigma  }_{ 31p } & { \sigma  }_{ 41p } \\ { \sigma  }_{ 12p } & { \sigma  }_{ 2 }^{ 2 } & { \sigma  }_{ 23p } & { \sigma  }_{ 24p } \\ { \sigma  }_{ 13p } & { \sigma  }_{ 12p } & { \sigma  }_{ 3 }^{ 2 } & { \sigma  }_{ 33p } \\ { \sigma  }_{ 14p } & { \sigma  }_{ 12p } & { \sigma  }_{ 12p } & { \sigma  }_{ 4 }^{ 2 } \end{matrix} \right] \quad
+$$
+
+### 6. Varying variances and varying covariances (model 6)
+
+This model corresponds to the mclust model "VVV". It allows the variances and the covariances to be freely estimated across profiles.
+
+It is specified with `variances = "varying"` and `covariances = "varying"`.
+
+Thus, it is the most complex model, with the potential to allow for understanding many aspects of the variables that are used to estimate the profiles and how they are related.
+However, it is less parsimonious than all of the other models, and the added parameters should be considered in light of how preferred this model is relative to those with more simple specifications.
+This model is sometimes referred to as a *class-varying unrestricted* parameterization.
+
+$$
+\left[ \begin{matrix} { \sigma  }_{ 1p }^{ 2 } & { \sigma  }_{ 21p } & { \sigma  }_{ 31p } & { \sigma  }_{ 41p } \\ { \sigma  }_{ 12p } & { \sigma  }_{ 2p }^{ 2 } & { \sigma  }_{ 23p } & { \sigma  }_{ 24p } \\ { \sigma  }_{ 13p } & { \sigma  }_{ 12p } & { \sigma  }_{ 3p }^{ 2 } & { \sigma  }_{ 33p } \\ { \sigma  }_{ 14p } & { \sigma  }_{ 12p } & { \sigma  }_{ 12p } & { \sigma  }_{ 4p }^{ 2 } \end{matrix} \right]
+$$
 
@@
 ````{r, just-one}
-curry_mac_mclust_j <- curry_mac %>% 
+urry_mac %>% 
   select(KinshipR:PossessionR) %>% 
   estimate_profiles(models = c(1, 2, 3, 6),
-                    n_profiles = 1:9) %>%
-  compare_solutions(statistics = c("AIC", "BIC"))
-
-write_rds(curry_mac_mclust_j, "cm_mclust_r.rds")
+                    n_profiles = 1:9)
 
@@
   estimate_profiles(models = c(1, 2, 3, 6),
-                    n_profiles = 1:9) %>%
-  compare_solutions(statistics = c("AIC", "BIC"))
-
-write_rds(curry_mac_mclust_j, "cm_mclust_j.rds")
+                    n_profiles = 1:9)
 
@@
                     n_profiles = 1:9,
-                    package = "mplus") %>%
-  compare_solutions(statistics = c("AIC", "BIC"))
+                    package = "mplus")
 
-write_rds(curry_mac_mclust_j, "cm_mclust_r.rds")
-
-curry_mac_mclust_j <- curry_mac %>% 
+curry_mac %>% 
   select(KinshipJ:PossessionJ) %>% 
@@
                     n_profiles = 1:9,
-                    package = "mplus") %>%
-  compare_solutions(statistics = c("AIC", "BIC"))
-
-write_rds(curry_mac_mclust_j, "cm_mclust_j.rds")
+                    package = "mplus")
 ```
@@
 
-```{r, m6p7-solution}
+```{r, m6p7-solution, eval = FALSE}
 curry_mac_m6p7 <- curry_mac %>% 
@@
 
-```{r, m6p2-solution}
+```{r, m6p2-solution, eval = FALSE}
 curry_mac_m6p2 <- curry_mac %>% 
@@
 
-``` {r, compare-mplus, eval = TRUE}
-curry_mac_mplus_solutions <- curry_mac %>% 
+``` {r, compare-mplus, eval = FALSE}
+curry_mac %>% 
   select(KinshipR:PossessionR, 
@@
   compare_solutions(statistics = c("AIC", "BIC"))
-
-write_rds(curry_mac_mplus_solutions, "cm_mplus.rds")
-
-# Best model according to AIC is Model 6 with 4 classes.
-# Best model according to BIC is Model 6 with 2 classes.
-
-# An analytic hierarchy process, based on the fit indices AIC, AWE, BIC, CLC, and KIC (Akogul & Erisoglu, 2017), suggests the best solution is Model 6 with 2 classes.
 ```
 
-```{r, m6ps-mplus, error = TRUE}
-curry_mac_m6p2_mplus <- curry_mac %>% 
+```{r, m6ps-mplus, eval = FALSE}
+curry_mac %>% 
   select(KinshipR:PossessionR, 
@@
 
-# equivalent to model 6
-
-plot_profiles(curry_mac_m6p2_mplus)
 ```
 
-```{r, m6p2-mclust}
-curry_mac_m6p2 <- curry_mac %>% 
+```{r, m6p2-mclust, eval = FALSE}
+curry_mac %>% 
   select(KinshipR:PossessionR, 
@@
 
-plot_profiles(curry_mac_m6p2)
 ```
@@
 
-``` {r, pisa-get-data}
-m <- pisaUSA15[1:100, ] %>%
-  select(broad_interest, enjoyment, self_efficacy) %>%
-  single_imputation() %>%
-  estimate_profiles(3:4)
-```
-
-```{r}
-get_data(m)
+``` {r}
+get_data()
 ```

### 2020-08-26 paper.rmd
update

@@
 library(tidyverse)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed, cache = TRUE)
+knitr::clean_cache()
+knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
 ```
@@
 
-``` {r, curry-mac-mclust}
+``` {r, curry-mac-mclust-example}
 m1 <- curry_mac %>% 
@@
 
-```{r}
+```{r, get-fit-m1}
 get_fit(m1)
@@
 We can use Mplus simply by changing the package argument for
-`estimate_profiles()` to "mplus" or "MplusAutomation" (either will work):
+`estimate_profiles()` to "mplus" or "MplusAutomation" (either will work; not run here):
 
@@
 
-``` {r, compare-mclust, eval = FALSE}
-curry_mac_mclust_solutions <- curry_mac %>% 
-  select(KinshipR:PossessionR, 
-         KinshipJ:PossessionJ) %>% 
+````{r, compare-mclust}
+compare_solutions_mclust <- curry_mac %>% 
+  select(KinshipJ:PossessionJ) %>% 
   estimate_profiles(models = c(1, 2, 3, 6),
-                    n_profiles = 1:9) %>%
-  compare_solutions(statistics = c("AIC", "BIC"))
-
-write_rds(curry_mac_mclust_solutions, "cm_mclust.rds")
-
-# Best model according to AIC is Model 6 with 7 classes.
-# Best model according to BIC is Model 6 with 2 classes.
-
-#An analytic hierarchy process, based on the fit indices AIC, AWE, BIC, CLC, and KIC (Akogul & Erisoglu, 2017), suggests the best solution is Model 6 with 2 classes.
+                    n_profiles = 1:9,
+                    package = "mclust")
 ```
 
-````{r, just-one}
-urry_mac %>% 
-  select(KinshipR:PossessionR) %>% 
-  estimate_profiles(models = c(1, 2, 3, 6),
-                    n_profiles = 1:9)
+```{r, eval = FALSE}
+write_rds(compare_solutions_mclust, "compare_solutions_mclust.rds")
+````
 
-curry_mac_mclust_j <- curry_mac %>% 
+```{r, compare-mplus}
+compare_solutions_mplus <- curry_mac %>% 
   select(KinshipJ:PossessionJ) %>% 
-  estimate_profiles(models = c(1, 2, 3, 6),
-                    n_profiles = 1:9)
-
-curry_mac_mclust_j <- curry_mac %>% 
-  select(KinshipR:PossessionR) %>% 
-  estimate_profiles(models = c(1, 2, 3, 6),
-                    n_profiles = 1:9,
-                    package = "mplus")
-
-curry_mac %>% 
-  select(KinshipJ:PossessionJ) %>% 
-  estimate_profiles(models = c(1, 2, 3, 6),
+  estimate_profiles(models = c(1, 2, 3, 4, 5, 6),
                     n_profiles = 1:9,
@@
 
-The results of this analysis suggests that the Model 6, two or seven class models are best.
-
-```{r, m6p7-solution, eval = FALSE}
-curry_mac_m6p7 <- curry_mac %>% 
-  select(KinshipR:PossessionR, 
-         KinshipJ:PossessionJ) %>% 
-  estimate_profiles(3, 
-                    variances = "varying",
-                    covariances = "varying")
-
-plot_profiles(curry_mac_m6p7)
+```{r, echo = FALSE}
+write_rds(compare_solutions_mplus, "compare_solutions_mplus.rds")
 ```
 
-```{r, m6p2-solution, eval = FALSE}
-curry_mac_m6p2 <- curry_mac %>% 
-  select(KinshipR:PossessionR, 
-         KinshipJ:PossessionJ) %>% 
-  estimate_profiles(3, 
-                    variances = "varying",
-                    covariances = "varying")
-
-plot_profiles(curry_mac_m6p7)
+```{r, compare-mplus-starts}
+compare_solutions_mplus_starts <- curry_mac %>% 
+  select(KinshipJ:PossessionJ) %>% 
+  estimate_profiles(models = c(1, 2, 3, 4, 5, 6),
+                    n_profiles = 1:9,
+                    package = "mplus",
+                    ANALYSIS = "starts = 100, 20;")
 ```
 
-We can also run the same process using mplus:
-
-``` {r, compare-mplus, eval = FALSE}
-curry_mac %>% 
-  select(KinshipR:PossessionR, 
-         KinshipJ:PossessionJ) %>% 
-  estimate_profiles(models = c(1, 2, 3, 6),
-                    n_profiles = 1:9, package = "mplus") %>%
-  compare_solutions(statistics = c("AIC", "BIC"))
+```{r, echo = FALSE}
+write_rds(compare_solutions_mplus_starts, "compare_solutions_mplus_starts.rds")
 ```
 
-```{r, m6ps-mplus, eval = FALSE}
-curry_mac %>% 
-  select(KinshipR:PossessionR, 
-         KinshipJ:PossessionJ) %>% 
-  estimate_profiles(n_profiles = 2, 
-                    variances = "varying",
-                    covariances = "varying", 
-                    package = "mplus")
-
-```
-
-```{r, m6p2-mclust, eval = FALSE}
-curry_mac %>% 
-  select(KinshipR:PossessionR, 
-         KinshipJ:PossessionJ) %>% 
-  estimate_profiles(n_profiles = 2, 
-                    variances = "varying",
-                    covariances = "varying")
-
-```
+The results of this analysis suggests that the Model 6, two or seven class models are best.
 
@@
 
-``` {r}
+``` {r, eval = FALSE}
 get_data()
@@
 
-``` {r}
+``` {r, eval = FALSE}
 get_data(m) %>% 

### 2020-08-27 paper.tex
update

--- 2020-01-16 paper.tex
@@
 \hypersetup{
-            pdftitle={Carrying out Latent Profile Analysis in Accordance With Best Practices: A Tutorial Using the tidyLPA R package},
+            pdftitle={Decisions Decisions Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package},
             pdfauthor={Joshua Rosenberg, Caspar van Lissa, Jennifer Schmidt, Patrick Beymer, Daniel Anderson, \& Matthew Schell},
@@
 
-\usepackage{etoolbox}
+\shorttitle{tidyLPA}
+\affiliation{
+\vspace{0.5cm}
+\textsuperscript{1} University of Tennessee, Knoxville\\\textsuperscript{2} Utrecht University\\\textsuperscript{3} Michigan State University\\\textsuperscript{4} University of Oregon\\\textsuperscript{5} }
+\keywords{Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust\newline\indent Word count: }
+\usepackage{csquotes}
+\usepackage{upgreek}
+\captionsetup{font=singlespacing,justification=justified}
+
+\usepackage{longtable}
+\usepackage{lscape}
+\usepackage{multirow}
+\usepackage{tabularx}
+\usepackage[flushleft]{threeparttable}
+\usepackage{threeparttablex}
+
+\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}
+
 \makeatletter
-\providecommand{\subtitle}[1]{% add subtitle to \maketitle
-  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
-}
+\newcommand\LastLTentrywidth{1em}
+\newlength\longtablewidth
+\setlength{\longtablewidth}{1in}
+\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}
+
+\DeclareDelayedFloatFlavor{ThreePartTable}{table}
+\DeclareDelayedFloatFlavor{lltable}{table}
+\DeclareDelayedFloatFlavor*{longtable}{table}
+\makeatletter
+\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
 \makeatother
 
-\title{Carrying out Latent Profile Analysis in Accordance With Best Practices: A Tutorial Using the tidyLPA R package}
+\title{Decisions Decisions Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package}
 \author{Joshua Rosenberg\textsuperscript{1}, Caspar van Lissa\textsuperscript{2}, Jennifer Schmidt\textsuperscript{3}, Patrick Beymer\textsuperscript{3}, Daniel Anderson\textsuperscript{4}, \& Matthew Schell\textsuperscript{3}}
@@
 \maketitle
+
+\begin{verbatim}
+## Warning: package 'tibble' was built under R version 3.6.2
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: package 'purrr' was built under R version 3.6.2
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: package 'dplyr' was built under R version 3.6.2
+\end{verbatim}
+
+\begin{verbatim}
+## NULL
+\end{verbatim}
 
@@
 covariances are fixed to zero (also described as the specification of the
-covariance matrix as ``diagonal,'' because only the diagonal components, or the
+covariance matrix as \enquote{diagonal,} because only the diagonal components, or the
 variances, are estimated).
@@
 model specifications, which are often described in terms of the geometric
-properties of the multivariate distributions being estimated (i.e., ``spherical,
-equal volume''), rather than in terms of whether and how the means, variances,
+properties of the multivariate distributions being estimated (i.e., \enquote{spherical,
+equal volume}), rather than in terms of whether and how the means, variances,
 and covariances are estimated. They also may use different default settings
@@
 
-One way that tidyLPA is designed to be easy to use is that it assumes a ``tidy''
+One way that tidyLPA is designed to be easy to use is that it assumes a \enquote{tidy}
 data structure (Wickham, 2014). This means that it emphasizes the use of a data
@@
 
+\begin{verbatim}
+## Warning: package 'tidyLPA' was built under R version 3.6.2
+\end{verbatim}
+
 \hypertarget{examples}{%
@@
 (using the \texttt{\%\textgreater{}\%} operator, loaded from the \texttt{dplyr} package). We pass the data to
-a function that selects relevant variables, and then to \texttt{estimate\_profiles}:
+a function that selects relevant variables, and then to \texttt{estimate\_profiles}.
+
+A simple summary of the analysis is printed to the console.
+The resulting object can be further passed down a pipeline to
+other functions, such as \texttt{compare\_solutions}, \texttt{get\_data}, and \texttt{get\_fit}.
+
+We can see a number of statistics for each of the models that were estimated, including the
+AIC (Aikake information criterion; based on -2 log-likelihood, and penalized by number of parameters) and BIC (BIC: Bayesian information criterion; based on -2 log-likelihood, and penalized by number of parameters adjusted by sample size).
+Both are \emph{penalized} likelihood statistics; the log-likelihood will always decrease with the addition of additional
+paremeters estimated by the model, whereas the AIC and BIC will account for the number of parameters; lower values of
+either indicate better fit.
+
+Entropy.
+\texttt{prob\_min} refers to the minimum of the diagonal of the average latent class probabilities for most likely class membership, by assigned class.
+The minimum should be as high as possible, reflecting greater classification certainty (cases are assigned to classes they have a high probability of belonging to; see Jung \& Wickrama, 2008); \texttt{prob\_max}
+refers to the maximum.
+Finally.
+\texttt{n\_min} refers to the proportion of the sample assigned to the smallest class (based on most likely class membership); \texttt{n\_max} to the largest.
+Finally, \texttt{BLRT\_p} refers to the p-value for the bootstrapped likelihood ratio test; a value greater than .05 indicates that the model fit better than one with one fewer class.
+Additional fit statistics (other penalized likelihood values, such as the CAIC, SABIC, and ICL), and the bootstrapped LRT
+test statistic, can be obtained with the \texttt{get\_fit()} function:
 
 \begin{verbatim}
-## tidyLPA analysis using mclust: 
-## 
-##  Model Classes AIC       BIC       Entropy prob_min prob_max n_min n_max BLRT_p
-##  1     3       158594.55 158898.38 0.82    0.91     0.92     0.15  0.53  0.01
+## # A tibble: 10 x 18
+##    Model Classes  LogLik    AIC    AWE    BIC   CAIC    CLC    KIC  SABIC
+##    <dbl>   <dbl>   <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>
+##  1     1       1 -81245. 1.63e5 1.63e5 1.63e5 1.63e5 1.62e5 1.63e5 1.63e5
+##  2     1       2 -79727. 1.60e5 1.60e5 1.60e5 1.60e5 1.59e5 1.60e5 1.60e5
+##  3     1       3 -79239. 1.59e5 1.59e5 1.59e5 1.59e5 1.58e5 1.59e5 1.59e5
+##  4     1       4 -79091. 1.58e5 1.59e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
+##  5     1       5 -79029. 1.58e5 1.60e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
+##  6     1       6 -78971. 1.58e5 1.60e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
+##  7     1       7 -78930. 1.58e5 1.60e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
+##  8     1       8 -78852. 1.58e5 1.60e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
+##  9     1       9 -78822. 1.58e5 1.60e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
+## 10     1      10 -78751. 1.58e5 1.60e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
+## # ... with 8 more variables: ICL <dbl>, Entropy <dbl>, prob_min <dbl>,
+## #   prob_max <dbl>, n_min <dbl>, n_max <dbl>, BLRT_val <dbl>, BLRT_p <dbl>
 \end{verbatim}
@@
 We can use Mplus simply by changing the package argument for
-\texttt{estimate\_profiles()}:
+\texttt{estimate\_profiles()} to \enquote{mplus} or \enquote{MplusAutomation} (either will work; not run here):
 
-A simple summary of the analysis is printed to the console.
-The resulting object can be further passed down a pipeline to
-other functions, such as \texttt{compare\_solutions}, \texttt{get\_data}, \texttt{get\_fit},
-etc. This is the ``tidy'' part, in that the function can be embedded in a tidy
-analysis pipeline.
+These are the very basics; we will return to what we do with this output later in the tutorial.
 
-If you have Mplus installed, you can call the version of this function that uses
-MPlus in the same way, by adding the argument \texttt{package\ =\ "MplusAutomation}.
+\hypertarget{specifying-the-model}{%
+\subsubsection{\texorpdfstring{Specifying the \emph{model}}{Specifying the model}}\label{specifying-the-model}}
 
-\hypertarget{plotting-the-profiles}{%
-\subsubsection{Plotting the profiles}\label{plotting-the-profiles}}
+We passed the number of profiles, but no details about the model we wished to estimate.
 
-We can plot the profiles by piping the output to \texttt{plot\_profiles()}.
+There are a number of different types of models for LPA that estimate different parameters; choosing
+between them is an important consideration, and one that has a bearing on how we will interpret the results we saw above.
 
-\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}
-
-\hypertarget{model-specification}{%
-\subsection{Model specification}\label{model-specification}}
-
-In addition to the number of profiles (specified with the \texttt{n\_profiles}
+Thus, in addition to the number of profiles (specified with the \texttt{n\_profiles}
 argument), the model can be specified in terms of whether and how the variable
-variances and covariances are estimated.
-
-The models are specified by passing arguments to the \texttt{variance} and \texttt{covariance}
+variances and covariances are estimated. The models are specified by passing arguments to the \texttt{variance} and \texttt{covariance}
 arguments. The possible values for these arguments are:
@@
 \item
-  \texttt{variances}: ``equal'' and ``zero''
+  \texttt{variances}: \enquote{equal} and \enquote{zero}
 \item
-  \texttt{covariances}: ``varying'', ``equal'', and ``zero''
+  \texttt{covariances}: \enquote{varying}, \enquote{equal}, and \enquote{zero}
 \end{itemize}
+
+In general, the approach to choosing the model is similar to choosing the number of profiles, requiring \textbf{deciding on the basis of evidence from multiple sources}, including information criteria, statistical tests, and concerns of interpretability and parsimony.
+The article by \href{https://www.sciencedirect.com/science/article/pii/S0361476X06000543}{Pastor and colleagues (2007)} has helpful information on the model specifications.
+Here, the six models that are possible to specify in LPA are described in terms of how the variables used to create the profiles are estimated.
+
+Note that \emph{p} represents different profiles and each parameterization is represented by a 4 x 4 covariance matrix and therefore would represent the parameterization for a four-profile solution.
+In all of the models, the means are estimated freely in the different profiles.
+Imagine that each row and column represents a different variable, i.e., the first row (and column) represents broad interest, the second enjoyment, the third self-efficacy, and the fourth another variable, i.e., future goals and plans.
+
+\hypertarget{equal-variances-and-covariances-fixed-to-0-model-1}{%
+\subsubsection{1. Equal variances, and covariances fixed to 0 (model 1)}\label{equal-variances-and-covariances-fixed-to-0-model-1}}
+
+In this model, which corresponds to the mclust model wit the name \enquote{EEI}, the variances are estimated to be equal across profiles, indicated by the absence of a p subscript for any of the diagonal elements of the matrix.
+The covariances are constrained to be zero, as indicated by the 0's between every combination of the variables.
+
+It is specified with \texttt{variances\ =\ "equal"} and \texttt{covariances\ =\ "zero"}.
+
+This model is highly constrained but also parsimonious: the profiles are estimated in such a way that the variables' variances are identical for each of the profiles, and the relationships between the variables are not estimated.
+In this way, less degrees of freedom are taken used to explain the observations that make up the data.
+However, estimating more parameters--as in the other models--may better explain the data, justifying the addition in complexity that their addition involves (and their reduction in degrees of freedom).
+This model is sometimes referred to as a \emph{class-invariant} parameterization.
+
+\[
+\left[ \begin{matrix} { \sigma  }_{ 1 }^{ 2 } & 0 & 0 & 0 \\ 0 & { \sigma  }_{ 2 }^{ 2 } & 0 & 0 \\ 0 & 0 & { \sigma  }_{ 3 }^{ 2 } & 0 \\ 0 & 0 & 0 & { \sigma  }_{ 4 }^{ 2 } \end{matrix} \right] 
+\]
+
+\hypertarget{varying-variances-and-covariances-fixed-to-0-model-2}{%
+\subsubsection{2. Varying variances and covariances fixed to 0 (model 2)}\label{varying-variances-and-covariances-fixed-to-0-model-2}}
+
+This model corresponds to the mclust model \enquote{VVI} and allows for the variances to be freely estimated across profiles. The covariances are constrained to zero.
+
+It is specified with \texttt{variances\ =\ "varying"} and \texttt{covariances\ =\ "zero"}.
+
+Thus, it is more flexible (and less parsimonious) than model 1, but in terms of the covariances, is more constrained than model 2.
+This model is sometimes referred to as a \emph{class-varying diagonal} parameterization.
+
+\[ 
+\left[ \begin{matrix} { \sigma  }_{ 1p }^{ 2 } & 0 & 0 & 0 \\ 0 & { \sigma  }_{ 2p }^{ 2 } & 0 & 0 \\ 0 & 0 & { \sigma  }_{ 3p }^{ 2 } & 0 \\ 0 & 0 & 0 & { \sigma  }_{ 4p }^{ 2 } \end{matrix} \right] 
+\]
+
+\hypertarget{equal-variances-and-equal-covariances-model-3}{%
+\subsubsection{3. Equal variances and equal covariances (model 3)}\label{equal-variances-and-equal-covariances-model-3}}
+
+This model corresponds to the mclust model \enquote{EEE}.
+In this model, the variances are still constrained to be the same across the profiles, although now the covariances are estimated (but like the variances, are constrained to be the same across profiles).
+
+It is specified with \texttt{variances\ =\ "equal"} and \texttt{covariances\ =\ "equal"}.
+
+Thus, this model is the first to estimate the covariance (or correlations) of the variables used to create the profiles, thus adding more information that can be used to better understand the characteristics of the profiles (and, potentially, better explain the data).
+This model is sometimes referred to as a \emph{class-invariant unrestricted} parameterization.
+
+\[
+\left[ \begin{matrix} { \sigma  }_{ 1 }^{ 2 } & { \sigma  }_{ 21 } & { \sigma  }_{ 31 } & { \sigma  }_{ 41 } \\ { \sigma  }_{ 12 } & { \sigma  }_{ 2 }^{ 2 } & { \sigma  }_{ 23 } & { \sigma  }_{ 24 } \\ { \sigma  }_{ 13 } & { \sigma  }_{ 12 } & { \sigma  }_{ 3 }^{ 2 } & { \sigma  }_{ 33 } \\ { \sigma  }_{ 14 } & { \sigma  }_{ 12 } & { \sigma  }_{ 12 } & { \sigma  }_{ 4 }^{ 2 } \end{matrix} \right]
+\]
+
+\hypertarget{varying-means-varying-variances-and-equal-covariances-model-4}{%
+\subsubsection{4. Varying means, varying variances, and equal covariances (model 4)}\label{varying-means-varying-variances-and-equal-covariances-model-4}}
+
+This model, which specifies for the variances to be freely estimated across the profiles and for the covariances to be estimated to be equal across profiles, extends model 3.
+
+It is specified with \texttt{variances\ =\ "varying"} and \texttt{covariances\ =\ "equal"}.
+
+Unfortunately, this model cannot be specified with mclust, though it can be with MPlus; this model \emph{can} be used with the functions to interface to MPlus described below.
+
+\[
+\left[ \begin{matrix} { \sigma  }_{ 1p }^{ 2 } & { \sigma  }_{ 21 } & { \sigma  }_{ 31 } & { \sigma  }_{ 41 } \\ { \sigma  }_{ 12 } & { \sigma  }_{ 2p }^{ 2 } & { \sigma  }_{ 23 } & { \sigma  }_{ 24 } \\ { \sigma  }_{ 13 } & { \sigma  }_{ 12 } & { \sigma  }_{ 3p }^{ 2 } & { \sigma  }_{ 33 } \\ { \sigma  }_{ 14 } & { \sigma  }_{ 12 } & { \sigma  }_{ 12 } & { \sigma  }_{ 4p }^{ 2 } \end{matrix} \right]
+\]
+
+\hypertarget{varying-means-equal-variances-and-varying-covariances-model-5}{%
+\subsubsection{5. Varying means, equal variances, and varying covariances (model 5)}\label{varying-means-equal-variances-and-varying-covariances-model-5}}
+
+This model specifies the variances to be equal across the profiles, but allows the covariances to be freely estimated across the profiles.
+
+It is specified with \texttt{variances\ =\ "equal"} and \texttt{covariances\ =\ "varying"}.
+
+Like model 4, this model cannot be specified with mclust, though it can be with MPlus. Again, this model \emph{can} be used with the functions to interface to MPlus described below.
+
+\[
+\left[ \begin{matrix} { \sigma  }_{ 1 }^{ 2 } & { \sigma  }_{ 21p } & { \sigma  }_{ 31p } & { \sigma  }_{ 41p } \\ { \sigma  }_{ 12p } & { \sigma  }_{ 2 }^{ 2 } & { \sigma  }_{ 23p } & { \sigma  }_{ 24p } \\ { \sigma  }_{ 13p } & { \sigma  }_{ 12p } & { \sigma  }_{ 3 }^{ 2 } & { \sigma  }_{ 33p } \\ { \sigma  }_{ 14p } & { \sigma  }_{ 12p } & { \sigma  }_{ 12p } & { \sigma  }_{ 4 }^{ 2 } \end{matrix} \right] \quad
+\]
+
+\hypertarget{varying-variances-and-varying-covariances-model-6}{%
+\subsubsection{6. Varying variances and varying covariances (model 6)}\label{varying-variances-and-varying-covariances-model-6}}
+
+This model corresponds to the mclust model \enquote{VVV}. It allows the variances and the covariances to be freely estimated across profiles.
+
+It is specified with \texttt{variances\ =\ "varying"} and \texttt{covariances\ =\ "varying"}.
+
+Thus, it is the most complex model, with the potential to allow for understanding many aspects of the variables that are used to estimate the profiles and how they are related.
+However, it is less parsimonious than all of the other models, and the added parameters should be considered in light of how preferred this model is relative to those with more simple specifications.
+This model is sometimes referred to as a \emph{class-varying unrestricted} parameterization.
+
+\[
+\left[ \begin{matrix} { \sigma  }_{ 1p }^{ 2 } & { \sigma  }_{ 21p } & { \sigma  }_{ 31p } & { \sigma  }_{ 41p } \\ { \sigma  }_{ 12p } & { \sigma  }_{ 2p }^{ 2 } & { \sigma  }_{ 23p } & { \sigma  }_{ 24p } \\ { \sigma  }_{ 13p } & { \sigma  }_{ 12p } & { \sigma  }_{ 3p }^{ 2 } & { \sigma  }_{ 33p } \\ { \sigma  }_{ 14p } & { \sigma  }_{ 12p } & { \sigma  }_{ 12p } & { \sigma  }_{ 4p }^{ 2 } \end{matrix} \right]
+\]
 
@@
 
-We can also run the same process using mplus:
+\begin{verbatim}
+## Warning: 
+## One or more analyses resulted in warnings!
+Examine these analyses carefully: model_1_class_6, model_1_class_7, model_1_class_8, model_1_class_9, model_2_class_5, model_2_class_6, model_2_class_7, model_2_class_8, model_2_class_9, model_3_class_1, model_3_class_2, model_3_class_3, model_3_class_4, model_3_class_5, model_3_class_6, model_3_class_7, model_3_class_8, model_3_class_9, model_4_class_1, model_4_class_2, model_4_class_3, model_4_class_4, model_4_class_5, model_4_class_6, model_4_class_7, model_4_class_8, model_4_class_9, model_5_class_1, model_5_class_2, model_5_class_3, model_5_class_4, model_5_class_5, model_5_class_6, model_5_class_7, model_5_class_8, model_5_class_9, model_6_class_3, model_6_class_4, model_6_class_5, model_6_class_6, model_6_class_7, model_6_class_8, model_6_class_9
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: 
+## One or more analyses resulted in warnings!
+Examine these analyses carefully: model_1_class_7, model_1_class_8, model_1_class_9, model_2_class_6, model_2_class_7, model_2_class_8, model_2_class_9, model_3_class_1, model_3_class_2, model_3_class_3, model_3_class_4, model_3_class_5, model_3_class_6, model_3_class_7, model_3_class_8, model_3_class_9, model_4_class_1, model_4_class_2, model_4_class_3, model_4_class_4, model_4_class_5, model_4_class_6, model_4_class_7, model_4_class_8, model_4_class_9, model_5_class_1, model_5_class_2, model_5_class_3, model_5_class_4, model_5_class_5, model_5_class_6, model_5_class_7, model_5_class_8, model_5_class_9, model_6_class_3, model_6_class_4, model_6_class_5, model_6_class_6, model_6_class_7, model_6_class_8, model_6_class_9
+\end{verbatim}
+
+The results of this analysis suggests that the Model 6, two or seven class models are best.
 
@@
 \texttt{get\_data()} returns the data:
-
-\begin{verbatim}
-## # A tibble: 700 x 9
-##    model_number classes_number broad_interest enjoyment self_efficacy Class
-##           <dbl>          <int>          <dbl>     <dbl>         <dbl> <dbl>
-##  1            1              3            3.8       4            1        1
-##  2            1              3            3         3            2.75     3
-##  3            1              3            1.8       2.8          3.38     3
-##  4            1              3            1.4       1            2.75     2
-##  5            1              3            1.8       2.2          2        3
-##  6            1              3            1.6       1.6          1.88     3
-##  7            1              3            3         3.8          2.25     1
-##  8            1              3            2.6       2.2          2        3
-##  9            1              3            1         2.8          2.62     3
-## 10            1              3            2.2       2            1.75     3
-## # … with 690 more rows, and 3 more variables: Class_prob <int>,
-## #   Probability <dbl>, id <int>
-\end{verbatim}
 
@@
 
-\begin{verbatim}
-## # A tibble: 700 x 9
-##    model_number classes_number broad_interest enjoyment self_efficacy Class
-##           <dbl>          <int>          <dbl>     <dbl>         <dbl> <dbl>
-##  1            1              3            3.8       4            1        1
-##  2            1              3            3         3            2.75     3
-##  3            1              3            1.8       2.8          3.38     3
-##  4            1              3            1.4       1            2.75     2
-##  5            1              3            1.8       2.2          2        3
-##  6            1              3            1.6       1.6          1.88     3
-##  7            1              3            3         3.8          2.25     1
-##  8            1              3            2.6       2.2          2        3
-##  9            1              3            1         2.8          2.62     3
-## 10            1              3            2.2       2            1.75     3
-## # … with 690 more rows, and 3 more variables: Class_prob <int>,
-## #   Probability <dbl>, id <int>
-\end{verbatim}
-
 \hypertarget{recommendations-regarding-analytic-choices}{%
@@
 
-\hypertarget{htmlwidget-a6cf205f9fe062a59d12}{}
+\hypertarget{htmlwidget-31a3e9100620ea87273d}{}
 
@@
 of, for example, tools to carry out highly-sophisticated Bayesian methods using
-the most cutting-edge tools (Bürkner \& others, 2017, @makowski2019). We think
+the most cutting-edge tools (Bürkner \& others, 2017, Makowski, Ben-Shachar, Chen, and Lüdecke (2019)). We think
 that tidyLPA speaks to this, too.
@@
 
+\leavevmode\hypertarget{ref-burkner2017}{}%
+Bürkner, P.-C., \& others. (2017). Brms: An r package for bayesian multilevel models using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.
+
+\leavevmode\hypertarget{ref-makowski2019}{}%
+Makowski, D., Ben-Shachar, M. S., Chen, S., \& Lüdecke, D. (2019). Indices of effect existence and significance in the bayesian framework. \emph{Frontiers in Psychology}, \emph{10}, 2767.
+
+\leavevmode\hypertarget{ref-R-tidyLPA}{}%
+Rosenberg, J.
+M., Beymer, P.
+N., Anderson, D.
+J., Lissa, C.
+van, \& Schmidt, J.
+A.
+(2018).
+TidyLPA: An r package to easily carry out latent profile analysis (lpa) using open-source or commercial software.
+\emph{Journal of Open Source Software}, \emph{3}(30), 978.
+doi:\href{https://doi.org/10.21105/joss.00978}{10.21105/joss.00978}
+
 \endgroup

### 2020-08-27 paper.rmd
update

@@
 
-````{r, compare-mclust}
+````{r, compare-mclust, eval= FALSE}
 compare_solutions_mclust <- curry_mac %>% 
@@
 
-```{r, eval = FALSE}
+```{r, eval = FALSE, echo = FALSE}
 write_rds(compare_solutions_mclust, "compare_solutions_mclust.rds")
-````
+```
 
-```{r, compare-mplus}
+```{r, echo = FALSE}
+compare_solutions_mclust <- read_rds("compare_solutions_mclust.rds")
+compare_solutions_mclust
+```
+
+```{r, compare-mplus, eval = FALSE}
 compare_solutions_mplus <- curry_mac %>% 
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, eval = FALSE}
 write_rds(compare_solutions_mplus, "compare_solutions_mplus.rds")
@@
 
-```{r, compare-mplus-starts}
+```{r, echo = FALSE}
+compare_solutions_mplus <- read_rds("compare_solutions_mplus.rds")
+compare_solutions_mplus
+```
+
+```{r, compare-mplus-starts, eval = FALSE}
 compare_solutions_mplus_starts <- curry_mac %>% 
@@
 
+```{r, echo = FALSE, eval = FALSE}
+write_rds(compare_solutions_mplus_starts, "compare_solutions_mplus_starts.rds")
+```
+
 ```{r, echo = FALSE}
-write_rds(compare_solutions_mplus_starts, "compare_solutions_mplus_starts.rds")
+compare_solutions_mplus <- read_rds("compare_solutions_mplus_starts.rds")
+compare_solutions_mplus
 ```

### 2020-10-12 paper.rmd
update article

--- 2020-08-26 paper.rmd
@@
 knitr::clean_cache()
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
+knitr::opts_chunk$set(cache.extra = knitr::rand_seed, warning = FALSE)
 ```
@@
 
+Note that here and throughout this paper we use the function `knitr::kable()` to format the output in an RMarkdown document, like the one we are using; this is not necessary to use unless you also wish to format the tidyLPA output in an RMarkdown document.
+
 ``` {r, curry-mac-mclust-example}
@@
          KinshipJ:PossessionJ) %>% 
-  estimate_profiles(1:10)
+  estimate_profiles(1:10) %>% 
+  knitr::kable()
 ```
@@
 
-```{r, eval = FALSE}
+```{r, eval = TRUE}
 write_rds(compare_solutions_mclust, "compare_solutions_mclust.rds")
@@
 t <- read_csv("GRoLTS-applicability-to-LPA.csv")
+
 tp <- t %>% 

### 2020-10-12 paper.tex

@@
-\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
 \PassOptionsToPackage{hyphens}{url}
 %
-\documentclass[man]{apa6}
+\documentclass[
+  english,
+  man]{apa6}
 \usepackage{lmodern}
@@
 \usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provides euro and other symbols
-\else % if luatex or xelatex
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
   \usepackage{unicode-math}
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
+\IfFileExists{microtype.sty}{% use microtype if available
   \usepackage[]{microtype}
@@
 }{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
   \IfFileExists{parskip.sty}{%
@@
     \setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\usepackage{hyperref}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \hypersetup{
   pdftitle={Decisions Decisions Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package},
-            pdfauthor={Joshua Rosenberg, Caspar van Lissa, Jennifer Schmidt, Patrick Beymer, Daniel Anderson, \& Matthew Schell},
+  pdflang={en-EN},
   pdfkeywords={Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage{graphicx,grffile}
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% Make \paragraph and \subparagraph free-standing
 \ifx\paragraph\undefined\else
@@
 \fi
-
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-
-\shorttitle{tidyLPA}
-\affiliation{
-\vspace{0.5cm}
-\textsuperscript{1} University of Tennessee, Knoxville\\\textsuperscript{2} Utrecht University\\\textsuperscript{3} Michigan State University\\\textsuperscript{4} University of Oregon\\\textsuperscript{5} }
-\keywords{Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust\newline\indent Word count: }
-\usepackage{csquotes}
+% Manuscript styling
 \usepackage{upgreek}
@@
 
+% Table formatting
 \usepackage{longtable}
 \usepackage{lscape}
-\usepackage{multirow}
-\usepackage{tabularx}
-\usepackage[flushleft]{threeparttable}
-\usepackage{threeparttablex}
+% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
+\usepackage{multirow}		% Table styling
+\usepackage{tabularx}		% Control Column width
+\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
+\usepackage{threeparttablex}            % Lets threeparttable work with longtable
 
+% Create new environments so endfloat can handle them
+% \newenvironment{ltable}
+%   {\begin{landscape}\begin{center}\begin{threeparttable}}
+%   {\end{threeparttable}\end{center}\end{landscape}}
 \newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}
 
+% Enables adjusting longtable caption width to table width
+% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
 \makeatletter
@@
 
+% \setlength{\parindent}{0.5in}
+% \setlength{\parskip}{0pt plus 0pt minus 0pt}
+
+% \usepackage{etoolbox}
+\makeatletter
+\patchcmd{\HyOrg@maketitle}
+  {\section{\normalfont\normalsize\abstractname}}
+  {\section*{\normalfont\normalsize\abstractname}}
+  {}{\typeout{Failed to patch abstract.}}
+\makeatother
+\shorttitle{tidyLPA}
+\author{Joshua Rosenberg\textsuperscript{1}, Caspar van Lissa\textsuperscript{2}, Jennifer Schmidt\textsuperscript{3}, Patrick Beymer\textsuperscript{3}, Daniel Anderson\textsuperscript{4}, \& Matthew Schell\textsuperscript{3}}
+\affiliation{
+\vspace{0.5cm}
+\textsuperscript{1} University of Tennessee, Knoxville\\\textsuperscript{2} Utrecht University\\\textsuperscript{3} Michigan State University\\\textsuperscript{4} University of Oregon\\\textsuperscript{5} }
+\authornote{
+
+Correspondence concerning this article should be addressed to Joshua Rosenberg, 1122 Volunteer Blvd., Knoxville, TN, 37996. E-mail: jmrosenberg@utk.edu}
+\keywords{Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust\newline\indent Word count: }
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 \makeatother
+\usepackage{csquotes}
+\ifxetex
+  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
+  \usepackage{polyglossia}
+  \setmainlanguage[]{english}
+\else
+  \usepackage[shorthands=off,main=english]{babel}
+\fi
 
 \title{Decisions Decisions Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package}
-\author{Joshua Rosenberg\textsuperscript{1}, Caspar van Lissa\textsuperscript{2}, Jennifer Schmidt\textsuperscript{3}, Patrick Beymer\textsuperscript{3}, Daniel Anderson\textsuperscript{4}, \& Matthew Schell\textsuperscript{3}}
+
 \date{}
-
-\authornote{
-
-Correspondence concerning this article should be addressed to Joshua Rosenberg, 1122 Volunteer Blvd., Knoxville, TN, 37996. E-mail: \href{mailto:jmrosenberg@utk.edu}{\nolinkurl{jmrosenberg@utk.edu}}}
-
-\abstract{
-
-}
 
@@
 \maketitle
-
-\begin{verbatim}
-## Warning: package 'tibble' was built under R version 3.6.2
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: package 'purrr' was built under R version 3.6.2
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: package 'dplyr' was built under R version 3.6.2
-\end{verbatim}
 
@@
 parameters and the distinct model parameterizations that can be estimated is the
-different terminology used. As one example, Scrucca et al. (2017) refer to these
+different terminology used. As one example, Scrucca et al.~(2017) refer to these
 parameterizations not in terms of whether and how parameters are estimated, but
@@
 You can also install the development version of tidyLPA from GitHub with:
-
-\begin{verbatim}
-## Warning: package 'tidyLPA' was built under R version 3.6.2
-\end{verbatim}
 
@@
 \begin{verbatim}
-## Warning: 
-## One or more analyses resulted in warnings!
-Examine these analyses carefully: model_1_class_6, model_1_class_7, model_1_class_8, model_1_class_9, model_2_class_5, model_2_class_6, model_2_class_7, model_2_class_8, model_2_class_9, model_3_class_1, model_3_class_2, model_3_class_3, model_3_class_4, model_3_class_5, model_3_class_6, model_3_class_7, model_3_class_8, model_3_class_9, model_4_class_1, model_4_class_2, model_4_class_3, model_4_class_4, model_4_class_5, model_4_class_6, model_4_class_7, model_4_class_8, model_4_class_9, model_5_class_1, model_5_class_2, model_5_class_3, model_5_class_4, model_5_class_5, model_5_class_6, model_5_class_7, model_5_class_8, model_5_class_9, model_6_class_3, model_6_class_4, model_6_class_5, model_6_class_6, model_6_class_7, model_6_class_8, model_6_class_9
+## tidyLPA analysis using mclust: 
+## 
+##  Model Classes AIC      BIC      Entropy prob_min prob_max n_min n_max BLRT_p
+##  1     1       79926.21 79999.54 1.00    1.00     1.00     1.00  1.00        
+##  1     2       79307.22 79422.47 0.59    0.86     0.89     0.44  0.56  0.01  
+##  1     3       79297.86 79455.02 0.54    0.45     0.87     0.10  0.45  0.01  
+##  1     4       79253.01 79452.07 0.51    0.53     0.82     0.13  0.45  0.01  
+##  1     5       79198.41 79439.38 0.54    0.49     0.82     0.05  0.43  0.01  
+##  1     6       79157.81 79440.69 0.52    0.56     0.70     0.05  0.26  0.01  
+##  1     7       79127.69 79452.48 0.52    0.62     0.67     0.04  0.23  0.01  
+##  1     8       79093.92 79460.61 0.56    0.58     0.72     0.04  0.23  0.01  
+##  1     9       79085.02 79493.62 0.58    0.50     0.73     0.02  0.21  0.02  
+##  2     1       79926.21 79999.54 1.00    1.00     1.00     1.00  1.00        
+##  2     2       79221.45 79373.37 0.60    0.87     0.88     0.49  0.51  0.01  
+##  2     3       79119.17 79349.67 0.55    0.68     0.83     0.28  0.41  0.01  
+##  2     4       79007.67 79316.75 0.57    0.60     0.82     0.13  0.38  0.01  
+##  2     5       78978.84 79366.48 0.56    0.60     0.82     0.10  0.30  0.01  
+##  2     6       78934.24 79400.47 0.60    0.59     0.83     0.09  0.32  0.01  
+##  2     7       78967.67 79512.47 0.56    0.58     0.81     0.07  0.22  0.97  
+##  2     8       78963.30 79586.68 0.57    0.55     0.82     0.05  0.24  0.13  
+##  2     9       78949.02 79650.98 0.59    0.55     0.79     0.04  0.23  0.04  
+##  3     1       79157.63 79340.98 1.00    1.00     1.00     1.00  1.00        
+##  3     2       79171.36 79396.62 0.31    0.76     0.77     0.49  0.51  0.08  
+##  3     3       79173.56 79440.72 0.37    0.45     0.79     0.10  0.53  0.12  
+##  3     4       79086.06 79395.13 0.49    0.46     0.85     0.13  0.58  0.01  
+##  3     5       79081.65 79432.62 0.46    0.49     0.79     0.09  0.48  0.06  
+##  3     6       79093.82 79486.70 0.40    0.50     0.69     0.09  0.24  0.58  
+##  3     7       79039.87 79474.67 0.49    0.37     0.72     0.05  0.34  0.01  
+##  3     8       79023.57 79500.27 0.51    0.44     0.73     0.05  0.33  0.02  
+##  3     9       79007.36 79525.97 0.55    0.50     0.75     0.02  0.33  0.01  
+##  6     1       79157.63 79340.98 1.00    1.00     1.00     1.00  1.00        
+##  6     2       78984.37 79356.30 0.42    0.80     0.84     0.49  0.51  0.01  
+##  6     3       78965.36 79525.87 0.50    0.67     0.80     0.19  0.43  0.01  
+##  6     4       78937.22 79686.33 0.50    0.63     0.77     0.18  0.31  0.01  
+##  6     5       78951.32 79889.01 0.52    0.64     0.74     0.13  0.29  0.38  
+##  6     6       78922.42 80048.70 0.61    0.62     0.79     0.11  0.26  0.04  
+##  6     7       78931.88 80246.75 0.62    0.65     0.83     0.09  0.23  0.43  
+##  6     8       78941.80 80445.25 0.64    0.63     0.82     0.08  0.19  0.50  
+##  6     9       78909.55 80601.58 0.66    0.62     0.83     0.06  0.20  0.12
 \end{verbatim}
@@
 \begin{verbatim}
-## Warning: 
-## One or more analyses resulted in warnings!
-Examine these analyses carefully: model_1_class_7, model_1_class_8, model_1_class_9, model_2_class_6, model_2_class_7, model_2_class_8, model_2_class_9, model_3_class_1, model_3_class_2, model_3_class_3, model_3_class_4, model_3_class_5, model_3_class_6, model_3_class_7, model_3_class_8, model_3_class_9, model_4_class_1, model_4_class_2, model_4_class_3, model_4_class_4, model_4_class_5, model_4_class_6, model_4_class_7, model_4_class_8, model_4_class_9, model_5_class_1, model_5_class_2, model_5_class_3, model_5_class_4, model_5_class_5, model_5_class_6, model_5_class_7, model_5_class_8, model_5_class_9, model_6_class_3, model_6_class_4, model_6_class_5, model_6_class_6, model_6_class_7, model_6_class_8, model_6_class_9
+## tidyLPA analysis using mplus: 
+## 
+##  Model Classes AIC      BIC      Entropy prob_min prob_max n_min n_max BLRT_p
+##  1     1       79926.21 79999.54 1.00    1.00     1.00     1.00  1.00        
+##  1     2       79305.15 79420.39 0.60    0.85     0.90     0.41  0.59  0.00  
+##  1     3       79255.92 79413.08 0.53    0.68     0.81     0.15  0.54  0.00  
+##  1     4       79185.44 79384.50 0.59    0.62     0.79     0.05  0.50  0.00  
+##  1     5       79122.99 79363.96 0.64    0.60     0.81     0.05  0.51  0.00  
+##  1     6       79102.55 79385.43 0.66    0.60     0.82     0.03  0.51  0.00  
+##  1     7       79096.03 79420.81 0.62    0.53     0.78     0.01  0.39  0.13  
+##  1     8       79068.33 79435.02 0.69    0.58     0.83     0.01  0.51  0.02  
+##  1     9       79065.91 79474.51 0.63    0.41     0.80     0.01  0.45  0.00  
+##  2     1       79926.21 79999.54 1.00    1.00     1.00     1.00  1.00        
+##  2     2       79220.63 79372.54 0.60    0.87     0.89     0.50  0.50  0.00  
+##  2     3       79100.80 79331.30 0.58    0.77     0.81     0.18  0.45  0.00  
+##  2     4       79000.42 79309.49 0.59    0.66     0.79     0.14  0.38  0.00  
+##  2     5       78946.87 79334.52 0.62    0.59     0.81     0.08  0.36  0.00  
+##  2     6       78926.11 79392.34 0.59    0.60     0.80     0.08  0.24  0.08  
+##  2     7       78902.41 79447.21 0.58    0.62     0.81     0.08  0.25  0.03  
+##  2     8       78900.48 79523.86 0.57    0.57     0.82     0.08  0.21  0.24  
+##  2     9       78884.99 79586.95 0.62    0.56     0.79     0.01  0.24  1.00  
+##  3     1       79157.63 79340.98 1.00    1.00     1.00     1.00  1.00        
+##  3     2       79090.94 79316.20 0.77    0.62     0.98     0.09  0.91  0.00  
+##  3     3       79084.97 79352.14 0.57    0.66     0.89     0.09  0.64  1.00  
+##  3     4       79047.05 79356.12 0.79    0.58     0.97     0.05  0.85  0.00  
+##  3     5       79023.67 79374.65 0.79    0.56     0.95     0.01  0.80  0.03  
+##  3     6       79017.46 79410.34 0.79    0.45     0.96     0.01  0.83  0.67  
+##  3     7       78998.02 79432.82 0.63    0.56     0.79     0.01  0.41  0.67  
+##  3     8       78992.36 79469.07 0.62    0.54     0.82     0.02  0.48  0.33  
+##  3     9       78976.63 79495.24 0.65    0.52     0.81     0.03  0.44  0.67  
+##  4     1       79157.63 79340.98 1.00    1.00     1.00     1.00  1.00        
+##  4     2       78988.37 79250.30 0.38    0.77     0.82     0.48  0.52  0.00  
+##  4     3       78942.17 79282.68 0.47    0.59     0.83     0.13  0.51  0.09  
+##  4     4       78916.22 79335.30 0.51    0.56     0.78     0.11  0.41  0.29  
+##  4     5       78887.30 79384.96 0.64    0.60     0.99     0.01  0.39  0.27  
+##  4     6       78876.11 79452.35 0.58    0.60     0.78     0.03  0.32  0.31  
+##  4     7       78865.12 79519.93 0.55    0.47     0.76     0.07  0.29  0.67  
+##  4     8       78862.98 79596.37 0.64    0.50     0.85     0.04  0.47  1.00  
+##  4     9       78859.93 79671.90 0.65    0.58     0.94     0.01  0.44  0.67  
+##  5     1       79157.63 79340.98 1.00    1.00     1.00     1.00  1.00        
+##  5     2       79090.89 79426.16 0.81    0.71     0.98     0.11  0.89  0.00  
+##  5     3       79057.51 79544.69 0.70    0.62     0.94     0.06  0.78  0.43  
+##  5     4       79024.23 79663.33 0.79    0.59     0.97     0.01  0.81  0.16  
+##  5     5       79001.55 79792.57 0.79    0.65     1.00     0.01  0.75  0.02  
+##  5     6       78960.55 79903.48 0.87    0.53     1.00     0.01  0.84  0.00  
+##  5     7       79005.61 80100.45 0.81    0.52     1.00     0.01  0.75  0.00  
+##  5     8       78958.75 80205.52 0.91    0.65     1.00     0.01  0.82  0.00  
+##  5     9       78911.08 80309.75 0.72    0.50     0.98     0.01  0.51  0.11  
+##  6     1       79157.63 79340.98 1.00    1.00     1.00     1.00  1.00        
+##  6     2       78980.55 79352.48 0.41    0.75     0.86     0.42  0.58  0.00  
+##  6     3       78946.26 79506.78 0.55    0.72     0.85     0.17  0.50  0.25  
+##  6     4       78921.55 79670.65 0.61    0.74     0.81     0.12  0.40  0.67  
+##  6     5       78905.25 79842.94 0.64    0.65     1.00     0.01  0.38  0.00  
+##  6     6       78889.78 80016.05 0.66    0.68     0.96     0.02  0.33  0.33  
+##  6     7       78880.06 80194.92 0.67    0.67     1.00     0.01  0.29  0.00  
+##  6     8       78844.83 80348.28 0.70    0.66     0.99     0.01  0.38  0.50  
+##  6     9       78871.79 80563.83 0.74    0.72     1.00     0.01  0.39  0.60
+\end{verbatim}
+
+\begin{verbatim}
+## tidyLPA analysis using mplus: 
+## 
+##  Model Classes AIC      BIC      Entropy prob_min prob_max n_min n_max BLRT_p
+##  1     1       79926.21 79999.54 1.00    1.00     1.00     1.00  1.00        
+##  1     2       79305.15 79420.39 0.60    0.85     0.90     0.41  0.59  0.00  
+##  1     3       79236.08 79393.23 0.68    0.58     0.90     0.05  0.55  0.00  
+##  1     4       79185.44 79384.50 0.59    0.62     0.79     0.05  0.50  0.00  
+##  1     5       79122.99 79363.96 0.64    0.60     0.81     0.05  0.51  0.00  
+##  1     6       79102.55 79385.43 0.66    0.60     0.83     0.03  0.51  0.00  
+##  1     7       79082.23 79407.02 0.69    0.60     0.83     0.01  0.51  0.00  
+##  1     8       79068.00 79434.69 0.65    0.51     0.78     0.01  0.43  0.00  
+##  1     9       79045.95 79454.55 0.65    0.56     0.79     0.02  0.42  0.00  
+##  2     1       79926.21 79999.54 1.00    1.00     1.00     1.00  1.00        
+##  2     2       79220.63 79372.54 0.60    0.87     0.89     0.50  0.50  0.00  
+##  2     3       79100.80 79331.30 0.58    0.77     0.81     0.18  0.45  0.00  
+##  2     4       79000.42 79309.49 0.59    0.66     0.79     0.14  0.38  0.00  
+##  2     5       78946.87 79334.52 0.62    0.59     0.81     0.08  0.36  0.00  
+##  2     6       78926.11 79392.34 0.59    0.60     0.80     0.08  0.24  0.07  
+##  2     7       78902.41 79447.21 0.58    0.62     0.81     0.08  0.25  0.00  
+##  2     8       78894.18 79517.57 0.62    0.55     0.80     0.03  0.34  0.67  
+##  2     9       78883.07 79585.03 0.62    0.60     0.83     0.04  0.16  1.00  
+##  3     1       79157.63 79340.98 1.00    1.00     1.00     1.00  1.00        
+##  3     2       79090.94 79316.20 0.77    0.62     0.98     0.09  0.91  0.00  
+##  3     3       79070.64 79337.81 0.85    0.63     0.98     0.01  0.90  0.00  
+##  3     4       79045.48 79354.55 0.81    0.61     0.97     0.03  0.85  0.00  
+##  3     5       79013.99 79364.97 0.79    0.62     0.96     0.01  0.80  0.00  
+##  3     6       79006.55 79399.44 0.76    0.50     0.94     0.02  0.76  0.24  
+##  3     7       78996.75 79431.55 0.68    0.61     0.83     0.01  0.43  0.18  
+##  3     8       78973.50 79450.20 0.79    0.45     0.95     0.01  0.78  0.00  
+##  3     9       78976.27 79494.88 0.62    0.54     0.84     0.01  0.41  1.00  
+##  4     1       79157.63 79340.98 1.00    1.00     1.00     1.00  1.00        
+##  4     2       78988.37 79250.30 0.38    0.77     0.82     0.48  0.52  0.00  
+##  4     3       78942.17 79282.68 0.47    0.59     0.83     0.13  0.51  0.05  
+##  4     4       78916.22 79335.30 0.51    0.56     0.78     0.11  0.41  0.38  
+##  4     5       78882.99 79380.65 0.55    0.58     0.99     0.01  0.36  0.27  
+##  4     6       78876.11 79452.35 0.58    0.60     0.78     0.03  0.32  0.38  
+##  4     7       78860.74 79515.56 0.60    0.60     0.97     0.02  0.29  0.43  
+##  4     8       78828.13 79561.52 0.66    0.54     0.97     0.01  0.35  0.33  
+##  4     9       78826.80 79638.77 0.65    0.53     0.97     0.01  0.38  0.00  
+##  5     1       79157.63 79340.98 1.00    1.00     1.00     1.00  1.00        
+##  5     2       79090.89 79426.16 0.81    0.71     0.98     0.11  0.89  0.00  
+##  5     3       79053.92 79541.10 0.68    0.68     0.92     0.13  0.72  0.25  
+##  5     4       79012.99 79652.08 0.93    0.71     0.99     0.01  0.90  0.04  
+##  5     5       78980.40 79771.42 0.78    0.67     1.00     0.01  0.72  0.10  
+##  5     6       78960.55 79903.48 0.87    0.53     1.00     0.01  0.84  0.00  
+##  5     7       78937.28 80032.13 0.88    0.63     1.00     0.01  0.79  0.24  
+##  5     8       78945.36 80192.12 0.75    0.66     1.00     0.01  0.53  0.24  
+##  5     9       78890.04 80288.71 0.77    0.50     1.00     0.01  0.56  0.00  
+##  6     1       79157.63 79340.98 1.00    1.00     1.00     1.00  1.00        
+##  6     2       78980.55 79352.48 0.41    0.75     0.86     0.42  0.58  0.00  
+##  6     3       78943.69 79504.21 0.51    0.64     0.88     0.20  0.58  0.67  
+##  6     4       78910.45 79659.56 0.64    0.73     0.92     0.03  0.49  0.10  
+##  6     5       78890.91 79828.60 0.61    0.56     0.85     0.08  0.47  0.36  
+##  6     6       78837.12 79963.39 0.73    0.74     0.96     0.01  0.47  0.00  
+##  6     7       78854.88 80169.74 0.68    0.58     0.94     0.02  0.42  0.67  
+##  6     8       78843.89 80347.34 0.81    0.72     1.00     0.01  0.58  0.00  
+##  6     9       78871.79 80563.83 0.74    0.72     1.00     0.01  0.39  0.60
 \end{verbatim}
@@
 
-\hypertarget{htmlwidget-31a3e9100620ea87273d}{}
+\includegraphics{paper_files/figure-latex/unnamed-chunk-14-1.pdf}
 
@@
 person-oriented approaches. To this end, presenting multiple models, as in
-Pastor et al. (2007), should be encouraged, rather than presenting one solution.
+Pastor et al.~(2007), should be encouraged, rather than presenting one solution.
 In addition, having multiple analysts review the solutions found is encouraged.
@@
 of, for example, tools to carry out highly-sophisticated Bayesian methods using
-the most cutting-edge tools (Bürkner \& others, 2017, Makowski, Ben-Shachar, Chen, and Lüdecke (2019)). We think
+the most cutting-edge tools (Bürkner \& others, 2017, p. @makowski2019). We think
 that tidyLPA speaks to this, too.
@@
 \emph{Journal of Open Source Software}, \emph{3}(30), 978.
-doi:\href{https://doi.org/10.21105/joss.00978}{10.21105/joss.00978}
-
-\leavevmode\hypertarget{ref-burkner2017}{}%
-Bürkner, P.-C., \& others. (2017). Brms: An r package for bayesian multilevel models using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.
-
-\leavevmode\hypertarget{ref-makowski2019}{}%
-Makowski, D., Ben-Shachar, M. S., Chen, S., \& Lüdecke, D. (2019). Indices of effect existence and significance in the bayesian framework. \emph{Frontiers in Psychology}, \emph{10}, 2767.
-
-\leavevmode\hypertarget{ref-R-tidyLPA}{}%
-Rosenberg, J.
-M., Beymer, P.
-N., Anderson, D.
-J., Lissa, C.
-van, \& Schmidt, J.
-A.
-(2018).
-TidyLPA: An r package to easily carry out latent profile analysis (lpa) using open-source or commercial software.
-\emph{Journal of Open Source Software}, \emph{3}(30), 978.
-doi:\href{https://doi.org/10.21105/joss.00978}{10.21105/joss.00978}
+\url{https://doi.org/10.21105/joss.00978}
 

### 2020-10-12 paper.rmd
update aper

@@
 title             : "Decisions Decisions Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package"
-shorttitle        : "tidyLPA"
+shorttitle        : "LPA Best Practices"
 
@@
 
-```{r analysis-preferences}
+```{r analysis-preferences, echo = FALSE}
 # Seed for random number generation
@@
 library(tidyverse)
+library(simstudy)
+
 knitr::clean_cache()
@@
 
-```{r, include = FALSE, eval = FALSE}
-library(simstudy)
-
+```{r, include = FALSE, eval = TRUE}
 def <- defData(varname = "x1", formula = 65, variance = 30)
@@
 
-```{r, echo = FALSE, eval = FALSE}
-
+```{r, include = FALSE, eval = TRUE}
 ggplot(dx, aes(x = xMix)) +
@@
 
-```{r, echo = FALSE, eval = FALSE}
+```{r, include = FALSE, eval = TRUE}
 dx %>% 
@@
 In this tutorial, we aim to describe one of the most commonly-used--and relevant
-to psychologists--application of the general mixture model, to cases for which
+to psychologists---application of the general mixture model, to cases for which
 all of the variables for which (relatively) homogeneous groups are identified
@@
 The goal of LPA is estimate the parameters for a number of distributions
-(typically multivariate) from a single dataset. Thus, such an approach is
+(typically multivariate) from a single data set. Thus, such an approach is
 model-based, and some descriptions in the literature refer to it as model-based
@@
 
-We wish to note that challenge facing the analyst using LPA is that these
+We note that challenge facing the analyst using LPA is that these
 parameters and the distinct model parameterizations that can be estimated is the
@@
 
-<!-- I think we move this to the discussion in light of best practices -->
-
-A note of caution is warranted about LPA in the context of their characteristics
-and strengths. Bauer (2007) notes that many samples of data can be usefully
-broken down into profiles, and that the addition of profiles will likely be
-suggested for reasons other than the samples coming from more than one
-distribution (i.e., due to non-normality in the variables measured). Bauer also
-cautions that profiles should not be reified; that profiles do not necessarily
-exist outside of the analysis that they should be interpreted more as useful
-interpretative devices. These cautions suggest that, in general, parsimony,
-interpretability, and a general sense that the profiles are not necessarily
-real, but are rather helpful analytic tools, should be both priorities for the
-analyst and the reader of studies using this approach.
-
 # Specific Software Tools
 
-## Other Tools
+## Tools other than tidyLPA
 
@@
 
-You can install tidyLPA from CRAN with:
+You can install tidyLPA from the Comprehensive R Archive Network, or CRAN, with:
 
@@
 
-You can also install the development version of tidyLPA from GitHub with:
+This is how we recommend most users install tidyLPA.
+
+You can also install the development version of tidyLPA, which may have new features that are not yet availalble on CRAN (because the version on CRAN is periodically updated) from GitHub with:
 
@@
 library(tidyLPA)
-library(dplyr)
+library(dplyr) # for preparing the data
 ```
@@
 
-Here, we use a simulated dataset, based on Curry et al., 2019, contains data on moral relevance and judgment across the seven domains of the *Morality As Cooperation* scale.
+Here, we use a simulated dataset contains data on moral relevance and judgment across the seven domains of the *Morality As Cooperation* scale. More on this scale can be found in Curry et al., (2019).
+
+Add one or two sentences here on the dataset?
 
@@
 
-A simple summary of the analysis is printed to the console. 
-The resulting object can be further passed down a pipeline to
-other functions, such as `compare_solutions`, `get_data`, and `get_fit`.
+A simple summary of the analysis is printed to the console. The resulting object
+can be further passed down a pipeline to other functions, such as
+`compare_solutions`, `get_data`, and `get_fit`.
 
-Note that here and throughout this paper we use the function `knitr::kable()` to format the output in an RMarkdown document, like the one we are using; this is not necessary to use unless you also wish to format the tidyLPA output in an RMarkdown document.
+Note that here and throughout this paper we use the function `knitr::kable()` to
+format the output in an RMarkdown document, like the one we are using; this is
+not necessary to use unless you also wish to format the tidyLPA output in an
+RMarkdown document.
 
@@
 We can see a number of statistics for each of the models that were estimated, including the 
-AIC (Aikake information criterion; based on -2 log-likelihood, and penalized by number of parameters) and BIC (BIC: Bayesian information criterion; based on -2 log-likelihood, and penalized by number of parameters adjusted by sample size).
+AIC (Aikake information criterion; based on -2 * the log-likelihood, and penalized by number of parameters) and BIC (BIC: Bayesian information criterion; based on -2 log-likelihood, and penalized by number of parameters adjusted by sample size).
 Both are _penalized_ likelihood statistics; the log-likelihood will always decrease with the addition of additional
@@
 ```{r, get-fit-m1}
-get_fit(m1)
+get_fit(m1) %>% 
+  knitr::kable()
 ```
@@
 Here is an example of specifying a model with varying variances and covariances
-(Model 6; not run here):
+(Model 6)
 
-``` {r, curry-mac-m6, eval = F}
+``` {r, curry-mac-m6, eval = TRUE}
 curry_mac %>% 
@@
 
-````{r, compare-mclust, eval= FALSE}
+````{r, compare-mclust, eval= TRUE}
 compare_solutions_mclust <- curry_mac %>% 
@@
                     n_profiles = 1:9,
-                    package = "mclust")
+                    package = "mclust") %>% 
+  knitr::kable()
 ```
@@
                     n_profiles = 1:9,
-                    package = "mplus")
+                    package = "mplus") %>% 
+  knitr::kable()
 ```
@@
 
-```{r, compare-mplus-starts, eval = FALSE}
+```{r, compare-mplus-starts, eval = TRUE}
 compare_solutions_mplus_starts <- curry_mac %>% 
@@
                     package = "mplus",
-                    ANALYSIS = "starts = 100, 20;")
+                    ANALYSIS = "starts = 100, 20;") %>% 
+  knitr::kable()
 ```
@@
 
+A note of caution is warranted about LPA in the context of their characteristics
+and strengths. Bauer (2007) notes that many samples of data can be usefully
+broken down into profiles, and that the addition of profiles will likely be
+suggested for reasons other than the samples coming from more than one
+distribution (i.e., due to non-normality in the variables measured). Bauer also
+cautions that profiles should not be reified; that profiles do not necessarily
+exist outside of the analysis that they should be interpreted more as useful
+interpretative devices. These cautions suggest that, in general, parsimony,
+interpretability, and a general sense that the profiles are not necessarily
+real, but are rather helpful analytic tools, should be both priorities for the
+analyst and the reader of studies using this approach.
+
+The GRoLTs checklist is a helpful way to think about how best practices for LPA;
+here, we show how to implement these using tidyLPA. First, here is the checklist:
+
 ## GRoLTs checklist
@@
 ```{r, echo = FALSE}
-t <- read_csv("GRoLTS-applicability-to-LPA.csv")
+t <- readr::read_csv("GRoLTS-applicability-to-LPA.csv")
 

### 2020-12-26 paper.rmd
update manuscript

@@
 knitr::clean_cache()
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed, warning = FALSE)
+knitr::opts_chunk$set(cache.extra = knitr::rand_seed, warning = FALSE, cache = TRUE)
 ```
@@
 
-# Specific Software Tools
+# Best Practices for Latent Profile Analysis
+
+# tidyLPA
 
@@
 
-## tidyLPA
+## The tidyLPA R package
 
@@
 
-Here, we use a simulated dataset contains data on moral relevance and judgment across the seven domains of the *Morality As Cooperation* scale. More on this scale can be found in Curry et al., (2019).
+Here, we use a simulated dataset contains data on broad interest, enjoyment, and self-efficacy from the PISA assessment.
+
+First, let's explore the data graphically using the {ggplot2} package:
+
+```{r}
+pisaUSA15 %>%
+  select(broad_interest, enjoyment, self_efficacy) %>% 
+  tidyr::gather(key, val) %>% 
+  ggplot(aes(x = val, group = key, fill = key)) + 
+  geom_density() + 
+  facet_wrap(~key) + 
+  theme_bw()
+```
+
+Here, we use the {psych} package to explore the distributional properties of the variables:
+
+```{r}
+pisaUSA15 %>%
+  select(broad_interest, enjoyment, self_efficacy) %>% 
+  psych::describe()
+```
 
@@
 ``` {r, curry-mac-mclust-example}
-m1 <- curry_mac %>% 
-  select(KinshipR:PossessionR, 
-         KinshipJ:PossessionJ) %>% 
-  estimate_profiles(1:10) %>% 
-  knitr::kable()
+pisaUSA15 %>%
+  select(broad_interest, enjoyment, self_efficacy) %>%
+  single_imputation() %>%
+  estimate_profiles(3)
 ```
+
+### Mplus via tidyLPA
+
+We can use Mplus simply by changing the package argument for
+`estimate_profiles()` to "mplus" or "MplusAutomation" (either will work; not run here):
+
+``` {r, curry-mac-mplus, eval = FALSE}
+pisaUSA15 %>%
+  select(broad_interest, enjoyment, self_efficacy) %>%
+  single_imputation() %>%
+  estimate_profiles(3, package = "mplus")
+```
+
+These are the very basics; we will return to what we do with this output later in the tutorial.
+
+### Fit
 
@@
 ```{r, get-fit-m1}
+pisaUSA15 %>%
+  select(broad_interest, enjoyment, self_efficacy) %>%
+  single_imputation() %>%
+  estimate_profiles(3) %>% 
   get_fit(m1) %>% 
@@
 
-### Mplus via tidyLPA
+## Comparing a wide range of solutions
 
-We can use Mplus simply by changing the package argument for
-`estimate_profiles()` to "mplus" or "MplusAutomation" (either will work; not run here):
+The function `compare_solutions()` compares the fit of several estimated models,
+with varying numbers of profiles and model specifications:
 
-``` {r, curry-mac-mplus, eval = FALSE}
-curry_mac %>% 
-  select(KinshipR:PossessionR, 
-         KinshipJ:PossessionJ) %>% 
-  estimate_profiles(1:10, package = "mplus")
+````{r, compare-mclust, eval= TRUE}
+compare_solutions_results <- pisaUSA15 %>%
+  select(broad_interest, enjoyment, self_efficacy) %>%
+  single_imputation() %>%
+  estimate_profiles(models = c(1, 2, 3, 6),
+                    n_profiles = 1:9)
+
+compare_solutions_results
+
+compare_solutions(compare_solutions_results)
 ```
-
-These are the very basics; we will return to what we do with this output later in the tutorial.
 
@@
 
-Here is an example of specifying a model with varying variances and covariances
-(Model 6)
-
-``` {r, curry-mac-m6, eval = TRUE}
-curry_mac %>% 
-  select(KinshipR:PossessionR, 
-         KinshipJ:PossessionJ) %>% 
-  estimate_profiles(3, 
-                    variances = "varying",
-                    covariances = "varying")
-```
-
-## Comparing a wide range of solutions
-
-The function `compare_solutions()` compares the fit of several estimated models,
-with varying numbers of profiles and model specifications:
-
-````{r, compare-mclust, eval= TRUE}
-compare_solutions_mclust <- curry_mac %>% 
-  select(KinshipJ:PossessionJ) %>% 
-  estimate_profiles(models = c(1, 2, 3, 6),
-                    n_profiles = 1:9,
-                    package = "mclust") %>% 
-  knitr::kable()
-```
-
-```{r, eval = FALSE, echo = FALSE}
-write_rds(compare_solutions_mclust, "compare_solutions_mclust.rds")
-```
-
-```{r, echo = FALSE}
-compare_solutions_mclust <- read_rds("compare_solutions_mclust.rds")
-compare_solutions_mclust
-```
-
-```{r, compare-mplus, eval = FALSE}
-compare_solutions_mplus <- curry_mac %>% 
-  select(KinshipJ:PossessionJ) %>% 
-  estimate_profiles(models = c(1, 2, 3, 4, 5, 6),
-                    n_profiles = 1:9,
-                    package = "mplus") %>% 
-  knitr::kable()
-```
-
-```{r, echo = FALSE, eval = FALSE}
-write_rds(compare_solutions_mplus, "compare_solutions_mplus.rds")
-```
-
-```{r, echo = FALSE}
-compare_solutions_mplus <- read_rds("compare_solutions_mplus.rds")
-compare_solutions_mplus
-```
-
-```{r, compare-mplus-starts, eval = TRUE}
-compare_solutions_mplus_starts <- curry_mac %>% 
-  select(KinshipJ:PossessionJ) %>% 
-  estimate_profiles(models = c(1, 2, 3, 4, 5, 6),
-                    n_profiles = 1:9,
-                    package = "mplus",
-                    ANALYSIS = "starts = 100, 20;") %>% 
-  knitr::kable()
-```
-
-```{r, echo = FALSE, eval = FALSE}
-write_rds(compare_solutions_mplus_starts, "compare_solutions_mplus_starts.rds")
-```
-
-```{r, echo = FALSE}
-compare_solutions_mplus <- read_rds("compare_solutions_mplus_starts.rds")
-compare_solutions_mplus
-```
-
-The results of this analysis suggests that the Model 6, two or seven class models are best.
-
 ## Accessing values in the output
@@
 
-``` {r, eval = FALSE}
-get_data()
+``` {r, get-data-for-solution, eval = FALSE}
+get_data(compare_solutions_mclust[[32]])
 ```
@@
 
-``` {r, eval = FALSE}
-get_data(m) %>% 
+``` {r, process-data, eval = FALSE}
+get_data(compare_solutions_mclust[[32]]) %>% 
   tidyr::gather(Class_prob, Probability, contains("CPROB"))
@@
 
-### Is information reported about the number of random start values and final iterations included?
-
-### Are the model comparison (and selection) tools described from a statistical perspective?
-
-### Are the total number of fitted models reported, including a one-class solution?
-
-### Are the number of cases per class reported for each model (absolute sample size, or proportion)?
-
-### If classification of cases in a trajectory is the goal, is entropy reported?
-
-### Are all of the estimates plotted?
-
-### Are the raw data plotted along with the estimates?
-
-### Are characteristics of the final class solution numerically described (i.e., means, SD/SE, n, CI, etc.)?
-
-### Are the syntax files available (either in the appendix, supplementary materials, or from the authors)?
+In LPA, we can determine which parameters - variance
+and covariance - are estimated to be the same or different across all profiles,
+and, in the case of covariance, whether it is fixed to zero. Of course, for the
+profiles we are most interested in, the mean is allowed to vary across profiles.
+The models with more parameters freely-estimated use more degrees of freedom;
+thus, similar to the number of profiles selected, the balance between adding
+more parameters (and having a more complex model) and interpretability and
+parsimony must be balanced. If a more complex model fits better and is
+interpretable and is justifiable in terms of the data collected, then it may be
+preferred to simpler models. If the two models are similar in terms of their
+fit, then the more simple, parsimonious model should be selected.
 
@@
 
-## How the Data are Centering or Scaled
+### Is information reported about the number of random start values and final iterations included?
 
-The data can be transformed by centering or scaling. Typically, these are done
-after the profiles are created, so that differences between profiles can be more
-easily explored. They can also be done prior to the analysis, which can be
-helpful for obtaining solutions when the variables are on very different
-scales.
-
-## How to Choose the Number of Profiles/Classes
+### Are the model comparison (and selection) tools described from a statistical perspective?
 
@@
 
-## How to Choose Options for the Clustering Procedure or to Select the Model Parameterization
+### Are the total number of fitted models reported, including a one-class solution?
 
-In LPA, we can determine which parameters - variance
-and covariance - are estimated to be the same or different across all profiles,
-and, in the case of covariance, whether it is fixed to zero. Of course, for the
-profiles we are most interested in, the mean is allowed to vary across profiles.
-The models with more parameters freely-estimated use more degrees of freedom;
-thus, similar to the number of profiles selected, the balance between adding
-more parameters (and having a more complex model) and interpretability and
-parsimony must be balanced. If a more complex model fits better and is
-interpretable and is justifiable in terms of the data collected, then it may be
-preferred to simpler models. If the two models are similar in terms of their
-fit, then the more simple, parsimonious model should be selected.
+### Are the number of cases per class reported for each model (absolute sample size, or proportion)?
 
-## More General Recommendations
+### If classification of cases in a trajectory is the goal, is entropy reported?
 
-This section highlights three general recommendations. First, scholars taking a
+### Are all of the estimates plotted?
+
+### Are the raw data plotted along with the estimates?
+
+### Are characteristics of the final class solution numerically described (i.e., means, SD/SE, n, CI, etc.)?
+
+### Are the syntax files available (either in the appendix, supplementary materials, or from the authors)?
+
+# Future Directions and Extensions
+
+First, scholars taking a
 person-oriented approach should emphasize reproducibility in carrying out

### 2020-12-26 paper.tex
update manuscript

--- 2020-08-27 paper.tex
@@
 
-\shorttitle{tidyLPA}
+\shorttitle{LPA Best Practices}
 \affiliation{
@@
 \begin{verbatim}
+## Warning: package 'ggplot2' was built under R version 3.6.2
+\end{verbatim}
+
+\begin{verbatim}
 ## Warning: package 'tibble' was built under R version 3.6.2
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: package 'tidyr' was built under R version 3.6.2
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: package 'readr' was built under R version 3.6.2
 \end{verbatim}
@@
 ## Warning: package 'dplyr' was built under R version 3.6.2
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: package 'simstudy' was built under R version 3.6.2
 \end{verbatim}
@@
 
-A note of caution is warranted about LPA in the context of their characteristics
-and strengths. Bauer (2007) notes that many samples of data can be usefully
-broken down into profiles, and that the addition of profiles will likely be
-suggested for reasons other than the samples coming from more than one
-distribution (i.e., due to non-normality in the variables measured). Bauer also
-cautions that profiles should not be reified; that profiles do not necessarily
-exist outside of the analysis that they should be interpreted more as useful
-interpretative devices. These cautions suggest that, in general, parsimony,
-interpretability, and a general sense that the profiles are not necessarily
-real, but are rather helpful analytic tools, should be both priorities for the
-analyst and the reader of studies using this approach.
-
 \hypertarget{specific-software-tools}{%
@@
 
-\hypertarget{other-tools}{%
-\subsection{Other Tools}\label{other-tools}}
+\hypertarget{tools-other-than-tidylpa}{%
+\subsection{Tools other than tidyLPA}\label{tools-other-than-tidylpa}}
 
@@
 
-You can install tidyLPA from CRAN with:
+You can install tidyLPA from the Comprehensive R Archive Network, or CRAN, with:
 
-You can also install the development version of tidyLPA from GitHub with:
+This is how we recommend most users install tidyLPA.
 
-\begin{verbatim}
-## Warning: package 'tidyLPA' was built under R version 3.6.2
-\end{verbatim}
+You can also install the development version of tidyLPA, which may have new features that are not yet availalble on CRAN (because the version on CRAN is periodically updated) from GitHub with:
 
@@
 
-Here, we use a simulated dataset, based on Curry et al., 2019, contains data on moral relevance and judgment across the seven domains of the \emph{Morality As Cooperation} scale.
+Here, we use a simulated dataset contains data on moral relevance and judgment across the seven domains of the \emph{Morality As Cooperation} scale.
+More on this scale can be found in Curry et al., (2019).
+
+Add one or two sentences here on the dataset?
 
@@
 
-A simple summary of the analysis is printed to the console.
-The resulting object can be further passed down a pipeline to
-other functions, such as \texttt{compare\_solutions}, \texttt{get\_data}, and \texttt{get\_fit}.
+A simple summary of the analysis is printed to the console. The resulting object
+can be further passed down a pipeline to other functions, such as
+\texttt{compare\_solutions}, \texttt{get\_data}, and \texttt{get\_fit}.
+
+Note that here and throughout this paper we use the function \texttt{knitr::kable()} to
+format the output in an RMarkdown document, like the one we are using; this is
+not necessary to use unless you also wish to format the tidyLPA output in an
+RMarkdown document.
 
 We can see a number of statistics for each of the models that were estimated, including the
-AIC (Aikake information criterion; based on -2 log-likelihood, and penalized by number of parameters) and BIC (BIC: Bayesian information criterion; based on -2 log-likelihood, and penalized by number of parameters adjusted by sample size).
+AIC (Aikake information criterion; based on -2 * the log-likelihood, and penalized by number of parameters) and BIC (BIC: Bayesian information criterion; based on -2 log-likelihood, and penalized by number of parameters adjusted by sample size).
 Both are \emph{penalized} likelihood statistics; the log-likelihood will always decrease with the addition of additional
@@
 
-\begin{verbatim}
-## # A tibble: 10 x 18
-##    Model Classes  LogLik    AIC    AWE    BIC   CAIC    CLC    KIC  SABIC
-##    <dbl>   <dbl>   <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>
-##  1     1       1 -81245. 1.63e5 1.63e5 1.63e5 1.63e5 1.62e5 1.63e5 1.63e5
-##  2     1       2 -79727. 1.60e5 1.60e5 1.60e5 1.60e5 1.59e5 1.60e5 1.60e5
-##  3     1       3 -79239. 1.59e5 1.59e5 1.59e5 1.59e5 1.58e5 1.59e5 1.59e5
-##  4     1       4 -79091. 1.58e5 1.59e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
-##  5     1       5 -79029. 1.58e5 1.60e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
-##  6     1       6 -78971. 1.58e5 1.60e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
-##  7     1       7 -78930. 1.58e5 1.60e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
-##  8     1       8 -78852. 1.58e5 1.60e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
-##  9     1       9 -78822. 1.58e5 1.60e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
-## 10     1      10 -78751. 1.58e5 1.60e5 1.59e5 1.59e5 1.58e5 1.58e5 1.58e5
-## # ... with 8 more variables: ICL <dbl>, Entropy <dbl>, prob_min <dbl>,
-## #   prob_max <dbl>, n_min <dbl>, n_max <dbl>, BLRT_val <dbl>, BLRT_p <dbl>
-\end{verbatim}
+\begin{tabular}{r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r}
+\hline
+Model & Classes & LogLik & AIC & AWE & BIC & CAIC & CLC & KIC & SABIC & ICL & Entropy & prob\_min & prob\_max & n\_min & n\_max & BLRT\_val & BLRT\_p\\
+\hline
+1 & 1 & -81245.00 & 162546.0 & 162977.4 & 162692.7 & 162720.7 & 162492.0 & 162577.0 & 162603.7 & -162692.7 & 1.0000000 & 1.0000000 & 1.0000000 & 1.0000000 & 1.0000000 & NA & NA\\
+\hline
+1 & 2 & -79727.50 & 159541.0 & 160204.8 & 159766.3 & 159809.3 & 159456.7 & 159587.0 & 159629.7 & -159925.3 & 0.8369494 & 0.9296991 & 0.9637189 & 0.3441092 & 0.6558908 & 3035.00346 & 0.009901\\
+\hline
+1 & 3 & -79239.27 & 158594.5 & 159490.6 & 158898.4 & 158956.4 & 158480.2 & 158655.5 & 158714.1 & -159175.7 & 0.8188766 & 0.9053581 & 0.9222054 & 0.1465517 & 0.5272989 & 976.44661 & 0.009901\\
+\hline
+1 & 4 & -79090.59 & 158327.2 & 159455.5 & 158709.6 & 158782.6 & 158182.7 & 158403.2 & 158477.7 & -159150.2 & 0.7763719 & 0.8393595 & 0.9127306 & 0.0732759 & 0.4655172 & 297.35974 & 0.009901\\
+\hline
+1 & 5 & -79029.50 & 158235.0 & 159595.6 & 158696.0 & 158784.0 & 158060.4 & 158326.0 & 158416.4 & -159398.5 & 0.7012112 & 0.6979575 & 0.8987374 & 0.0675287 & 0.2406609 & 122.19715 & 0.009901\\
+\hline
+1 & 6 & -78971.30 & 158148.6 & 159741.3 & 158688.2 & 158791.2 & 157944.0 & 158254.6 & 158361.0 & -159456.7 & 0.6954594 & 0.6929238 & 0.9230731 & 0.0603448 & 0.3060345 & 116.39215 & 0.009901\\
+\hline
+1 & 7 & -78930.08 & 158096.2 & 159921.0 & 158714.3 & 158832.3 & 157861.5 & 158217.2 & 158339.5 & -159550.6 & 0.6933730 & 0.5964096 & 0.9236077 & 0.0610632 & 0.2550287 & 82.44963 & 0.009901\\
+\hline
+1 & 8 & -78851.80 & 157969.6 & 160026.5 & 158666.3 & 158799.3 & 157705.1 & 158105.6 & 158243.8 & -159391.4 & 0.7477064 & 0.6432919 & 0.9219048 & 0.0423851 & 0.2902299 & 156.55019 & 0.009901\\
+\hline
+1 & 9 & -78821.74 & 157939.5 & 160228.6 & 158714.8 & 158862.8 & 157645.0 & 158090.5 & 158244.6 & -159455.8 & 0.7515427 & 0.6071032 & 0.9456474 & 0.0402299 & 0.2938218 & 60.12578 & 0.009901\\
+\hline
+1 & 10 & -78751.27 & 157828.5 & 160349.8 & 158682.4 & 158845.4 & 157504.1 & 157994.5 & 158164.6 & -159420.9 & 0.7625067 & 0.6463008 & 0.9396876 & 0.0337644 & 0.2909483 & 140.92881 & 0.009901\\
+\hline
+\end{tabular}
 
@@
 Here is an example of specifying a model with varying variances and covariances
-(Model 6; not run here):
+(Model 6)
+
+\begin{verbatim}
+## tidyLPA analysis using mclust: 
+## 
+##  Model Classes AIC       BIC       Entropy prob_min prob_max n_min n_max BLRT_p
+##  6     3       156872.22 158752.84 0.67    0.82     0.87     0.23  0.41  0.01
+\end{verbatim}
 
@@
 \begin{verbatim}
-## Warning: 
-## One or more analyses resulted in warnings!
-Examine these analyses carefully: model_1_class_6, model_1_class_7, model_1_class_8, model_1_class_9, model_2_class_5, model_2_class_6, model_2_class_7, model_2_class_8, model_2_class_9, model_3_class_1, model_3_class_2, model_3_class_3, model_3_class_4, model_3_class_5, model_3_class_6, model_3_class_7, model_3_class_8, model_3_class_9, model_4_class_1, model_4_class_2, model_4_class_3, model_4_class_4, model_4_class_5, model_4_class_6, model_4_class_7, model_4_class_8, model_4_class_9, model_5_class_1, model_5_class_2, model_5_class_3, model_5_class_4, model_5_class_5, model_5_class_6, model_5_class_7, model_5_class_8, model_5_class_9, model_6_class_3, model_6_class_4, model_6_class_5, model_6_class_6, model_6_class_7, model_6_class_8, model_6_class_9
+## tidyLPA analysis using mclust: 
+## 
+##  Model Classes AIC       BIC       Entropy prob_min prob_max n_min n_max BLRT_p
+##  1     1       139226.33 139352.05 1.00    1.00     1.00     1.00  1.00        
+##  1     2       136511.15 136704.97 0.83    0.92     0.97     0.34  0.66  0.01  
+##  1     3       135651.76 135913.69 0.81    0.90     0.92     0.14  0.53  0.01  
+##  1     4       135391.00 135721.03 0.77    0.83     0.91     0.07  0.47  0.01  
+##  1     5       135326.12 135724.24 0.72    0.57     0.91     0.07  0.39  0.01  
+##  1     6       135281.64 135747.86 0.68    0.56     0.92     0.07  0.28  0.01  
+##  1     7       135218.07 135752.39 0.67    0.59     0.91     0.07  0.29  0.01  
+##  1     8       135135.62 135738.04 0.70    0.59     0.92     0.05  0.31  0.01  
+##  1     9       135098.58 135769.10 0.70    0.57     0.93     0.05  0.30  0.01  
+##  2     1       139226.33 139352.05 1.00    1.00     1.00     1.00  1.00        
+##  2     2       136245.76 136502.44 0.81    0.92     0.96     0.39  0.61  0.01  
+##  2     3       135337.03 135724.68 0.80    0.91     0.91     0.25  0.49  0.01  
+##  2     4       135103.11 135621.72 0.75    0.82     0.92     0.17  0.37  0.01  
+##  2     5       134840.49 135490.07 0.75    0.76     0.89     0.13  0.31  0.01  
+##  2     6       134688.81 135469.35 0.75    0.71     0.89     0.06  0.29  0.01  
+##  2     7       134579.47 135490.97 0.75    0.64     0.90     0.06  0.23  0.01  
+##  2     8       134482.30 135524.76 0.75    0.71     0.90     0.06  0.18  0.01  
+##  2     9       134434.76 135608.19 0.75    0.65     0.91     0.06  0.17  0.01  
+##  3     1       134912.97 135384.44 1.00    1.00     1.00     1.00  1.00        
+##  3     2       134836.15 135375.72 0.69    0.78     0.96     0.22  0.78  0.01  
+##  3     3       134862.13 135469.80 0.41    0.16     0.88     0.07  0.69  0.76  
+##  3     4       134728.94 135404.71 0.50    0.31     0.83     0.10  0.63  0.01  
+##  3     5       134615.77 135359.63 0.58    0.43     0.79     0.07  0.56  0.01  
+##  3     6       134696.35 135508.32 0.52    0.43     0.81     0.06  0.39  0.71  
+##  3     7       134589.15 135469.21 0.54    0.36     0.84     0.06  0.37  0.01  
+##  3     8       134604.12 135552.29 0.54    0.40     0.84     0.04  0.33  0.77  
+##  3     9       134568.05 135584.32 0.56    0.31     0.83     0.03  0.34  0.01  
+##  6     1       134912.97 135384.44 1.00    1.00     1.00     1.00  1.00        
+##  6     2       134223.62 135171.78 0.62    0.87     0.90     0.48  0.52  0.01  
+##  6     3       134151.67 135576.55 0.68    0.80     0.88     0.18  0.47  0.01  
+##  6     4       134089.83 135991.41 0.68    0.77     0.87     0.17  0.36  0.01  
+##  6     5       134092.59 136470.87 0.71    0.77     0.89     0.06  0.33  0.34  
+##  6     6       134088.85 136943.83 0.72    0.75     0.90     0.06  0.29  0.43  
+##  6     7       134021.25 137352.94 0.74    0.77     0.88     0.09  0.20  0.01  
+##  6     8       134038.51 137846.90 0.77    0.79     0.98     0.03  0.19  0.58  
+##  6     9       134067.02 138352.11 0.78    0.77     0.96     0.04  0.18  0.78
 \end{verbatim}
@@
 \begin{verbatim}
-## Warning: 
-## One or more analyses resulted in warnings!
-Examine these analyses carefully: model_1_class_7, model_1_class_8, model_1_class_9, model_2_class_6, model_2_class_7, model_2_class_8, model_2_class_9, model_3_class_1, model_3_class_2, model_3_class_3, model_3_class_4, model_3_class_5, model_3_class_6, model_3_class_7, model_3_class_8, model_3_class_9, model_4_class_1, model_4_class_2, model_4_class_3, model_4_class_4, model_4_class_5, model_4_class_6, model_4_class_7, model_4_class_8, model_4_class_9, model_5_class_1, model_5_class_2, model_5_class_3, model_5_class_4, model_5_class_5, model_5_class_6, model_5_class_7, model_5_class_8, model_5_class_9, model_6_class_3, model_6_class_4, model_6_class_5, model_6_class_6, model_6_class_7, model_6_class_8, model_6_class_9
+## Compare tidyLPA solutions:
+## 
+##  Model Classes BIC       
+##  1     1       139352.049
+##  1     2       136704.970
+##  1     3       135913.687
+##  1     4       135721.026
+##  1     5       135724.243
+##  1     6       135747.862
+##  1     7       135752.395
+##  1     8       135738.044
+##  1     9       135769.104
+##  2     1       139352.049
+##  2     2       136502.441
+##  2     3       135724.680
+##  2     4       135621.724
+##  2     5       135490.065
+##  2     6       135469.349
+##  2     7       135490.971
+##  2     8       135524.760
+##  2     9       135608.188
+##  3     1       135384.436
+##  3     2       135375.718
+##  3     3       135469.796
+##  3     4       135404.708
+##  3     5       135359.634
+##  3     6       135508.318
+##  3     7       135469.214
+##  3     8       135552.290
+##  3     9       135584.321
+##  6     1       135384.436
+##  6     2       135171.784
+##  6     3       135576.546
+##  6     4       135991.409
+##  6     5       136470.871
+##  6     6       136943.826
+##  6     7       137352.936
+##  6     8       137846.896
+##  6     9       138352.109
+## 
+## Best model according to BIC is Model 6 with 2 classes.
+## 
+## An analytic hierarchy process, based on the fit indices AIC, AWE, BIC, CLC, and KIC (Akogul & Erisoglu, 2017), suggests the best solution is Model 6 with 2 classes.
 \end{verbatim}
 
-The results of this analysis suggests that the Model 6, two or seven class models are best.
+\includegraphics{paper_files/figure-latex/plot-one-solution-1.pdf}
 
@@
 
+A note of caution is warranted about LPA in the context of their characteristics
+and strengths. Bauer (2007) notes that many samples of data can be usefully
+broken down into profiles, and that the addition of profiles will likely be
+suggested for reasons other than the samples coming from more than one
+distribution (i.e., due to non-normality in the variables measured). Bauer also
+cautions that profiles should not be reified; that profiles do not necessarily
+exist outside of the analysis that they should be interpreted more as useful
+interpretative devices. These cautions suggest that, in general, parsimony,
+interpretability, and a general sense that the profiles are not necessarily
+real, but are rather helpful analytic tools, should be both priorities for the
+analyst and the reader of studies using this approach.
+
+The GRoLTs checklist is a helpful way to think about how best practices for LPA;
+here, we show how to implement these using tidyLPA. First, here is the checklist:
+
 \hypertarget{grolts-checklist}{%
@@
 
-\hypertarget{htmlwidget-31a3e9100620ea87273d}{}
+\hypertarget{htmlwidget-675d126ab2afaf44720a}{}
 

### 2020-12-28 paper.rmd
update

@@
 library(papaja)
-```
-
-```{r analysis-preferences, echo = FALSE}
 # Seed for random number generation
@@
 
+## Recommendations Regarding Analytic Choices
+
+A note of caution is warranted about LPA in the context of their characteristics
+and strengths. Bauer (2007) notes that many samples of data can be usefully
+broken down into profiles, and that the addition of profiles will likely be
+suggested for reasons other than the samples coming from more than one
+distribution (i.e., due to non-normality in the variables measured). Bauer also
+cautions that profiles should not be reified; that profiles do not necessarily
+exist outside of the analysis that they should be interpreted more as useful
+interpretative devices. These cautions suggest that, in general, parsimony,
+interpretability, and a general sense that the profiles are not necessarily
+real, but are rather helpful analytic tools, should be both priorities for the
+analyst and the reader of studies using this approach.
+
+The GRoLTs checklist is a helpful way to think about how best practices for LPA;
+here, we show how to implement these using tidyLPA. First, here is the checklist:
+
+## GRoLTs checklist
+
+```{r, echo = FALSE}
+t <- readr::read_csv("GRoLTS-applicability-to-LPA.csv")
+
+tp <- t %>% 
+  filter(use == 1) %>% 
+  mutate(`LPA Number` = 1:nrow(.)) %>% 
+  select(`GRoLTS Number` = Number, `LPA Number`, `Checklist Item`)
+
+tp %>% 
+  DT::datatable()
+```
+
+We walk through each of these in turn.
+
+### Is the missing data mechanism reported?
+
+### Is a description provided of what variables are related to attrition/missing data?
+
+### Is a description provided of how missing data in the analyses were dealt with?
+
+### Is information about the distribution of the observed variables included?
+
+### Is the software mentioned?
+
+### Are alternative specifications of the between-class differences in variance–covariance matrix structure considered and clearly documented?
+
+In LPA, we can determine which parameters - variance
+and covariance - are estimated to be the same or different across all profiles,
+and, in the case of covariance, whether it is fixed to zero. Of course, for the
+profiles we are most interested in, the mean is allowed to vary across profiles.
+The models with more parameters freely-estimated use more degrees of freedom;
+thus, similar to the number of profiles selected, the balance between adding
+more parameters (and having a more complex model) and interpretability and
+parsimony must be balanced. If a more complex model fits better and is
+interpretable and is justifiable in terms of the data collected, then it may be
+preferred to simpler models. If the two models are similar in terms of their
+fit, then the more simple, parsimonious model should be selected.
+
+There are a number of analytic choices that need to be made when carrying out
+person-oriented analyses. Because such person-oriented approaches are often more
+subjective (in practice) than other approaches (Linnenbrink-Garcia and
+Wormington, 2017), there is no one rule for determining the solution obtained.
+This solution is obtained on the basis of multiple decisions, such as the number
+of profiles selected or the modeling decisions such as what specific options are
+used for the cluster analysis (i.e., the distance metric used to calculate the
+similarity of the observations as part of the Ward’s hierarchical clustering) or
+what parameters are estimated and how as part of LPA.
+
+Given the subjectivity involved, it is important that researchers be transparent
+and work as part of a team to obtain clustering solutions. Transparency about
+the design and analytic choices is important so that readers can appropriately
+interpret the report. Researchers can enhance transparency and reproducibility
+by sharing detailed descriptions of methodology and document it through the use
+of syntax (and, if possible, data) that we share with others. Working as part of
+a team can help to serve as a check on several of the choices researchers make,
+such as over-fitting or under-fitting the model to the data. Each decision
+depends on multiple factors and balancing tensions. We discuss each of the key
+decisions listed in an analysis.
+
+### Is information reported about the number of random start values and final iterations included?
+
+### Are the model comparison (and selection) tools described from a statistical perspective?
+
+In the case of choosing the number of profiles (and the specification of the
+model / profile solution), multiple criteria, such as the BIC or the proportion
+of variance explained are recommended for decision-making, but also
+interpretability in light of theory, parsimony, and evidence from
+cross-validation should be considered.
+
+### Are the total number of fitted models reported, including a one-class solution?
+
+### Are the number of cases per class reported for each model (absolute sample size, or proportion)?
+
+### If classification of cases in a trajectory is the goal, is entropy reported?
+
+### Are all of the estimates plotted?
+
+### Are the raw data plotted along with the estimates?
+
+### Are characteristics of the final class solution numerically described (i.e., means, SD/SE, n, CI, etc.)?
+
+### Are the syntax files available (either in the appendix, supplementary materials, or from the authors)?
+
+# Future Directions and Extensions
+
+First, scholars taking a
+person-oriented approach should emphasize reproducibility in carrying out
+analyses. This is in part due to the exploratory and interpretative nature of
+person-oriented approaches. To this end, presenting multiple models, as in
+Pastor et al. (2007), should be encouraged, rather than presenting one solution.
+In addition, having multiple analysts review the solutions found is encouraged.
+As part of this recommendation, we suggest that researchers consider how their
+analyses can be reproduced by other analysts outside of the research team:
+Sharing code and data is an important part of this work. Also related,
+researchers should consider how profiles are replicated across samples.
+
+A second general recommendation considers more flexibly incorporating time, when
+data are collected across multiple time points, into analyses. ISOA groups all
+time points and doesn’t make distinctions. Other approaches perform analysis
+separately at different timepoints (Corpus & Wormington, 2014). Some integrate
+time as a part of the profiles, i.e. growth mixture modeling (groups of
+patterns), i.e. within-person growth modeling, where there are individual growth
+patterns. Research to date has yet to consider additional challenges in applying
+person centered approaches to longitudinal data. For instance, Schmidt et al
+(2018) use an Experience Sampling Method (ESM) approach to collecting data and
+used a person-centered approach to generate profiles of students in science
+class. This work does not account for student-level effects. In other words, it
+did not model the shared variance of multiple observations of the same student.
+
+Third, best practices in within-person or longitudinal research call for
+modeling the nesting structure. However, researchers have yet to successfully
+incorporate this practice into the person centered approach. One way to approach
+this is to use cross-classified mixed effects models, as in Strati, Schmidt, and
+Maier (2017) and in Rosenberg (2018). In such approaches, dependencies in terms
+of, for example, individuals responding to (ESM) surveys at the same time and
+repeated responses being associated with the same individuals can both be
+modeled, although the effects of these two sources of dependencies are not
+nested as in very common uses of multi-level models, but rather are
+cross-classified. West, Welch, & Galecki (2014) have a description of the use of
+multi-level models with cross-classified data and tools (including those freely
+available through R) that can be used to estimate them.
+
 # tidyLPA
@@
   select(broad_interest, enjoyment, self_efficacy) %>% 
-  psych::describe()
+  psych::describe() %>% 
+  knitr::kable()
 ```
@@
 
-``` {r, curry-mac-mclust-example}
+``` {r, curry-mac-mclust-example, echo = TRUE}
 pisaUSA15 %>%
@@
 
-``` {r, curry-mac-mplus, eval = FALSE}
+``` {r, curry-mac-mplus, echo = TRUE}
 pisaUSA15 %>%
@@
   estimate_profiles(3) %>% 
-  get_fit(m1) %>% 
-  knitr::kable()
+  get_fit(m1)
 ```
@@
 ````{r, compare-mclust, eval= TRUE}
-compare_solutions_results <- pisaUSA15 %>%
+many_analyses <- pisaUSA15 %>%
   select(broad_interest, enjoyment, self_efficacy) %>%
   single_imputation() %>%
-  estimate_profiles(models = c(1, 2, 3, 6),
-                    n_profiles = 1:9)
+  estimate_profiles(models = 1,
+                    n_profiles = 1:6)
 
-compare_solutions_results
+many_analyses
 
-compare_solutions(compare_solutions_results)
+compare_solutions(many_analyses)
+
+many_analyses_1_3 <- pisaUSA15 %>%
+  select(broad_interest, enjoyment, self_efficacy) %>%
+  single_imputation() %>%
+  estimate_profiles(models = c(1, 3),
+                    n_profiles = 1:6)
+
+many_analyses_1_3
+
+compare_solutions(many_analyses_1_3)
+```
+
+It appears that the four and five-class solutions fit best. Let's inspect those, specifically:
+
+```{r}
+models_4_5 <- pisaUSA15[1:100, ] %>%
+  select(broad_interest, enjoyment, self_efficacy) %>%
+  single_imputation() %>%
+  estimate_profiles(models = 1,
+                    n_profiles = c(4, 5))
+
+models_4_5 %>% 
+  plot_profiles(add_line = TRUE)
+```
+
+We can also create a faceted plot of density plots for an object of class 'tidyLPA'.
+For each variable, a Total density plot will be shown, along with separate density plots for each latent class, where cases are weighted by the posterior probability of being assigned to that class:
+
+```{r}
+models_4_5 %>% 
+  plot_profiles(add_line = TRUE)
+```
+
+Last, we can create a faceted plot of two-dimensional correlation plots and unidimensional density plots for an object of class 'tidyProfile':
+
+```{r}
+models_4_5 %>% 
+  plot_bivariate()
+```
+
+```{r, eval = FALSE, echo = FALSE}
+many_analyses_mplus <- pisaUSA15 %>%
+  select(broad_interest, enjoyment, self_efficacy) %>%
+  single_imputation() %>%
+  estimate_profiles(models = c(1),
+                    n_profiles = 1:7,
+                    package = "mplus")
+
+many_analyses_mplus
+
+compare_solutions(many_analyses_mplus)
 ```
@@
 
-# Recommendations Regarding Analytic Choices
-
-A note of caution is warranted about LPA in the context of their characteristics
-and strengths. Bauer (2007) notes that many samples of data can be usefully
-broken down into profiles, and that the addition of profiles will likely be
-suggested for reasons other than the samples coming from more than one
-distribution (i.e., due to non-normality in the variables measured). Bauer also
-cautions that profiles should not be reified; that profiles do not necessarily
-exist outside of the analysis that they should be interpreted more as useful
-interpretative devices. These cautions suggest that, in general, parsimony,
-interpretability, and a general sense that the profiles are not necessarily
-real, but are rather helpful analytic tools, should be both priorities for the
-analyst and the reader of studies using this approach.
-
-The GRoLTs checklist is a helpful way to think about how best practices for LPA;
-here, we show how to implement these using tidyLPA. First, here is the checklist:
-
-## GRoLTs checklist
-
-```{r, echo = FALSE}
-t <- readr::read_csv("GRoLTS-applicability-to-LPA.csv")
-
-tp <- t %>% 
-  filter(use == 1) %>% 
-  mutate(`LPA Number` = 1:nrow(.)) %>% 
-  select(`GRoLTS Number` = Number, `LPA Number`, `Checklist Item`)
-
-tp %>% 
-  DT::datatable()
-```
-
-We walk through each of these in turn.
-
-### Is the missing data mechanism reported?
-
-### Is a description provided of what variables are related to attrition/missing data?
-
-### Is a description provided of how missing data in the analyses were dealt with?
-
-### Is information about the distribution of the observed variables included?
-
-### Is the software mentioned?
-
-### Are alternative specifications of the between-class differences in variance–covariance matrix structure considered and clearly documented?
-
-In LPA, we can determine which parameters - variance
-and covariance - are estimated to be the same or different across all profiles,
-and, in the case of covariance, whether it is fixed to zero. Of course, for the
-profiles we are most interested in, the mean is allowed to vary across profiles.
-The models with more parameters freely-estimated use more degrees of freedom;
-thus, similar to the number of profiles selected, the balance between adding
-more parameters (and having a more complex model) and interpretability and
-parsimony must be balanced. If a more complex model fits better and is
-interpretable and is justifiable in terms of the data collected, then it may be
-preferred to simpler models. If the two models are similar in terms of their
-fit, then the more simple, parsimonious model should be selected.
-
-There are a number of analytic choices that need to be made when carrying out
-person-oriented analyses. Because such person-oriented approaches are often more
-subjective (in practice) than other approaches (Linnenbrink-Garcia and
-Wormington, 2017), there is no one rule for determining the solution obtained.
-This solution is obtained on the basis of multiple decisions, such as the number
-of profiles selected or the modeling decisions such as what specific options are
-used for the cluster analysis (i.e., the distance metric used to calculate the
-similarity of the observations as part of the Ward’s hierarchical clustering) or
-what parameters are estimated and how as part of LPA.
-
-Given the subjectivity involved, it is important that researchers be transparent
-and work as part of a team to obtain clustering solutions. Transparency about
-the design and analytic choices is important so that readers can appropriately
-interpret the report. Researchers can enhance transparency and reproducibility
-by sharing detailed descriptions of methodology and document it through the use
-of syntax (and, if possible, data) that we share with others. Working as part of
-a team can help to serve as a check on several of the choices researchers make,
-such as over-fitting or under-fitting the model to the data. Each decision
-depends on multiple factors and balancing tensions. We discuss each of the key
-decisions listed in an analysis.
-
-### Is information reported about the number of random start values and final iterations included?
-
-### Are the model comparison (and selection) tools described from a statistical perspective?
-
-In the case of choosing the number of profiles (and the specification of the
-model / profile solution), multiple criteria, such as the BIC or the proportion
-of variance explained are recommended for decision-making, but also
-interpretability in light of theory, parsimony, and evidence from
-cross-validation should be considered.
-
-### Are the total number of fitted models reported, including a one-class solution?
-
-### Are the number of cases per class reported for each model (absolute sample size, or proportion)?
-
-### If classification of cases in a trajectory is the goal, is entropy reported?
-
-### Are all of the estimates plotted?
-
-### Are the raw data plotted along with the estimates?
-
-### Are characteristics of the final class solution numerically described (i.e., means, SD/SE, n, CI, etc.)?
-
-### Are the syntax files available (either in the appendix, supplementary materials, or from the authors)?
-
-# Future Directions and Extensions
-
-First, scholars taking a
-person-oriented approach should emphasize reproducibility in carrying out
-analyses. This is in part due to the exploratory and interpretative nature of
-person-oriented approaches. To this end, presenting multiple models, as in
-Pastor et al. (2007), should be encouraged, rather than presenting one solution.
-In addition, having multiple analysts review the solutions found is encouraged.
-As part of this recommendation, we suggest that researchers consider how their
-analyses can be reproduced by other analysts outside of the research team:
-Sharing code and data is an important part of this work. Also related,
-researchers should consider how profiles are replicated across samples.
-
-A second general recommendation considers more flexibly incorporating time, when
-data are collected across multiple time points, into analyses. ISOA groups all
-time points and doesn’t make distinctions. Other approaches perform analysis
-separately at different timepoints (Corpus & Wormington, 2014). Some integrate
-time as a part of the profiles, i.e. growth mixture modeling (groups of
-patterns), i.e. within-person growth modeling, where there are individual growth
-patterns. Research to date has yet to consider additional challenges in applying
-person centered approaches to longitudinal data. For instance, Schmidt et al
-(2018) use an Experience Sampling Method (ESM) approach to collecting data and
-used a person-centered approach to generate profiles of students in science
-class. This work does not account for student-level effects. In other words, it
-did not model the shared variance of multiple observations of the same student.
-
-Third, best practices in within-person or longitudinal research call for
-modeling the nesting structure. However, researchers have yet to successfully
-incorporate this practice into the person centered approach. One way to approach
-this is to use cross-classified mixed effects models, as in Strati, Schmidt, and
-Maier (2017) and in Rosenberg (2018). In such approaches, dependencies in terms
-of, for example, individuals responding to (ESM) surveys at the same time and
-repeated responses being associated with the same individuals can both be
-modeled, although the effects of these two sources of dependencies are not
-nested as in very common uses of multi-level models, but rather are
-cross-classified. West, Welch, & Galecki (2014) have a description of the use of
-multi-level models with cross-classified data and tools (including those freely
-available through R) that can be used to estimate them.
-
 # Implications

### 2020-12-28 paper.tex
update

@@
 \urlstyle{same}  % don't use monospace font for urls
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
 \usepackage{graphicx,grffile}
@@
 \maketitle
-
-\begin{verbatim}
-## Warning: package 'ggplot2' was built under R version 3.6.2
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: package 'tibble' was built under R version 3.6.2
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: package 'tidyr' was built under R version 3.6.2
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: package 'readr' was built under R version 3.6.2
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: package 'purrr' was built under R version 3.6.2
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: package 'dplyr' was built under R version 3.6.2
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: package 'simstudy' was built under R version 3.6.2
-\end{verbatim}
-
-\begin{verbatim}
-## NULL
-\end{verbatim}
 
@@
 
-\hypertarget{specific-software-tools}{%
-\section{Specific Software Tools}\label{specific-software-tools}}
+\hypertarget{best-practices-for-latent-profile-analysis}{%
+\section{Best Practices for Latent Profile Analysis}\label{best-practices-for-latent-profile-analysis}}
+
+\hypertarget{recommendations-regarding-analytic-choices}{%
+\subsection{Recommendations Regarding Analytic Choices}\label{recommendations-regarding-analytic-choices}}
+
+A note of caution is warranted about LPA in the context of their characteristics
+and strengths. Bauer (2007) notes that many samples of data can be usefully
+broken down into profiles, and that the addition of profiles will likely be
+suggested for reasons other than the samples coming from more than one
+distribution (i.e., due to non-normality in the variables measured). Bauer also
+cautions that profiles should not be reified; that profiles do not necessarily
+exist outside of the analysis that they should be interpreted more as useful
+interpretative devices. These cautions suggest that, in general, parsimony,
+interpretability, and a general sense that the profiles are not necessarily
+real, but are rather helpful analytic tools, should be both priorities for the
+analyst and the reader of studies using this approach.
+
+The GRoLTs checklist is a helpful way to think about how best practices for LPA;
+here, we show how to implement these using tidyLPA. First, here is the checklist:
+
+\hypertarget{grolts-checklist}{%
+\subsection{GRoLTs checklist}\label{grolts-checklist}}
+
+\hypertarget{htmlwidget-6725a3bb4d9d4a47b614}{}
+
+We walk through each of these in turn.
+
+\hypertarget{is-the-missing-data-mechanism-reported}{%
+\subsubsection{Is the missing data mechanism reported?}\label{is-the-missing-data-mechanism-reported}}
+
+\hypertarget{is-a-description-provided-of-what-variables-are-related-to-attritionmissing-data}{%
+\subsubsection{Is a description provided of what variables are related to attrition/missing data?}\label{is-a-description-provided-of-what-variables-are-related-to-attritionmissing-data}}
+
+\hypertarget{is-a-description-provided-of-how-missing-data-in-the-analyses-were-dealt-with}{%
+\subsubsection{Is a description provided of how missing data in the analyses were dealt with?}\label{is-a-description-provided-of-how-missing-data-in-the-analyses-were-dealt-with}}
+
+\hypertarget{is-information-about-the-distribution-of-the-observed-variables-included}{%
+\subsubsection{Is information about the distribution of the observed variables included?}\label{is-information-about-the-distribution-of-the-observed-variables-included}}
+
+\hypertarget{is-the-software-mentioned}{%
+\subsubsection{Is the software mentioned?}\label{is-the-software-mentioned}}
+
+\hypertarget{are-alternative-specifications-of-the-between-class-differences-in-variancecovariance-matrix-structure-considered-and-clearly-documented}{%
+\subsubsection{Are alternative specifications of the between-class differences in variance--covariance matrix structure considered and clearly documented?}\label{are-alternative-specifications-of-the-between-class-differences-in-variancecovariance-matrix-structure-considered-and-clearly-documented}}
+
+In LPA, we can determine which parameters - variance
+and covariance - are estimated to be the same or different across all profiles,
+and, in the case of covariance, whether it is fixed to zero. Of course, for the
+profiles we are most interested in, the mean is allowed to vary across profiles.
+The models with more parameters freely-estimated use more degrees of freedom;
+thus, similar to the number of profiles selected, the balance between adding
+more parameters (and having a more complex model) and interpretability and
+parsimony must be balanced. If a more complex model fits better and is
+interpretable and is justifiable in terms of the data collected, then it may be
+preferred to simpler models. If the two models are similar in terms of their
+fit, then the more simple, parsimonious model should be selected.
+
+There are a number of analytic choices that need to be made when carrying out
+person-oriented analyses. Because such person-oriented approaches are often more
+subjective (in practice) than other approaches (Linnenbrink-Garcia and
+Wormington, 2017), there is no one rule for determining the solution obtained.
+This solution is obtained on the basis of multiple decisions, such as the number
+of profiles selected or the modeling decisions such as what specific options are
+used for the cluster analysis (i.e., the distance metric used to calculate the
+similarity of the observations as part of the Ward's hierarchical clustering) or
+what parameters are estimated and how as part of LPA.
+
+Given the subjectivity involved, it is important that researchers be transparent
+and work as part of a team to obtain clustering solutions. Transparency about
+the design and analytic choices is important so that readers can appropriately
+interpret the report. Researchers can enhance transparency and reproducibility
+by sharing detailed descriptions of methodology and document it through the use
+of syntax (and, if possible, data) that we share with others. Working as part of
+a team can help to serve as a check on several of the choices researchers make,
+such as over-fitting or under-fitting the model to the data. Each decision
+depends on multiple factors and balancing tensions. We discuss each of the key
+decisions listed in an analysis.
+
+\hypertarget{is-information-reported-about-the-number-of-random-start-values-and-final-iterations-included}{%
+\subsubsection{Is information reported about the number of random start values and final iterations included?}\label{is-information-reported-about-the-number-of-random-start-values-and-final-iterations-included}}
+
+\hypertarget{are-the-model-comparison-and-selection-tools-described-from-a-statistical-perspective}{%
+\subsubsection{Are the model comparison (and selection) tools described from a statistical perspective?}\label{are-the-model-comparison-and-selection-tools-described-from-a-statistical-perspective}}
+
+In the case of choosing the number of profiles (and the specification of the
+model / profile solution), multiple criteria, such as the BIC or the proportion
+of variance explained are recommended for decision-making, but also
+interpretability in light of theory, parsimony, and evidence from
+cross-validation should be considered.
+
+\hypertarget{are-the-total-number-of-fitted-models-reported-including-a-one-class-solution}{%
+\subsubsection{Are the total number of fitted models reported, including a one-class solution?}\label{are-the-total-number-of-fitted-models-reported-including-a-one-class-solution}}
+
+\hypertarget{are-the-number-of-cases-per-class-reported-for-each-model-absolute-sample-size-or-proportion}{%
+\subsubsection{Are the number of cases per class reported for each model (absolute sample size, or proportion)?}\label{are-the-number-of-cases-per-class-reported-for-each-model-absolute-sample-size-or-proportion}}
+
+\hypertarget{if-classification-of-cases-in-a-trajectory-is-the-goal-is-entropy-reported}{%
+\subsubsection{If classification of cases in a trajectory is the goal, is entropy reported?}\label{if-classification-of-cases-in-a-trajectory-is-the-goal-is-entropy-reported}}
+
+\hypertarget{are-all-of-the-estimates-plotted}{%
+\subsubsection{Are all of the estimates plotted?}\label{are-all-of-the-estimates-plotted}}
+
+\hypertarget{are-the-raw-data-plotted-along-with-the-estimates}{%
+\subsubsection{Are the raw data plotted along with the estimates?}\label{are-the-raw-data-plotted-along-with-the-estimates}}
+
+\hypertarget{are-characteristics-of-the-final-class-solution-numerically-described-i.e.-means-sdse-n-ci-etc.}{%
+\subsubsection{Are characteristics of the final class solution numerically described (i.e., means, SD/SE, n, CI, etc.)?}\label{are-characteristics-of-the-final-class-solution-numerically-described-i.e.-means-sdse-n-ci-etc.}}
+
+\hypertarget{are-the-syntax-files-available-either-in-the-appendix-supplementary-materials-or-from-the-authors}{%
+\subsubsection{Are the syntax files available (either in the appendix, supplementary materials, or from the authors)?}\label{are-the-syntax-files-available-either-in-the-appendix-supplementary-materials-or-from-the-authors}}
+
+\hypertarget{future-directions-and-extensions}{%
+\section{Future Directions and Extensions}\label{future-directions-and-extensions}}
+
+First, scholars taking a
+person-oriented approach should emphasize reproducibility in carrying out
+analyses. This is in part due to the exploratory and interpretative nature of
+person-oriented approaches. To this end, presenting multiple models, as in
+Pastor et al. (2007), should be encouraged, rather than presenting one solution.
+In addition, having multiple analysts review the solutions found is encouraged.
+As part of this recommendation, we suggest that researchers consider how their
+analyses can be reproduced by other analysts outside of the research team:
+Sharing code and data is an important part of this work. Also related,
+researchers should consider how profiles are replicated across samples.
+
+A second general recommendation considers more flexibly incorporating time, when
+data are collected across multiple time points, into analyses. ISOA groups all
+time points and doesn't make distinctions. Other approaches perform analysis
+separately at different timepoints (Corpus \& Wormington, 2014). Some integrate
+time as a part of the profiles, i.e.~growth mixture modeling (groups of
+patterns), i.e.~within-person growth modeling, where there are individual growth
+patterns. Research to date has yet to consider additional challenges in applying
+person centered approaches to longitudinal data. For instance, Schmidt et al
+(2018) use an Experience Sampling Method (ESM) approach to collecting data and
+used a person-centered approach to generate profiles of students in science
+class. This work does not account for student-level effects. In other words, it
+did not model the shared variance of multiple observations of the same student.
+
+Third, best practices in within-person or longitudinal research call for
+modeling the nesting structure. However, researchers have yet to successfully
+incorporate this practice into the person centered approach. One way to approach
+this is to use cross-classified mixed effects models, as in Strati, Schmidt, and
+Maier (2017) and in Rosenberg (2018). In such approaches, dependencies in terms
+of, for example, individuals responding to (ESM) surveys at the same time and
+repeated responses being associated with the same individuals can both be
+modeled, although the effects of these two sources of dependencies are not
+nested as in very common uses of multi-level models, but rather are
+cross-classified. West, Welch, \& Galecki (2014) have a description of the use of
+multi-level models with cross-classified data and tools (including those freely
+available through R) that can be used to estimate them.
+
+\hypertarget{tidylpa}{%
+\section{tidyLPA}\label{tidylpa}}
 
@@
 
-\hypertarget{tidylpa}{%
-\subsection{tidyLPA}\label{tidylpa}}
+\hypertarget{the-tidylpa-r-package}{%
+\subsection{The tidyLPA R package}\label{the-tidylpa-r-package}}
 
@@
 
-Here, we use a simulated dataset contains data on moral relevance and judgment across the seven domains of the \emph{Morality As Cooperation} scale.
-More on this scale can be found in Curry et al., (2019).
+Here, we use a simulated dataset contains data on broad interest, enjoyment, and self-efficacy from the PISA assessment.
+
+First, let's explore the data graphically using the \{ggplot2\} package:
+
+\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1.pdf}
+
+Here, we use the \{psych\} package to explore the distributional properties of the variables:
+
+\begin{tabular}{l|r|r|r|r|r|r|r|r|r|r|r|r|r}
+\hline
+  & vars & n & mean & sd & median & trimmed & mad & min & max & range & skew & kurtosis & se\\
+\hline
+broad\_interest & 1 & 5454 & 2.652246 & 0.7772383 & 2.8 & 2.683902 & 0.59304 & 1 & 5 & 4 & -0.2767312 & -0.0261463 & 0.0105244\\
+\hline
+enjoyment & 2 & 5520 & 2.780263 & 0.7509386 & 3.0 & 2.819973 & 0.59304 & 1 & 4 & 3 & -0.5076839 & 0.1652595 & 0.0101073\\
+\hline
+self\_efficacy & 3 & 5409 & 2.136430 & 0.6729143 & 2.0 & 2.112135 & 0.63540 & 1 & 4 & 3 & 0.4376698 & 0.0705830 & 0.0091496\\
+\hline
+\end{tabular}
 
@@
 
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{pisaUSA15 }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(broad_interest, enjoyment, self_efficacy) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{single_imputation}\NormalTok{() }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{estimate_profiles}\NormalTok{(}\DecValTok{3}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## tidyLPA analysis using mclust: 
+## 
+##  Model Classes AIC      BIC      Entropy prob_min prob_max n_min n_max BLRT_p
+##  1     3       34815.05 34908.15 0.75    0.77     0.94     0.16  0.67  0.01
+\end{verbatim}
+
+\hypertarget{mplus-via-tidylpa}{%
+\subsubsection{Mplus via tidyLPA}\label{mplus-via-tidylpa}}
+
+We can use Mplus simply by changing the package argument for
+\texttt{estimate\_profiles()} to \enquote{mplus} or \enquote{MplusAutomation} (either will work; not run here):
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{pisaUSA15 }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(broad_interest, enjoyment, self_efficacy) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{single_imputation}\NormalTok{() }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{estimate_profiles}\NormalTok{(}\DecValTok{3}\NormalTok{, }\DataTypeTok{package =} \StringTok{"mplus"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## tidyLPA analysis using mplus: 
+## 
+##  Model Classes AIC      BIC      Entropy prob_min prob_max n_min n_max BLRT_p
+##  1     3       34804.95 34898.05 0.77    0.79     0.94     0.16  0.67  0.00
+\end{verbatim}
+
+These are the very basics; we will return to what we do with this output later in the tutorial.
+
+\hypertarget{fit}{%
+\subsubsection{Fit}\label{fit}}
+
 We can see a number of statistics for each of the models that were estimated, including the
@@
 
-\begin{tabular}{r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r}
-\hline
-Model & Classes & LogLik & AIC & AWE & BIC & CAIC & CLC & KIC & SABIC & ICL & Entropy & prob\_min & prob\_max & n\_min & n\_max & BLRT\_val & BLRT\_p\\
-\hline
-1 & 1 & -81245.00 & 162546.0 & 162977.4 & 162692.7 & 162720.7 & 162492.0 & 162577.0 & 162603.7 & -162692.7 & 1.0000000 & 1.0000000 & 1.0000000 & 1.0000000 & 1.0000000 & NA & NA\\
-\hline
-1 & 2 & -79727.50 & 159541.0 & 160204.8 & 159766.3 & 159809.3 & 159456.7 & 159587.0 & 159629.7 & -159925.3 & 0.8369494 & 0.9296991 & 0.9637189 & 0.3441092 & 0.6558908 & 3035.00346 & 0.009901\\
-\hline
-1 & 3 & -79239.27 & 158594.5 & 159490.6 & 158898.4 & 158956.4 & 158480.2 & 158655.5 & 158714.1 & -159175.7 & 0.8188766 & 0.9053581 & 0.9222054 & 0.1465517 & 0.5272989 & 976.44661 & 0.009901\\
-\hline
-1 & 4 & -79090.59 & 158327.2 & 159455.5 & 158709.6 & 158782.6 & 158182.7 & 158403.2 & 158477.7 & -159150.2 & 0.7763719 & 0.8393595 & 0.9127306 & 0.0732759 & 0.4655172 & 297.35974 & 0.009901\\
-\hline
-1 & 5 & -79029.50 & 158235.0 & 159595.6 & 158696.0 & 158784.0 & 158060.4 & 158326.0 & 158416.4 & -159398.5 & 0.7012112 & 0.6979575 & 0.8987374 & 0.0675287 & 0.2406609 & 122.19715 & 0.009901\\
-\hline
-1 & 6 & -78971.30 & 158148.6 & 159741.3 & 158688.2 & 158791.2 & 157944.0 & 158254.6 & 158361.0 & -159456.7 & 0.6954594 & 0.6929238 & 0.9230731 & 0.0603448 & 0.3060345 & 116.39215 & 0.009901\\
-\hline
-1 & 7 & -78930.08 & 158096.2 & 159921.0 & 158714.3 & 158832.3 & 157861.5 & 158217.2 & 158339.5 & -159550.6 & 0.6933730 & 0.5964096 & 0.9236077 & 0.0610632 & 0.2550287 & 82.44963 & 0.009901\\
-\hline
-1 & 8 & -78851.80 & 157969.6 & 160026.5 & 158666.3 & 158799.3 & 157705.1 & 158105.6 & 158243.8 & -159391.4 & 0.7477064 & 0.6432919 & 0.9219048 & 0.0423851 & 0.2902299 & 156.55019 & 0.009901\\
-\hline
-1 & 9 & -78821.74 & 157939.5 & 160228.6 & 158714.8 & 158862.8 & 157645.0 & 158090.5 & 158244.6 & -159455.8 & 0.7515427 & 0.6071032 & 0.9456474 & 0.0402299 & 0.2938218 & 60.12578 & 0.009901\\
-\hline
-1 & 10 & -78751.27 & 157828.5 & 160349.8 & 158682.4 & 158845.4 & 157504.1 & 157994.5 & 158164.6 & -159420.9 & 0.7625067 & 0.6463008 & 0.9396876 & 0.0337644 & 0.2909483 & 140.92881 & 0.009901\\
-\hline
-\end{tabular}
+\begin{verbatim}
+## # A tibble: 1 x 18
+##   Model Classes  LogLik    AIC    AWE    BIC   CAIC    CLC    KIC  SABIC     ICL
+##   <dbl>   <dbl>   <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>   <dbl>
+## 1     1       3 -17386. 34800. 35055. 34893. 34907. 34774. 34817. 34849. -36424.
+## # ... with 7 more variables: Entropy <dbl>, prob_min <dbl>, prob_max <dbl>,
+## #   n_min <dbl>, n_max <dbl>, BLRT_val <dbl>, BLRT_p <dbl>
+\end{verbatim}
 
-\hypertarget{mplus-via-tidylpa}{%
-\subsubsection{Mplus via tidyLPA}\label{mplus-via-tidylpa}}
+\hypertarget{comparing-a-wide-range-of-solutions}{%
+\subsection{Comparing a wide range of solutions}\label{comparing-a-wide-range-of-solutions}}
 
-We can use Mplus simply by changing the package argument for
-\texttt{estimate\_profiles()} to \enquote{mplus} or \enquote{MplusAutomation} (either will work; not run here):
+The function \texttt{compare\_solutions()} compares the fit of several estimated models,
+with varying numbers of profiles and model specifications:
 
-These are the very basics; we will return to what we do with this output later in the tutorial.
+\begin{verbatim}
+## tidyLPA analysis using mclust: 
+## 
+##  Model Classes AIC      BIC      Entropy prob_min prob_max n_min n_max BLRT_p
+##  1     1       37962.61 38002.51 1.00    1.00     1.00     1.00  1.00        
+##  1     2       35495.33 35561.83 0.74    0.83     0.96     0.24  0.76  0.01  
+##  1     3       34850.99 34944.09 0.75    0.76     0.94     0.16  0.67  0.01  
+##  1     4       32990.63 33110.33 0.93    0.94     0.97     0.07  0.54  0.01  
+##  1     5       32652.78 32799.09 0.89    0.79     0.99     0.07  0.46  0.01  
+##  1     6       32631.88 32804.79 0.76    0.44     0.99     0.07  0.34  0.01
+\end{verbatim}
+
+\begin{verbatim}
+## Compare tidyLPA solutions:
+## 
+##  Model Classes BIC      
+##  1     1       38002.509
+##  1     2       35561.831
+##  1     3       34944.094
+##  1     4       33110.333
+##  1     5       32799.090
+##  1     6       32804.788
+## 
+## Best model according to BIC is Model 1 with 5 classes.
+## 
+## An analytic hierarchy process, based on the fit indices AIC, AWE, BIC, CLC, and KIC (Akogul & Erisoglu, 2017), suggests the best solution is Model 1 with 6 classes.
+\end{verbatim}
 
@@
 
-Here is an example of specifying a model with varying variances and covariances
-(Model 6)
-
-\begin{verbatim}
-## tidyLPA analysis using mclust: 
-## 
-##  Model Classes AIC       BIC       Entropy prob_min prob_max n_min n_max BLRT_p
-##  6     3       156872.22 158752.84 0.67    0.82     0.87     0.23  0.41  0.01
-\end{verbatim}
-
-\hypertarget{comparing-a-wide-range-of-solutions}{%
-\subsection{Comparing a wide range of solutions}\label{comparing-a-wide-range-of-solutions}}
-
-The function \texttt{compare\_solutions()} compares the fit of several estimated models,
-with varying numbers of profiles and model specifications:
-
-\begin{verbatim}
-## tidyLPA analysis using mclust: 
-## 
-##  Model Classes AIC       BIC       Entropy prob_min prob_max n_min n_max BLRT_p
-##  1     1       139226.33 139352.05 1.00    1.00     1.00     1.00  1.00        
-##  1     2       136511.15 136704.97 0.83    0.92     0.97     0.34  0.66  0.01  
-##  1     3       135651.76 135913.69 0.81    0.90     0.92     0.14  0.53  0.01  
-##  1     4       135391.00 135721.03 0.77    0.83     0.91     0.07  0.47  0.01  
-##  1     5       135326.12 135724.24 0.72    0.57     0.91     0.07  0.39  0.01  
-##  1     6       135281.64 135747.86 0.68    0.56     0.92     0.07  0.28  0.01  
-##  1     7       135218.07 135752.39 0.67    0.59     0.91     0.07  0.29  0.01  
-##  1     8       135135.62 135738.04 0.70    0.59     0.92     0.05  0.31  0.01  
-##  1     9       135098.58 135769.10 0.70    0.57     0.93     0.05  0.30  0.01  
-##  2     1       139226.33 139352.05 1.00    1.00     1.00     1.00  1.00        
-##  2     2       136245.76 136502.44 0.81    0.92     0.96     0.39  0.61  0.01  
-##  2     3       135337.03 135724.68 0.80    0.91     0.91     0.25  0.49  0.01  
-##  2     4       135103.11 135621.72 0.75    0.82     0.92     0.17  0.37  0.01  
-##  2     5       134840.49 135490.07 0.75    0.76     0.89     0.13  0.31  0.01  
-##  2     6       134688.81 135469.35 0.75    0.71     0.89     0.06  0.29  0.01  
-##  2     7       134579.47 135490.97 0.75    0.64     0.90     0.06  0.23  0.01  
-##  2     8       134482.30 135524.76 0.75    0.71     0.90     0.06  0.18  0.01  
-##  2     9       134434.76 135608.19 0.75    0.65     0.91     0.06  0.17  0.01  
-##  3     1       134912.97 135384.44 1.00    1.00     1.00     1.00  1.00        
-##  3     2       134836.15 135375.72 0.69    0.78     0.96     0.22  0.78  0.01  
-##  3     3       134862.13 135469.80 0.41    0.16     0.88     0.07  0.69  0.76  
-##  3     4       134728.94 135404.71 0.50    0.31     0.83     0.10  0.63  0.01  
-##  3     5       134615.77 135359.63 0.58    0.43     0.79     0.07  0.56  0.01  
-##  3     6       134696.35 135508.32 0.52    0.43     0.81     0.06  0.39  0.71  
-##  3     7       134589.15 135469.21 0.54    0.36     0.84     0.06  0.37  0.01  
-##  3     8       134604.12 135552.29 0.54    0.40     0.84     0.04  0.33  0.77  
-##  3     9       134568.05 135584.32 0.56    0.31     0.83     0.03  0.34  0.01  
-##  6     1       134912.97 135384.44 1.00    1.00     1.00     1.00  1.00        
-##  6     2       134223.62 135171.78 0.62    0.87     0.90     0.48  0.52  0.01  
-##  6     3       134151.67 135576.55 0.68    0.80     0.88     0.18  0.47  0.01  
-##  6     4       134089.83 135991.41 0.68    0.77     0.87     0.17  0.36  0.01  
-##  6     5       134092.59 136470.87 0.71    0.77     0.89     0.06  0.33  0.34  
-##  6     6       134088.85 136943.83 0.72    0.75     0.90     0.06  0.29  0.43  
-##  6     7       134021.25 137352.94 0.74    0.77     0.88     0.09  0.20  0.01  
-##  6     8       134038.51 137846.90 0.77    0.79     0.98     0.03  0.19  0.58  
-##  6     9       134067.02 138352.11 0.78    0.77     0.96     0.04  0.18  0.78
-\end{verbatim}
-
-\begin{verbatim}
-## Compare tidyLPA solutions:
-## 
-##  Model Classes BIC       
-##  1     1       139352.049
-##  1     2       136704.970
-##  1     3       135913.687
-##  1     4       135721.026
-##  1     5       135724.243
-##  1     6       135747.862
-##  1     7       135752.395
-##  1     8       135738.044
-##  1     9       135769.104
-##  2     1       139352.049
-##  2     2       136502.441
-##  2     3       135724.680
-##  2     4       135621.724
-##  2     5       135490.065
-##  2     6       135469.349
-##  2     7       135490.971
-##  2     8       135524.760
-##  2     9       135608.188
-##  3     1       135384.436
-##  3     2       135375.718
-##  3     3       135469.796
-##  3     4       135404.708
-##  3     5       135359.634
-##  3     6       135508.318
-##  3     7       135469.214
-##  3     8       135552.290
-##  3     9       135584.321
-##  6     1       135384.436
-##  6     2       135171.784
-##  6     3       135576.546
-##  6     4       135991.409
-##  6     5       136470.871
-##  6     6       136943.826
-##  6     7       137352.936
-##  6     8       137846.896
-##  6     9       138352.109
-## 
-## Best model according to BIC is Model 6 with 2 classes.
-## 
-## An analytic hierarchy process, based on the fit indices AIC, AWE, BIC, CLC, and KIC (Akogul & Erisoglu, 2017), suggests the best solution is Model 6 with 2 classes.
-\end{verbatim}
-
-\includegraphics{paper_files/figure-latex/plot-one-solution-1.pdf}
-
 \hypertarget{accessing-values-in-the-output}{%
@@
 function from the \textbf{tidyr} package can be used, e.g.:
-
-\hypertarget{recommendations-regarding-analytic-choices}{%
-\section{Recommendations Regarding Analytic Choices}\label{recommendations-regarding-analytic-choices}}
-
-A note of caution is warranted about LPA in the context of their characteristics
-and strengths. Bauer (2007) notes that many samples of data can be usefully
-broken down into profiles, and that the addition of profiles will likely be
-suggested for reasons other than the samples coming from more than one
-distribution (i.e., due to non-normality in the variables measured). Bauer also
-cautions that profiles should not be reified; that profiles do not necessarily
-exist outside of the analysis that they should be interpreted more as useful
-interpretative devices. These cautions suggest that, in general, parsimony,
-interpretability, and a general sense that the profiles are not necessarily
-real, but are rather helpful analytic tools, should be both priorities for the
-analyst and the reader of studies using this approach.
-
-The GRoLTs checklist is a helpful way to think about how best practices for LPA;
-here, we show how to implement these using tidyLPA. First, here is the checklist:
-
-\hypertarget{grolts-checklist}{%
-\subsection{GRoLTs checklist}\label{grolts-checklist}}
-
-\hypertarget{htmlwidget-675d126ab2afaf44720a}{}
-
-We walk through each of these in turn.
-
-\hypertarget{is-the-missing-data-mechanism-reported}{%
-\subsubsection{Is the missing data mechanism reported?}\label{is-the-missing-data-mechanism-reported}}
-
-\hypertarget{is-a-description-provided-of-what-variables-are-related-to-attritionmissing-data}{%
-\subsubsection{Is a description provided of what variables are related to attrition/missing data?}\label{is-a-description-provided-of-what-variables-are-related-to-attritionmissing-data}}
-
-\hypertarget{is-a-description-provided-of-how-missing-data-in-the-analyses-were-dealt-with}{%
-\subsubsection{Is a description provided of how missing data in the analyses were dealt with?}\label{is-a-description-provided-of-how-missing-data-in-the-analyses-were-dealt-with}}
-
-\hypertarget{is-information-about-the-distribution-of-the-observed-variables-included}{%
-\subsubsection{Is information about the distribution of the observed variables included?}\label{is-information-about-the-distribution-of-the-observed-variables-included}}
-
-\hypertarget{is-the-software-mentioned}{%
-\subsubsection{Is the software mentioned?}\label{is-the-software-mentioned}}
-
-\hypertarget{are-alternative-specifications-of-the-between-class-differences-in-variancecovariance-matrix-structure-considered-and-clearly-documented}{%
-\subsubsection{Are alternative specifications of the between-class differences in variance--covariance matrix structure considered and clearly documented?}\label{are-alternative-specifications-of-the-between-class-differences-in-variancecovariance-matrix-structure-considered-and-clearly-documented}}
-
-\hypertarget{is-information-reported-about-the-number-of-random-start-values-and-final-iterations-included}{%
-\subsubsection{Is information reported about the number of random start values and final iterations included?}\label{is-information-reported-about-the-number-of-random-start-values-and-final-iterations-included}}
-
-\hypertarget{are-the-model-comparison-and-selection-tools-described-from-a-statistical-perspective}{%
-\subsubsection{Are the model comparison (and selection) tools described from a statistical perspective?}\label{are-the-model-comparison-and-selection-tools-described-from-a-statistical-perspective}}
-
-\hypertarget{are-the-total-number-of-fitted-models-reported-including-a-one-class-solution}{%
-\subsubsection{Are the total number of fitted models reported, including a one-class solution?}\label{are-the-total-number-of-fitted-models-reported-including-a-one-class-solution}}
-
-\hypertarget{are-the-number-of-cases-per-class-reported-for-each-model-absolute-sample-size-or-proportion}{%
-\subsubsection{Are the number of cases per class reported for each model (absolute sample size, or proportion)?}\label{are-the-number-of-cases-per-class-reported-for-each-model-absolute-sample-size-or-proportion}}
-
-\hypertarget{if-classification-of-cases-in-a-trajectory-is-the-goal-is-entropy-reported}{%
-\subsubsection{If classification of cases in a trajectory is the goal, is entropy reported?}\label{if-classification-of-cases-in-a-trajectory-is-the-goal-is-entropy-reported}}
-
-\hypertarget{are-all-of-the-estimates-plotted}{%
-\subsubsection{Are all of the estimates plotted?}\label{are-all-of-the-estimates-plotted}}
-
-\hypertarget{are-the-raw-data-plotted-along-with-the-estimates}{%
-\subsubsection{Are the raw data plotted along with the estimates?}\label{are-the-raw-data-plotted-along-with-the-estimates}}
-
-\hypertarget{are-characteristics-of-the-final-class-solution-numerically-described-i.e.-means-sdse-n-ci-etc.}{%
-\subsubsection{Are characteristics of the final class solution numerically described (i.e., means, SD/SE, n, CI, etc.)?}\label{are-characteristics-of-the-final-class-solution-numerically-described-i.e.-means-sdse-n-ci-etc.}}
-
-\hypertarget{are-the-syntax-files-available-either-in-the-appendix-supplementary-materials-or-from-the-authors}{%
-\subsubsection{Are the syntax files available (either in the appendix, supplementary materials, or from the authors)?}\label{are-the-syntax-files-available-either-in-the-appendix-supplementary-materials-or-from-the-authors}}
-
-There are a number of analytic choices that need to be made when carrying out
-person-oriented analyses. Because such person-oriented approaches are often more
-subjective (in practice) than other approaches (Linnenbrink-Garcia and
-Wormington, 2017), there is no one rule for determining the solution obtained.
-This solution is obtained on the basis of multiple decisions, such as the number
-of profiles selected or the modeling decisions such as what specific options are
-used for the cluster analysis (i.e., the distance metric used to calculate the
-similarity of the observations as part of the Ward's hierarchical clustering) or
-what parameters are estimated and how as part of LPA.
-
-Given the subjectivity involved, it is important that researchers be transparent
-and work as part of a team to obtain clustering solutions. Transparency about
-the design and analytic choices is important so that readers can appropriately
-interpret the report. Researchers can enhance transparency and reproducibility
-by sharing detailed descriptions of methodology and document it through the use
-of syntax (and, if possible, data) that we share with others. Working as part of
-a team can help to serve as a check on several of the choices researchers make,
-such as over-fitting or under-fitting the model to the data. Each decision
-depends on multiple factors and balancing tensions. We discuss each of the key
-decisions listed in an analysis.
-
-\hypertarget{how-the-data-are-centering-or-scaled}{%
-\subsection{How the Data are Centering or Scaled}\label{how-the-data-are-centering-or-scaled}}
-
-The data can be transformed by centering or scaling. Typically, these are done
-after the profiles are created, so that differences between profiles can be more
-easily explored. They can also be done prior to the analysis, which can be
-helpful for obtaining solutions when the variables are on very different
-scales.
-
-\hypertarget{how-to-choose-the-number-of-profilesclasses}{%
-\subsection{How to Choose the Number of Profiles/Classes}\label{how-to-choose-the-number-of-profilesclasses}}
-
-In the case of choosing the number of profiles (and the specification of the
-model / profile solution), multiple criteria, such as the BIC or the proportion
-of variance explained are recommended for decision-making, but also
-interpretability in light of theory, parsimony, and evidence from
-cross-validation should be considered.
-
-\hypertarget{how-to-choose-options-for-the-clustering-procedure-or-to-select-the-model-parameterization}{%
-\subsection{How to Choose Options for the Clustering Procedure or to Select the Model Parameterization}\label{how-to-choose-options-for-the-clustering-procedure-or-to-select-the-model-parameterization}}
-
-In LPA, we can determine which parameters - variance
-and covariance - are estimated to be the same or different across all profiles,
-and, in the case of covariance, whether it is fixed to zero. Of course, for the
-profiles we are most interested in, the mean is allowed to vary across profiles.
-The models with more parameters freely-estimated use more degrees of freedom;
-thus, similar to the number of profiles selected, the balance between adding
-more parameters (and having a more complex model) and interpretability and
-parsimony must be balanced. If a more complex model fits better and is
-interpretable and is justifiable in terms of the data collected, then it may be
-preferred to simpler models. If the two models are similar in terms of their
-fit, then the more simple, parsimonious model should be selected.
-
-\hypertarget{more-general-recommendations}{%
-\subsection{More General Recommendations}\label{more-general-recommendations}}
-
-This section highlights three general recommendations. First, scholars taking a
-person-oriented approach should emphasize reproducibility in carrying out
-analyses. This is in part due to the exploratory and interpretative nature of
-person-oriented approaches. To this end, presenting multiple models, as in
-Pastor et al. (2007), should be encouraged, rather than presenting one solution.
-In addition, having multiple analysts review the solutions found is encouraged.
-As part of this recommendation, we suggest that researchers consider how their
-analyses can be reproduced by other analysts outside of the research team:
-Sharing code and data is an important part of this work. Also related,
-researchers should consider how profiles are replicated across samples.
-
-A second general recommendation considers more flexibly incorporating time, when
-data are collected across multiple time points, into analyses. ISOA groups all
-time points and doesn't make distinctions. Other approaches perform analysis
-separately at different timepoints (Corpus \& Wormington, 2014). Some integrate
-time as a part of the profiles, i.e.~growth mixture modeling (groups of
-patterns), i.e.~within-person growth modeling, where there are individual growth
-patterns. Research to date has yet to consider additional challenges in applying
-person centered approaches to longitudinal data. For instance, Schmidt et al
-(2018) use an Experience Sampling Method (ESM) approach to collecting data and
-used a person-centered approach to generate profiles of students in science
-class. This work does not account for student-level effects. In other words, it
-did not model the shared variance of multiple observations of the same student.
-
-Third, best practices in within-person or longitudinal research call for
-modeling the nesting structure. However, researchers have yet to successfully
-incorporate this practice into the person centered approach. One way to approach
-this is to use cross-classified mixed effects models, as in Strati, Schmidt, and
-Maier (2017) and in Rosenberg (2018). In such approaches, dependencies in terms
-of, for example, individuals responding to (ESM) surveys at the same time and
-repeated responses being associated with the same individuals can both be
-modeled, although the effects of these two sources of dependencies are not
-nested as in very common uses of multi-level models, but rather are
-cross-classified. West, Welch, \& Galecki (2014) have a description of the use of
-multi-level models with cross-classified data and tools (including those freely
-available through R) that can be used to estimate them.
 

### 2020-12-29 paper.rmd
format paper

@@
 
-In statistics classes, textbooks, and workshops, an example like the following is common:
+In statistics classes, textbooks, and workshops, an example like the following
+is common:
 
-> Grades are normally distributed, with $\mu$ = 75, $\sigma$ = 5.
+> Grades are normally distributed, with $\\mu$ = 75, $\\sigma$ = 5.
 
-Students, may (reasonably) ask, though, are grades *really* normally distributed?
+Students, may (reasonably) ask, though, are grades *really* normally
+distributed?
 
-```{r, include = FALSE, eval = TRUE}
+``` {r, include = FALSE, eval = TRUE}
 def <- defData(varname = "x1", formula = 65, variance = 30)
@@
                dist = "mixture")
+```
 
-set.seed(2716)
-dx <- genData(1000, def)
+set.seed(2716) dx \<- genData(1000, def)
+
+``` 
 ```
@@
 
-```{r, include = FALSE, eval = TRUE}
+``` {r, include = FALSE, eval = TRUE}
 ggplot(dx, aes(x = xMix)) +
@@
 
-```{r, include = FALSE, eval = TRUE}
+``` {r, include = FALSE, eval = TRUE}
 dx %>% 
@@
 
-We note that challenge facing the analyst using LPA is that these
-parameters and the distinct model parameterizations that can be estimated is the
-different terminology used. As one example, Scrucca et al. (2017) refer to these
+We note that challenge facing the analyst using LPA is that these parameters and
+the distinct model parameterizations that can be estimated is the different
+terminology used. As one example, Scrucca et al. (2017) refer to these
 parameterizations not in terms of whether and how parameters are estimated, but
@@
 The GRoLTs checklist is a helpful way to think about how best practices for LPA;
-here, we show how to implement these using tidyLPA. First, here is the checklist:
+here, we show how to implement these using tidyLPA. First, here is the
+checklist:
 
@@
 
-```{r, echo = FALSE}
+``` {r, echo = FALSE}
 t <- readr::read_csv("GRoLTS-applicability-to-LPA.csv")
+```
 
-tp <- t %>% 
-  filter(use == 1) %>% 
-  mutate(`LPA Number` = 1:nrow(.)) %>% 
+tp \<- t %\>% filter(use == 1) %\>% mutate(`LPA Number` = 1:nrow(.)) %\>%
 select(`GRoLTS Number` = Number, `LPA Number`, `Checklist Item`)
 
-tp %>% 
-  DT::datatable()
+tp %\>% DT::datatable()
+
+``` 
 ```
@@
 
-In LPA, we can determine which parameters - variance
-and covariance - are estimated to be the same or different across all profiles,
-and, in the case of covariance, whether it is fixed to zero. Of course, for the
-profiles we are most interested in, the mean is allowed to vary across profiles.
-The models with more parameters freely-estimated use more degrees of freedom;
-thus, similar to the number of profiles selected, the balance between adding
-more parameters (and having a more complex model) and interpretability and
-parsimony must be balanced. If a more complex model fits better and is
-interpretable and is justifiable in terms of the data collected, then it may be
-preferred to simpler models. If the two models are similar in terms of their
-fit, then the more simple, parsimonious model should be selected.
+In LPA, we can determine which parameters - variance and covariance - are
+estimated to be the same or different across all profiles, and, in the case of
+covariance, whether it is fixed to zero. Of course, for the profiles we are most
+interested in, the mean is allowed to vary across profiles. The models with more
+parameters freely-estimated use more degrees of freedom; thus, similar to the
+number of profiles selected, the balance between adding more parameters (and
+having a more complex model) and interpretability and parsimony must be
+balanced. If a more complex model fits better and is interpretable and is
+justifiable in terms of the data collected, then it may be preferred to simpler
+models. If the two models are similar in terms of their fit, then the more
+simple, parsimonious model should be selected.
 
@@
 
-First, scholars taking a
-person-oriented approach should emphasize reproducibility in carrying out
-analyses. This is in part due to the exploratory and interpretative nature of
-person-oriented approaches. To this end, presenting multiple models, as in
-Pastor et al. (2007), should be encouraged, rather than presenting one solution.
-In addition, having multiple analysts review the solutions found is encouraged.
-As part of this recommendation, we suggest that researchers consider how their
-analyses can be reproduced by other analysts outside of the research team:
-Sharing code and data is an important part of this work. Also related,
-researchers should consider how profiles are replicated across samples.
+First, scholars taking a person-oriented approach should emphasize
+reproducibility in carrying out analyses. This is in part due to the exploratory
+and interpretative nature of person-oriented approaches. To this end, presenting
+multiple models, as in Pastor et al. (2007), should be encouraged, rather than
+presenting one solution. In addition, having multiple analysts review the
+solutions found is encouraged. As part of this recommendation, we suggest that
+researchers consider how their analyses can be reproduced by other analysts
+outside of the research team: Sharing code and data is an important part of this
+work. Also related, researchers should consider how profiles are replicated
+across samples.
 
@@
 same maximum likelihood estimation procedure, they are benchmarked to produce
-the same output. Also, note that we have described the model
-specifications with descriptions of what is estimated in terms of the variances
-and covariances, the common names for the models (i.e., class-varying
-unrestricted), and the covariance matrix associated with the parameterizationfor
-the six models that are possible to be estimated on the website for tidyLPA (see
-here).
+the same output. Also, note that we have described the model specifications with
+descriptions of what is estimated in terms of the variances and covariances, the
+common names for the models (i.e., class-varying unrestricted), and the
+covariance matrix associated with the parameterizationfor the six models that
+are possible to be estimated on the website for tidyLPA (see here).
 
@@
 
-You can also install the development version of tidyLPA, which may have new features that are not yet availalble on CRAN (because the version on CRAN is periodically updated) from GitHub with:
+You can also install the development version of tidyLPA, which may have new
+features that are not yet availalble on CRAN (because the version on CRAN is
+periodically updated) from GitHub with:
 
@@
 
-Here, we use a simulated dataset contains data on broad interest, enjoyment, and self-efficacy from the PISA assessment.
+Here, we use a simulated dataset contains data on broad interest, enjoyment, and
+self-efficacy from the PISA assessment.
 
@@
 
-```{r}
+``` {r}
 pisaUSA15 %>%
@@
 
-Here, we use the {psych} package to explore the distributional properties of the variables:
+Here, we use the {psych} package to explore the distributional properties of the
+variables:
 
-```{r}
+``` {r}
 pisaUSA15 %>%
@@
 We can use Mplus simply by changing the package argument for
-`estimate_profiles()` to "mplus" or "MplusAutomation" (either will work; not run here):
+`estimate_profiles()` to "mplus" or "MplusAutomation" (either will work; not run
+here):
 
@@
 
-These are the very basics; we will return to what we do with this output later in the tutorial.
+These are the very basics; we will return to what we do with this output later
+in the tutorial.
 
@@
 
-We can see a number of statistics for each of the models that were estimated, including the 
-AIC (Aikake information criterion; based on -2 * the log-likelihood, and penalized by number of parameters) and BIC (BIC: Bayesian information criterion; based on -2 log-likelihood, and penalized by number of parameters adjusted by sample size).
-Both are _penalized_ likelihood statistics; the log-likelihood will always decrease with the addition of additional
-paremeters estimated by the model, whereas the AIC and BIC will account for the number of parameters; lower values of
-either indicate better fit.
+We can see a number of statistics for each of the models that were estimated,
+including the AIC (Aikake information criterion; based on -2 \* the
+log-likelihood, and penalized by number of parameters) and BIC (BIC: Bayesian
+information criterion; based on -2 log-likelihood, and penalized by number of
+parameters adjusted by sample size). Both are *penalized* likelihood statistics;
+the log-likelihood will always decrease with the addition of additional
+paremeters estimated by the model, whereas the AIC and BIC will account for the
+number of parameters; lower values of either indicate better fit.
 
-Entropy.
-`prob_min` refers to the minimum of the diagonal of the average latent class probabilities for most likely class membership, by assigned class.
-The minimum should be as high as possible, reflecting greater classification certainty (cases are assigned to classes they have a high probability of belonging to; see Jung & Wickrama, 2008); `prob_max`
-refers to the maximum.
-Finally.
-`n_min` refers to the proportion of the sample assigned to the smallest class (based on most likely class membership); `n_max` to the largest.
-Finally, `BLRT_p` refers to the p-value for the bootstrapped likelihood ratio test; a value greater than .05 indicates that the model fit better than one with one fewer class.
-Additional fit statistics (other penalized likelihood values, such as the CAIC, SABIC, and ICL), and the bootstrapped LRT 
-test statistic, can be obtained with the `get_fit()` function:
+Entropy. `prob_min` refers to the minimum of the diagonal of the average latent
+class probabilities for most likely class membership, by assigned class. The
+minimum should be as high as possible, reflecting greater classification
+certainty (cases are assigned to classes they have a high probability of
+belonging to; see Jung & Wickrama, 2008); `prob_max` refers to the maximum.
+Finally. `n_min` refers to the proportion of the sample assigned to the smallest
+class (based on most likely class membership); `n_max` to the largest. Finally,
+`BLRT_p` refers to the p-value for the bootstrapped likelihood ratio test; a
+value greater than .05 indicates that the model fit better than one with one
+fewer class. Additional fit statistics (other penalized likelihood values, such
+as the CAIC, SABIC, and ICL), and the bootstrapped LRT test statistic, can be
+obtained with the `get_fit()` function:
 
-```{r, get-fit-m1}
+``` {r, get-fit-m1}
 pisaUSA15 %>%
@@
 
-````{r, compare-mclust, eval= TRUE}
+``` {r, compare-mclust, eval= TRUE}
 many_analyses <- pisaUSA15 %>%
@@
                     n_profiles = 1:6)
-
-many_analyses
-
-compare_solutions(many_analyses)
-
-many_analyses_1_3 <- pisaUSA15 %>%
-  select(broad_interest, enjoyment, self_efficacy) %>%
-  single_imputation() %>%
-  estimate_profiles(models = c(1, 3),
-                    n_profiles = 1:6)
-
-many_analyses_1_3
-
-compare_solutions(many_analyses_1_3)
 ```
 
-It appears that the four and five-class solutions fit best. Let's inspect those, specifically:
+many\_analyses
 
-```{r}
+compare\_solutions(many\_analyses)
+
+many\_analyses\_1\_3 \<- pisaUSA15 %\>% select(broad\_interest, enjoyment,
+self\_efficacy) %\>% single\_imputation() %\>% estimate\_profiles(models =
+c(1, 3), n\_profiles = 1:6)
+
+many\_analyses\_1\_3
+
+compare\_solutions(many\_analyses\_1\_3)
+
+``` 
+```
+
+It appears that the four and five-class solutions fit best. Let's inspect those,
+specifically:
+
+``` {r}
 models_4_5 <- pisaUSA15[1:100, ] %>%
@@
                     n_profiles = c(4, 5))
+```
 
+models\_4\_5 %\>% plot\_profiles(add\_line = TRUE)
+
+``` 
+```
+
+We can also create a faceted plot of density plots for an object of class
+'tidyLPA'. For each variable, a Total density plot will be shown, along with
+separate density plots for each latent class, where cases are weighted by the
+posterior probability of being assigned to that class:
+
+``` {r}
 models_4_5 %>% 
@@
 
-We can also create a faceted plot of density plots for an object of class 'tidyLPA'.
-For each variable, a Total density plot will be shown, along with separate density plots for each latent class, where cases are weighted by the posterior probability of being assigned to that class:
+Last, we can create a faceted plot of two-dimensional correlation plots and
+unidimensional density plots for an object of class 'tidyProfile':
 
-```{r}
-models_4_5 %>% 
-  plot_profiles(add_line = TRUE)
-```
-
-Last, we can create a faceted plot of two-dimensional correlation plots and unidimensional density plots for an object of class 'tidyProfile':
-
-```{r}
+``` {r}
 models_4_5 %>% 
@@
 
-```{r, eval = FALSE, echo = FALSE}
+``` {r, eval = FALSE, echo = FALSE}
 many_analyses_mplus <- pisaUSA15 %>%
@@
                     package = "mplus")
-
-many_analyses_mplus
-
-compare_solutions(many_analyses_mplus)
 ```
 
-### Specifying the _model_
+many\_analyses\_mplus
 
-We passed the number of profiles, but no details about the model we wished to estimate.
+compare\_solutions(many\_analyses\_mplus)
 
-There are a number of different types of models for LPA that estimate different parameters; choosing 
-between them is an important consideration, and one that has a bearing on how we will interpret the results we saw above.
+``` 
+```
+
+### Specifying the *model*
+
+We passed the number of profiles, but no details about the model we wished to
+estimate.
+
+There are a number of different types of models for LPA that estimate different
+parameters; choosing between them is an important consideration, and one that
+has a bearing on how we will interpret the results we saw above.
 
@@
 argument), the model can be specified in terms of whether and how the variable
-variances and covariances are estimated. The models are specified by passing arguments to the `variance` and `covariance`
-arguments. The possible values for these arguments are:
+variances and covariances are estimated. The models are specified by passing
+arguments to the `variance` and `covariance` arguments. The possible values for
+these arguments are:
 
@@
 
-In general, the approach to choosing the model is similar to choosing the number of profiles, requiring **deciding on the basis of evidence from multiple sources**, including information criteria, statistical tests, and concerns of interpretability and parsimony.
-The article by [Pastor and colleagues (2007)](https://www.sciencedirect.com/science/article/pii/S0361476X06000543) has helpful information on the model specifications.
-Here, the six models that are possible to specify in LPA are described in terms of how the variables used to create the profiles are estimated.
+In general, the approach to choosing the model is similar to choosing the number
+of profiles, requiring **deciding on the basis of evidence from multiple
+sources**, including information criteria, statistical tests, and concerns of
+interpretability and parsimony. The article by [Pastor and colleagues
+(2007)](https://www.sciencedirect.com/science/article/pii/S0361476X06000543) has
+helpful information on the model specifications. Here, the six models that are
+possible to specify in LPA are described in terms of how the variables used to
+create the profiles are estimated.
 
-Note that *p* represents different profiles and each parameterization is represented by a 4 x 4 covariance matrix and therefore would represent the parameterization for a four-profile solution.
-In all of the models, the means are estimated freely in the different profiles.
-Imagine that each row and column represents a different variable, i.e., the first row (and column) represents broad interest, the second enjoyment, the third self-efficacy, and the fourth another variable, i.e., future goals and plans.
+Note that *p* represents different profiles and each parameterization is
+represented by a 4 x 4 covariance matrix and therefore would represent the
+parameterization for a four-profile solution. In all of the models, the means
+are estimated freely in the different profiles. Imagine that each row and column
+represents a different variable, i.e., the first row (and column) represents
+broad interest, the second enjoyment, the third self-efficacy, and the fourth
+another variable, i.e., future goals and plans.
 
-### 1. Equal variances, and covariances fixed to 0 (model 1)
+### 1\. Equal variances, and covariances fixed to 0 (model 1)
 
-In this model, which corresponds to the mclust model wit the name "EEI", the variances are estimated to be equal across profiles, indicated by the absence of a p subscript for any of the diagonal elements of the matrix.
-The covariances are constrained to be zero, as indicated by the 0's between every combination of the variables.
+In this model, which corresponds to the mclust model wit the name "EEI", the
+variances are estimated to be equal across profiles, indicated by the absence of
+a p subscript for any of the diagonal elements of the matrix. The covariances
+are constrained to be zero, as indicated by the 0's between every combination of
+the variables.
 
@@
 
-This model is highly constrained but also parsimonious: the profiles are estimated in such a way that the variables' variances are identical for each of the profiles, and the relationships between the variables are not estimated.
-In this way, less degrees of freedom are taken used to explain the observations that make up the data.
-However, estimating more parameters--as in the other models--may better explain the data, justifying the addition in complexity that their addition involves (and their reduction in degrees of freedom).
-This model is sometimes referred to as a *class-invariant* parameterization.
+This model is highly constrained but also parsimonious: the profiles are
+estimated in such a way that the variables' variances are identical for each of
+the profiles, and the relationships between the variables are not estimated. In
+this way, less degrees of freedom are taken used to explain the observations
+that make up the data. However, estimating more parameters--as in the other
+models--may better explain the data, justifying the addition in complexity that
+their addition involves (and their reduction in degrees of freedom). This model
+is sometimes referred to as a *class-invariant* parameterization.
 
-$$
-\left[ \begin{matrix} { \sigma  }_{ 1 }^{ 2 } & 0 & 0 & 0 \\ 0 & { \sigma  }_{ 2 }^{ 2 } & 0 & 0 \\ 0 & 0 & { \sigma  }_{ 3 }^{ 2 } & 0 \\ 0 & 0 & 0 & { \sigma  }_{ 4 }^{ 2 } \end{matrix} \right] 
-$$
+$$ \\left[ \\begin{matrix} { \\sigma }*{ 1 }^{ 2 } & 0 & 0 & 0 \\ 0 & { \\sigma
+}*{ 2 }^{ 2 } & 0 & 0 \\ 0 & 0 & { \\sigma }*{ 3 }^{ 2 } & 0 \\ 0 & 0 & 0 & {
+\\sigma }*{ 4 }^{ 2 } \\end{matrix} \\right] $$
 
-### 2. Varying variances and covariances fixed to 0 (model 2)
+### 2\. Varying variances and covariances fixed to 0 (model 2)
 
-This model corresponds to the mclust model "VVI" and allows for the variances to be freely estimated across profiles. The covariances are constrained to zero.
+This model corresponds to the mclust model "VVI" and allows for the variances to
+be freely estimated across profiles. The covariances are constrained to zero.
 
@@
 
-Thus, it is more flexible (and less parsimonious) than model 1, but in terms of the covariances, is more constrained than model 2.
-This model is sometimes referred to as a *class-varying diagonal* parameterization.
+Thus, it is more flexible (and less parsimonious) than model 1, but in terms of
+the covariances, is more constrained than model 2. This model is sometimes
+referred to as a *class-varying diagonal* parameterization.
 
-$$ 
-\left[ \begin{matrix} { \sigma  }_{ 1p }^{ 2 } & 0 & 0 & 0 \\ 0 & { \sigma  }_{ 2p }^{ 2 } & 0 & 0 \\ 0 & 0 & { \sigma  }_{ 3p }^{ 2 } & 0 \\ 0 & 0 & 0 & { \sigma  }_{ 4p }^{ 2 } \end{matrix} \right] 
-$$
+$$ \\left[ \\begin{matrix} { \\sigma }*{ 1p }^{ 2 } & 0 & 0 & 0 \\ 0 & {
+\\sigma }*{ 2p }^{ 2 } & 0 & 0 \\ 0 & 0 & { \\sigma }*{ 3p }^{ 2 } & 0 \\ 0 & 0
+& 0 & { \\sigma }*{ 4p }^{ 2 } \\end{matrix} \\right] $$
 
-### 3. Equal variances and equal covariances (model 3)
+### 3\. Equal variances and equal covariances (model 3)
 
-This model corresponds to the mclust model "EEE".
-In this model, the variances are still constrained to be the same across the profiles, although now the covariances are estimated (but like the variances, are constrained to be the same across profiles).
+This model corresponds to the mclust model "EEE". In this model, the variances
+are still constrained to be the same across the profiles, although now the
+covariances are estimated (but like the variances, are constrained to be the
+same across profiles).
 
@@
 
-Thus, this model is the first to estimate the covariance (or correlations) of the variables used to create the profiles, thus adding more information that can be used to better understand the characteristics of the profiles (and, potentially, better explain the data).
-This model is sometimes referred to as a *class-invariant unrestricted* parameterization.
+Thus, this model is the first to estimate the covariance (or correlations) of
+the variables used to create the profiles, thus adding more information that can
+be used to better understand the characteristics of the profiles (and,
+potentially, better explain the data). This model is sometimes referred to as a
+*class-invariant unrestricted* parameterization.
 
-$$
-\left[ \begin{matrix} { \sigma  }_{ 1 }^{ 2 } & { \sigma  }_{ 21 } & { \sigma  }_{ 31 } & { \sigma  }_{ 41 } \\ { \sigma  }_{ 12 } & { \sigma  }_{ 2 }^{ 2 } & { \sigma  }_{ 23 } & { \sigma  }_{ 24 } \\ { \sigma  }_{ 13 } & { \sigma  }_{ 12 } & { \sigma  }_{ 3 }^{ 2 } & { \sigma  }_{ 33 } \\ { \sigma  }_{ 14 } & { \sigma  }_{ 12 } & { \sigma  }_{ 12 } & { \sigma  }_{ 4 }^{ 2 } \end{matrix} \right]
-$$
+$$ \\left[ \\begin{matrix} { \\sigma }*{ 1 }^{ 2 } & { \\sigma }*{ 21 } & {
+\\sigma }*{ 31 } & { \\sigma }*{ 41 } \\ { \\sigma }*{ 12 } & { \\sigma }*{ 2
+}^{ 2 } & { \\sigma }*{ 23 } & { \\sigma }*{ 24 } \\ { \\sigma }*{ 13 } & {
+\\sigma }*{ 12 } & { \\sigma }*{ 3 }^{ 2 } & { \\sigma }*{ 33 } \\ { \\sigma
+}*{ 14 } & { \\sigma }*{ 12 } & { \\sigma }*{ 12 } & { \\sigma }*{ 4 }^{ 2 }
+\\end{matrix} \\right] $$
 
-### 4. Varying means, varying variances, and equal covariances (model 4)
+### 4\. Varying means, varying variances, and equal covariances (model 4)
 
-This model, which specifies for the variances to be freely estimated across the profiles and for the covariances to be estimated to be equal across profiles, extends model 3.
+This model, which specifies for the variances to be freely estimated across the
+profiles and for the covariances to be estimated to be equal across profiles,
+extends model 3.
 
@@
 
-Unfortunately, this model cannot be specified with mclust, though it can be with MPlus; this model *can* be used with the functions to interface to MPlus described below.
+Unfortunately, this model cannot be specified with mclust, though it can be with
+MPlus; this model *can* be used with the functions to interface to MPlus
+described below.
 
-$$
-\left[ \begin{matrix} { \sigma  }_{ 1p }^{ 2 } & { \sigma  }_{ 21 } & { \sigma  }_{ 31 } & { \sigma  }_{ 41 } \\ { \sigma  }_{ 12 } & { \sigma  }_{ 2p }^{ 2 } & { \sigma  }_{ 23 } & { \sigma  }_{ 24 } \\ { \sigma  }_{ 13 } & { \sigma  }_{ 12 } & { \sigma  }_{ 3p }^{ 2 } & { \sigma  }_{ 33 } \\ { \sigma  }_{ 14 } & { \sigma  }_{ 12 } & { \sigma  }_{ 12 } & { \sigma  }_{ 4p }^{ 2 } \end{matrix} \right]
-$$
+$$ \\left[ \\begin{matrix} { \\sigma }*{ 1p }^{ 2 } & { \\sigma }*{ 21 } & {
+\\sigma }*{ 31 } & { \\sigma }*{ 41 } \\ { \\sigma }*{ 12 } & { \\sigma }*{ 2p
+}^{ 2 } & { \\sigma }*{ 23 } & { \\sigma }*{ 24 } \\ { \\sigma }*{ 13 } & {
+\\sigma }*{ 12 } & { \\sigma }*{ 3p }^{ 2 } & { \\sigma }*{ 33 } \\ { \\sigma
+}*{ 14 } & { \\sigma }*{ 12 } & { \\sigma }*{ 12 } & { \\sigma }*{ 4p }^{ 2 }
+\\end{matrix} \\right] $$
 
-### 5. Varying means, equal variances, and varying covariances (model 5)
+### 5\. Varying means, equal variances, and varying covariances (model 5)
 
-This model specifies the variances to be equal across the profiles, but allows the covariances to be freely estimated across the profiles.
+This model specifies the variances to be equal across the profiles, but allows
+the covariances to be freely estimated across the profiles.
 
@@
 
-Like model 4, this model cannot be specified with mclust, though it can be with MPlus. Again, this model *can* be used with the functions to interface to MPlus described below.
+Like model 4, this model cannot be specified with mclust, though it can be with
+MPlus. Again, this model *can* be used with the functions to interface to MPlus
+described below.
 
-$$
-\left[ \begin{matrix} { \sigma  }_{ 1 }^{ 2 } & { \sigma  }_{ 21p } & { \sigma  }_{ 31p } & { \sigma  }_{ 41p } \\ { \sigma  }_{ 12p } & { \sigma  }_{ 2 }^{ 2 } & { \sigma  }_{ 23p } & { \sigma  }_{ 24p } \\ { \sigma  }_{ 13p } & { \sigma  }_{ 12p } & { \sigma  }_{ 3 }^{ 2 } & { \sigma  }_{ 33p } \\ { \sigma  }_{ 14p } & { \sigma  }_{ 12p } & { \sigma  }_{ 12p } & { \sigma  }_{ 4 }^{ 2 } \end{matrix} \right] \quad
-$$
+$$ \\left[ \\begin{matrix} { \\sigma }*{ 1 }^{ 2 } & { \\sigma }*{ 21p } & {
+\\sigma }*{ 31p } & { \\sigma }*{ 41p } \\ { \\sigma }*{ 12p } & { \\sigma }*{ 2
+}^{ 2 } & { \\sigma }*{ 23p } & { \\sigma }*{ 24p } \\ { \\sigma }*{ 13p } & {
+\\sigma }*{ 12p } & { \\sigma }*{ 3 }^{ 2 } & { \\sigma }*{ 33p } \\ { \\sigma
+}*{ 14p } & { \\sigma }*{ 12p } & { \\sigma }*{ 12p } & { \\sigma }*{ 4 }^{ 2 }
+\\end{matrix} \\right] \\quad $$
 
-### 6. Varying variances and varying covariances (model 6)
+### 6\. Varying variances and varying covariances (model 6)
 
-This model corresponds to the mclust model "VVV". It allows the variances and the covariances to be freely estimated across profiles.
+This model corresponds to the mclust model "VVV". It allows the variances and
+the covariances to be freely estimated across profiles.
 
@@
 
-Thus, it is the most complex model, with the potential to allow for understanding many aspects of the variables that are used to estimate the profiles and how they are related.
-However, it is less parsimonious than all of the other models, and the added parameters should be considered in light of how preferred this model is relative to those with more simple specifications.
-This model is sometimes referred to as a *class-varying unrestricted* parameterization.
+Thus, it is the most complex model, with the potential to allow for
+understanding many aspects of the variables that are used to estimate the
+profiles and how they are related. However, it is less parsimonious than all of
+the other models, and the added parameters should be considered in light of how
+preferred this model is relative to those with more simple specifications. This
+model is sometimes referred to as a *class-varying unrestricted*
+parameterization.
 
-$$
-\left[ \begin{matrix} { \sigma  }_{ 1p }^{ 2 } & { \sigma  }_{ 21p } & { \sigma  }_{ 31p } & { \sigma  }_{ 41p } \\ { \sigma  }_{ 12p } & { \sigma  }_{ 2p }^{ 2 } & { \sigma  }_{ 23p } & { \sigma  }_{ 24p } \\ { \sigma  }_{ 13p } & { \sigma  }_{ 12p } & { \sigma  }_{ 3p }^{ 2 } & { \sigma  }_{ 33p } \\ { \sigma  }_{ 14p } & { \sigma  }_{ 12p } & { \sigma  }_{ 12p } & { \sigma  }_{ 4p }^{ 2 } \end{matrix} \right]
-$$
+$$ \\left[ \\begin{matrix} { \\sigma }*{ 1p }^{ 2 } & { \\sigma }*{ 21p } & {
+\\sigma }*{ 31p } & { \\sigma }*{ 41p } \\ { \\sigma }*{ 12p } & { \\sigma
+}*{ 2p }^{ 2 } & { \\sigma }*{ 23p } & { \\sigma }*{ 24p } \\ { \\sigma }*{ 13p
+} & { \\sigma }*{ 12p } & { \\sigma }*{ 3p }^{ 2 } & { \\sigma }*{ 33p } \\ {
+\\sigma }*{ 14p } & { \\sigma }*{ 12p } & { \\sigma }*{ 12p } & { \\sigma }*{ 4p
+}^{ 2 } \\end{matrix} \\right] $$
 
@@
 
-A step away from the tutorial side of this work, we wish to note some implications 
-for the developers of statistical software for psychological scientists.
+A step away from the tutorial side of this work, we wish to note some
+implications for the developers of statistical software for psychological
+scientists.
 
@@
 of, for example, tools to carry out highly-sophisticated Bayesian methods using
-the most cutting-edge tools [@burkner2017, @makowski2019]. We think
-that tidyLPA speaks to this, too.
+the most cutting-edge tools [@burkner2017, @makowski2019]. We think that
+tidyLPA speaks to this, too.
 

### 2020-12-29 paper.rmd
re-run analysis from top-to-bottom

--- 2020-12-26 paper.rmd
@@
   - name          : "Patrick Beymer"
-    affiliation   : "3"
+    affiliation   : "5"
   - name          : "Daniel Anderson"
@@
     institution   : "Michigan State University"
+  - id            : "3"
+    institution   : "University of Oregon"
   - id            : "4"
-    institution   : "University of Oregon"
-  - id            : "5"
+    institution   : "University of Wisconsin, Madison"
 
@@
 keywords          : "Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust"
-wordcount         : ""
+wordcount         : "suggested to be no more than 3,000"
 
@@
 library(papaja)
-```
-
-```{r analysis-preferences, echo = FALSE}
 # Seed for random number generation
-set.seed(42)
+set.seed(20201229)
 library(tidyverse)
@@
 
-In statistics classes, textbooks, and workshops, an example like the following is common:
+In statistics classes, textbooks, and workshops, an example like the following
+is common:
 
@@
 
-Students, may (reasonably) ask, though, are grades *really* normally distributed?
-
-```{r, include = FALSE, eval = TRUE}
+``` {r, include = FALSE, eval = TRUE}
 def <- defData(varname = "x1", formula = 65, variance = 30)
@@
                dist = "mixture")
-
-set.seed(2716)
 dx <- genData(1000, def)
@@
 
-Are their one distribution?
+Is there one distribution?
 
-```{r, include = FALSE, eval = TRUE}
+``` {r, echo = FALSE, eval = TRUE}
 ggplot(dx, aes(x = xMix)) +
@@
 
-Or two?
+Or, are there two?
 
-```{r, include = FALSE, eval = TRUE}
+``` {r, echo = FALSE, eval = TRUE}
 dx %>% 
@@
 
-This kind of distribution (that is *bimodal*) is not exclusive to grades. For
-teachers, psychologists, researchers, and friends and family members, people are
-highly-complex and not easily able to be characterized by *one* characteristic
-or personality trait--and its distribution. In the social sciences, broadly, and
-in psychology, particularly, a statistical method that can be used to describe
-how people, in their individual particularities, may have similarities on the
-basis of some set of measures through which they can be grouped in meaningful,
-distinctive ways. This approach, which has a provenance in developmental (or
+This kind of distribution (one that is *bimodal*) is not exclusive to grades.
+For teachers, psychologists, researchers, and even friends and family members,
+people are highly-complex and not easily able to be characterized by *one*
+characteristic or personality trait--and its distribution.
+
+In the social sciences, broadly, and in the psychological sciences, in
+particular, a statistical method that can be used to describe how people, in
+their individual particularities, may have similarities on the basis of some set
+of measures through which they can be grouped in meaningful, distinctive ways.
+
+This approach, which we view as having a provenance in developmental (or
 *person-oriented*) approaches (Bergman & El-Khouri, 1997; Magnusson &
@@
 
-In this tutorial, we aim to describe one of the most commonly-used--and relevant
-to psychologists---application of the general mixture model, to cases for which
-all of the variables for which (relatively) homogeneous groups are identified
-from among a (relatively) heterogeneous sample are continuos, *latent profile
-analysis* (LPA). We note that scholars using the approach, especially in
-developmental and educational psychology, use cluster analytic methods. Such an
-approach has some similarities with LPA (CITE paper showing how k-means is
-similar to the simplest model specification of LPA?), though LPA also has a
-number of benefits (Linnenbrink-Garcia & Wormington, 2017), and we we focus on
-LPA in this paper.
+In this tutorial, we aim to describe one of the most commonly-used---and
+relevant to psychologists---application of the general mixture model, to cases
+for which all of the variables for which (relatively) homogeneous groups are
+identified from among a (relatively) heterogeneous sample are continuos, *latent
+profile analysis* (LPA). After describing the method and some examples of its
+use, we provide a tutorial for carrying out LPA in the context of a
+freely-available, open-source statistical software package that we have
+developed and supported over the past three years for R (R Core Team, 2019),
+tidyLPA. Finally, we offer some ideas about best practices and informed
+recommendations for researchers aiming to use LPA in their applied work, and
+conclude with reflections on the role of statistical software--especially
+software that is freely-available, open-source, and highly-performant--in the
+psychological sciences.
 
-After describing the method and some examples of its use, we provide a tutorial
-for carrying out LPA in the context of a freely-available, open-source
-statistical software package we created for R (R Core Team, 2019), tidyLPA.
-Finally, we offer some ideas about best practices and informed recommendations
-for researchers aiming to use LPA in their applied work, and conclude with
-reflections on the role of statistical software--especially software that is
-freely-available, open-source, and highly-performant--in the psychological
-sciences.
-
-# Latent Profile Analysis
+# Latent Profile Analysis and the Need for Efficient and Reproducible Analyses
 
@@
 analytic approaches is that LPA is model-based; instead of using algorithms to
-group together cases, LPA seeks to estimate parameters - in terms of variances
-and covariances and how they are the same or different across profiles - that
+group together cases, LPA seeks to estimate parameters---in terms of variances
+and covariances and how they are the same or different across profiles---that
 best characterize the different distributions. Then, this approach seeks to
-assign to each observation a probability that the observation is a sample from
-the population associated with each profile (or mixture component).
+estimate a probability that the observation is a sample from the population
+associated with each profile (or mixture component) for each observation in the
+dataset.
 
@@
 which can be freely-estimated, estimated to be equal, or fixed to be zero.
-
-We note that challenge facing the analyst using LPA is that these
-parameters and the distinct model parameterizations that can be estimated is the
-different terminology used. As one example, Scrucca et al. (2017) refer to these
-parameterizations not in terms of whether and how parameters are estimated, but
-rather in terms of the geometric properties of the distributions that result
-from particular parameterizations. Muthen and Muthen (1997-2017) and others
-(Pastor et al., 2007) commonly refer to local independence to mean that the
-covariances are fixed to zero (also described as the specification of the
-covariance matrix as “diagonal,” because only the diagonal components, or the
-variances, are estimated).
 
@@
 decrease the degrees of freedom), we are more likely to fit the data, but less
-likely to be able to replicate the model with a second set of data. In other
-words, more parameters may mean a loss of external validity. As we progress
-toward more complex models (with increasingly complex parameterization), then we
-are more likely to fit the data better. More information on the model
-parameterizations are discussed in the context of the software tool tidyLPA that
-we have developed.
+likely to be able to replicate the model with a second set of data. As we
+progress toward more complex models (with increasingly complex
+parameterization), then we are more likely to fit the data better.
+In all, this flexibility associated with LPA also has a more pragmatic cost. It
+can be very difficult to efficiently---and reproducibly---estimate and compare a
+number of models. This cost means that analysts may focus on a subset of models.
+It may also mean that unintentional errors are introduced through copying and
+pasting the output of model estimations. Part of why we developed tidyLPA was to
+make the model estimation process more efficient and more reproducible. As we
+describe later, we do this using R, which has a number of benefits, including
+being open-source, powerful (in terms of pre-processing and using the output of
+models), and increasingly widely-used in the psychological sciences. We also
+provide an interface to perhaps the most widely-used software for LPA, MPlus. In
+this way, tidyLPA can both allow researchers to estimate models they already
+estimate using MPlus more efficiently and reproducibly, and, for those without
+access to this software, to do so entirely using R and the open-source R package
+mclust.
 
-# Best Practices for Latent Profile Analysis
+# Existing Software Options
 
-# tidyLPA
-
-## Tools other than tidyLPA
+There are a number of software options available that provide context for why we 
+developed tidyLPA.
 
@@
 RMixMod, which uses compiled C++ code) and others in terms of ease-of-use (i.e.,
-the plot methods built-in to RMixMod, mclust, and other tools). However, they
-also have some drawbacks, in that it can be difficult to translate between the
-model specifications, which are often described in terms of the geometric
+the plot methods built-in to RMixMod, mclust, and other tools).
+
+However, these existing R package also have some drawbacks in addition to positive features, in that it can be difficult to translate between the model specifications, which are often described in terms of the geometric
 properties of the multivariate distributions being estimated (i.e., “spherical,
@@
 algorithm, which can make comparing results across tools challenging.
-
-## The tidyLPA R package
 
@@
 same maximum likelihood estimation procedure, they are benchmarked to produce
-the same output. Also, note that we have described the model
-specifications with descriptions of what is estimated in terms of the variances
-and covariances, the common names for the models (i.e., class-varying
-unrestricted), and the covariance matrix associated with the parameterizationfor
-the six models that are possible to be estimated on the website for tidyLPA (see
-here).
+the same output. Also, note that we have described the model specifications with
+descriptions of what is estimated in terms of the variances and covariances, the
+common names for the models (i.e., class-varying unrestricted), and the
+covariance matrix associated with the parameterizationfor the six models that
+are possible to be estimated.
 
+# A Walkthrough Using the tidyLPA R package
+
+Here, we provide a walkthrough of an analysis using the tidyLPA R package.
 ## Installation
@@
 
-``` {r, eval = F}
+``` {r, eval = F, echo = TRUE}
 install.packages("tidyLPA")
@@
 
-You can also install the development version of tidyLPA, which may have new features that are not yet availalble on CRAN (because the version on CRAN is periodically updated) from GitHub with:
-
-``` {r gh-installation, eval = FALSE}
+You can also install the development version of tidyLPA, which may have new
+features that are not yet availalble on CRAN (because the version on CRAN is
+periodically updated) from GitHub with:
+``` {r gh-installation, eval = FALSE, echo = TRUE}
 # install.packages("devtools")
@@
 ```
-
-``` {r, message = F}
+``` {r, message = F, echo = TRUE}
 library(tidyLPA)
@@
 ```
+## Plotting and Describing the Data
 
-## Examples
-
-Here, we use a simulated dataset contains data on broad interest, enjoyment, and self-efficacy from the PISA assessment.
-
+Here, we use a simulated dataset contains data on broad (general, and not situation- or topic-specific) interest, sense of enjoyment, and self-efficacy from the PISA assessment.
 First, let's explore the data graphically using the {ggplot2} package:
-
-```{r}
+```{r, echo = TRUE}
 pisaUSA15 %>%
@@
 ```
-
-Here, we use the {psych} package to explore the distributional properties of the variables:
-
-```{r}
+We see that the data broadly _appears_ to be normally distributed, as the following statistics from the use of the psych package suggest:
+``` {r, echo = TRUE}
 pisaUSA15 %>%
@@
 
-Add one or two sentences here on the dataset?
+## Using tidyLPA directly and exclusively within R
 
-### Mclust via tidyLPA
+First, we show how to use tidyLPA within R---without any need to use or purchase MPlus.
 
@@
 
-Note that here and throughout this paper we use the function `knitr::kable()` to
-format the output in an RMarkdown document, like the one we are using; this is
-not necessary to use unless you also wish to format the tidyLPA output in an
-RMarkdown document.
-
-``` {r, curry-mac-mclust-example}
+``` {r, curry-mac-mclust-example, echo = TRUE}
 pisaUSA15 %>%
@@
 
-### Mplus via tidyLPA
+## Using tidyLPA to estimate models efficiently and reproducibly in Mplus
 
 We can use Mplus simply by changing the package argument for
-`estimate_profiles()` to "mplus" or "MplusAutomation" (either will work; not run here):
+`estimate_profiles()` to "mplus" or "MplusAutomation" (either will work; not run
+here):
 
-``` {r, curry-mac-mplus, eval = FALSE}
+``` {r, curry-mac-mplus, echo = TRUE}
 pisaUSA15 %>%
@@
 
-These are the very basics; we will return to what we do with this output later in the tutorial.
+These are the very basics; we will return to what we do with this output later
+in the tutorial.
 
-### Fit
+## Determining the Fit of an Estimated Model
 
-We can see a number of statistics for each of the models that were estimated, including the 
-AIC (Aikake information criterion; based on -2 * the log-likelihood, and penalized by number of parameters) and BIC (BIC: Bayesian information criterion; based on -2 log-likelihood, and penalized by number of parameters adjusted by sample size).
-Both are _penalized_ likelihood statistics; the log-likelihood will always decrease with the addition of additional
-paremeters estimated by the model, whereas the AIC and BIC will account for the number of parameters; lower values of
-either indicate better fit.
+We can see a number of statistics for each of the models that were estimated,
+including the AIC (Aikake information criterion; based on -2 \* the
+log-likelihood, and penalized by number of parameters) and BIC (BIC: Bayesian
+information criterion; based on -2 log-likelihood, and penalized by number of
+parameters adjusted by sample size). Both are *penalized* likelihood statistics;
+the log-likelihood will always decrease with the addition of additional
+paremeters estimated by the model, whereas the AIC and BIC will account for the
+number of parameters; lower values of either indicate better fit.
 
-Entropy.
-`prob_min` refers to the minimum of the diagonal of the average latent class probabilities for most likely class membership, by assigned class.
-The minimum should be as high as possible, reflecting greater classification certainty (cases are assigned to classes they have a high probability of belonging to; see Jung & Wickrama, 2008); `prob_max`
-refers to the maximum.
-Finally.
-`n_min` refers to the proportion of the sample assigned to the smallest class (based on most likely class membership); `n_max` to the largest.
-Finally, `BLRT_p` refers to the p-value for the bootstrapped likelihood ratio test; a value greater than .05 indicates that the model fit better than one with one fewer class.
-Additional fit statistics (other penalized likelihood values, such as the CAIC, SABIC, and ICL), and the bootstrapped LRT 
-test statistic, can be obtained with the `get_fit()` function:
+Entropy. `prob_min` refers to the minimum of the diagonal of the average latent
+class probabilities for most likely class membership, by assigned class. The
+minimum should be as high as possible, reflecting greater classification
+certainty (cases are assigned to classes they have a high probability of
+belonging to; see Jung & Wickrama, 2008); `prob_max` refers to the maximum.
+Finally. `n_min` refers to the proportion of the sample assigned to the smallest
+class (based on most likely class membership); `n_max` to the largest. Finally,
+`BLRT_p` refers to the p-value for the bootstrapped likelihood ratio test; a
+value greater than .05 indicates that the model fit better than one with one
+fewer class. Additional fit statistics (other penalized likelihood values, such
+as the CAIC, SABIC, and ICL), and the bootstrapped LRT test statistic, can be
+obtained with the `get_fit()` function (processed with the tidyr R package to be in "long" rather than "wide" format):
 
-```{r, get-fit-m1}
+``` {r, get-fit-m1, echo = TRUE}
+library(tidyr)
+
 pisaUSA15 %>%
@@
   get_fit(m1) %>% 
-  knitr::kable()
+  gather(stat, val)
 ```
 
-## Comparing a wide range of solutions
+## Comparing the Fit for Many Estimated Models
 
@@
 
-````{r, compare-mclust, eval= TRUE}
-compare_solutions_results <- pisaUSA15 %>%
+````{r, compare-mclust, eval = TRUE, echo = TRUE}
+many_analyses_1_3 <- pisaUSA15 %>%
   select(broad_interest, enjoyment, self_efficacy) %>%
   single_imputation() %>%
-  estimate_profiles(models = c(1, 2, 3, 6),
-                    n_profiles = 1:9)
+  estimate_profiles(models = c(1, 3),
+                    n_profiles = 1:5)
 
-compare_solutions_results
-
-compare_solutions(compare_solutions_results)
+compare_solutions(many_analyses_1_3)
 ```
 
-### Specifying the _model_
+## Inspecting and Plotting the best-fitting estimated models
 
-We passed the number of profiles, but no details about the model we wished to estimate.
+It appears that the four and five-class solutions fit best. Let's inspect those, specifically:
 
-There are a number of different types of models for LPA that estimate different parameters; choosing 
-between them is an important consideration, and one that has a bearing on how we will interpret the results we saw above.
+```{r, echo = TRUE, eval = TRUE}
+models_4_5 <- pisaUSA15[1:100, ] %>%
+  select(broad_interest, enjoyment, self_efficacy) %>%
+  single_imputation() %>%
+  estimate_profiles(models = 1,
+                    n_profiles = c(4, 5))
 
-Thus, in addition to the number of profiles (specified with the `n_profiles`
-argument), the model can be specified in terms of whether and how the variable
-variances and covariances are estimated. The models are specified by passing arguments to the `variance` and `covariance`
-arguments. The possible values for these arguments are:
+models_4_5 %>% 
+  plot_profiles(add_line = TRUE)
+```
+
+Here, it appears that the addition of a fifth class does not add a great deal relative to the four-class solution.
+Accordingly, we consider the four-class solution to be the solution we interpret and present, while heeding the recommendations about reporting which other models we estimated, how we chose to focus on the two best-fitting models we did, and why we presented the four-class, class-invariant set of estimates.
+
+Were we to wish to inspect _only_ that chosen model, we could estimate the model again, replacing `n_profiles = c(4, 5)` with `n_profiles = 4`.
+Alternatively, we could use list-indexing within R to index only that model.
+Note that the object `models_4_5` below includes the four and five-class estimated models, in that order.
+Using `[[1]]` indexes---selects---only the first of the two models, whereas `[[2]]` indexes the second.
+We index the four-class solution using `[[1]]`, and save it to a new object.
+
+```{r}
+model_4 <- models_4_5[[1]]
+```
+
+Using this single estimated model, we can also create a faceted plot of density plots for an object of class 'tidyLPA'.
+For each variable, a Total density plot will be shown, where cases are weighted by the posterior probability of being assigned to that class:
+
+```{r}
+model_4 %>% 
+  plot_profiles(add_line = TRUE)
+```
+
+## Visualizing and Examining a Chosen Solution Further
+Last, we can create a faceted plot of two-dimensional correlation plots and unidimensional density plots for a single set of model estimates:
+```{r}
+model_4 %>% 
+  plot_bivariate()
+```
+We can use `get_data()` to return data in wide format when applied to an object
+of class tidyProfile (one element of a tidyLPA object)
+
+`get_data()` returns long format when applied to a
+tidyLPA object containing multiple tidyProfile analyses (because then the wide
+format does not make sense).
+
+To transform data in the wide format into the long format, the `gather()`
+function from the **tidyr** package can be used:
+
+``` {r, process-data, eval = TRUE, echo = TRUE}
+get_data(model_4) %>% 
+  tidyr::gather(Class_prob, Probability, contains("CPROB"))
+```
+
+# Best Practices for Latent Profile Analysis
+
+In addition to the pragmatic challenges facing the analyst carrying out LPA, there 
+are a number of statistical issues that anaylsts face. Past research is informative in
+this respect; Bauer
+
+## Recommendations Regarding Analytic Choices
+
+A note of caution is warranted about LPA in the context of their characteristics
+and strengths. Bauer (2007) notes that many samples of data can be usefully
+broken down into profiles, and that the addition of profiles will likely be
+suggested for reasons other than the samples coming from more than one
+distribution (i.e., due to non-normality in the variables measured). Bauer also
+cautions that profiles should not be reified; that profiles do not necessarily
+exist outside of the analysis that they should be interpreted more as useful
+interpretative devices. These cautions suggest that, in general, parsimony,
+interpretability, and a general sense that the profiles are not necessarily
+real, but are rather helpful analytic tools, should be both priorities for the
+analyst and the reader of studies using this approach.
+
+The GRoLTs checklist, developed for a particular use of the general mixture
+model---a growth mixture model---is a helpful way to think about how best
+practices for LPA [@van2017grolts]. We interpret them in light of LPA; note that
+the full checklist is not applicable (as, for example, guidelines about
+trajectories over time are not pertinent to LPA), and we only present those that
+are.
+
+First, here is the checklist:
+
+## GRoLTs checklist
+
+``` {r, echo = FALSE}
+t <- readr::read_csv("GRoLTS-applicability-to-LPA.csv")
+
+tp <- t %>% filter(use == 1) %>% mutate(`LPA Number` = 1:nrow(.)) %>%
+  select( `LPA Number`, `GRoLTS Number` = Number, `Checklist Item`)
+
+tp %>% DT::datatable()
+```
+
+We walk through each of these in turn.
+
+### Is the missing data mechanism reported?
+
+As missing data can have a bearing on the estimates, an important consideration is whether or not the missing data is associated with the levels of the variables---or not.
+Consider consulting general guidelines on missing data, including @enders2010applied, @little2019statistical, and @allison2001missing.
+### Is a description provided of what variables are related to attrition/missing data?
+
+Note whether any demographic or other variables are related to missing data.
+
+### Is a description provided of how missing data in the analyses were dealt with?
+
+Because LPA cannot accomodate any missing values (and so uses listwise deletion) for this recommendation, tidyLPA has a function for addressing missing data that are based upon single imputation (with two methods available, one using a contemporary, machine learning-based procedure).
+
+### Is information about the distribution of the observed variables included?
+
+The variables can be plotted and described prior to the analyses.
+While not built-in, we show later how this can be done using the ggplot2 R package, and the `describe()` function from the psych R package.
+
+### Is the software mentioned?
+
+The name and version of the software can be mentioned.
+Within R, a citation including the version can be generated for any R package, including tidyLPA, by using the `citation()` function (e.g., `citation("tidyLPA")).
+
+### Are alternative specifications of the between-class differences in variance–covariance matrix structure considered and clearly documented?
+
+In LPA, we can determine which parameters - variance and covariance - are
+estimated to be the same or different across all profiles, and, in the case of
+covariance, whether it is fixed to zero. Of course, for the profiles we are most
+interested in, the mean is allowed to vary across profiles.
+
+In general, the approach to choosing the model is similar to choosing the number
+of profiles, requiring **deciding on the basis of evidence from multiple
+sources**, including information criteria, statistical tests, and concerns of
+interpretability and parsimony. The article by [Pastor and colleagues
+(2007)](https://www.sciencedirect.com/science/article/pii/S0361476X06000543) has
+helpful information on the model specifications.
+
+Here, the six models that are possible to specify in LPA are described in terms of how the variables used to create the profiles are estimated. In tidyLPA, the models are specified by passing
+arguments to the `variance` and `covariance` arguments. The possible values for
+these arguments are:
 
@@
 
-In general, the approach to choosing the model is similar to choosing the number of profiles, requiring **deciding on the basis of evidence from multiple sources**, including information criteria, statistical tests, and concerns of interpretability and parsimony.
-The article by [Pastor and colleagues (2007)](https://www.sciencedirect.com/science/article/pii/S0361476X06000543) has helpful information on the model specifications.
-Here, the six models that are possible to specify in LPA are described in terms of how the variables used to create the profiles are estimated.
+These models are described in Appendix A.
+
+### Is information reported about the number of random start values and final iterations included?
+
+MPlus uses random starts to initialize the estimation, whereas mclust (the R package) does not.
+Thus, for users of MPlus, the number of random starts and final iterations should be reported.
+By default, tidyLPA uses the MPlus defaults.
+Information on the number of random starts can be found in the model object returned by the tidyLPA function `estimate_profiles()`.
+Different random starts can be passed by passing arguments to the `ANALYSIS` code for MPlus, e.g.:
+
+```{r, eval = FALSE}
+estimate_profiles(3, package = "Mplus", 
+                  ANALYSIS = "starts = 100, 20;")
+```
+
+### Are the model comparison (and selection) tools described from a statistical perspective?
+
+In the case of choosing the number of profiles (and the specification of the
+model / profile solution), multiple criteria, such as the BIC or the proportion
+of variance explained are recommended for decision-making, but also
+interpretability in light of theory, parsimony, and evidence from
+cross-validation should be considered.
+
+We provide a number of functions for model comparison. The `compare_solutions()` profile
+takes as an argument a particular measure of fit (defaulting to the BIC), and returns information
+for all of the models estimated. It also implements an ensemble, analytic hierarchiy approach, that uses weighted values of a number of fit indices; see [@akogul2017approach] for more.
+
+### Are the total number of fitted models reported, including a one-class solution?
+
+The number of fitted models can easily be reported based on the tidyLPA code run.
+For example, for the following code, models with 1-6 profiles, and the first (class invariant) and third (class invariant, unrestricted) model parameterizations are estimated:
+
+```{r, eval = FALSE}
+estimate_profiles(n_profiles = 1:6, 
+                  models = c(1, 3))
+```
+
+### Are the number of cases per class reported for each model (absolute sample size, or proportion)?
+
+Determining the number of cases per class can easily be done by using the `get_data()` function from tidyLPA, passing the estimated model object from tidyLPA.; see the `Class` column (as well as the posterior probability estimates for every class for every observation).
+
+### Are all of the estimates plotted?
+
+This is a recommendation which can be difficult to achieve using existing software. tidyLPA provides two functions, `plot_profiles()` and `plot_bivariate()`, to plot the estimates.
+
+### Are the raw data plotted along with the estimates?
+
+The above-mentioned tidyLPA functions also plot the raw data with the estimates (by default).
+
+### Are characteristics of the final class solution numerically described (i.e., means, SD/SE, n, CI, etc.)?
+
+These characteristics can be obtained by using the `get_estimates()` function, passing the estimated model object from tidyLPA.
+
+### Are the syntax files available (either in the appendix, supplementary materials, or from the authors)?
+
+The R files used for the analyses can easily be included in online repositories and supplementary online materials (as `.R` or `.Rmd` files).
+
+# Summary and Conclusion
+
+A step away from the tutorial side of this work, we wish to note some
+implications for the developers of statistical software for psychological
+scientists.
+
+First, we note that developing a tool that emphasized both rigor and access has
+been, so far, well-received. This suggests that there is an audience (or a
+market) for tools to carry out sophisticated tools that may be otherwise
+out-of-reach. Tools do not need to be hard to use, a point made by developers
+of, for example, tools to carry out highly-sophisticated Bayesian methods using
+the most cutting-edge tools [@burkner2017, @makowski2019]. We think that
+tidyLPA speaks to this, too.
+
+Another implication concerns how we developed tidyLPA. We used a modern software
+development workflow, including git/GitHub, a suite of tests, and continuous
+integration tools. We also note that we emphasized communicating that we were
+developing the software; when it was updated; and how we made major updates.
+
+There are a number of analytic choices that need to be made when carrying out
+person-oriented analyses. Because such person-oriented approaches are often more
+subjective (in practice) than other approaches (Linnenbrink-Garcia and
+Wormington, 2017), there is no one rule for determining the solution obtained.
+This solution is obtained on the basis of multiple decisions, such as the number
+of profiles selected or the modeling decisions such as what specific options are
+used for the cluster analysis (i.e., the distance metric used to calculate the
+similarity of the observations as part of the Ward’s hierarchical clustering) or
+what parameters are estimated and how as part of LPA.
+
+Given the subjectivity involved, it is important that researchers be transparent
+and work as part of a team to obtain clustering solutions. Transparency about
+the design and analytic choices is important so that readers can appropriately
+interpret the report. Researchers can enhance transparency and reproducibility
+by sharing detailed descriptions of methodology and document it through the use
+of syntax (and, if possible, data) that we share with others. Working as part of
+a team can help to serve as a check on several of the choices researchers make,
+such as over-fitting or under-fitting the model to the data. Each decision
+depends on multiple factors and balancing tensions. We discuss each of the key
+decisions listed in an analysis.
+
+LPA can be used to evaluate how psychological constructs are experienced (and
+can be analyzed) together and at once. Though described in contrast to a
+variable-centered approach, scholars have pointed out how person-oriented
+approaches, including LPA, are complementary to variable-centered analyses
+(Marsh, Ludtke, Trautwein, & Morin, 2009). Our aim was to describe how
+researchers may get started in an informed way as researchers seek to understand
+how individuals interact, behave, and learn in ways that embraces the complexity
+of these experiences.
+
+# Appendix A
 
@@
 
-If no values are specified for these, then the variances are constrained to be
-equal across classes, and covariances are fixed to 0 (conditional independence
-of the indicators).
-
-These arguments allow for four models to be specified:
-
-- Equal variances and covariances fixed to 0 (Model 1)
-- Varying variances and covariances fixed to 0 (Model 2)
-- Equal variances and equal covariances (Model 3)
-- Varying variances and varying covariances (Model 6)
-
-Two additional models (Models 4 and 5) can be fit using MPlus. More information
-on the models can be found in the
-[vignette](https://data-edu.github.io/tidyLPA/articles/Introduction_to_tidyLPA.html).
-
-## Accessing values in the output
-
-A few helper functions are available to make it easier to work with the output
-of an analysis.
-
-`get_data()` returns the data:
-
-``` {r, get-data-for-solution, eval = FALSE}
-get_data(compare_solutions_mclust[[32]])
-```
-
-We note that `get_data()` returns data in wide format when applied to an object
-of class tidyProfile (one element of a tidyLPA object), or when applied to a
-tidyLPA object of length one. `get_data()` returns long format when applied to a
-tidyLPA object containing multiple tidyProfile analyses (because then the wide
-format does not make sense).
-
-To transform data in the wide format into the long format, the `gather()`
-function from the **tidyr** package can be used, e.g.:
-
-``` {r, process-data, eval = FALSE}
-get_data(compare_solutions_mclust[[32]]) %>% 
-  tidyr::gather(Class_prob, Probability, contains("CPROB"))
-```
-
-# Recommendations Regarding Analytic Choices
-
-A note of caution is warranted about LPA in the context of their characteristics
-and strengths. Bauer (2007) notes that many samples of data can be usefully
-broken down into profiles, and that the addition of profiles will likely be
-suggested for reasons other than the samples coming from more than one
-distribution (i.e., due to non-normality in the variables measured). Bauer also
-cautions that profiles should not be reified; that profiles do not necessarily
-exist outside of the analysis that they should be interpreted more as useful
-interpretative devices. These cautions suggest that, in general, parsimony,
-interpretability, and a general sense that the profiles are not necessarily
-real, but are rather helpful analytic tools, should be both priorities for the
-analyst and the reader of studies using this approach.
-
-The GRoLTs checklist is a helpful way to think about how best practices for LPA;
-here, we show how to implement these using tidyLPA. First, here is the checklist:
-
-## GRoLTs checklist
-
-```{r, echo = FALSE}
-t <- readr::read_csv("GRoLTS-applicability-to-LPA.csv")
-
-tp <- t %>% 
-  filter(use == 1) %>% 
-  mutate(`LPA Number` = 1:nrow(.)) %>% 
-  select(`GRoLTS Number` = Number, `LPA Number`, `Checklist Item`)
-
-tp %>% 
-  DT::datatable()
-```
-
-We walk through each of these in turn.
-
-### Is the missing data mechanism reported?
-
-### Is a description provided of what variables are related to attrition/missing data?
-
-### Is a description provided of how missing data in the analyses were dealt with?
-
-### Is information about the distribution of the observed variables included?
-
-### Is the software mentioned?
-
-### Are alternative specifications of the between-class differences in variance–covariance matrix structure considered and clearly documented?
-
-In LPA, we can determine which parameters - variance
-and covariance - are estimated to be the same or different across all profiles,
-and, in the case of covariance, whether it is fixed to zero. Of course, for the
-profiles we are most interested in, the mean is allowed to vary across profiles.
-The models with more parameters freely-estimated use more degrees of freedom;
-thus, similar to the number of profiles selected, the balance between adding
-more parameters (and having a more complex model) and interpretability and
-parsimony must be balanced. If a more complex model fits better and is
-interpretable and is justifiable in terms of the data collected, then it may be
-preferred to simpler models. If the two models are similar in terms of their
-fit, then the more simple, parsimonious model should be selected.
-
-There are a number of analytic choices that need to be made when carrying out
-person-oriented analyses. Because such person-oriented approaches are often more
-subjective (in practice) than other approaches (Linnenbrink-Garcia and
-Wormington, 2017), there is no one rule for determining the solution obtained.
-This solution is obtained on the basis of multiple decisions, such as the number
-of profiles selected or the modeling decisions such as what specific options are
-used for the cluster analysis (i.e., the distance metric used to calculate the
-similarity of the observations as part of the Ward’s hierarchical clustering) or
-what parameters are estimated and how as part of LPA.
-
-Given the subjectivity involved, it is important that researchers be transparent
-and work as part of a team to obtain clustering solutions. Transparency about
-the design and analytic choices is important so that readers can appropriately
-interpret the report. Researchers can enhance transparency and reproducibility
-by sharing detailed descriptions of methodology and document it through the use
-of syntax (and, if possible, data) that we share with others. Working as part of
-a team can help to serve as a check on several of the choices researchers make,
-such as over-fitting or under-fitting the model to the data. Each decision
-depends on multiple factors and balancing tensions. We discuss each of the key
-decisions listed in an analysis.
-
-### Is information reported about the number of random start values and final iterations included?
-
-### Are the model comparison (and selection) tools described from a statistical perspective?
-
-In the case of choosing the number of profiles (and the specification of the
-model / profile solution), multiple criteria, such as the BIC or the proportion
-of variance explained are recommended for decision-making, but also
-interpretability in light of theory, parsimony, and evidence from
-cross-validation should be considered.
-
-### Are the total number of fitted models reported, including a one-class solution?
-
-### Are the number of cases per class reported for each model (absolute sample size, or proportion)?
-
-### If classification of cases in a trajectory is the goal, is entropy reported?
-
-### Are all of the estimates plotted?
-
-### Are the raw data plotted along with the estimates?
-
-### Are characteristics of the final class solution numerically described (i.e., means, SD/SE, n, CI, etc.)?
-
-### Are the syntax files available (either in the appendix, supplementary materials, or from the authors)?
-
-# Future Directions and Extensions
-
-First, scholars taking a
-person-oriented approach should emphasize reproducibility in carrying out
-analyses. This is in part due to the exploratory and interpretative nature of
-person-oriented approaches. To this end, presenting multiple models, as in
-Pastor et al. (2007), should be encouraged, rather than presenting one solution.
-In addition, having multiple analysts review the solutions found is encouraged.
-As part of this recommendation, we suggest that researchers consider how their
-analyses can be reproduced by other analysts outside of the research team:
-Sharing code and data is an important part of this work. Also related,
-researchers should consider how profiles are replicated across samples.
-
-A second general recommendation considers more flexibly incorporating time, when
-data are collected across multiple time points, into analyses. ISOA groups all
-time points and doesn’t make distinctions. Other approaches perform analysis
-separately at different timepoints (Corpus & Wormington, 2014). Some integrate
-time as a part of the profiles, i.e. growth mixture modeling (groups of
-patterns), i.e. within-person growth modeling, where there are individual growth
-patterns. Research to date has yet to consider additional challenges in applying
-person centered approaches to longitudinal data. For instance, Schmidt et al
-(2018) use an Experience Sampling Method (ESM) approach to collecting data and
-used a person-centered approach to generate profiles of students in science
-class. This work does not account for student-level effects. In other words, it
-did not model the shared variance of multiple observations of the same student.
-
-Third, best practices in within-person or longitudinal research call for
-modeling the nesting structure. However, researchers have yet to successfully
-incorporate this practice into the person centered approach. One way to approach
-this is to use cross-classified mixed effects models, as in Strati, Schmidt, and
-Maier (2017) and in Rosenberg (2018). In such approaches, dependencies in terms
-of, for example, individuals responding to (ESM) surveys at the same time and
-repeated responses being associated with the same individuals can both be
-modeled, although the effects of these two sources of dependencies are not
-nested as in very common uses of multi-level models, but rather are
-cross-classified. West, Welch, & Galecki (2014) have a description of the use of
-multi-level models with cross-classified data and tools (including those freely
-available through R) that can be used to estimate them.
-
-# Implications
-
-A step away from the tutorial side of this work, we wish to note some implications 
-for the developers of statistical software for psychological scientists.
-
-First, we note that developing a tool that emphasized both rigor and access has
-been, so far, well-received. This suggests that there is an audience (or a
-market) for tools to carry out sophisticated tools that may be otherwise
-out-of-reach. Tools do not need to be hard to use, a point made by developers
-of, for example, tools to carry out highly-sophisticated Bayesian methods using
-the most cutting-edge tools [@burkner2017, @makowski2019]. We think
-that tidyLPA speaks to this, too.
-
-Another implication concerns how we developed tidyLPA. We used a modern software
-development workflow, including git/GitHub, a suite of tests, and continuous
-integration tools. We also note that we emphasized communicating that we were
-developing the software; when it was updated; and how we made major updates.
-
-# Conclusion
-
-Person-oriented analysis ia way to consider how psychological constructs are
-experienced (and can be analyzed) together and at once. Though described in
-contrast to a variable-centered approach, scholars have pointed out how
-person-oriented approaches are complementary to variable-centered analyses
-(Marsh, Ludtke, Trautwein, & Morin, 2009). A person-oriented approach can help
-us to consider multiple variables together and at once and in a more dynamic
-way, reflected in the methodological approaches for cluster analysis and LPA
-that identify profiles of individuals responses.
-
-This manuscript provided an outline of how to get started with person-oriented
-analyses in an informed way. We provided a general overview of the methodology
-and described tools to carry out such an analysis. We also described specific
-tools, emphasizing freely-available open-source options that we have developed.
-Because of the inherently exploratory nature of person-oriented analysis,
-carrying out the analysis in a trustworthy and open way is particularly
-important. In this way, the interpretative aspect of settling on a solution
-shares some features of quantitative and qualitative research: The systematic
-nature of quantitative research methods (focused upon agreed-upon criteria such
-as likelihood-ratio tests) and qualitative research methods (focused upon the
-trustworthiness of both the analysis and the analyst) are important to consider
-when carrying out person-oriented analysis. Lastly, we made some general
-recommendations for future directions--and also highlighted some situations for
-which person-oriented approaches may not be the best and some cautions raised in
-past research regarding how such approaches are used.
-
-In conclusion, as use of person-oriented approaches expand, new questions and
-opportunities for carrying out research in a more holistic, dynamic way will be
-presented. Analyzing constructs together and at once is appealing to
-researchers, particularly those carrying out research in fields such as
-education for which communicating findings to stakeholders in a way that has the
-chance to impact practice is important. Our aim was not to suggest that such an
-approach is always the goal or should always be carried out, but rather to
-describe how researchers may get started in an informed way as researchers seek
-to understand how individuals interact, behave, and learn in ways that embraces
-the complexity of these experiences.
-
-\newpage
-
 # References

### 2020-12-29 paper.tex
re-run analysis from top-to-bottom

@@
 \vspace{0.5cm}
-\textsuperscript{1} University of Tennessee, Knoxville\\\textsuperscript{2} Utrecht University\\\textsuperscript{3} Michigan State University\\\textsuperscript{4} University of Oregon\\\textsuperscript{5} }
-\keywords{Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust\newline\indent Word count: }
+\textsuperscript{1} University of Tennessee, Knoxville\\\textsuperscript{2} Utrecht University\\\textsuperscript{3} Michigan State University\\\textsuperscript{3} University of Oregon\\\textsuperscript{4} University of Wisconsin, Madison}
+\keywords{Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust\newline\indent Word count: suggested to be no more than 3,000}
 \usepackage{csquotes}
@@
 \title{Decisions Decisions Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package}
-\author{Joshua Rosenberg\textsuperscript{1}, Caspar van Lissa\textsuperscript{2}, Jennifer Schmidt\textsuperscript{3}, Patrick Beymer\textsuperscript{3}, Daniel Anderson\textsuperscript{4}, \& Matthew Schell\textsuperscript{3}}
+\author{Joshua Rosenberg\textsuperscript{1}, Caspar van Lissa\textsuperscript{2}, Jennifer Schmidt\textsuperscript{3}, Patrick Beymer\textsuperscript{5}, Daniel Anderson\textsuperscript{4}, \& Matthew Schell\textsuperscript{3}}
 \date{}
@@
 
-Students, may (reasonably) ask, though, are grades \emph{really} normally distributed?
+Is there one distribution?
 
-Are their one distribution?
+\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
 
-Or two?
+Or, are there two?
 
-This kind of distribution (that is \emph{bimodal}) is not exclusive to grades. For
-teachers, psychologists, researchers, and friends and family members, people are
-highly-complex and not easily able to be characterized by \emph{one} characteristic
-or personality trait--and its distribution. In the social sciences, broadly, and
-in psychology, particularly, a statistical method that can be used to describe
-how people, in their individual particularities, may have similarities on the
-basis of some set of measures through which they can be grouped in meaningful,
-distinctive ways. This approach, which has a provenance in developmental (or
+\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}
+
+This kind of distribution (one that is \emph{bimodal}) is not exclusive to grades.
+For teachers, psychologists, researchers, and even friends and family members,
+people are highly-complex and not easily able to be characterized by \emph{one}
+characteristic or personality trait--and its distribution.
+
+In the social sciences, broadly, and in the psychological sciences, in
+particular, a statistical method that can be used to describe how people, in
+their individual particularities, may have similarities on the basis of some set
+of measures through which they can be grouped in meaningful, distinctive ways.
+
+This approach, which we view as having a provenance in developmental (or
 \emph{person-oriented}) approaches (Bergman \& El-Khouri, 1997; Magnusson \&
@@
 
-In this tutorial, we aim to describe one of the most commonly-used--and relevant
-to psychologists---application of the general mixture model, to cases for which
-all of the variables for which (relatively) homogeneous groups are identified
-from among a (relatively) heterogeneous sample are continuos, \emph{latent profile
-analysis} (LPA). We note that scholars using the approach, especially in
-developmental and educational psychology, use cluster analytic methods. Such an
-approach has some similarities with LPA (CITE paper showing how k-means is
-similar to the simplest model specification of LPA?), though LPA also has a
-number of benefits (Linnenbrink-Garcia \& Wormington, 2017), and we we focus on
-LPA in this paper.
+In this tutorial, we aim to describe one of the most commonly-used---and
+relevant to psychologists---application of the general mixture model, to cases
+for which all of the variables for which (relatively) homogeneous groups are
+identified from among a (relatively) heterogeneous sample are continuos, \emph{latent
+profile analysis} (LPA). After describing the method and some examples of its
+use, we provide a tutorial for carrying out LPA in the context of a
+freely-available, open-source statistical software package that we have
+developed and supported over the past three years for R (R Core Team, 2019),
+tidyLPA. Finally, we offer some ideas about best practices and informed
+recommendations for researchers aiming to use LPA in their applied work, and
+conclude with reflections on the role of statistical software--especially
+software that is freely-available, open-source, and highly-performant--in the
+psychological sciences.
 
-After describing the method and some examples of its use, we provide a tutorial
-for carrying out LPA in the context of a freely-available, open-source
-statistical software package we created for R (R Core Team, 2019), tidyLPA.
-Finally, we offer some ideas about best practices and informed recommendations
-for researchers aiming to use LPA in their applied work, and conclude with
-reflections on the role of statistical software--especially software that is
-freely-available, open-source, and highly-performant--in the psychological
-sciences.
-
-\hypertarget{latent-profile-analysis}{%
-\section{Latent Profile Analysis}\label{latent-profile-analysis}}
+\hypertarget{latent-profile-analysis-and-the-need-for-efficient-and-reproducible-analyses}{%
+\section{Latent Profile Analysis and the Need for Efficient and Reproducible Analyses}\label{latent-profile-analysis-and-the-need-for-efficient-and-reproducible-analyses}}
 
@@
 which can be freely-estimated, estimated to be equal, or fixed to be zero.
-
-We note that challenge facing the analyst using LPA is that these
-parameters and the distinct model parameterizations that can be estimated is the
-different terminology used. As one example, Scrucca et al. (2017) refer to these
-parameterizations not in terms of whether and how parameters are estimated, but
-rather in terms of the geometric properties of the distributions that result
-from particular parameterizations. Muthen and Muthen (1997-2017) and others
-(Pastor et al., 2007) commonly refer to local independence to mean that the
-covariances are fixed to zero (also described as the specification of the
-covariance matrix as \enquote{diagonal,} because only the diagonal components, or the
-variances, are estimated).
 
@@
 decrease the degrees of freedom), we are more likely to fit the data, but less
-likely to be able to replicate the model with a second set of data. In other
-words, more parameters may mean a loss of external validity. As we progress
-toward more complex models (with increasingly complex parameterization), then we
-are more likely to fit the data better. More information on the model
-parameterizations are discussed in the context of the software tool tidyLPA that
-we have developed.
+likely to be able to replicate the model with a second set of data. As we
+progress toward more complex models (with increasingly complex
+parameterization), then we are more likely to fit the data better.
+In all, this flexibility associated with LPA also has a more pragmatic cost. It
+can be very difficult to efficiently---and reproducibly---estimate and compare a
+number of models. This cost means that analysts may focus on a subset of models.
+It may also mean that unintentional errors are introduced through copying and
+pasting the output of model estimations. Part of why we developed tidyLPA was to
+make the model estimation process more efficient and more reproducible. As we
+describe later, we do this using R, which has a number of benefits, including
+being open-source, powerful (in terms of pre-processing and using the output of
+models), and increasingly widely-used in the psychological sciences. We also
+provide an interface to perhaps the most widely-used software for LPA, MPlus. In
+this way, tidyLPA can both allow researchers to estimate models they already
+estimate using MPlus more efficiently and reproducibly, and, for those without
+access to this software, to do so entirely using R and the open-source R package
+mclust.
 
-\hypertarget{best-practices-for-latent-profile-analysis}{%
-\section{Best Practices for Latent Profile Analysis}\label{best-practices-for-latent-profile-analysis}}
+\hypertarget{existing-software-options}{%
+\section{Existing Software Options}\label{existing-software-options}}
 
-\hypertarget{recommendations-regarding-analytic-choices}{%
-\subsection{Recommendations Regarding Analytic Choices}\label{recommendations-regarding-analytic-choices}}
-
-A note of caution is warranted about LPA in the context of their characteristics
-and strengths. Bauer (2007) notes that many samples of data can be usefully
-broken down into profiles, and that the addition of profiles will likely be
-suggested for reasons other than the samples coming from more than one
-distribution (i.e., due to non-normality in the variables measured). Bauer also
-cautions that profiles should not be reified; that profiles do not necessarily
-exist outside of the analysis that they should be interpreted more as useful
-interpretative devices. These cautions suggest that, in general, parsimony,
-interpretability, and a general sense that the profiles are not necessarily
-real, but are rather helpful analytic tools, should be both priorities for the
-analyst and the reader of studies using this approach.
-
-The GRoLTs checklist is a helpful way to think about how best practices for LPA;
-here, we show how to implement these using tidyLPA. First, here is the checklist:
-
-\hypertarget{grolts-checklist}{%
-\subsection{GRoLTs checklist}\label{grolts-checklist}}
-
-\hypertarget{htmlwidget-6725a3bb4d9d4a47b614}{}
-
-We walk through each of these in turn.
-
-\hypertarget{is-the-missing-data-mechanism-reported}{%
-\subsubsection{Is the missing data mechanism reported?}\label{is-the-missing-data-mechanism-reported}}
-
-\hypertarget{is-a-description-provided-of-what-variables-are-related-to-attritionmissing-data}{%
-\subsubsection{Is a description provided of what variables are related to attrition/missing data?}\label{is-a-description-provided-of-what-variables-are-related-to-attritionmissing-data}}
-
-\hypertarget{is-a-description-provided-of-how-missing-data-in-the-analyses-were-dealt-with}{%
-\subsubsection{Is a description provided of how missing data in the analyses were dealt with?}\label{is-a-description-provided-of-how-missing-data-in-the-analyses-were-dealt-with}}
-
-\hypertarget{is-information-about-the-distribution-of-the-observed-variables-included}{%
-\subsubsection{Is information about the distribution of the observed variables included?}\label{is-information-about-the-distribution-of-the-observed-variables-included}}
-
-\hypertarget{is-the-software-mentioned}{%
-\subsubsection{Is the software mentioned?}\label{is-the-software-mentioned}}
-
-\hypertarget{are-alternative-specifications-of-the-between-class-differences-in-variancecovariance-matrix-structure-considered-and-clearly-documented}{%
-\subsubsection{Are alternative specifications of the between-class differences in variance--covariance matrix structure considered and clearly documented?}\label{are-alternative-specifications-of-the-between-class-differences-in-variancecovariance-matrix-structure-considered-and-clearly-documented}}
-
-In LPA, we can determine which parameters - variance
-and covariance - are estimated to be the same or different across all profiles,
-and, in the case of covariance, whether it is fixed to zero. Of course, for the
-profiles we are most interested in, the mean is allowed to vary across profiles.
-The models with more parameters freely-estimated use more degrees of freedom;
-thus, similar to the number of profiles selected, the balance between adding
-more parameters (and having a more complex model) and interpretability and
-parsimony must be balanced. If a more complex model fits better and is
-interpretable and is justifiable in terms of the data collected, then it may be
-preferred to simpler models. If the two models are similar in terms of their
-fit, then the more simple, parsimonious model should be selected.
-
-There are a number of analytic choices that need to be made when carrying out
-person-oriented analyses. Because such person-oriented approaches are often more
-subjective (in practice) than other approaches (Linnenbrink-Garcia and
-Wormington, 2017), there is no one rule for determining the solution obtained.
-This solution is obtained on the basis of multiple decisions, such as the number
-of profiles selected or the modeling decisions such as what specific options are
-used for the cluster analysis (i.e., the distance metric used to calculate the
-similarity of the observations as part of the Ward's hierarchical clustering) or
-what parameters are estimated and how as part of LPA.
-
-Given the subjectivity involved, it is important that researchers be transparent
-and work as part of a team to obtain clustering solutions. Transparency about
-the design and analytic choices is important so that readers can appropriately
-interpret the report. Researchers can enhance transparency and reproducibility
-by sharing detailed descriptions of methodology and document it through the use
-of syntax (and, if possible, data) that we share with others. Working as part of
-a team can help to serve as a check on several of the choices researchers make,
-such as over-fitting or under-fitting the model to the data. Each decision
-depends on multiple factors and balancing tensions. We discuss each of the key
-decisions listed in an analysis.
-
-\hypertarget{is-information-reported-about-the-number-of-random-start-values-and-final-iterations-included}{%
-\subsubsection{Is information reported about the number of random start values and final iterations included?}\label{is-information-reported-about-the-number-of-random-start-values-and-final-iterations-included}}
-
-\hypertarget{are-the-model-comparison-and-selection-tools-described-from-a-statistical-perspective}{%
-\subsubsection{Are the model comparison (and selection) tools described from a statistical perspective?}\label{are-the-model-comparison-and-selection-tools-described-from-a-statistical-perspective}}
-
-In the case of choosing the number of profiles (and the specification of the
-model / profile solution), multiple criteria, such as the BIC or the proportion
-of variance explained are recommended for decision-making, but also
-interpretability in light of theory, parsimony, and evidence from
-cross-validation should be considered.
-
-\hypertarget{are-the-total-number-of-fitted-models-reported-including-a-one-class-solution}{%
-\subsubsection{Are the total number of fitted models reported, including a one-class solution?}\label{are-the-total-number-of-fitted-models-reported-including-a-one-class-solution}}
-
-\hypertarget{are-the-number-of-cases-per-class-reported-for-each-model-absolute-sample-size-or-proportion}{%
-\subsubsection{Are the number of cases per class reported for each model (absolute sample size, or proportion)?}\label{are-the-number-of-cases-per-class-reported-for-each-model-absolute-sample-size-or-proportion}}
-
-\hypertarget{if-classification-of-cases-in-a-trajectory-is-the-goal-is-entropy-reported}{%
-\subsubsection{If classification of cases in a trajectory is the goal, is entropy reported?}\label{if-classification-of-cases-in-a-trajectory-is-the-goal-is-entropy-reported}}
-
-\hypertarget{are-all-of-the-estimates-plotted}{%
-\subsubsection{Are all of the estimates plotted?}\label{are-all-of-the-estimates-plotted}}
-
-\hypertarget{are-the-raw-data-plotted-along-with-the-estimates}{%
-\subsubsection{Are the raw data plotted along with the estimates?}\label{are-the-raw-data-plotted-along-with-the-estimates}}
-
-\hypertarget{are-characteristics-of-the-final-class-solution-numerically-described-i.e.-means-sdse-n-ci-etc.}{%
-\subsubsection{Are characteristics of the final class solution numerically described (i.e., means, SD/SE, n, CI, etc.)?}\label{are-characteristics-of-the-final-class-solution-numerically-described-i.e.-means-sdse-n-ci-etc.}}
-
-\hypertarget{are-the-syntax-files-available-either-in-the-appendix-supplementary-materials-or-from-the-authors}{%
-\subsubsection{Are the syntax files available (either in the appendix, supplementary materials, or from the authors)?}\label{are-the-syntax-files-available-either-in-the-appendix-supplementary-materials-or-from-the-authors}}
-
-\hypertarget{future-directions-and-extensions}{%
-\section{Future Directions and Extensions}\label{future-directions-and-extensions}}
-
-First, scholars taking a
-person-oriented approach should emphasize reproducibility in carrying out
-analyses. This is in part due to the exploratory and interpretative nature of
-person-oriented approaches. To this end, presenting multiple models, as in
-Pastor et al. (2007), should be encouraged, rather than presenting one solution.
-In addition, having multiple analysts review the solutions found is encouraged.
-As part of this recommendation, we suggest that researchers consider how their
-analyses can be reproduced by other analysts outside of the research team:
-Sharing code and data is an important part of this work. Also related,
-researchers should consider how profiles are replicated across samples.
-
-A second general recommendation considers more flexibly incorporating time, when
-data are collected across multiple time points, into analyses. ISOA groups all
-time points and doesn't make distinctions. Other approaches perform analysis
-separately at different timepoints (Corpus \& Wormington, 2014). Some integrate
-time as a part of the profiles, i.e.~growth mixture modeling (groups of
-patterns), i.e.~within-person growth modeling, where there are individual growth
-patterns. Research to date has yet to consider additional challenges in applying
-person centered approaches to longitudinal data. For instance, Schmidt et al
-(2018) use an Experience Sampling Method (ESM) approach to collecting data and
-used a person-centered approach to generate profiles of students in science
-class. This work does not account for student-level effects. In other words, it
-did not model the shared variance of multiple observations of the same student.
-
-Third, best practices in within-person or longitudinal research call for
-modeling the nesting structure. However, researchers have yet to successfully
-incorporate this practice into the person centered approach. One way to approach
-this is to use cross-classified mixed effects models, as in Strati, Schmidt, and
-Maier (2017) and in Rosenberg (2018). In such approaches, dependencies in terms
-of, for example, individuals responding to (ESM) surveys at the same time and
-repeated responses being associated with the same individuals can both be
-modeled, although the effects of these two sources of dependencies are not
-nested as in very common uses of multi-level models, but rather are
-cross-classified. West, Welch, \& Galecki (2014) have a description of the use of
-multi-level models with cross-classified data and tools (including those freely
-available through R) that can be used to estimate them.
-
-\hypertarget{tidylpa}{%
-\section{tidyLPA}\label{tidylpa}}
-
-\hypertarget{tools-other-than-tidylpa}{%
-\subsection{Tools other than tidyLPA}\label{tools-other-than-tidylpa}}
+There are a number of software options available that provide context for why we
+developed tidyLPA.
 
@@
 RMixMod, which uses compiled C++ code) and others in terms of ease-of-use (i.e.,
-the plot methods built-in to RMixMod, mclust, and other tools). However, they
-also have some drawbacks, in that it can be difficult to translate between the
-model specifications, which are often described in terms of the geometric
+the plot methods built-in to RMixMod, mclust, and other tools).
+
+However, these existing R package also have some drawbacks in addition to positive features, in that it can be difficult to translate between the model specifications, which are often described in terms of the geometric
 properties of the multivariate distributions being estimated (i.e., \enquote{spherical,
@@
 algorithm, which can make comparing results across tools challenging.
-
-\hypertarget{the-tidylpa-r-package}{%
-\subsection{The tidyLPA R package}\label{the-tidylpa-r-package}}
 
@@
 same maximum likelihood estimation procedure, they are benchmarked to produce
-the same output. Also, note that we have described the model
-specifications with descriptions of what is estimated in terms of the variances
-and covariances, the common names for the models (i.e., class-varying
-unrestricted), and the covariance matrix associated with the parameterizationfor
-the six models that are possible to be estimated on the website for tidyLPA (see
-here).
+the same output. Also, note that we have described the model specifications with
+descriptions of what is estimated in terms of the variances and covariances, the
+common names for the models (i.e., class-varying unrestricted), and the
+covariance matrix associated with the parameterizationfor the six models that
+are possible to be estimated.
+
+\hypertarget{a-walkthrough-using-the-tidylpa-r-package}{%
+\section{A Walkthrough Using the tidyLPA R package}\label{a-walkthrough-using-the-tidylpa-r-package}}
+
+Here, we provide a walkthrough of an analysis using the tidyLPA R package.
 
@@
 
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{install.packages}\NormalTok{(}\StringTok{"tidyLPA"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
 This is how we recommend most users install tidyLPA.
 
-You can also install the development version of tidyLPA, which may have new features that are not yet availalble on CRAN (because the version on CRAN is periodically updated) from GitHub with:
+You can also install the development version of tidyLPA, which may have new
+features that are not yet availalble on CRAN (because the version on CRAN is
+periodically updated) from GitHub with:
 
-\hypertarget{examples}{%
-\subsection{Examples}\label{examples}}
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{# install.packages("devtools")}
+\NormalTok{devtools}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"data-edu/tidyLPA"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
-Here, we use a simulated dataset contains data on broad interest, enjoyment, and self-efficacy from the PISA assessment.
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(tidyLPA)}
+\KeywordTok{library}\NormalTok{(dplyr) }\CommentTok{# for preparing the data}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{plotting-and-describing-the-data}{%
+\subsection{Plotting and Describing the Data}\label{plotting-and-describing-the-data}}
+
+Here, we use a simulated dataset contains data on broad (general, and not situation- or topic-specific) interest, sense of enjoyment, and
+self-efficacy from the PISA assessment.
 
@@
 
-\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1.pdf}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{pisaUSA15 }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(broad_interest, enjoyment, self_efficacy) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\NormalTok{tidyr}\OperatorTok{::}\KeywordTok{gather}\NormalTok{(key, val) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ val, }\DataTypeTok{group =}\NormalTok{ key, }\DataTypeTok{fill =}\NormalTok{ key)) }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{geom_density}\NormalTok{() }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{key) }\OperatorTok{+}\StringTok{ }
+\StringTok{  }\KeywordTok{theme_bw}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
 
-Here, we use the \{psych\} package to explore the distributional properties of the variables:
+\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1.pdf}
 
-\begin{tabular}{l|r|r|r|r|r|r|r|r|r|r|r|r|r}
-\hline
-  & vars & n & mean & sd & median & trimmed & mad & min & max & range & skew & kurtosis & se\\
-\hline
-broad\_interest & 1 & 5454 & 2.652246 & 0.7772383 & 2.8 & 2.683902 & 0.59304 & 1 & 5 & 4 & -0.2767312 & -0.0261463 & 0.0105244\\
-\hline
-enjoyment & 2 & 5520 & 2.780263 & 0.7509386 & 3.0 & 2.819973 & 0.59304 & 1 & 4 & 3 & -0.5076839 & 0.1652595 & 0.0101073\\
-\hline
-self\_efficacy & 3 & 5409 & 2.136430 & 0.6729143 & 2.0 & 2.112135 & 0.63540 & 1 & 4 & 3 & 0.4376698 & 0.0705830 & 0.0091496\\
-\hline
-\end{tabular}
+We see that the data broadly \emph{appears} to be normally distributed, as the following statistics from the use of the psych package suggest:
 
-Add one or two sentences here on the dataset?
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{pisaUSA15 }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(broad_interest, enjoyment, self_efficacy) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\NormalTok{psych}\OperatorTok{::}\KeywordTok{describe}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
 
-\hypertarget{mclust-via-tidylpa}{%
-\subsubsection{Mclust via tidyLPA}\label{mclust-via-tidylpa}}
+\begin{verbatim}
+##                vars    n mean   sd median trimmed  mad min max range  skew
+## broad_interest    1 5454 2.65 0.78    2.8    2.68 0.59   1   5     4 -0.28
+## enjoyment         2 5520 2.78 0.75    3.0    2.82 0.59   1   4     3 -0.51
+## self_efficacy     3 5409 2.14 0.67    2.0    2.11 0.64   1   4     3  0.44
+##                kurtosis   se
+## broad_interest    -0.03 0.01
+## enjoyment          0.17 0.01
+## self_efficacy      0.07 0.01
+\end{verbatim}
+
+\hypertarget{using-tidylpa-directly-and-exclusively-within-r}{%
+\subsection{Using tidyLPA directly and exclusively within R}\label{using-tidylpa-directly-and-exclusively-within-r}}
+
+First, we show how to use tidyLPA within R---without any need to use or purchase MPlus.
 
@@
 \texttt{compare\_solutions}, \texttt{get\_data}, and \texttt{get\_fit}.
-
-Note that here and throughout this paper we use the function \texttt{knitr::kable()} to
-format the output in an RMarkdown document, like the one we are using; this is
-not necessary to use unless you also wish to format the tidyLPA output in an
-RMarkdown document.
 
@@
 ##  Model Classes AIC      BIC      Entropy prob_min prob_max n_min n_max BLRT_p
-##  1     3       34815.05 34908.15 0.75    0.77     0.94     0.16  0.67  0.01
+##  1     3       34805.16 34898.27 0.74    0.75     0.94     0.16  0.67  0.01
 \end{verbatim}
 
-\hypertarget{mplus-via-tidylpa}{%
-\subsubsection{Mplus via tidyLPA}\label{mplus-via-tidylpa}}
+\hypertarget{using-tidylpa-to-estimate-models-efficiently-and-reproducibly-in-mplus}{%
+\subsection{Using tidyLPA to estimate models efficiently and reproducibly in Mplus}\label{using-tidylpa-to-estimate-models-efficiently-and-reproducibly-in-mplus}}
 
 We can use Mplus simply by changing the package argument for
-\texttt{estimate\_profiles()} to \enquote{mplus} or \enquote{MplusAutomation} (either will work; not run here):
+\texttt{estimate\_profiles()} to \enquote{mplus} or \enquote{MplusAutomation} (either will work; not run
+here):
 
@@
 ##  Model Classes AIC      BIC      Entropy prob_min prob_max n_min n_max BLRT_p
-##  1     3       34804.95 34898.05 0.77    0.79     0.94     0.16  0.67  0.00
+##  1     3       34809.14 34902.24 0.77    0.79     0.94     0.16  0.67  0.00
 \end{verbatim}
 
-These are the very basics; we will return to what we do with this output later in the tutorial.
+These are the very basics; we will return to what we do with this output later
+in the tutorial.
 
-\hypertarget{fit}{%
-\subsubsection{Fit}\label{fit}}
+\hypertarget{determining-the-fit-of-an-estimated-model}{%
+\subsection{Determining the Fit of an Estimated Model}\label{determining-the-fit-of-an-estimated-model}}
 
-We can see a number of statistics for each of the models that were estimated, including the
-AIC (Aikake information criterion; based on -2 * the log-likelihood, and penalized by number of parameters) and BIC (BIC: Bayesian information criterion; based on -2 log-likelihood, and penalized by number of parameters adjusted by sample size).
-Both are \emph{penalized} likelihood statistics; the log-likelihood will always decrease with the addition of additional
-paremeters estimated by the model, whereas the AIC and BIC will account for the number of parameters; lower values of
-either indicate better fit.
+We can see a number of statistics for each of the models that were estimated,
+including the AIC (Aikake information criterion; based on -2 * the
+log-likelihood, and penalized by number of parameters) and BIC (BIC: Bayesian
+information criterion; based on -2 log-likelihood, and penalized by number of
+parameters adjusted by sample size). Both are \emph{penalized} likelihood statistics;
+the log-likelihood will always decrease with the addition of additional
+paremeters estimated by the model, whereas the AIC and BIC will account for the
+number of parameters; lower values of either indicate better fit.
 
-Entropy.
-\texttt{prob\_min} refers to the minimum of the diagonal of the average latent class probabilities for most likely class membership, by assigned class.
-The minimum should be as high as possible, reflecting greater classification certainty (cases are assigned to classes they have a high probability of belonging to; see Jung \& Wickrama, 2008); \texttt{prob\_max}
-refers to the maximum.
-Finally.
-\texttt{n\_min} refers to the proportion of the sample assigned to the smallest class (based on most likely class membership); \texttt{n\_max} to the largest.
-Finally, \texttt{BLRT\_p} refers to the p-value for the bootstrapped likelihood ratio test; a value greater than .05 indicates that the model fit better than one with one fewer class.
-Additional fit statistics (other penalized likelihood values, such as the CAIC, SABIC, and ICL), and the bootstrapped LRT
-test statistic, can be obtained with the \texttt{get\_fit()} function:
+Entropy. \texttt{prob\_min} refers to the minimum of the diagonal of the average latent
+class probabilities for most likely class membership, by assigned class. The
+minimum should be as high as possible, reflecting greater classification
+certainty (cases are assigned to classes they have a high probability of
+belonging to; see Jung \& Wickrama, 2008); \texttt{prob\_max} refers to the maximum.
+Finally. \texttt{n\_min} refers to the proportion of the sample assigned to the smallest
+class (based on most likely class membership); \texttt{n\_max} to the largest. Finally,
+\texttt{BLRT\_p} refers to the p-value for the bootstrapped likelihood ratio test; a
+value greater than .05 indicates that the model fit better than one with one
+fewer class. Additional fit statistics (other penalized likelihood values, such
+as the CAIC, SABIC, and ICL), and the bootstrapped LRT test statistic, can be
+obtained with the \texttt{get\_fit()} function:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{pisaUSA15 }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(broad_interest, enjoyment, self_efficacy) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{single_imputation}\NormalTok{() }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{estimate_profiles}\NormalTok{(}\DecValTok{3}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{get_fit}\NormalTok{(m1)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 ##   <dbl>   <dbl>   <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>   <dbl>
-## 1     1       3 -17386. 34800. 35055. 34893. 34907. 34774. 34817. 34849. -36424.
+## 1     1       3 -17399. 34826. 35080. 34919. 34933. 34799. 34843. 34874. -36442.
 ## # ... with 7 more variables: Entropy <dbl>, prob_min <dbl>, prob_max <dbl>,
@@
 
-\hypertarget{comparing-a-wide-range-of-solutions}{%
-\subsection{Comparing a wide range of solutions}\label{comparing-a-wide-range-of-solutions}}
+\hypertarget{comparing-the-fit-for-many-estimated-models}{%
+\subsection{Comparing the Fit for Many Estimated Models}\label{comparing-the-fit-for-many-estimated-models}}
 
@@
 
-\begin{verbatim}
-## tidyLPA analysis using mclust: 
-## 
-##  Model Classes AIC      BIC      Entropy prob_min prob_max n_min n_max BLRT_p
-##  1     1       37962.61 38002.51 1.00    1.00     1.00     1.00  1.00        
-##  1     2       35495.33 35561.83 0.74    0.83     0.96     0.24  0.76  0.01  
-##  1     3       34850.99 34944.09 0.75    0.76     0.94     0.16  0.67  0.01  
-##  1     4       32990.63 33110.33 0.93    0.94     0.97     0.07  0.54  0.01  
-##  1     5       32652.78 32799.09 0.89    0.79     0.99     0.07  0.46  0.01  
-##  1     6       32631.88 32804.79 0.76    0.44     0.99     0.07  0.34  0.01
-\end{verbatim}
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{many_analyses_}\DecValTok{1}\NormalTok{_}\DecValTok{3}\NormalTok{ <-}\StringTok{ }\NormalTok{pisaUSA15 }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(broad_interest, enjoyment, self_efficacy) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{single_imputation}\NormalTok{() }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{estimate_profiles}\NormalTok{(}\DataTypeTok{models =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{3}\NormalTok{),}
+                    \DataTypeTok{n_profiles =} \DecValTok{1}\OperatorTok{:}\DecValTok{5}\NormalTok{)}
+
+\KeywordTok{compare_solutions}\NormalTok{(many_analyses_}\DecValTok{1}\NormalTok{_}\DecValTok{3}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
@@
 ## 
-##  Model Classes BIC      
-##  1     1       38002.509
-##  1     2       35561.831
-##  1     3       34944.094
-##  1     4       33110.333
-##  1     5       32799.090
-##  1     6       32804.788
+##  Model Classes BIC      Warnings
+##  1     1       38089.12         
+##  1     2       35632.66         
+##  1     3       34991.63         
+##  1     4       33145.73         
+##  1     5       32830.99         
+##  3     1       35262.73         
+##  3     2       34836.77         
+##  3     3       34868.19         
+##  3     4       34600.33 Warning 
+##  3     5       34391.97         
 ## 
@@
 ## 
-## An analytic hierarchy process, based on the fit indices AIC, AWE, BIC, CLC, and KIC (Akogul & Erisoglu, 2017), suggests the best solution is Model 1 with 6 classes.
+## An analytic hierarchy process, based on the fit indices AIC, AWE, BIC, CLC, and KIC (Akogul & Erisoglu, 2017), suggests the best solution is Model 1 with 5 classes.
 \end{verbatim}
 
-\hypertarget{specifying-the-model}{%
-\subsubsection{\texorpdfstring{Specifying the \emph{model}}{Specifying the model}}\label{specifying-the-model}}
+\hypertarget{inspecting-and-plotting-the-best-fitting-estimated-models}{%
+\subsection{Inspecting and Plotting the best-fitting estimated models}\label{inspecting-and-plotting-the-best-fitting-estimated-models}}
 
-We passed the number of profiles, but no details about the model we wished to estimate.
+It appears that the four and five-class solutions fit best. Let's inspect those, specifically:
 
-There are a number of different types of models for LPA that estimate different parameters; choosing
-between them is an important consideration, and one that has a bearing on how we will interpret the results we saw above.
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{models_}\DecValTok{4}\NormalTok{_}\DecValTok{5}\NormalTok{ <-}\StringTok{ }\NormalTok{pisaUSA15[}\DecValTok{1}\OperatorTok{:}\DecValTok{100}\NormalTok{, ] }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{select}\NormalTok{(broad_interest, enjoyment, self_efficacy) }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{single_imputation}\NormalTok{() }\OperatorTok{%>%}
+\StringTok{  }\KeywordTok{estimate_profiles}\NormalTok{(}\DataTypeTok{models =} \DecValTok{1}\NormalTok{,}
+                    \DataTypeTok{n_profiles =} \KeywordTok{c}\NormalTok{(}\DecValTok{4}\NormalTok{, }\DecValTok{5}\NormalTok{))}
 
-Thus, in addition to the number of profiles (specified with the \texttt{n\_profiles}
-argument), the model can be specified in terms of whether and how the variable
-variances and covariances are estimated. The models are specified by passing arguments to the \texttt{variance} and \texttt{covariance}
-arguments. The possible values for these arguments are:
+\NormalTok{models_}\DecValTok{4}\NormalTok{_}\DecValTok{5} \OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{plot_profiles}\NormalTok{(}\DataTypeTok{add_line =} \OtherTok{TRUE}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}
+
+Here, it appears that the addition of a fifth class does not add a great deal relative to the four-class solution.
+Accordingly, we consider the four-class solution to be the solution we interpret and present, while heeding the recommendations about reporting which other models we estimated, how we chose to focus on the two best-fitting models we did, and why we presented the four-class, class-invariant set of estimates.
+
+Were we to wish to inspect \emph{only} that chosen model, we could estimate the model again, replacing \texttt{n\_profiles\ =\ c(4,\ 5)} with \texttt{n\_profiles\ =\ 4}.
+Alternatively, we could use list-indexing within R to index only that model.
+Note that the object \texttt{models\_4\_5} below includes the four and five-class estimated models, in that order.
+Using \texttt{{[}{[}1{]}{]}} indexes---selects---only the first of the two models, whereas \texttt{{[}{[}2{]}{]}} indexes the second.
+We index the four-class solution using \texttt{{[}{[}1{]}{]}}, and save it to a new object.
+
+Using this single estimated model, we can also create a faceted plot of density plots for an object of class \enquote{tidyLPA}.
+For each variable, a Total density plot will be shown, where cases are weighted by the posterior probability of being assigned to that class:
+
+\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1.pdf}
+
+\hypertarget{visualizing-and-examining-a-chosen-solution-further}{%
+\subsection{Visualizing and Examining a Chosen Solution Further}\label{visualizing-and-examining-a-chosen-solution-further}}
+
+Last, we can create a faceted plot of two-dimensional correlation plots and unidimensional density plots for a single set of model estimates:
+
+\includegraphics{paper_files/figure-latex/unnamed-chunk-11-1.pdf}
+
+We can use \texttt{get\_data()} to return data in wide format when applied to an object
+of class tidyProfile (one element of a tidyLPA object)
+
+\texttt{get\_data()} returns long format when applied to a
+tidyLPA object containing multiple tidyProfile analyses (because then the wide
+format does not make sense).
+
+To transform data in the wide format into the long format, the \texttt{gather()}
+function from the \textbf{tidyr} package can be used:
+
+\begin{verbatim}
+## # A tibble: 400 x 8
+##    model_number classes_number broad_interest enjoyment self_efficacy Class
+##           <dbl>          <dbl>          <dbl>     <dbl>         <dbl> <dbl>
+##  1            1              4            3.8       4            1        1
+##  2            1              4            3         3            2.75     2
+##  3            1              4            1.8       2.8          3.38     2
+##  4            1              4            1.4       1            2.75     3
+##  5            1              4            1.8       2.2          2        4
+##  6            1              4            1.6       1.6          1.88     4
+##  7            1              4            3         3.8          2.25     1
+##  8            1              4            2.6       2.2          2        4
+##  9            1              4            1         2.8          2.62     4
+## 10            1              4            2.2       2            1.75     4
+## # ... with 390 more rows, and 2 more variables: Class_prob <chr>,
+## #   Probability <dbl>
+\end{verbatim}
+
+\hypertarget{best-practices-for-latent-profile-analysis}{%
+\section{Best Practices for Latent Profile Analysis}\label{best-practices-for-latent-profile-analysis}}
+
+In addition to the pragmatic challenges facing the analyst carrying out LPA, there
+are a number of statistical issues that anaylsts face. Past research is informative in
+this respect; Bauer
+
+\hypertarget{recommendations-regarding-analytic-choices}{%
+\subsection{Recommendations Regarding Analytic Choices}\label{recommendations-regarding-analytic-choices}}
+
+A note of caution is warranted about LPA in the context of their characteristics
+and strengths. Bauer (2007) notes that many samples of data can be usefully
+broken down into profiles, and that the addition of profiles will likely be
+suggested for reasons other than the samples coming from more than one
+distribution (i.e., due to non-normality in the variables measured). Bauer also
+cautions that profiles should not be reified; that profiles do not necessarily
+exist outside of the analysis that they should be interpreted more as useful
+interpretative devices. These cautions suggest that, in general, parsimony,
+interpretability, and a general sense that the profiles are not necessarily
+real, but are rather helpful analytic tools, should be both priorities for the
+analyst and the reader of studies using this approach.
+
+The GRoLTs checklist, developed for a particular use of the general mixture
+model---a growth mixture model---is a helpful way to think about how best
+practices for LPA (Van De Schoot, Sijbrandij, Winter, Depaoli, \& Vermunt, 2017). We interpret them in light of LPA; note that
+the full checklist is not applicable (as, for example, guidelines about
+trajectories over time are not pertinent to LPA), and we only present those that
+are.
+
+First, here is the checklist:
+
+\hypertarget{grolts-checklist}{%
+\subsection{GRoLTs checklist}\label{grolts-checklist}}
+
+\hypertarget{htmlwidget-fffc7926638efd89a6a8}{}
+
+We walk through each of these in turn.
+
+\hypertarget{is-the-missing-data-mechanism-reported}{%
+\subsubsection{Is the missing data mechanism reported?}\label{is-the-missing-data-mechanism-reported}}
+
+As missing data can have a bearing on the estimates, an important consideration is whether or not the missing data is associated with the levels of the variables---or not.
+Consider consulting general guidelines on missing data, including Enders (2010), Little \& Rubin (2019), and Allison (2001).
+\#\#\# Is a description provided of what variables are related to attrition/missing data?
+
+Note whether any demographic or other variables are related to missing data.
+
+\hypertarget{is-a-description-provided-of-how-missing-data-in-the-analyses-were-dealt-with}{%
+\subsubsection{Is a description provided of how missing data in the analyses were dealt with?}\label{is-a-description-provided-of-how-missing-data-in-the-analyses-were-dealt-with}}
+
+Because LPA cannot accomodate any missing values (and so uses listwise deletion) for this recommendation, tidyLPA has a function for addressing missing data that are based upon single imputation (with two methods available, one using a contemporary, machine learning-based procedure).
+
+\hypertarget{is-information-about-the-distribution-of-the-observed-variables-included}{%
+\subsubsection{Is information about the distribution of the observed variables included?}\label{is-information-about-the-distribution-of-the-observed-variables-included}}
+
+The variables can be plotted and described prior to the analyses.
+While not built-in, we show later how this can be done using the ggplot2 R package, and the \texttt{describe()} function from the psych R package.
+
+\hypertarget{is-the-software-mentioned}{%
+\subsubsection{Is the software mentioned?}\label{is-the-software-mentioned}}
+
+The name and version of the software can be mentioned.
+Within R, a citation including the version can be generated for any R package, including tidyLPA, by using the \texttt{citation()} function (e.g., `citation(\enquote{tidyLPA})).
+
+\hypertarget{are-alternative-specifications-of-the-between-class-differences-in-variancecovariance-matrix-structure-considered-and-clearly-documented}{%
+\subsubsection{Are alternative specifications of the between-class differences in variance--covariance matrix structure considered and clearly documented?}\label{are-alternative-specifications-of-the-between-class-differences-in-variancecovariance-matrix-structure-considered-and-clearly-documented}}
+
+In LPA, we can determine which parameters - variance and covariance - are
+estimated to be the same or different across all profiles, and, in the case of
+covariance, whether it is fixed to zero. Of course, for the profiles we are most
+interested in, the mean is allowed to vary across profiles.
+
+In general, the approach to choosing the model is similar to choosing the number
+of profiles, requiring \textbf{deciding on the basis of evidence from multiple
+sources}, including information criteria, statistical tests, and concerns of
+interpretability and parsimony. The article by \href{https://www.sciencedirect.com/science/article/pii/S0361476X06000543}{Pastor and colleagues
+(2007)} has
+helpful information on the model specifications.
+
+Here, the six models that are possible to specify in LPA are described in terms of how the variables used to create the profiles are estimated. In tidyLPA, the models are specified by passing
+arguments to the \texttt{variance} and \texttt{covariance} arguments. The possible values for
+these arguments are:
 
@@
 
-In general, the approach to choosing the model is similar to choosing the number of profiles, requiring \textbf{deciding on the basis of evidence from multiple sources}, including information criteria, statistical tests, and concerns of interpretability and parsimony.
-The article by \href{https://www.sciencedirect.com/science/article/pii/S0361476X06000543}{Pastor and colleagues (2007)} has helpful information on the model specifications.
-Here, the six models that are possible to specify in LPA are described in terms of how the variables used to create the profiles are estimated.
+These models are described in Appendix A.
+
+\hypertarget{is-information-reported-about-the-number-of-random-start-values-and-final-iterations-included}{%
+\subsubsection{Is information reported about the number of random start values and final iterations included?}\label{is-information-reported-about-the-number-of-random-start-values-and-final-iterations-included}}
+
+MPlus uses random starts to initialize the estimation, whereas mclust (the R package) does not.
+Thus, for users of MPlus, the number of random starts and final iterations should be reported.
+By default, tidyLPA uses the MPlus defaults.
+Information on the number of random starts can be found in the model object returned by the tidyLPA function \texttt{estimate\_profiles()}.
+Different random starts can be passed by passing arguments to the \texttt{ANALYSIS} code for MPlus, e.g.:
+
+\hypertarget{are-the-model-comparison-and-selection-tools-described-from-a-statistical-perspective}{%
+\subsubsection{Are the model comparison (and selection) tools described from a statistical perspective?}\label{are-the-model-comparison-and-selection-tools-described-from-a-statistical-perspective}}
+
+In the case of choosing the number of profiles (and the specification of the
+model / profile solution), multiple criteria, such as the BIC or the proportion
+of variance explained are recommended for decision-making, but also
+interpretability in light of theory, parsimony, and evidence from
+cross-validation should be considered.
+
+We provide a number of functions for model comparison. The \texttt{compare\_solutions()} profile
+takes as an argument a particular measure of fit (defaulting to the BIC), and returns information
+for all of the models estimated. It also implements an ensemble, analytic hierarchiy approach, that uses weighted values of a number of fit indices; see (Akogul \& Erisoglu, 2017) for more.
+
+\hypertarget{are-the-total-number-of-fitted-models-reported-including-a-one-class-solution}{%
+\subsubsection{Are the total number of fitted models reported, including a one-class solution?}\label{are-the-total-number-of-fitted-models-reported-including-a-one-class-solution}}
+
+The number of fitted models can easily be reported based on the tidyLPA code run.
+For example, for the following code, models with 1-6 profiles, and the first (class invariant) and third (class invariant, unrestricted) model parameterizations are estimated:
+
+\hypertarget{are-the-number-of-cases-per-class-reported-for-each-model-absolute-sample-size-or-proportion}{%
+\subsubsection{Are the number of cases per class reported for each model (absolute sample size, or proportion)?}\label{are-the-number-of-cases-per-class-reported-for-each-model-absolute-sample-size-or-proportion}}
+
+Determining the number of cases per class can easily be done by using the \texttt{get\_data()} function from tidyLPA, passing the estimated model object from tidyLPA.; see the \texttt{Class} column (as well as the posterior probability estimates for every class for every observation).
+
+\hypertarget{are-all-of-the-estimates-plotted}{%
+\subsubsection{Are all of the estimates plotted?}\label{are-all-of-the-estimates-plotted}}
+
+This is a recommendation which can be difficult to achieve using existing software. tidyLPA provides two functions, \texttt{plot\_profiles()} and \texttt{plot\_bivariate()}, to plot the estimates.
+
+\hypertarget{are-the-raw-data-plotted-along-with-the-estimates}{%
+\subsubsection{Are the raw data plotted along with the estimates?}\label{are-the-raw-data-plotted-along-with-the-estimates}}
+
+The above-mentioned tidyLPA functions also plot the raw data with the estimates (by default).
+
+\hypertarget{are-characteristics-of-the-final-class-solution-numerically-described-i.e.-means-sdse-n-ci-etc.}{%
+\subsubsection{Are characteristics of the final class solution numerically described (i.e., means, SD/SE, n, CI, etc.)?}\label{are-characteristics-of-the-final-class-solution-numerically-described-i.e.-means-sdse-n-ci-etc.}}
+
+These characteristics can be obtained by using the \texttt{get\_estimates()} function, passing the estimated model object from tidyLPA.
+
+\hypertarget{are-the-syntax-files-available-either-in-the-appendix-supplementary-materials-or-from-the-authors}{%
+\subsubsection{Are the syntax files available (either in the appendix, supplementary materials, or from the authors)?}\label{are-the-syntax-files-available-either-in-the-appendix-supplementary-materials-or-from-the-authors}}
+
+The R files used for the analyses can easily be included in online repositories and supplementary online materials (as \texttt{.R} or \texttt{.Rmd} files).
+
+\hypertarget{summary-and-conclusion}{%
+\section{Summary and Conclusion}\label{summary-and-conclusion}}
+
+A step away from the tutorial side of this work, we wish to note some
+implications for the developers of statistical software for psychological
+scientists.
+
+First, we note that developing a tool that emphasized both rigor and access has
+been, so far, well-received. This suggests that there is an audience (or a
+market) for tools to carry out sophisticated tools that may be otherwise
+out-of-reach. Tools do not need to be hard to use, a point made by developers
+of, for example, tools to carry out highly-sophisticated Bayesian methods using
+the most cutting-edge tools (Bürkner \& others, 2017, Makowski, Ben-Shachar, Chen, and Lüdecke (2019)). We think that
+tidyLPA speaks to this, too.
+
+Another implication concerns how we developed tidyLPA. We used a modern software
+development workflow, including git/GitHub, a suite of tests, and continuous
+integration tools. We also note that we emphasized communicating that we were
+developing the software; when it was updated; and how we made major updates.
+
+There are a number of analytic choices that need to be made when carrying out
+person-oriented analyses. Because such person-oriented approaches are often more
+subjective (in practice) than other approaches (Linnenbrink-Garcia and
+Wormington, 2017), there is no one rule for determining the solution obtained.
+This solution is obtained on the basis of multiple decisions, such as the number
+of profiles selected or the modeling decisions such as what specific options are
+used for the cluster analysis (i.e., the distance metric used to calculate the
+similarity of the observations as part of the Ward's hierarchical clustering) or
+what parameters are estimated and how as part of LPA.
+
+Given the subjectivity involved, it is important that researchers be transparent
+and work as part of a team to obtain clustering solutions. Transparency about
+the design and analytic choices is important so that readers can appropriately
+interpret the report. Researchers can enhance transparency and reproducibility
+by sharing detailed descriptions of methodology and document it through the use
+of syntax (and, if possible, data) that we share with others. Working as part of
+a team can help to serve as a check on several of the choices researchers make,
+such as over-fitting or under-fitting the model to the data. Each decision
+depends on multiple factors and balancing tensions. We discuss each of the key
+decisions listed in an analysis.
+
+LPA can be used to evaluate how psychological constructs are experienced (and
+can be analyzed) together and at once. Though described in contrast to a
+variable-centered approach, scholars have pointed out how person-oriented
+approaches, including LPA, are complementary to variable-centered analyses
+(Marsh, Ludtke, Trautwein, \& Morin, 2009). Our aim was to describe how
+researchers may get started in an informed way as researchers seek to understand
+how individuals interact, behave, and learn in ways that embraces the complexity
+of these experiences.
+
+\hypertarget{appendix-a}{%
+\section{Appendix A}\label{appendix-a}}
 
@@
 
-If no values are specified for these, then the variances are constrained to be
-equal across classes, and covariances are fixed to 0 (conditional independence
-of the indicators).
-
-These arguments allow for four models to be specified:
-
-\begin{itemize}
-\tightlist
-\item
-  Equal variances and covariances fixed to 0 (Model 1)
-\item
-  Varying variances and covariances fixed to 0 (Model 2)
-\item
-  Equal variances and equal covariances (Model 3)
-\item
-  Varying variances and varying covariances (Model 6)
-\end{itemize}
-
-Two additional models (Models 4 and 5) can be fit using MPlus. More information
-on the models can be found in the
-\href{https://data-edu.github.io/tidyLPA/articles/Introduction_to_tidyLPA.html}{vignette}.
-
-\hypertarget{accessing-values-in-the-output}{%
-\subsection{Accessing values in the output}\label{accessing-values-in-the-output}}
-
-A few helper functions are available to make it easier to work with the output
-of an analysis.
-
-\texttt{get\_data()} returns the data:
-
-We note that \texttt{get\_data()} returns data in wide format when applied to an object
-of class tidyProfile (one element of a tidyLPA object), or when applied to a
-tidyLPA object of length one. \texttt{get\_data()} returns long format when applied to a
-tidyLPA object containing multiple tidyProfile analyses (because then the wide
-format does not make sense).
-
-To transform data in the wide format into the long format, the \texttt{gather()}
-function from the \textbf{tidyr} package can be used, e.g.:
-
-\hypertarget{implications}{%
-\section{Implications}\label{implications}}
-
-A step away from the tutorial side of this work, we wish to note some implications
-for the developers of statistical software for psychological scientists.
-
-First, we note that developing a tool that emphasized both rigor and access has
-been, so far, well-received. This suggests that there is an audience (or a
-market) for tools to carry out sophisticated tools that may be otherwise
-out-of-reach. Tools do not need to be hard to use, a point made by developers
-of, for example, tools to carry out highly-sophisticated Bayesian methods using
-the most cutting-edge tools (Bürkner \& others, 2017, Makowski, Ben-Shachar, Chen, and Lüdecke (2019)). We think
-that tidyLPA speaks to this, too.
-
-Another implication concerns how we developed tidyLPA. We used a modern software
-development workflow, including git/GitHub, a suite of tests, and continuous
-integration tools. We also note that we emphasized communicating that we were
-developing the software; when it was updated; and how we made major updates.
-
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
-
-Person-oriented analysis ia way to consider how psychological constructs are
-experienced (and can be analyzed) together and at once. Though described in
-contrast to a variable-centered approach, scholars have pointed out how
-person-oriented approaches are complementary to variable-centered analyses
-(Marsh, Ludtke, Trautwein, \& Morin, 2009). A person-oriented approach can help
-us to consider multiple variables together and at once and in a more dynamic
-way, reflected in the methodological approaches for cluster analysis and LPA
-that identify profiles of individuals responses.
-
-This manuscript provided an outline of how to get started with person-oriented
-analyses in an informed way. We provided a general overview of the methodology
-and described tools to carry out such an analysis. We also described specific
-tools, emphasizing freely-available open-source options that we have developed.
-Because of the inherently exploratory nature of person-oriented analysis,
-carrying out the analysis in a trustworthy and open way is particularly
-important. In this way, the interpretative aspect of settling on a solution
-shares some features of quantitative and qualitative research: The systematic
-nature of quantitative research methods (focused upon agreed-upon criteria such
-as likelihood-ratio tests) and qualitative research methods (focused upon the
-trustworthiness of both the analysis and the analyst) are important to consider
-when carrying out person-oriented analysis. Lastly, we made some general
-recommendations for future directions--and also highlighted some situations for
-which person-oriented approaches may not be the best and some cautions raised in
-past research regarding how such approaches are used.
-
-In conclusion, as use of person-oriented approaches expand, new questions and
-opportunities for carrying out research in a more holistic, dynamic way will be
-presented. Analyzing constructs together and at once is appealing to
-researchers, particularly those carrying out research in fields such as
-education for which communicating findings to stakeholders in a way that has the
-chance to impact practice is important. Our aim was not to suggest that such an
-approach is always the goal or should always be carried out, but rather to
-describe how researchers may get started in an informed way as researchers seek
-to understand how individuals interact, behave, and learn in ways that embraces
-the complexity of these experiences.
-
-\newpage
-
 \hypertarget{references}{%
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-akogul2017approach}{}%
+Akogul, S., \& Erisoglu, M. (2017). An approach for determining the number of clusters in a model-based cluster analysis. \emph{Entropy}, \emph{19}(9), 452.
+
+\leavevmode\hypertarget{ref-allison2001missing}{}%
+Allison, P. D. (2001). \emph{Missing data} (Vol. 136). Sage publications.
+
 \leavevmode\hypertarget{ref-burkner2017}{}%
 Bürkner, P.-C., \& others. (2017). Brms: An r package for bayesian multilevel models using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.
+
+\leavevmode\hypertarget{ref-enders2010applied}{}%
+Enders, C. K. (2010). \emph{Applied missing data analysis}. Guilford press.
+
+\leavevmode\hypertarget{ref-little2019statistical}{}%
+Little, R. J., \& Rubin, D. B. (2019). \emph{Statistical analysis with missing data} (Vol. 793). John Wiley \& Sons.
 
@@
 
+\leavevmode\hypertarget{ref-van2017grolts}{}%
+Van De Schoot, R., Sijbrandij, M., Winter, S.
+D., Depaoli, S., \& Vermunt, J.
+K.
+(2017).
+The grolts-checklist: Guidelines for reporting on latent trajectory studies.
+\emph{Structural Equation Modeling: A Multidisciplinary Journal}, \emph{24}(3), 451--467.
+
+\leavevmode\hypertarget{ref-akogul2017approach}{}%
+Akogul, S., \& Erisoglu, M. (2017). An approach for determining the number of clusters in a model-based cluster analysis. \emph{Entropy}, \emph{19}(9), 452.
+
+\leavevmode\hypertarget{ref-allison2001missing}{}%
+Allison, P. D. (2001). \emph{Missing data} (Vol. 136). Sage publications.
+
 \leavevmode\hypertarget{ref-burkner2017}{}%
 Bürkner, P.-C., \& others. (2017). Brms: An r package for bayesian multilevel models using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.
+
+\leavevmode\hypertarget{ref-enders2010applied}{}%
+Enders, C. K. (2010). \emph{Applied missing data analysis}. Guilford press.
+
+\leavevmode\hypertarget{ref-little2019statistical}{}%
+Little, R. J., \& Rubin, D. B. (2019). \emph{Statistical analysis with missing data} (Vol. 793). John Wiley \& Sons.
 
@@
 
+\leavevmode\hypertarget{ref-van2017grolts}{}%
+Van De Schoot, R., Sijbrandij, M., Winter, S.
+D., Depaoli, S., \& Vermunt, J.
+K.
+(2017).
+The grolts-checklist: Guidelines for reporting on latent trajectory studies.
+\emph{Structural Equation Modeling: A Multidisciplinary Journal}, \emph{24}(3), 451--467.
+
 \endgroup

### 2020-12-29 paper.tex
clean up repo

@@
 Here, we provide a walkthrough of an analysis using the tidyLPA R package.
-
-\hypertarget{installation}{%
-\subsection{Installation}\label{installation}}
+\#\# Installation
 
@@
 
-Here, we use a simulated dataset contains data on broad (general, and not situation- or topic-specific) interest, sense of enjoyment, and
-self-efficacy from the PISA assessment.
-
+Here, we use a simulated dataset contains data on broad (general, and not situation- or topic-specific) interest, sense of enjoyment, and self-efficacy from the PISA assessment.
 First, let's explore the data graphically using the \{ggplot2\} package:
@@
 \includegraphics{paper_files/figure-latex/unnamed-chunk-6-1.pdf}
-
 We see that the data broadly \emph{appears} to be normally distributed, as the following statistics from the use of the psych package suggest:
@@
 as the CAIC, SABIC, and ICL), and the bootstrapped LRT test statistic, can be
-obtained with the \texttt{get\_fit()} function:
+obtained with the \texttt{get\_fit()} function (processed with the tidyr R package to be in \enquote{long} rather than \enquote{wide} format):
 
@@
 \begin{Highlighting}[]
+\KeywordTok{library}\NormalTok{(tidyr)}
+
 \NormalTok{pisaUSA15 }\OperatorTok{%>%}
@@
 \StringTok{  }\KeywordTok{estimate_profiles}\NormalTok{(}\DecValTok{3}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
-\StringTok{  }\KeywordTok{get_fit}\NormalTok{(m1)}
+\StringTok{  }\KeywordTok{get_fit}\NormalTok{(m1) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\KeywordTok{gather}\NormalTok{(stat, val)}
 \end{Highlighting}
@@
 \begin{verbatim}
-## # A tibble: 1 x 18
-##   Model Classes  LogLik    AIC    AWE    BIC   CAIC    CLC    KIC  SABIC     ICL
-##   <dbl>   <dbl>   <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>   <dbl>
-## 1     1       3 -17399. 34826. 35080. 34919. 34933. 34799. 34843. 34874. -36442.
-## # ... with 7 more variables: Entropy <dbl>, prob_min <dbl>, prob_max <dbl>,
-## #   n_min <dbl>, n_max <dbl>, BLRT_val <dbl>, BLRT_p <dbl>
+## # A tibble: 18 x 2
+##    stat              val
+##    <chr>           <dbl>
+##  1 Model         1      
+##  2 Classes       3      
+##  3 LogLik   -17399.     
+##  4 AIC       34826.     
+##  5 AWE       35080.     
+##  6 BIC       34919.     
+##  7 CAIC      34933.     
+##  8 CLC       34799.     
+##  9 KIC       34843.     
+## 10 SABIC     34874.     
+## 11 ICL      -36442.     
+## 12 Entropy       0.744  
+## 13 prob_min      0.750  
+## 14 prob_max      0.936  
+## 15 n_min         0.161  
+## 16 n_max         0.669  
+## 17 BLRT_val    649.     
+## 18 BLRT_p        0.00990
 \end{verbatim}
@@
 Last, we can create a faceted plot of two-dimensional correlation plots and unidimensional density plots for a single set of model estimates:
-
 \includegraphics{paper_files/figure-latex/unnamed-chunk-11-1.pdf}
-
 We can use \texttt{get\_data()} to return data in wide format when applied to an object
@@
 function from the \textbf{tidyr} package can be used:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\KeywordTok{get_data}\NormalTok{(model_}\DecValTok{4}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\NormalTok{tidyr}\OperatorTok{::}\KeywordTok{gather}\NormalTok{(Class_prob, Probability, }\KeywordTok{contains}\NormalTok{(}\StringTok{"CPROB"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
 

### 2020-12-29 paper.rmd
clean up repo
fix typos

@@
 ---
-title             : "Decisions Decisions Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package"
-shorttitle        : "LPA Best Practices"
+title             : "Decisions. Decisions, Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package"
+shorttitle        : "Using tidyLPA in accordance with best practices"
 
@@
     affiliation   : "2"
+  - name          : "Patrick Beymer"
+    affiliation   : "5"
   - name          : "Jennifer Schmidt"
     affiliation   : "3"
-  - name          : "Patrick Beymer"
-    affiliation   : "5"
   - name          : "Daniel Anderson"
@@
 keywords          : "Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust"
-wordcount         : "suggested to be no more than 3,000"
+wordcount         : "5,385"
 
@@
 
-In statistics classes, textbooks, and workshops, an example like the following
-is common:
+In statistics classes and workshops, an example like the following
+is commonly used:
 
@@
 for which all of the variables for which (relatively) homogeneous groups are
-identified from among a (relatively) heterogeneous sample are continuos, *latent
+identified from among a (relatively) heterogeneous sample are continuous, *latent
 profile analysis* (LPA). After describing the method and some examples of its
@@
 associated with each profile (or mixture component) for each observation in the
-dataset.
+data set.
 
@@
 software to MPlus. The functions in tidyLPA that use mclust have been
-benchmarked to MPlus for a series of simple models (with small datasets and for
+benchmarked to MPlus for a series of simple models (with small data sets and for
 models with small numbers of profiles. This R Markdown output contains
@@
 benchmark. Note that most of the output is identical, though there are some
-differences in the hundreths decimal places for some. Because of differences in
+differences in the hundredths decimal places for some. Because of differences in
 settings for the EM algorithm and particularly for the start values (random
@@
 common names for the models (i.e., class-varying unrestricted), and the
-covariance matrix associated with the parameterizationfor the six models that
+covariance matrix associated with the parameterization for the six models that
 are possible to be estimated.
@@
 You can also install the development version of tidyLPA, which may have new
-features that are not yet availalble on CRAN (because the version on CRAN is
+features that are not yet available on CRAN (because the version on CRAN is
 periodically updated) from GitHub with:
@@
 
-Here, we use a simulated dataset contains data on broad (general, and not situation- or topic-specific) interest, sense of enjoyment, and self-efficacy from the PISA assessment.
+Here, we use a simulated data set contains data on broad (general, and not situation- or topic-specific) interest, sense of enjoyment, and self-efficacy from the PISA assessment.
 First, let's explore the data graphically using the {ggplot2} package:
@@
 the log-likelihood will always decrease with the addition of additional
-paremeters estimated by the model, whereas the AIC and BIC will account for the
+parameters estimated by the model, whereas the AIC and BIC will account for the
 number of parameters; lower values of either indicate better fit.
@@
 In addition to the pragmatic challenges facing the analyst carrying out LPA, there 
-are a number of statistical issues that anaylsts face. Past research is informative in
+are a number of statistical issues that analysts face. Past research is informative in
 this respect; Bauer
@@
 tp <- t %>% filter(use == 1) %>% mutate(`LPA Number` = 1:nrow(.)) %>%
-  select( `LPA Number`, `GRoLTS Number` = Number, `Checklist Item`)
+  select( `LPA Number`, `GRoLTs Number` = Number, `Checklist Item`)
 
@@
 
-Because LPA cannot accomodate any missing values (and so uses listwise deletion) for this recommendation, tidyLPA has a function for addressing missing data that are based upon single imputation (with two methods available, one using a contemporary, machine learning-based procedure).
+Because LPA cannot accommodate any missing values (and so uses listwise deletion) for this recommendation, tidyLPA has a function for addressing missing data that are based upon single imputation (with two methods available, one using a contemporary, machine learning-based procedure).
 
@@
 takes as an argument a particular measure of fit (defaulting to the BIC), and returns information
-for all of the models estimated. It also implements an ensemble, analytic hierarchiy approach, that uses weighted values of a number of fit indices; see [@akogul2017approach] for more.
+for all of the models estimated. It also implements an ensemble, analytic hierarchy approach, that uses weighted values of a number of fit indices; see [@akogul2017approach] for more.
 
@@
 # References
-```{r create_r-references}
-r_refs(file = "r-references.bib")
-```
+
+\newpage
 

### 2020-12-29 paper.tex
fix typos

@@
 %
-\documentclass[man]{apa6}
+\documentclass[english,man]{apa6}
 \usepackage{lmodern}
@@
 \hypersetup{
-            pdftitle={Decisions Decisions Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package},
-            pdfauthor={Joshua Rosenberg, Caspar van Lissa, Jennifer Schmidt, Patrick Beymer, Daniel Anderson, \& Matthew Schell},
+            pdftitle={Decisions. Decisions, Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package},
+            pdfauthor={Joshua Rosenberg1, Caspar van Lissa2, Patrick Beymer5, Jennifer Schmidt3, Daniel Anderson4, \& Matthew Schell3},
             pdfkeywords={Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust},
@@
 \setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
 
@@
 
-\shorttitle{LPA Best Practices}
-\affiliation{
-\vspace{0.5cm}
-\textsuperscript{1} University of Tennessee, Knoxville\\\textsuperscript{2} Utrecht University\\\textsuperscript{3} Michigan State University\\\textsuperscript{3} University of Oregon\\\textsuperscript{4} University of Wisconsin, Madison}
-\keywords{Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust\newline\indent Word count: suggested to be no more than 3,000}
-\usepackage{csquotes}
+% Manuscript styling
 \usepackage{upgreek}
@@
 
+% \setlength{\parindent}{0.5in}
+% \setlength{\parskip}{0pt plus 0pt minus 0pt}
+
+% \usepackage{etoolbox}
+\makeatletter
+\patchcmd{\HyOrg@maketitle}
+  {\section{\normalfont\normalsize\abstractname}}
+  {\section*{\normalfont\normalsize\abstractname}}
+  {}{\typeout{Failed to patch abstract.}}
+\patchcmd{\HyOrg@maketitle}
+  {\section{\protect\normalfont{\@title}}}
+  {\section*{\protect\normalfont{\@title}}}
+  {}{\typeout{Failed to patch title.}}
+\makeatother
+\shorttitle{Using tidyLPA in accordance with best practices}
+\keywords{Latent Profile Analysis, mixture models, finite mixture models, tutorial, R, MPlus, mclust\newline\indent Word count: 5,385}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 \makeatother
+\usepackage{csquotes}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[shorthands=off,main=english]{babel}
+\else
+  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
+  \usepackage{polyglossia}
+  \setmainlanguage[]{english}
+\fi
 
-\title{Decisions Decisions Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package}
-\author{Joshua Rosenberg\textsuperscript{1}, Caspar van Lissa\textsuperscript{2}, Jennifer Schmidt\textsuperscript{3}, Patrick Beymer\textsuperscript{5}, Daniel Anderson\textsuperscript{4}, \& Matthew Schell\textsuperscript{3}}
+\title{Decisions. Decisions, Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package}
+\author{Joshua Rosenberg\textsuperscript{1}, Caspar van Lissa\textsuperscript{2}, Patrick Beymer\textsuperscript{5}, Jennifer Schmidt\textsuperscript{3}, Daniel Anderson\textsuperscript{4}, \& Matthew Schell\textsuperscript{3}}
 \date{}
 
-\authornote{
-
-Correspondence concerning this article should be addressed to Joshua Rosenberg, 1122 Volunteer Blvd., Knoxville, TN, 37996. E-mail: \href{mailto:jmrosenberg@utk.edu}{\nolinkurl{jmrosenberg@utk.edu}}}
-
-\abstract{
-
-}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} University of Tennessee, Knoxville\\\textsuperscript{2} Utrecht University\\\textsuperscript{3} Michigan State University\\\textsuperscript{3} University of Oregon\\\textsuperscript{4} University of Wisconsin, Madison}
 
@@
 for which all of the variables for which (relatively) homogeneous groups are
-identified from among a (relatively) heterogeneous sample are continuos, \emph{latent
+identified from among a (relatively) heterogeneous sample are continuous, \emph{latent
 profile analysis} (LPA). After describing the method and some examples of its
@@
 
-Here, we use a simulated dataset contains data on broad (general, and not situation- or topic-specific) interest, sense of enjoyment, and self-efficacy from the PISA assessment.
+Here, we use a simulated data set contains data on broad (general, and not situation- or topic-specific) interest, sense of enjoyment, and self-efficacy from the PISA assessment.
 First, let's explore the data graphically using the \{ggplot2\} package:
@@
 In addition to the pragmatic challenges facing the analyst carrying out LPA, there
-are a number of statistical issues that anaylsts face. Past research is informative in
+are a number of statistical issues that analysts face. Past research is informative in
 this respect; Bauer
@@
 As missing data can have a bearing on the estimates, an important consideration is whether or not the missing data is associated with the levels of the variables---or not.
-Consider consulting general guidelines on missing data, including Enders (2010), Little \& Rubin (2019), and Allison (2001).
+Consider consulting general guidelines on missing data, including Enders (2010), Little and Rubin (2019), and Allison (2001).
 \#\#\# Is a description provided of what variables are related to attrition/missing data?
@@
 takes as an argument a particular measure of fit (defaulting to the BIC), and returns information
-for all of the models estimated. It also implements an ensemble, analytic hierarchiy approach, that uses weighted values of a number of fit indices; see (Akogul \& Erisoglu, 2017) for more.
+for all of the models estimated. It also implements an ensemble, analytic hierarchy approach, that uses weighted values of a number of fit indices; see (Akogul \& Erisoglu, 2017) for more.
 
@@
 of, for example, tools to carry out highly-sophisticated Bayesian methods using
-the most cutting-edge tools (Bürkner \& others, 2017, Makowski, Ben-Shachar, Chen, and Lüdecke (2019)). We think that
+the most cutting-edge tools (Bürkner \& others, 2017, @makowski2019). We think that
 tidyLPA speaks to this, too.
@@
 
+\newpage
+
 \begingroup
@@
 \emph{Journal of Open Source Software}, \emph{3}(30), 978.
-doi:\href{https://doi.org/10.21105/joss.00978}{10.21105/joss.00978}
-
-\leavevmode\hypertarget{ref-van2017grolts}{}%
-Van De Schoot, R., Sijbrandij, M., Winter, S.
-D., Depaoli, S., \& Vermunt, J.
-K.
-(2017).
-The grolts-checklist: Guidelines for reporting on latent trajectory studies.
-\emph{Structural Equation Modeling: A Multidisciplinary Journal}, \emph{24}(3), 451--467.
-
-\leavevmode\hypertarget{ref-akogul2017approach}{}%
-Akogul, S., \& Erisoglu, M. (2017). An approach for determining the number of clusters in a model-based cluster analysis. \emph{Entropy}, \emph{19}(9), 452.
-
-\leavevmode\hypertarget{ref-allison2001missing}{}%
-Allison, P. D. (2001). \emph{Missing data} (Vol. 136). Sage publications.
-
-\leavevmode\hypertarget{ref-burkner2017}{}%
-Bürkner, P.-C., \& others. (2017). Brms: An r package for bayesian multilevel models using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.
-
-\leavevmode\hypertarget{ref-enders2010applied}{}%
-Enders, C. K. (2010). \emph{Applied missing data analysis}. Guilford press.
-
-\leavevmode\hypertarget{ref-little2019statistical}{}%
-Little, R. J., \& Rubin, D. B. (2019). \emph{Statistical analysis with missing data} (Vol. 793). John Wiley \& Sons.
-
-\leavevmode\hypertarget{ref-makowski2019}{}%
-Makowski, D., Ben-Shachar, M. S., Chen, S., \& Lüdecke, D. (2019). Indices of effect existence and significance in the bayesian framework. \emph{Frontiers in Psychology}, \emph{10}, 2767.
-
-\leavevmode\hypertarget{ref-R-tidyLPA}{}%
-Rosenberg, J.
-M., Beymer, P.
-N., Anderson, D.
-J., Lissa, C.
-van, \& Schmidt, J.
-A.
-(2018).
-TidyLPA: An r package to easily carry out latent profile analysis (lpa) using open-source or commercial software.
-\emph{Journal of Open Source Software}, \emph{3}(30), 978.
-doi:\href{https://doi.org/10.21105/joss.00978}{10.21105/joss.00978}
+\url{https://doi.org/10.21105/joss.00978}
 
